### 2018-08-09 docs/scheduler.tex

\documentclass[]{article}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{hyperref}

\newcommand{\mc}{\mathcal}
\newcommand{\bb}{\mathbb}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

%opening
\title{Application of Network Flow and Linear Programming to Scheduling of Clinicians}
\author{David Landsman}

\begin{document}
\maketitle

\section{Problem} \label{problem}
A group of clinicians are working on-call at a clinic each day of the year, including weekends and holidays, at multiple divisions. When a clinician is assigned to work during the week, they work from 8 A.M. on Monday to 5 P.M. on Friday. During weekends, a clinician works the complement period of time, that is, Friday 5 P.M. to Monday 8 A.M. Each clinician can request not to work certain weeks and weekends of the year. We refer to this as the clinician's \textit{time-off}.

\subsection{Constraints} \label{constraints}
We formalize the constraints of the problem to be able to describe them mathematically later on.
	\begin{enumerate}
		\item Each division needs to have one and only one clinician that covers a every block
		\item Every weekend needs to have one and only one clinician that covers it
		\item For each division, each clinician can only work between the minimum and maximum number of weeks they are allowed (note: these may be different for different divisions)
		\item Each clinician should work a given number of long weekends. Roughly, this is the number of long weekends divided by the number of clinicians
		\item A clinician cannot work two consecutive blocks (either in the same division, or in different divisions)
		\item A clinician cannot work two consecutive weekends
	\end{enumerate}

\subsection{Objectives} \label{objectives}
Since it is not always possible to accommodate all time-off requests from all clinicians, we decided to use this metric as an objective instead of a constraint. 
	\begin{enumerate}
		\item Minimize the number of blocks assigned to a clinician that they requested as time-off
		\item Minimize the number of weekends assigned to a clinician that they requested as time-off
		\item Maximize the adjacency between blocks and weekends assigned for a given clinician. In particular, if a clinician is assigned to work a given block in a given division, we will prefer it if the clinician also works in the weekend between the weeks of the block.
	\end{enumerate}

\section{Linear Programming Formulation}
The problem described in \ref{problem} corresponds to an Assignment Problem between clinicians and blocks/weekends. Such a problem can be presented as an Integer Linear Program (ILP).  \\ \\
We begin by defining the sets, constants, and indices that will be used in this formulation.
	\begin{align*}
		\mc{D}\coloneqq &\text{ the set of all divisions, indexed by } i \\
		\mc{C}\coloneqq &\text{ the set of all clinicians, indexed by } j \\
		\mc{B}\coloneqq &\text{ the set of all blocks, indexed by } k \\
		\mc{W}\coloneqq &\text{ the set of all weekends, indexed by } l \\
		\mc{W} \supset \mc{L}\coloneqq &\text{ the set of all long weekends} \\
		\mc{B} \supset \mc{S}_j\coloneqq &\text{ the set of blocks clinician } j \text{ requested off} \\
		\mc{W} \supset \mc{T}_j\coloneqq &\text{ the set of weekends clinician } j \text{ requested off} \\
		m^i_j \coloneqq &\text{ the minimum number of blocks clinician } j \text{ should work in division } i \\
		M^i_j \coloneqq &\text{ the maximum number of blocks clinician } j \text{ should work in division } i
	\end{align*}
We also define the set of 0-1 variables $X^i_{j,k}$ and $Y_{j,l}$ that we will solve for in the ILP.
	\begin{alignat}{2}
		&X^i_{j,k} \in \{0, 1\}: &&\text{ clinician } j \text{ covers block } k \text{ for division } i \\
		&Y_{j, l} \in \{0, 1\}: &&\text{ clinician } j \text{ covers weekend } l
	\end{alignat}
Now we can write the constraints in \ref{constraints} mathematically.
	\begin{enumerate}
		\item 
			\begin{equation}
				\sum_j X^i_{j, k} = 1 \text{ for each } i, k
			\end{equation}
		\item
			\begin{equation}
				\sum_j Y_{j, l} = 1 \text{ for each } l
			\end{equation}
		\item 
			\begin{equation}
				m^i_j \leq \sum_{k} X^i_{j, k} \leq M^i_j \text{ for each } i, j
			\end{equation}
		\item
			\begin{equation}
				\floor*{\frac{|\mc{L}|}{|\mc{C}|}} \leq \sum_{l \in \mc{L}} Y_{j, l} \leq \ceil*{\frac{|\mc{L}|}{|\mc{C}|}}
			\end{equation}
		\item
			\begin{equation}
				\sum_i \left(X^i_{j, k} + X^i_{j, k+1}\right) \leq 1 \text{ for each } j,k \text{ where } k \leq |\mc{B}| - 1
			\end{equation}
		\item
			\begin{equation}
				Y_{j, l} + Y_{j, l+1} \leq 1 \text{ for each } j, l \text{ where } l \leq |\mc{W}| - 1
			\end{equation}
	\end{enumerate}
Note that in ILP we can only have a single objective function. Since we have multiple objectives in \ref{objectives}, we will need to combine them in an appropriate manner. Assuming each objective can be summarized as a linear function of the variables defined above, say $Obj_n\left(X^i_{j, k}, Y_{j, l}\right)$, $1 \leq n \leq N$, then we shall define the overall objective function as a linear combination of $Obj_n$'s:
	\begin{equation}
		Obj\left(X^i_{j, k}, Y_{j, l}\right) \coloneqq \frac{Obj_1 + \ldots + Obj_N}{N} 
	\end{equation}
This ensures that all objectives have an equal weight of $\frac{1}{N}$. It is also necessary to choose whether we minimize or maximize $Obj$, so we will need to convert the minimization objectives. In particular, we can define the following objectives:
	\begin{enumerate}
		\item We will maximize the number of blocks assigned to a clinician that are \textit{not} requested for time-off:
			\begin{equation}
				Obj_1\left(X^i_{j, k}, Y_{j, l}\right) \coloneqq \sum_i \sum_j \sum_k
					\begin{cases}
					X^i_{j, k} &\text{ if } k \in \mc{S}_j \\
					-X^i_{j, k} &\text{ otherwise }
					\end{cases}
			\end{equation}
		\item We will maximize the number of weekends assigned to a clinician that are \textit{not} requested for time-off
			\begin{equation}
				Obj_2\left(X^i_{j, k}, Y_{j, l}\right) \coloneqq \sum_j \sum_l 
					\begin{cases}
					Y_{j, l} &\text{ if } l \in \mc{T}_j \\
					-Y_{j, l} &\text{ otherwise }
					\end{cases}
			\end{equation}
		\item We want to maximize the number of adjacent blocks and weekends. Since our variables are 0-1, we can accomplish this by maximizing the product of $X^i_{j, k} \cdot Y_{j, l}$ where $k$ and $l$ are adjacent. When both variables are assigned a value of 1, they contribute a single unit to the objective. Otherwise, they contribute nothing. \\
		However, this objective is \textit{not} linear. To make it linear, we introduce a set of helper variables:
			\begin{equation}
				Z^i_{j, k} \in \{0, 1\}: \text{represents the product }X^i_{j, k} \cdot Y_{j, 2k+1}
			\end{equation}
		with constraints:
			\begin{align}
				Z^i_{j, k} &\leq X^i_{j, k} \\
				Z^i_{j, k} &\leq Y_{j, 2k+1} \text{ for each } i
			\end{align}
		This gives us the third objective:
			\begin{equation}
				Obj_3\left(Z^i_{j, k}\right) \coloneqq \sum_i \sum_j \sum_k Z^i_{j, k}
			\end{equation}
	\end{enumerate}
	Putting all the components above together, we get a linear program defined as:
		\begin{equation}
			\begin{split}
			&\text{maximize } \frac{Obj_1 + Obj_2 + Obj_3}{3} \\
			&\text{subject to } (3)\text{-}(8)
			\end{split}
		\end{equation}
\end{document}

### 2019-01-02 docs/scheduler.tex
Cleanup notation in latex writeup

@@
 \usepackage{hyperref}
+\usepackage{bbm}
 
@@
 
+\section{Linear Programming Approach}
+We will attempt to solve \ref{problem} using an integer linear program.
+The assignment of a clinician from a certain division to a certain week or weekend is captured in a 0-1 variable.
+Constraints on the assignments are captured using (in-)equalities.
+
+\subsection{Constants}
+	\begin{align*}
+		\{1, \ldots, D\} : &\text{ the set of all divisions, indexed by } d \\
+		\{1, \ldots, C\} : &\text{ the set of all clinicians, indexed by } c \\
+		\{1, \ldots, B\} : &\text{ the set of all blocks, indexed by } b \\
+		\{1, \ldots, W\} : &\text{ the set of all weekends, indexed by } w \\
+		\{1, \ldots, W\} \supset \mc{L} : &\text{ the set of all long weekends} \\
+		\{1, \ldots, B\} \supset\mc{S}_c : &\text{ the set of blocks clinician $c$ requested off} \\
+		\{1, \ldots, W\} \supset\mc{T}_c : &\text{ the set of weekends clinician $c$ requested off} \\
+		m_{c, d} : &\text{ the minimum number of blocks clinician $c$ should work in division $d$} \\
+		M_{c, d} : &\text{ the maximum number of blocks clinician $c$ should work in division $d$}
+	\end{align*}
+
+\subsection{Variables}
+	\begin{align}
+		X_{c, b, d} \in \{0, 1\}: &\text{ clinician $c$ covers block $b$ for division $d$} \\
+		Y_{c, w} \in \{0, 1\}: &\text{ clinician $c$ covers weekend $w$}
+	\end{align}
+Note that weekends are not distinguished according to division, since the clinician assigned covers all divisions.
+
 \subsection{Constraints} \label{constraints}
-We formalize the constraints of the problem to be able to describe them mathematically later on.
-	\begin{enumerate}
-		\item Each division needs to have one and only one clinician that covers a every block
-		\item Every weekend needs to have one and only one clinician that covers it
-		\item For each division, each clinician can only work between the minimum and maximum number of weeks they are allowed (note: these may be different for different divisions)
-		\item Each clinician should work a given number of long weekends. Roughly, this is the number of long weekends divided by the number of clinicians
-		\item A clinician cannot work two consecutive blocks (either in the same division, or in different divisions)
+	\begin{itemize}
+		\item Each division needs to have exactly one clinician that covers any given block
+			\begin{equation}
+				\sum_{c} X_{c, b, d} = 1 \text{ for each } b, d
+			\end{equation}
+		\item Every weekend needs to have exactly one clinician that covers it
+			\begin{equation}
+				\sum_{c} Y_{c, w} = 1 \text{ for each } w
+			\end{equation}
+		\item In a given division, each clinician can only work between the minimum and maximum number of weeks they are allowed (note: these may be different for different divisions)
+			\begin{equation}
+				m_{c, d} \leq \sum_{d} X_{c, b, d} \leq M_{c, d} \text{ for each } c, d
+			\end{equation}			
+		\item Long weekends should be roughly equally distributed between clinicians.
+			\begin{equation}
+				\floor*{\frac{|\mc{L}|}{C}} \leq \sum_{w \in \mc{L}} Y_{c, w} \leq \ceil*{\frac{|\mc{L}|}{C}} \text{ for each } c
+			\end{equation}
+		Each clinician should work a given number of long weekends. Roughly, this is the number of long weekends divided by the number of clinicians
+		\item A clinician cannot work two consecutive blocks (across all divisions)
+			\begin{equation}
+				\sum_d \left(X_{c, b, d} + X_{c, b+1, d}\right) \leq 1 \text{ for each $c$ and each $b \leq B - 1$}
+			\end{equation}
 		\item A clinician cannot work two consecutive weekends
-	\end{enumerate}
+			\begin{equation}
+				Y_{c, w} + Y_{c, w+1} \leq 1 \text{ for each $c$ and each $w \leq W - 1$}
+			\end{equation}
+	\end{itemize}
 
 \subsection{Objectives} \label{objectives}
-Since it is not always possible to accommodate all time-off requests from all clinicians, we decided to use this metric as an objective instead of a constraint. 
-	\begin{enumerate}
+Our goal in the linear program is to accommodate clinician time-off requests as much as possible. Moreover, we try to assign weekends close to blocks, if possible.
+	\begin{itemize}
 		\item Minimize the number of blocks assigned to a clinician that they requested as time-off
+			\begin{equation}
+				\label{eqn_block_obj}
+				\min \sum_{c, b, d} (-1)^{ I(b \in \mc{S}_c) } \cdot X_{c, b, d}
+			\end{equation}
+		where $I(\cdot)$ is the indicator function.
 		\item Minimize the number of weekends assigned to a clinician that they requested as time-off
+			\begin{equation}
+				\label{eqn_weekend_obj}
+				\min \sum_{c, w} (-1)^{ I (w \in \mc{T}_c)} \cdot Y_{c, w}
+			\end{equation}
 		\item Maximize the adjacency between blocks and weekends assigned for a given clinician.
 In particular, if a clinician is assigned to work a given block in a given division, we will prefer it if the clinician also works in the weekend between the weeks of the block.
-	\end{enumerate}
+\\ \\
+		In our formulation, this corresponds to maximizing the sum of products of $X_{c, b, d}$ and $Y_{c, w}$ where $w$ is a weekend adjacent to block $b$ (say, $w = 2b+1$, if each block covers 2 weeks).
+To accomplish this, we define helper variables:
+			\begin{equation}
+				Z_{c, b, d} \in \{0, 1\} : \text{represents product $X_{c, b, d}\cdot Y_{c, 2b+1}$}
+			\end{equation}
+		with constraints
+			\begin{align}
+				Z_{c, b, d} &\leq X_{c, b, d} \\
+				Z_{c, b, d} &\leq Y_{c, 2b+1} \text{ for each division $d$}
+			\end{align}
+		Thus, the objective is:
+			\begin{equation}
+				\label{eqn_adjacency_obj}
+				\max \sum_{c, b, d} Z_{c, b, d} = \min \left(- \sum_{c, b, d} Z_{c, b, d}\right)
+			\end{equation}
+	\end{itemize}
 
-\section{Linear Programming Formulation}
-The problem described in \ref{problem} corresponds to an Assignment Problem between clinicians and blocks/weekends. Such a problem can be presented as an Integer Linear Program (ILP).  \\ \\
-We begin by defining the sets, constants, and indices that will be used in this formulation.
-	\begin{align*}
-		\mc{D}\coloneqq &\text{ the set of all divisions, indexed by } i \\
-		\mc{C}\coloneqq &\text{ the set of all clinicians, indexed by } j \\
-		\mc{B}\coloneqq &\text{ the set of all blocks, indexed by } k \\
-		\mc{W}\coloneqq &\text{ the set of all weekends, indexed by } l \\
-		\mc{W} \supset \mc{L}\coloneqq &\text{ the set of all long weekends} \\
-		\mc{B} \supset \mc{S}_j\coloneqq &\text{ the set of blocks clinician } j \text{ requested off} \\
-		\mc{W} \supset \mc{T}_j\coloneqq &\text{ the set of weekends clinician } j \text{ requested off} \\
-		m^i_j \coloneqq &\text{ the minimum number of blocks clinician } j \text{ should work in division } i \\
-		M^i_j \coloneqq &\text{ the maximum number of blocks clinician } j \text{ should work in division } i
-	\end{align*}
-We also define the set of 0-1 variables $X^i_{j,k}$ and $Y_{j,l}$ that we will solve for in the ILP.
-	\begin{alignat}{2}
-		&X^i_{j,k} \in \{0, 1\}: &&\text{ clinician } j \text{ covers block } k \text{ for division } i \\
-		&Y_{j, l} \in \{0, 1\}: &&\text{ clinician } j \text{ covers weekend } l
-	\end{alignat}
-Now we can write the constraints in \ref{constraints} mathematically.
-	\begin{enumerate}
-		\item 
-			\begin{equation}
-				\sum_j X^i_{j, k} = 1 \text{ for each } i, k
-			\end{equation}
-		\item
-			\begin{equation}
-				\sum_j Y_{j, l} = 1 \text{ for each } l
-			\end{equation}
-		\item 
-			\begin{equation}
-				m^i_j \leq \sum_{k} X^i_{j, k} \leq M^i_j \text{ for each } i, j
-			\end{equation}
-		\item
-			\begin{equation}
-				\floor*{\frac{|\mc{L}|}{|\mc{C}|}} \leq \sum_{l \in \mc{L}} Y_{j, l} \leq \ceil*{\frac{|\mc{L}|}{|\mc{C}|}}
-			\end{equation}
-		\item
-			\begin{equation}
-				\sum_i \left(X^i_{j, k} + X^i_{j, k+1}\right) \leq 1 \text{ for each } j,k \text{ where } k \leq |\mc{B}| - 1
-			\end{equation}
-		\item
-			\begin{equation}
-				Y_{j, l} + Y_{j, l+1} \leq 1 \text{ for each } j, l \text{ where } l \leq |\mc{W}| - 1
-			\end{equation}
-	\end{enumerate}
-Note that in ILP we can only have a single objective function.
-Since we have multiple objectives in \ref{objectives}, we will need to combine them in an appropriate manner.
-Assuming each objective can be summarized as a linear function of the variables defined above, say $Obj_n\left(X^i_{j, k}, Y_{j, l}\right)$, $1 \leq n \leq N$, then we shall define the overall objective function as a linear combination of $Obj_n$'s:
-	\begin{equation}
-		Obj\left(X^i_{j, k}, Y_{j, l}\right) \coloneqq \frac{Obj_1 + \ldots + Obj_N}{N} 
-	\end{equation}
-This ensures that all objectives have an equal weight of $\frac{1}{N}$.
-It is also necessary to choose whether we minimize or maximize $Obj$, so we will need to convert the minimization objectives.
-In particular, we can define the following objectives:
-	\begin{enumerate}
-		\item We will maximize the number of blocks assigned to a clinician that are \textit{not} requested for time-off:
-			\begin{equation}
-				Obj_1\left(X^i_{j, k}, Y_{j, l}\right) \coloneqq \sum_i \sum_j \sum_k
-					\begin{cases}
-					X^i_{j, k} &\text{ if } k \in \mc{S}_j \\
-					-X^i_{j, k} &\text{ otherwise }
-					\end{cases}
-			\end{equation}
-		\item We will maximize the number of weekends assigned to a clinician that are \textit{not} requested for time-off
-			\begin{equation}
-				Obj_2\left(X^i_{j, k}, Y_{j, l}\right) \coloneqq \sum_j \sum_l 
-					\begin{cases}
-					Y_{j, l} &\text{ if } l \in \mc{T}_j \\
-					-Y_{j, l} &\text{ otherwise }
-					\end{cases}
-			\end{equation}
-		\item We want to maximize the number of adjacent blocks and weekends.
-Since our variables are 0-1, we can accomplish this by maximizing the product of $X^i_{j, k} \cdot Y_{j, l}$ where $k$ and $l$ are adjacent.
-When both variables are assigned a value of 1, they contribute a single unit to the objective.
-Otherwise, they contribute nothing.
-\\
-		However, this objective is \textit{not} linear. To make it linear, we introduce a set of helper variables:
-			\begin{equation}
-				Z^i_{j, k} \in \{0, 1\}: \text{represents the product }X^i_{j, k} \cdot Y_{j, 2k+1}
-			\end{equation}
-		with constraints:
-			\begin{align}
-				Z^i_{j, k} &\leq X^i_{j, k} \\
-				Z^i_{j, k} &\leq Y_{j, 2k+1} \text{ for each } i
-			\end{align}
-		This gives us the third objective:
-			\begin{equation}
-				Obj_3\left(Z^i_{j, k}\right) \coloneqq \sum_i \sum_j \sum_k Z^i_{j, k}
-			\end{equation}
-	\end{enumerate}
-	Putting all the components above together, we get a linear program defined as:
-		\begin{equation}
-			\begin{split}
-			&\text{maximize } \frac{Obj_1 + Obj_2 + Obj_3}{3} \\
-			&\text{subject to } (3)\text{-}(8)
-			\end{split}
-		\end{equation}
+We can optimize the objectives simultaneously by optimizing a weighted sum of the objective functions. We chose to weigh each objective equally.
 \end{document}

### 2019-01-04 src/ui/mainwindow.py
Display config file path in tab title
Remove "loaded/not loaded" config status from scheduler tab

Closes #11

@@ -36,16 +36,21 @@ class MainWindow(QMainWindow, Ui_MainWindow):
     def configuration(self, value):
         if type(value) is dict:
             self._configuration = value
-            configStatus = len(self._configuration.keys()) > 0
-            self.configFileLabel.setText('Configuration: {}'.format(
-                'loaded' if configStatus else 'not loaded'
-            ))
         else:
             raise TypeError(value)
 
-    def setupConfigurationTab(self):
-        self.configuration = {}
+    @property
+    def configPath(self):
+        return self._configPath
 
+    @configPath.setter
+    def configPath(self, value):
+        if value:
+            self._configPath = value
+            self.tabWidget.setTabText(0, 'Configuration ({})'.format(self._configPath))
+            self.tabWidget.setTabText(1, 'Scheduler ({})'.format(self._configPath))
+
+    def setupConfigurationTab(self):
         # treeview setup
         self.model = TreeModel(None)
         self.treeEditorDelegate = TreeEditDelegate()
@@ -59,6 +64,8 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         self.actionNew_Clinician.triggered.connect(self.createNewClinician)
         self.actionEdit_Clinician.triggered.connect(self.editClinician)
         self.actionDelete_Clinician.triggered.connect(self.deleteClinician)
+
+        self.newConfig()
 
     def setupSchedulerTab(self):
         # action setup
@@ -79,6 +86,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         if path != '':
             try:
                 with open(path, 'r') as f:
+                    self.configPath = path
                     self.configuration = json.load(f)
                     UiHelper.syncTreeView(self.treeView, self.model, self.configuration)
 
@@ -96,6 +104,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
                     data = {}
                     UiHelper.convertTreeToDict(self.model.rootItem, data)
                     json.dump(data, f)
+                    self.configPath = fileName
 
             except Exception as ex:
                 QMessageBox.critical(self, "Unable to save file", str(ex))
@@ -103,6 +112,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
     def newConfig(self):
         # clear data
         self.configuration = {}
+        self.configPath = 'new config'
         
         # build treeview
         UiHelper.syncTreeView(self.treeView, self.model, self.configuration)

### 2019-01-08 src/ui/mainwindow.py
Initialize private properties before calling

@@ -22,6 +22,10 @@ from .ui_mainwindow import Ui_MainWindow
 class MainWindow(QMainWindow, Ui_MainWindow):
     def __init__(self, *args, **kwargs):
         super(MainWindow, self).__init__(*args, **kwargs)
+
+        self._configuration = {}
+        self._configPath = ''
+
         self.setupUi(self)
         self.setupConfigurationTab()
         self.setupSchedulerTab()

### 2019-01-08 src/services/scheduler.py
Toggle constraints in scheduler service

@@ -224,7 +224,7 @@ class Scheduler:
     schedule based on the supplied clinician and division data.
     """
 
-    def __init__(self, logger, num_blocks, clin_data={}, request_dict=[], holidays=[]):
+    def __init__(self, logger, num_blocks, clin_data={}, request_dict=[], holidays=[], constraints=[]):
         self.num_blocks = num_blocks
         self.num_weekends = num_blocks * BLOCK_SIZE
 
@@ -233,10 +233,12 @@ class Scheduler:
         self.holiday_map = {}
         self.long_weekends = []
         self._logger = logger
+        self.constraints = []
         
         self.set_long_weekends(holidays)
         self.set_data(clin_data)
         self.set_timeoff(request_dict)
+        self.set_constraints(constraints)
 
     def generate(self, verbose=False, shuffle=False):
         self.setup_solver()
@@ -330,15 +332,6 @@ class Scheduler:
                     max_=div_object['max']
                 )
 
-    def read_config(self, config_path):
-        """
-        Retrieves static data about clinicians and divisions from a 
-        config file.
-        """
-        with open(config_path, 'r') as f:
-            data = json.load(f)
-            self.set_data(data)
-
     def set_timeoff(self, request_dict):
         """
         Populates timeoff for each clinician in self.clinicians based
@@ -390,15 +383,21 @@ class Scheduler:
                 self.holiday_map[date] = date.isocalendar()[1]
         
         self.long_weekends = list(self.holiday_map.values())
-    
+   
+    def set_constraints(self, constraint_dict):
+        for key in constraint_dict:
+            self.constraints.append(Scheduler.CONSTRAINT_MAPPING[key])
+ 
     def setup_problem(self, shuffle=False):
         """
         Constructs an LP program based on the clinician, division data.
         """
         self.problem = pulp.LpProblem('scheduler', sense=pulp.LpMaximize)
 
-        for clinician in self.clinicians.values(): clinician.reset()
-        for division in self.divisions.values(): division.reset()
+        for clinician in self.clinicians.values():
+            clinician.reset()
+        for division in self.divisions.values():
+            division.reset()
 
         divisions = list(self.divisions.values())
         clinicians = list(self.clinicians.values())
@@ -406,14 +405,10 @@ class Scheduler:
         if shuffle: random.shuffle(clinicians)
 
         self._build_clinician_variables(divisions, clinicians)
[…]

### 2019-01-08 src/ui/mainwindow.py
Read toggled constraints and pass to service

@@ -234,11 +234,16 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
         # requests = []
         # holidays = []
+        for row in range(self.constraintsForm.rowCount()):
+            widget = self.constraintsForm.itemAt(row, QFormLayout.FieldRole).widget()
+            if widget.isChecked():
+                constraints.append(widget.objectName().replace('CheckBox', ''))
 
         # init scheduler with all the given data
         schedule = scheduler \
             .Scheduler(logger=self._logger, num_blocks=numBlocks, clin_data=self.configuration, \
-                 request_dict=self._requests, holidays=self._holidays) \
+                 request_dict=self._requests, holidays=self._holidays,
+                 constraints=constraints) \
             .generate(verbose=verbose, shuffle=shuffle)
         if schedule is None:
             self._logger.write_line('Could not generate schedule! Try adjusting min/max values in the configuration tab.', level='ERROR')

### 2019-01-10 docs/source/conf.py
Make docs/ a Sphinx documentation folder

@@ -1,0 +1,159 @@
+# -*- coding: utf-8 -*-
+#
+# Configuration file for the Sphinx documentation builder.
+#
+# This file does only contain a selection of the most common options. For a
+# full list see the documentation:
+# http://www.sphinx-doc.org/en/master/config
+
+# -- Path setup --------------------------------------------------------------
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+#
+# import os
+# import sys
+# sys.path.insert(0, os.path.abspath('.'))
+
+
+# -- Project information -----------------------------------------------------
+
+project = 'Clinician Scheduler'
+copyright = '2019, David Landsman'
+author = 'David Landsman'
+
+# The short X.Y version
+version = ''
+# The full version, including alpha/beta/rc tags
+release = '1.0'
+
+
+# -- General configuration ---------------------------------------------------
+
+# If your documentation needs a minimal Sphinx version, state it here.
+#
+# needs_sphinx = '1.0'
+
+# Add any Sphinx extension module names here, as strings. They can be
+# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
+# ones.
+extensions = [
+    'sphinx.ext.githubpages',
+]
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ['_templates']
+
+# The suffix(es) of source filenames.
+# You can specify multiple suffix as a list of string:
+#
+# source_suffix = ['.rst', '.md']
+source_suffix = '.rst'
+
+# The master toctree document.
+master_doc = 'index'
+
+# The language for content autogenerated by Sphinx. Refer to documentation
+# for a list of supported languages.
+#
+# This is also used if you do content translation via gettext catalogs.
+# Usually you set "language" from the command line for these cases.
+language = None
+
+# List of patterns, relative to source directory, that match files and
+# directories to ignore when looking for source files.
+# This pattern also affects html_static_path and html_extra_path .
+exclude_patterns = []
+
+# The name of the Pygments (syntax highlighting) style to use.
+pygments_style = 'sphinx'
+
+
+# -- Options for HTML output -------------------------------------------------
+
+# The theme to use for HTML and HTML Help pages.  See the documentation for
+# a list of builtin themes.
+#
+html_theme = 'alabaster'
+
[…]

### 2019-01-10 docs/source/conf.py
Add instructions for generating credentials

@@ -39,7 +39,7 @@ release = '1.0'
 # extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
 # ones.
 extensions = [
-    'sphinx.ext.githubpages',
+    'sphinx.ext.githubpages'
 ]
 
 # Add any paths that contain templates here, relative to this directory.

### 2019-01-22 src/ui/delegates.py
Fix bug where clinician was duplicated on edit
Closes #19.

Also,
* separate handling of adding and editing a clinician in dialog
* remove ability to overwrite existing clinician in dialog
* remove ability to edit treeview in main window; editing should only be done through dialog window

@@ -16,21 +16,3 @@ class ValidationDelegate(QItemDelegate):
         editor = QLineEdit(parent)
         editor.setValidator(self.validator())
         return editor
-
-class TreeEditDelegate(QItemDelegate):
-    """
-    Custom delegate that copies data into editor during initialization
-    """
-
-    def __init__(self, parent=None):
-        QItemDelegate.__init__(self, parent)
-
-    def createEditor(self, parent, option, index):
-        editor = QLineEdit(parent)
-        return editor
-
-    def setEditorData(self, editor, index):
-        if index.isValid():
-            editor.setText(index.data())
-        else:
-            QItemDelegate.setEditorData(editor, index)

### 2019-01-22 src/ui/dialog.py

@@ -8,9 +8,10 @@ from ui.ui_dialog import Ui_Dialog
 
 
 class DialogWindow(QDialog, Ui_Dialog):
-    def __init__(self, *args, **kwargs):
+    def __init__(self, isEdit=False, *args, **kwargs):
         super(DialogWindow, self).__init__(*args, **kwargs)
         self.setupUi(self)
+        self.isEdit = isEdit
 
         # make sure min/max columns only accept integer values
         # note: need class to "own" delegates, otherwise they will be GC'd and cause a segfault
@@ -37,6 +38,9 @@ class DialogWindow(QDialog, Ui_Dialog):
         self.clinDict = clinDict
 
     def setClinician(self, data):
+        # store current clinician name in case it changes
+        if self.isEdit: self.oldName = data['name']
+    
         self.nameLineEdit.setText(data['name'])
         self.emailLineEdit.setText(data['email'])
         
@@ -69,34 +73,88 @@ class DialogWindow(QDialog, Ui_Dialog):
         return True
 
     def accept(self):
+        """
+        Accept changes and update configuration.
+        """
+
         # validate data
         if not self.checkInput():
             return
 
-        # create new clinician in self.data
-        clinName = self.nameLineEdit.text()
+        ret = False
+
+        if self.isEdit: 
+            ret = self.acceptEdit()
+        else: 
+            ret = self.acceptNew()
+
+        if ret: QDialog.accept(self)
+
+    def acceptEdit(self):
+        """
+        Updates configuration based on edits made to existing clinician.
+
+        Returns True iff changes were accepted.
+        """
+
+        clinician = self.extractFormData()
+
+        # check if the newly chosen name already exists in the config
+        if self.oldName != clinician['name'] and clinician['name'] in self.clinDict:
+            QMessageBox.critical(
+                self,
+                "Name Error",
+                "A clinician with this name already exists in the configuration file. Please choose a different name.",
+                QMessageBox.Ok
+            )
+
+            return False
+
+        # remove old data and update config
+        del self.clinDict[self.oldName]
+        self.clinDict[clinician['name']] = clinician
+
+        return True
+
+    def acceptNew(self):
+        """
+        Updates configuration for newly added clinician.
+        
+        Returns True iff changes were accepted.
+        """
+
+        clinician = self.extractFormData()
+
+        # check if such a name is already stored in config
+        if clinician['name'] in self.clinDict:
+            QMessageBox.critical(
+                self,
+                "Name Error",
[…]

### 2019-01-22 src/ui/mainwindow.py

@@ -14,7 +14,6 @@ from helpers.excelhelper import ExcelHelper
 from helpers.uihelper import UiHelper
 from services import scheduler
 
-from .delegates import TreeEditDelegate
 from .dialog import DialogWindow
 from .models import TreeModel
 from .ui_mainwindow import Ui_MainWindow
@@ -61,8 +60,6 @@ class MainWindow(QMainWindow, Ui_MainWindow):
     def setupConfigurationTab(self):
         # treeview setup
         self.model = TreeModel(None)
-        self.treeEditorDelegate = TreeEditDelegate()
-        self.treeView.setItemDelegate(self.treeEditorDelegate)
         self.treeView.setModel(self.model)
 
         # action setup
@@ -143,7 +140,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         UiHelper.syncTreeView(self.treeView, self.model, self.configuration)
 
     def createNewClinician(self):
-        dialog = DialogWindow()
+        dialog = DialogWindow(isEdit=False)
         dialog.setClinicianDictionary(self.configuration)
         dialog.exec_()
 
@@ -158,7 +155,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
             clinData["name"] = selectedItem.itemData[0]
 
             # transfer clinData to edit dialog
-            dialog = DialogWindow()
+            dialog = DialogWindow(isEdit=True)
             dialog.setClinicianDictionary(self.configuration)
             dialog.setClinician(clinData)
             dialog.exec_()
@@ -172,7 +169,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
             reply = QMessageBox.question(
                 self, 
                 "Delete Warning", 
-                "This action will delete the current data stored for {}. Are you sure you want to continue?".format(
+                "This action will delete the current data stored for \"{}\". Are you sure you want to continue?".format(
                     clinName)
                 , QMessageBox.Yes | QMessageBox.No)
             if reply == QMessageBox.No: return

### 2019-01-22 src/ui/ui_mainwindow.py

@@ -1,6 +1,6 @@
 # -*- coding: utf-8 -*-
 
-# Form implementation generated from reading ui file 'src\designer\mainwindow.ui'
+# Form implementation generated from reading ui file 'designer\mainwindow.ui'
 #
 # Created by: PyQt5 UI code generator 5.11.2
 #
@@ -51,7 +51,7 @@ class Ui_MainWindow(object):
         self.horizontalLayout.addWidget(self.deleteClinicianButton)
         self.verticalLayout.addWidget(self.configActions)
         self.treeView = QtWidgets.QTreeView(self.configurationTab)
-        self.treeView.setEditTriggers(QtWidgets.QAbstractItemView.EditKeyPressed|QtWidgets.QAbstractItemView.SelectedClicked)
+        self.treeView.setEditTriggers(QtWidgets.QAbstractItemView.NoEditTriggers)
         self.treeView.setAlternatingRowColors(True)
         self.treeView.setSelectionBehavior(QtWidgets.QAbstractItemView.SelectItems)
         self.treeView.setExpandsOnDoubleClick(True)

### 2019-01-22 src/ui/mainwindow.py
Automatically resize treeview columns when expanding items
Closes #23.

@@ -63,6 +63,8 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         self.treeView.setModel(self.model)
 
         # action setup
+        self.treeView.expanded.connect(self.autoSizeTreeView)
+
         self.openConfigButton.clicked.connect(self.openConfig)
         self.saveConfigButton.clicked.connect(self.saveConfig)
         self.newConfigButton.clicked.connect(self.newConfig)
@@ -99,6 +101,10 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         # add new config path in brackets, if no brackets yet
         newText = re.sub(r'\((?:.*)\)|(\w+)(.*)', r'\1 ({})'.format(self._configPath), text)
         self.tabWidget.setTabText(selectedTabIdx, newText)
+
+    def autoSizeTreeView(self):
+        for i in range(self.model.rootItem.columnCount()):
+            self.treeView.resizeColumnToContents(i)
 
     def openConfig(self):
         path, _ = QFileDialog.getOpenFileName(

### 2019-01-24 docs/build/latex/ClinicianScheduler.tex
Add pdf version of manual (sphinx-build latex)

%% Generated by Sphinx.
\def\sphinxdocclass{report}
\documentclass[letterpaper,10pt,english]{sphinxmanual}
\ifdefined\pdfpxdimen
   \let\sphinxpxdimen\pdfpxdimen\else\newdimen\sphinxpxdimen
\fi \sphinxpxdimen=.75bp\relax

\PassOptionsToPackage{warn}{textcomp}
\usepackage[utf8]{inputenc}
\ifdefined\DeclareUnicodeCharacter
 \ifdefined\DeclareUnicodeCharacterAsOptional
  \DeclareUnicodeCharacter{"00A0}{\nobreakspace}
  \DeclareUnicodeCharacter{"2500}{\sphinxunichar{2500}}
  \DeclareUnicodeCharacter{"2502}{\sphinxunichar{2502}}
  \DeclareUnicodeCharacter{"2514}{\sphinxunichar{2514}}
  \DeclareUnicodeCharacter{"251C}{\sphinxunichar{251C}}
  \DeclareUnicodeCharacter{"2572}{\textbackslash}
 \else
  \DeclareUnicodeCharacter{00A0}{\nobreakspace}
  \DeclareUnicodeCharacter{2500}{\sphinxunichar{2500}}
  \DeclareUnicodeCharacter{2502}{\sphinxunichar{2502}}
  \DeclareUnicodeCharacter{2514}{\sphinxunichar{2514}}
  \DeclareUnicodeCharacter{251C}{\sphinxunichar{251C}}
  \DeclareUnicodeCharacter{2572}{\textbackslash}
 \fi
\fi
\usepackage{cmap}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amstext}
\usepackage{babel}
\usepackage{times}
\usepackage[Bjarne]{fncychap}
\usepackage{sphinx}

\usepackage{geometry}

% Include hyperref last.
\usepackage{hyperref}
% Fix anchor placement for figures with captions.
\usepackage{hypcap}% it must be loaded after hyperref.
% Set up styles of URL: it should be placed after hyperref.
\urlstyle{same}

\addto\captionsenglish{\renewcommand{\figurename}{Fig.}}
\addto\captionsenglish{\renewcommand{\tablename}{Table}}
\addto\captionsenglish{\renewcommand{\literalblockname}{Listing}}

\addto\captionsenglish{\renewcommand{\literalblockcontinuedname}{continued from previous page}}
\addto\captionsenglish{\renewcommand{\literalblockcontinuesname}{continues on next page}}

\addto\extrasenglish{\def\pageautorefname{page}}

\title{Clinician Scheduler Documentation}
\date{Jan 24, 2019}
\release{1.0}
\author{David Landsman}
\newcommand{\sphinxlogo}{\vbox{}}
\renewcommand{\releasename}{Release}
\makeindex

\begin{document}

\maketitle
\sphinxtableofcontents
\phantomsection\label{\detokenize{index::doc}}

\begin{sphinxShadowBox}
\sphinxstyletopictitle{Contents:}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{index:id2}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{index:id3}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}

\end{itemize}

\item {} 
\phantomsection\label{\detokenize{index:id4}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{index:id5}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{index:id6}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}

\item {} 
\phantomsection\label{\detokenize{index:id7}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}

\item {} 
\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}

\item {} 
\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:saving-the-configuration-to-a-file}]{\sphinxcrossref{Saving the configuration to a file}}}

\item {} 
\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}

\end{itemize}

\item {} 
\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}

\item {} 
\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}

\item {} 
\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}

\end{itemize}

\item {} 
\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:id1}]{\sphinxcrossref{Loading a configuration file}}}

\item {} 
\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}

\item {} 
\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}

\item {} 
\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}

\end{itemize}

\end{itemize}

\end{itemize}
\end{sphinxShadowBox}

The clinician scheduler allows you to automatically generate and publish
schedules that satisfy common constraints in an on-call system, while taking
into account the preferences of clinicians.

\chapter{Setup}
\label{\detokenize{index:setup}}
Download and unzip the compressed file to an appropriate folder.

\section{Generating Google API credentials}
\label{\detokenize{index:generating-google-api-credentials}}
\begin{sphinxadmonition}{note}{Note:}
This process should only be done the very first time you are running
the program. If you have previously generated a credential file using
this process you should be able to re-use it. Just make sure that the
credential file is placed in the same folder as the executable (\sphinxcode{\sphinxupquote{scheduler.exe}}).
\end{sphinxadmonition}

The application uses Google calendar for retrieving clinician
time-off requests and long weekend information, as well as uploading the
generated schedule to the calendar. These operations require the use
of Google API credentials, which can be generated as follows.
\begin{enumerate}
\item {} 
Sign into \sphinxurl{https://console.developers.google.com}.

\item {} 
Click on the project selector at the top left of the page.

\end{enumerate}

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step2\_select\_a\_project.png}{\sphinxincludegraphics{{step2_select_a_project}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}
\begin{enumerate}
\setcounter{enumi}{2}
\item {} 
Click on \sphinxtitleref{New Project}.

\end{enumerate}

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step3\_new\_project.png}{\sphinxincludegraphics{{step3_new_project}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}
\begin{enumerate}
\setcounter{enumi}{3}
\item {} 
Enter “Clinician Scheduler” as the \sphinxtitleref{Project Name} and click \sphinxtitleref{Create}.

\end{enumerate}

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step4\_create\_project.png}{\sphinxincludegraphics{{step4_create_project}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

5. Now you should see the dashboard for the Clinician Scheduler project.
You will need to enable the Google calendar API. Click on \sphinxtitleref{Enable APIs and Services}.

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step5\_enable\_apis\_and\_services.png}{\sphinxincludegraphics{{step5_enable_apis_and_services}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}
\begin{enumerate}
\setcounter{enumi}{5}
\item {} 
Search for “Google calendar API” using the search bar, and select it.

\end{enumerate}

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step6\_search\_calendar\_api.png}{\sphinxincludegraphics{{step6_search_calendar_api}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}
\begin{enumerate}
\setcounter{enumi}{6}
\item {} 
Click \sphinxtitleref{Enable}.

\end{enumerate}

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step7\_enable\_calendar\_api.png}{\sphinxincludegraphics{{step7_enable_calendar_api}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

8. Now you should see the overview page for the Google calendar API.
To generate the credentials, click \sphinxtitleref{Create credentials}.

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step8\_create\_credentials.png}{\sphinxincludegraphics{{step8_create_credentials}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

9. On the credentials form, choose “Google Calendar API” for \sphinxtitleref{Which API are you using?},
then “Other UI (e.g. Windows, CLI tool)” for \sphinxtitleref{Where will you be calling the API from?}
and “User data” for \sphinxtitleref{What data will you be accessing?}. Then click on \sphinxtitleref{What credentials do I need?}.

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step9\_credentials\_form.png}{\sphinxincludegraphics{{step9_credentials_form}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}
\begin{enumerate}
\setcounter{enumi}{9}
\item {} 
Enter “Client” for \sphinxtitleref{Name} and click on \sphinxtitleref{Create OAuth client ID}.

\end{enumerate}

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step10\_oauth\_client\_id.png}{\sphinxincludegraphics{{step10_oauth_client_id}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

11. Choose the email address associated to your account for \sphinxtitleref{Email address}
and enter “Clinician Scheduler” for \sphinxtitleref{Product name shown to users}, then
click \sphinxtitleref{Continue}.

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step11\_oauth\_consent\_screen.png}{\sphinxincludegraphics{{step11_oauth_consent_screen}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

12. Your credentials are now generated! Make sure to download and save them
in the same location that you unzipped the application, so that the
credential file and the executable file (\sphinxcode{\sphinxupquote{scheduler.exe}}) are in the same folder.

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/google\_credentials/step12\_download\_credentials.png}{\sphinxincludegraphics{{step12_download_credentials}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

\begin{sphinxadmonition}{attention}{Attention:}
Make sure the credential file is saved as \sphinxcode{\sphinxupquote{credentials.json}} (rename it, if necessary),
or otherwise the application will not be able to recognize it!
\end{sphinxadmonition}

\chapter{Usage}
\label{\detokenize{index:usage}}

\section{Clinician Configuration}
\label{\detokenize{index:clinician-configuration}}
Before we can generate a schedule, we need to create a configuration file
that specifies which clinicians are available, and how many weeks each
clinician should fulfill.

\subsection{Adding a new clinician}
\label{\detokenize{index:adding-a-new-clinician}}
1. From the configuration tab, click \sphinxtitleref{New Clinician}. You should see a
form for supplying details.

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/add\_clinician/step1\_new\_clinician.png}{\sphinxincludegraphics{{step1_new_clinician}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

2. Fill out the name, email (optional), and divisions that the clinician
will be covering. To add a division you can click on \sphinxtitleref{Add} and a new row
will be added to the table which you can fill out. You can set the minimum
and maximum number of blocks that a clinician can work in a given division.

\begin{sphinxadmonition}{note}{Note:}
A single block corresponds to two weeks.
\end{sphinxadmonition}

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/add\_clinician/step2\_add\_division.png}{\sphinxincludegraphics{{step2_add_division}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}
\begin{enumerate}
\setcounter{enumi}{2}
\item {} 
To delete a row from the table, select the row and then click \sphinxtitleref{Remove}.

\end{enumerate}

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/add\_clinician/step3\_remove\_division.png}{\sphinxincludegraphics{{step3_remove_division}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

4. When you are finished entering the data for the clinician, click \sphinxtitleref{Ok}.
You should now see a new entry in the main table for that clinician.

\begin{figure}[htbp]
\centering
\sphinxhref{\_static/images/add\_clinician/step4\_add\_clinician.png}{\sphinxincludegraphics{{step4_add_clinician}.png}}\end{figure}

\begin{DUlineblock}{0em}
\item[] 
\end{DUlineblock}

\subsection{Deleting an existing clinician}
\label{\detokenize{index:deleting-an-existing-clinician}}

\subsection{Editing an existing clinician}
\label{\detokenize{index:editing-an-existing-clinician}}

\subsection{Saving the configuration to a file}
\label{\detokenize{index:saving-the-configuration-to-a-file}}

\subsection{Loading a configuration file}
\label{\detokenize{index:loading-a-configuration-file}}

\section{Google Calendar Configuration}
\label{\detokenize{index:google-calendar-configuration}}

\subsection{Creating a calendar}
\label{\detokenize{index:creating-a-calendar}}

\subsection{Adding holiday events}
\label{\detokenize{index:adding-holiday-events}}

\subsection{Adding clinician requests}
\label{\detokenize{index:adding-clinician-requests}}

\section{Scheduling}
\label{\detokenize{index:scheduling}}

\subsection{Loading a configuration file}
\label{\detokenize{index:id1}}

\subsection{Generating a schedule}
\label{\detokenize{index:generating-a-schedule}}

\subsection{Exporting a schedule}
\label{\detokenize{index:exporting-a-schedule}}

\subsection{Publishing a schedule to Google Calendar}
\label{\detokenize{index:publishing-a-schedule-to-google-calendar}}

\renewcommand{\indexname}{Index}
\printindex
\end{document}

### 2019-01-24 src/ui/mainwindow.py
Disable event retrieval when calendar id not supplied
Closes #24.

@@ -31,6 +31,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         self.setupConfigurationTab()
         self.setupSchedulerTab()
 
+        self.handleCalendarIdChange(self.gCalLineEdit.text())
         self.tabWidget.setCurrentIndex(0)
 
         self.show()
@@ -76,6 +77,8 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
     def setupSchedulerTab(self):
         # action setup
+        self.gCalLineEdit.textChanged[str].connect(self.handleCalendarIdChange)
+
         self.loadButton.clicked.connect(self.openConfig)
         self.generateScheduleButton.clicked.connect(self.generateSchedule)
         self.exportScheduleButton.clicked.connect(self.exportSchedule)
@@ -85,6 +88,20 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
         # misc vars
         self.holidayMap = {}
+
+    def handleCalendarIdChange(self, text):
+        if len(text) == 0:
+            # first, uncheck "retrieve" checkboxes
+            self.retrieveTimeOffRequestsCheckBox.setChecked(False)
+            self.retrieveLongWeekendsCheckBox.setChecked(False)
+            # then disable them altogether
+            self.retrieveTimeOffRequestsCheckBox.setDisabled(True)
+            self.retrieveLongWeekendsCheckBox.setDisabled(True)
+
+        else:
+            # enable "retrieve" checkboxes
+            self.retrieveTimeOffRequestsCheckBox.setDisabled(False)
+            self.retrieveLongWeekendsCheckBox.setDisabled(False)
 
     def updateTabText(self, selectedTabIdx):
         # clear (configPath) from all other tabs

### 2019-01-24 docs/build/latex/ClinicianScheduler.tex
Work on `Scheduling` section and add `Authentication` section in manual

@@
 \item {} 
-\phantomsection\label{\detokenize{index:id2}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
+\phantomsection\label{\detokenize{index:id7}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id3}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
+\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id4}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
+\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id5}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
+\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
+
+\item {} 
+\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id6}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id7}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
+\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
+\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:saving-the-configuration-to-a-file}]{\sphinxcrossref{Saving the configuration to a file}}}
+\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:saving-the-configuration-to-a-file}]{\sphinxcrossref{Saving the configuration to a file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
+\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
+\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
+\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
+\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
+\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:id1}]{\sphinxcrossref{Loading a configuration file}}}
+\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:id4}]{\sphinxcrossref{Loading a configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
+\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
+\phantomsection\label{\detokenize{index:id24}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
+\phantomsection\label{\detokenize{index:id25}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
 
@@
 
+\section{Authentication}
+\label{\detokenize{index:authentication}}\label{\detokenize{index:id1}}
+When the scheduler first makes a connection to Google calendar, it needs
+to create an authentication token, which will be used to simplify future
+connections.
+
+This requires you to allow the scheduler application to access a calendar
+that you specified. You may encounter this when using any functionality that
+interacts with Google calendar, for example in {\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{\DUrole{std,std-ref}{Generating a schedule}}}}
+or in {\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{\DUrole{std,std-ref}{Publishing a schedule to Google Calendar}}}}.
+
+1. The scheduler should have automatically opened the access request page
+in a browser window, in which case you can skip to step 2. \sphinxstylestrong{Otherwise},
+locate the command prompt for the scheduler. You should see instructions
+to open the access request printed on the command prompt. Follow these instructions
+in order to open the access request page manually.
+
+2. Follow the instructions on the access request page in order to allow
+the scheduler application to read/write from/to the calendar you specified.
+\sphinxstylestrong{(Optional)} If you opened the access request manually in step 1, you will
+receive an authentication code which needs to be pasted into the command prompt
+
+3. Once you have completed the authentication process, the scheduler will
+automatically resume its functionality.
+
 \section{Clinician Configuration}
-\label{\detokenize{index:clinician-configuration}}
+\label{\detokenize{index:clinician-configuration}}\label{\detokenize{index:id2}}
 Before we can generate a schedule, we need to create a configuration file
@@
 \subsection{Adding clinician requests}
-\label{\detokenize{index:adding-clinician-requests}}
+\label{\detokenize{index:adding-clinician-requests}}\label{\detokenize{index:id3}}
 
@@
 \subsection{Loading a configuration file}
-\label{\detokenize{index:id1}}
+\label{\detokenize{index:id4}}
+From the scheduling tab, click on \sphinxtitleref{Load} and select the configuration
+file that you generated in {\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
+
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/scheduling/load\_config/load\_config.png}{\sphinxincludegraphics{{load_config}.png}}\end{figure}
+
+\begin{DUlineblock}{0em}
+\item[] 
+\end{DUlineblock}
 
 \subsection{Generating a schedule}
-\label{\detokenize{index:generating-a-schedule}}
+\label{\detokenize{index:generating-a-schedule}}\label{\detokenize{index:id5}}\begin{enumerate}
+\item {} 
+Enter the calendar year for the schedule that you want to generate.
+
+\end{enumerate}
+
+2. \sphinxstylestrong{(Optional)} In order to retrieve the time-off requests that were populated
+in Google calendar in {\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding clinician requests}}}}, we need to specify the
+calendar ID. Open the calendar in your browser, and navigate to the
+\sphinxtitleref{Settings and sharing} page.
+On the setting page, navigate to the \sphinxtitleref{Integrate calendar} section, and copy
+the value under \sphinxtitleref{Calendar ID} to your clipboard.
+Paste this value into the \sphinxtitleref{Google Calendar ID} textbox on the \sphinxtitleref{Scheduler}
+tab.
+
+3. \sphinxstylestrong{(Optional)} If you supplied a calendar ID in step 2, you can configure
+the options \sphinxtitleref{Retrieve Time-off Requests} and \sphinxtitleref{Retrieve Long Weekends} as necessary.
+Enabling \sphinxtitleref{Retrieve Time-off Requests} will read the time-off calendar events
+from the calendar specified in \sphinxtitleref{Calendar ID}, while enabling
+\sphinxtitleref{Retrieve Long Weekends} will read the holiday events from that calendar.
+
+4. \sphinxstylestrong{(Optional)} In the case that you only need to generate a schedule for
+a subset of the calendar year, you can select the amount of blocks you need
+to generate by specifying a value in \sphinxtitleref{Number of Blocks}. By default, the
+scheduler will generate a full schedule, corresponding to 26 blocks.
+
+5. Click on \sphinxtitleref{Generate}, and after a few moments you should see a preview of
+the generated schedule in the table. \sphinxstylestrong{(Optional)} See {\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{\DUrole{std,std-ref}{Authentication}}}}
+in case you encounter problems with Google calendar at this stage.
+
+\begin{sphinxadmonition}{note}{Note:}
+It is possible that the scheduler will not be able to come up with a
+schedule that satisfies your constraints. You can try adjusting
+some constraints by changing the minimum and maximum number of blocks of
+clinicians in the configuration file. See {\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
+for more information on changing the configuration file.
+\end{sphinxadmonition}
 
@@
 \subsection{Publishing a schedule to Google Calendar}
-\label{\detokenize{index:publishing-a-schedule-to-google-calendar}}
+\label{\detokenize{index:publishing-a-schedule-to-google-calendar}}\label{\detokenize{index:id6}}
 

### 2019-01-29 docs/build/latex/ClinicianScheduler.tex
Fix multi-line numbered lists
Also add some instructions for excel export

@@
 \title{Clinician Scheduler Documentation}
-\date{Jan 24, 2019}
+\date{Jan 29, 2019}
 \release{1.0}
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id7}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
+\phantomsection\label{\detokenize{index:id6}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
+\phantomsection\label{\detokenize{index:id7}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
+\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
+\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
+\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
+\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
+\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:saving-the-configuration-to-a-file}]{\sphinxcrossref{Saving the configuration to a file}}}
+\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:saving-the-configuration-to-a-file}]{\sphinxcrossref{Saving the configuration to a file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
+\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
+\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
+\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
+\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
+\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:id4}]{\sphinxcrossref{Loading a configuration file}}}
+\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
+\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id24}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
-
-\item {} 
-\phantomsection\label{\detokenize{index:id25}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
+\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
 
@@
 \end{DUlineblock}
+\begin{enumerate}
+\setcounter{enumi}{4}
+\item {} 
+Now you should see the dashboard for the Clinician Scheduler project.
+You will need to enable the Google calendar API. Click on \sphinxtitleref{Enable APIs and Services}.
 
-5. Now you should see the dashboard for the Clinician Scheduler project.
-You will need to enable the Google calendar API. Click on \sphinxtitleref{Enable APIs and Services}.
+\end{enumerate}
 
@@
 \end{DUlineblock}
+\begin{enumerate}
+\setcounter{enumi}{7}
+\item {} 
+Now you should see the overview page for the Google calendar API.
+To generate the credentials, click \sphinxtitleref{Create credentials}.
 
-8. Now you should see the overview page for the Google calendar API.
-To generate the credentials, click \sphinxtitleref{Create credentials}.
+\end{enumerate}
 
@@
 \end{DUlineblock}
-
-9. On the credentials form, choose “Google Calendar API” for \sphinxtitleref{Which API are you using?},
+\begin{enumerate}
+\setcounter{enumi}{8}
+\item {} 
+On the credentials form, choose “Google Calendar API” for \sphinxtitleref{Which API are you using?},
 then “Other UI (e.g. Windows, CLI tool)” for \sphinxtitleref{Where will you be calling the API from?}
 and “User data” for \sphinxtitleref{What data will you be accessing?}. Then click on \sphinxtitleref{What credentials do I need?}.
+
+\end{enumerate}
 
@@
 \end{DUlineblock}
-
-11. Choose the email address associated to your account for \sphinxtitleref{Email address}
+\begin{enumerate}
+\setcounter{enumi}{10}
+\item {} 
+Choose the email address associated to your account for \sphinxtitleref{Email address}
 and enter “Clinician Scheduler” for \sphinxtitleref{Product name shown to users}, then
 click \sphinxtitleref{Continue}.
+
+\end{enumerate}
 
@@
 \end{DUlineblock}
+\begin{enumerate}
+\setcounter{enumi}{11}
+\item {} 
+Your credentials are now generated! Make sure to download and save
+them in the same location that you unzipped the application, so that
+the credential file and the executable file (\sphinxcode{\sphinxupquote{scheduler.exe}}) are in the same folder.
 
-12. Your credentials are now generated! Make sure to download and save them
-in the same location that you unzipped the application, so that the
-credential file and the executable file (\sphinxcode{\sphinxupquote{scheduler.exe}}) are in the same folder.
+\end{enumerate}
 
@@
 or in {\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{\DUrole{std,std-ref}{Publishing a schedule to Google Calendar}}}}.
-
-1. The scheduler should have automatically opened the access request page
+\begin{enumerate}
+\item {} 
+The scheduler should have automatically opened the access request page
 in a browser window, in which case you can skip to step 2. \sphinxstylestrong{Otherwise},
@@
 
-2. Follow the instructions on the access request page in order to allow
+\end{enumerate}
+\begin{enumerate}
+\setcounter{enumi}{1}
+\item {} 
+Follow the instructions on the access request page in order to allow
 the scheduler application to read/write from/to the calendar you specified.
+
 \sphinxstylestrong{(Optional)} If you opened the access request manually in step 1, you will
@@
 
-3. Once you have completed the authentication process, the scheduler will
+\end{enumerate}
+\begin{enumerate}
+\setcounter{enumi}{2}
+\item {} 
+Once you have completed the authentication process, the scheduler will
 automatically resume its functionality.
+
+\end{enumerate}
 
@@
 \end{DUlineblock}
-
-2. Fill out the name, email (optional), and divisions that the clinician
+\begin{enumerate}
+\setcounter{enumi}{1}
+\item {} 
+Fill out the name, email (optional), and divisions that the clinician
 will be covering. To add a division you can click on \sphinxtitleref{Add} and a new row
@@
 and maximum number of blocks that a clinician can work in a given division.
+
+\end{enumerate}
 
@@
 \end{DUlineblock}
+\begin{enumerate}
+\setcounter{enumi}{3}
+\item {} 
+When you are finished entering the data for the clinician, click \sphinxtitleref{Ok}.
+You should now see a new entry in the main table for that clinician.
 
-4. When you are finished entering the data for the clinician, click \sphinxtitleref{Ok}.
-You should now see a new entry in the main table for that clinician.
+\end{enumerate}
 
@@
 
-\subsection{Loading a configuration file}
-\label{\detokenize{index:id4}}
+\subsection{Generating a schedule}
+\label{\detokenize{index:generating-a-schedule}}\label{\detokenize{index:id4}}
+Once you have created a configuration file, you can switch over to the
+\sphinxtitleref{Scheduler} tab of the application in order to generate a schedule.
+\begin{enumerate}
+\item {} 
 From the scheduling tab, click on \sphinxtitleref{Load} and select the configuration
@@
 
-\begin{figure}[htbp]
-\centering
-\sphinxhref{\_static/images/scheduling/load\_config/load\_config.png}{\sphinxincludegraphics{{load_config}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
-
-\subsection{Generating a schedule}
-\label{\detokenize{index:generating-a-schedule}}\label{\detokenize{index:id5}}\begin{enumerate}
+\end{enumerate}
+\begin{enumerate}
+\setcounter{enumi}{1}
 \item {} 
@@
 \end{enumerate}
-
-2. \sphinxstylestrong{(Optional)} In order to retrieve the time-off requests that were populated
+\begin{enumerate}
+\setcounter{enumi}{2}
+\item {} 
+\sphinxstylestrong{(Optional)} In order to retrieve the time-off requests that were populated
 in Google calendar in {\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding clinician requests}}}}, we need to specify the
@@
 
-3. \sphinxstylestrong{(Optional)} If you supplied a calendar ID in step 2, you can configure
+\end{enumerate}
+\begin{enumerate}
+\setcounter{enumi}{3}
+\item {} 
+\sphinxstylestrong{(Optional)} If you supplied a calendar ID in step 3, you can configure
 the options \sphinxtitleref{Retrieve Time-off Requests} and \sphinxtitleref{Retrieve Long Weekends} as necessary.
@@
 
-4. \sphinxstylestrong{(Optional)} In the case that you only need to generate a schedule for
+\end{enumerate}
+\begin{enumerate}
+\setcounter{enumi}{4}
+\item {} 
+\sphinxstylestrong{(Optional)} In the case that you only need to generate a schedule for
 a subset of the calendar year, you can select the amount of blocks you need
@@
 
-5. Click on \sphinxtitleref{Generate}, and after a few moments you should see a preview of
-the generated schedule in the table. \sphinxstylestrong{(Optional)} See {\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{\DUrole{std,std-ref}{Authentication}}}}
-in case you encounter problems with Google calendar at this stage.
+\end{enumerate}
+\begin{enumerate}
+\setcounter{enumi}{5}
+\item {} 
+Click on \sphinxtitleref{Generate}, and after a few moments you should see a preview of
+the generated schedule in the table.
+
+\sphinxstylestrong{(Optional)} See {\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{\DUrole{std,std-ref}{Authentication}}}} in case you encounter problems
+with Google calendar at this stage.
+
+\end{enumerate}
 
@@
 \label{\detokenize{index:exporting-a-schedule}}
+If you are satisfied with the generated schedule, you can choose to export
+it as an Excel file. There are two format options: \sphinxtitleref{Yearly Export} and
+\sphinxtitleref{Monthly Export}.
+
+Selecting the \sphinxtitleref{Yearly Export} option will generate an excel file with a single
+sheet, displaying the clinicians that are covering a particular division
+for a given week or weekend. It is very similar to the table output in
+the application itself.
+
+Selecting the \sphinxtitleref{Monthly Export} option will generate a more detailed breakdown
+of the schedule, with a separate sheet for every month, detailing which
+clinician covers which division on which day.
 
 \subsection{Publishing a schedule to Google Calendar}
-\label{\detokenize{index:publishing-a-schedule-to-google-calendar}}\label{\detokenize{index:id6}}
+\label{\detokenize{index:publishing-a-schedule-to-google-calendar}}\label{\detokenize{index:id5}}
 

### 2019-01-29 docs/build/latex/ClinicianScheduler.tex
Add info on creating/saving/loading config file
Also, move GCal section to setup

@@
 \begin{sphinxShadowBox}
-\sphinxstyletopictitle{Contents:}
+\sphinxstyletopictitle{Table of Contents}
 \begin{itemize}
@@
 
-\end{itemize}
+\item {} 
+\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
+\begin{itemize}
+\item {} 
+\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
+\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
 
-\item {} 
-\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
-
-\item {} 
-\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
-
-\item {} 
-\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:saving-the-configuration-to-a-file}]{\sphinxcrossref{Saving the configuration to a file}}}
-
-\item {} 
-\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+\end{itemize}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
+\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
+\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
+\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
+\begin{itemize}
+\item {} 
+\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
+\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
+
+\item {} 
+\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+
+\item {} 
+\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+
+\item {} 
+\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
+
+\item {} 
+\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
+\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
+\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
+\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
+\phantomsection\label{\detokenize{index:id24}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
+
+\item {} 
+\phantomsection\label{\detokenize{index:id25}}{\hyperref[\detokenize{index:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
 
@@
 \label{\detokenize{index:setup}}
-Download and unzip the compressed file to an appropriate folder.
 
@@
 
+\section{Google Calendar Configuration}
+\label{\detokenize{index:google-calendar-configuration}}
+
+\subsection{Creating a calendar}
+\label{\detokenize{index:creating-a-calendar}}
+
+\subsection{Adding holiday events}
+\label{\detokenize{index:adding-holiday-events}}
+
+\subsection{Adding clinician requests}
+\label{\detokenize{index:adding-clinician-requests}}\label{\detokenize{index:id1}}
+
 \chapter{Usage}
@@
 \section{Authentication}
-\label{\detokenize{index:authentication}}\label{\detokenize{index:id1}}
+\label{\detokenize{index:authentication}}\label{\detokenize{index:id2}}
 When the scheduler first makes a connection to Google calendar, it needs
@@
 \section{Clinician Configuration}
-\label{\detokenize{index:clinician-configuration}}\label{\detokenize{index:id2}}
+\label{\detokenize{index:clinician-configuration}}\label{\detokenize{index:id3}}
 Before we can generate a schedule, we need to create a configuration file
@@
 clinician should fulfill.
+
+\subsection{Creating a new configuration file}
+\label{\detokenize{index:creating-a-new-configuration-file}}
+By default, you will get a blank configuration file when you launch
+the program. If you would like to discard the changes you have made and
+start a configuration file from scratch, simply click on \sphinxtitleref{New Config}.
+
+\textless{}image placeholder\textgreater{}
+
+\begin{sphinxadmonition}{warning}{Warning:}
+Unsaved changes to a configuration file will be discarded upon clicking
+\sphinxtitleref{New Config}.
+\end{sphinxadmonition}
+
+\subsection{Saving the configuration file}
+\label{\detokenize{index:saving-the-configuration-file}}
+When you are ready to save the configuration you created, click on
+\sphinxtitleref{Save Config} and choose a place to save your file. Make note of the name
+and directory of the file so you could load it in future runs.
+
+\textless{}image placeholder\textgreater{}
+
+\subsection{Loading a configuration file}
+\label{\detokenize{index:loading-a-configuration-file}}
+If you would like to open a previously created configuration file, simply
+click on \sphinxtitleref{Open Config}, navigate to the location of the configuration file,
+and select it.
+
+\textless{}image placeholder\textgreater{}
 
@@
 \label{\detokenize{index:deleting-an-existing-clinician}}
+To delete an existing clinician, simply select a row corresponding
+to the clinician in the table and click on \sphinxtitleref{Delete Clinician}.
+
+\textless{}image placeholder\textgreater{}
 
@@
 \label{\detokenize{index:editing-an-existing-clinician}}
+To edit the information of a clinician, select a row corresponding
+to the clinician in the table and click on \sphinxtitleref{Edit Clinician}. You should
+see a dialog window where you can change the information.
 
-\subsection{Saving the configuration to a file}
-\label{\detokenize{index:saving-the-configuration-to-a-file}}
-
-\subsection{Loading a configuration file}
-\label{\detokenize{index:loading-a-configuration-file}}
-
-\section{Google Calendar Configuration}
-\label{\detokenize{index:google-calendar-configuration}}
-
-\subsection{Creating a calendar}
-\label{\detokenize{index:creating-a-calendar}}
-
-\subsection{Adding holiday events}
-\label{\detokenize{index:adding-holiday-events}}
-
-\subsection{Adding clinician requests}
-\label{\detokenize{index:adding-clinician-requests}}\label{\detokenize{index:id3}}
+\textless{}image placeholder\textgreater{}
 
@@
 \end{enumerate}
+
+\textless{}image placeholder\textgreater{}
 \begin{enumerate}
@@
 \end{enumerate}
+
+\textless{}image placeholder\textgreater{}
 
@@
 
+\textless{}image placeholder\textgreater{}
+
 Selecting the \sphinxtitleref{Yearly Export} option will generate an excel file with a single
@@
 
+\textless{}image placeholder\textgreater{}
+
 Selecting the \sphinxtitleref{Monthly Export} option will generate a more detailed breakdown
@@
 
+\textless{}image placeholder\textgreater{}
+
 \subsection{Publishing a schedule to Google Calendar}
 \label{\detokenize{index:publishing-a-schedule-to-google-calendar}}\label{\detokenize{index:id5}}
+If you are satisfied with the generated schedule, you can publish it
+to Google calendar, by clicking on \sphinxtitleref{Publish}.
+
+\textless{}image placeholder\textgreater{}
+
+\begin{sphinxadmonition}{warning}{Warning:}
+This process might take some time, especially if the number of events
+to be published is quite large.
+\end{sphinxadmonition}
+
+\subsection{Clearing a published schedule}
+\label{\detokenize{index:clearing-a-published-schedule}}
+If you want to remove all the published events created by the application,
+simply click on \sphinxtitleref{Clear}.
+
+\textless{}image placeholder\textgreater{}
+
+\begin{sphinxadmonition}{warning}{Warning:}
+This process might take some time, especially if the number of published
+events is quite large.
+\end{sphinxadmonition}
 

### 2019-01-31 docs/build/latex/ClinicianScheduler.tex
Add screenshots for auth + config section in manual

@@
 \title{Clinician Scheduler Documentation}
-\date{Jan 29, 2019}
+\date{Jan 31, 2019}
 \release{1.0}
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id6}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
+\phantomsection\label{\detokenize{index:id7}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id7}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
+\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
+\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
+\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
+\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
+\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
+\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
+\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
+\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
+\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
+\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
+\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
+\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
+\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
+\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
+\phantomsection\label{\detokenize{index:id24}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id24}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
+\phantomsection\label{\detokenize{index:id25}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id25}}{\hyperref[\detokenize{index:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
+\phantomsection\label{\detokenize{index:id26}}{\hyperref[\detokenize{index:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
 
@@
 \end{enumerate}
+
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/authentication/open\_in\_browser.png}{\sphinxincludegraphics{{open_in_browser}.png}}\end{figure}
+
+\begin{DUlineblock}{0em}
+\item[] 
+\end{DUlineblock}
 \begin{enumerate}
@@
 
+\end{enumerate}
+
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/authentication/access\_request.png}{\sphinxincludegraphics{{access_request}.png}}\end{figure}
+
+\begin{DUlineblock}{0em}
+\item[] 
+\end{DUlineblock}
+\begin{quote}
+
 \sphinxstylestrong{(Optional)} If you opened the access request manually in step 1, you will
 receive an authentication code which needs to be pasted into the command prompt
+\end{quote}
 
-\end{enumerate}
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/authentication/access\_request\_auth\_code.png}{\sphinxincludegraphics{{access_request_auth_code}.png}}\end{figure}
+
+\begin{DUlineblock}{0em}
+\item[] 
+\end{DUlineblock}
 \begin{enumerate}
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/configuration/new\_config.png}{\sphinxincludegraphics{{new_config}.png}}\end{figure}
 
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/configuration/save\_config.png}{\sphinxincludegraphics{{save_config}.png}}\end{figure}
 
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/configuration/load\_config.png}{\sphinxincludegraphics{{load_config}.png}}\end{figure}
 
 \subsection{Adding a new clinician}
-\label{\detokenize{index:adding-a-new-clinician}}
+\label{\detokenize{index:adding-a-new-clinician}}\label{\detokenize{index:id4}}
 1. From the configuration tab, click \sphinxtitleref{New Clinician}. You should see a
@@
 \centering
-\sphinxhref{\_static/images/add\_clinician/step1\_new\_clinician.png}{\sphinxincludegraphics{{step1_new_clinician}.png}}\end{figure}
+\sphinxhref{\_static/images/configuration/add\_clinician/step1\_new\_clinician.png}{\sphinxincludegraphics{{step1_new_clinician}.png}}\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/add\_clinician/step2\_add\_division.png}{\sphinxincludegraphics{{step2_add_division}.png}}\end{figure}
+\sphinxhref{\_static/images/configuration/add\_clinician/step2\_add\_division.png}{\sphinxincludegraphics{{step2_add_division}.png}}\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/add\_clinician/step3\_remove\_division.png}{\sphinxincludegraphics{{step3_remove_division}.png}}\end{figure}
+\sphinxhref{\_static/images/configuration/add\_clinician/step3\_remove\_division.png}{\sphinxincludegraphics{{step3_remove_division}.png}}\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/add\_clinician/step4\_add\_clinician.png}{\sphinxincludegraphics{{step4_add_clinician}.png}}\end{figure}
+\sphinxhref{\_static/images/configuration/add\_clinician/step4\_add\_clinician.png}{\sphinxincludegraphics{{step4_add_clinician}.png}}\end{figure}
 
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/configuration/delete\_clinician.png}{\sphinxincludegraphics{{delete_clinician}.png}}\end{figure}
 
@@
 to the clinician in the table and click on \sphinxtitleref{Edit Clinician}. You should
-see a dialog window where you can change the information.
+see a dialog window where you can change the information. For more
+details on how to enter data in the edit dialog, see {\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding a new clinician}}}}.
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[htbp]
+\centering
+\sphinxhref{\_static/images/configuration/edit\_clinician.png}{\sphinxincludegraphics{{edit_clinician}.png}}\end{figure}
 
@@
 \subsection{Generating a schedule}
-\label{\detokenize{index:generating-a-schedule}}\label{\detokenize{index:id4}}
+\label{\detokenize{index:generating-a-schedule}}\label{\detokenize{index:id5}}
 Once you have created a configuration file, you can switch over to the
@@
 \subsection{Publishing a schedule to Google Calendar}
-\label{\detokenize{index:publishing-a-schedule-to-google-calendar}}\label{\detokenize{index:id5}}
+\label{\detokenize{index:publishing-a-schedule-to-google-calendar}}\label{\detokenize{index:id6}}
 If you are satisfied with the generated schedule, you can publish it

### 2019-01-31 docs/build/latex/ClinicianScheduler.tex
Add images to "generating a schedule" instructions

@@
 
-\begin{figure}[htbp]
+\begin{figure}[h!]
 \centering
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step1\_load\_config.png}{\sphinxincludegraphics{{step1_load_config}.png}}\end{figure}
 \begin{enumerate}
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step2\_calendar\_year.png}{\sphinxincludegraphics{{step2_calendar_year}.png}}\end{figure}
 \begin{enumerate}
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step3\_calendar\_id.png}{\sphinxincludegraphics{{step3_calendar_id}.png}}\end{figure}
 \begin{enumerate}
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step4\_retrieve\_events.png}{\sphinxincludegraphics{{step4_retrieve_events}.png}}\end{figure}
 \begin{enumerate}
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step5\_number\_of\_blocks.png}{\sphinxincludegraphics{{step5_number_of_blocks}.png}}\end{figure}
 \begin{enumerate}
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step6\_generate\_schedule.png}{\sphinxincludegraphics{{step6_generate_schedule}.png}}\end{figure}
 

### 2019-01-31 docs/source/conf.py

@@ -123,6 +123,7 @@ latex_elements = {
     # Latex figure (float) alignment
     #
     # 'figure_align': 'htbp',
+    'figure_align': 'h!'
 }
 
 # Grouping the document tree into LaTeX files. List of tuples

### 2019-02-05 docs/build/latex/ClinicianScheduler.tex
Finish "publishing", "exporting" and "google calendar" sections + images

@@
 \title{Clinician Scheduler Documentation}
-\date{Jan 31, 2019}
+\date{Feb 05, 2019}
 \release{1.0}
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id7}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
+\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
+\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
+\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
+\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
+\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
+\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
+\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
+\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
+\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
+\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
+\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
+\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
+\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
+\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
+\phantomsection\label{\detokenize{index:id24}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id24}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
+\phantomsection\label{\detokenize{index:id25}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id25}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
+\phantomsection\label{\detokenize{index:id26}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id26}}{\hyperref[\detokenize{index:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
+\phantomsection\label{\detokenize{index:id27}}{\hyperref[\detokenize{index:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
 
@@
 \sphinxhref{\_static/images/google\_credentials/step2\_select\_a\_project.png}{\sphinxincludegraphics{{step2_select_a_project}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step3\_new\_project.png}{\sphinxincludegraphics{{step3_new_project}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step4\_create\_project.png}{\sphinxincludegraphics{{step4_create_project}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step5\_enable\_apis\_and\_services.png}{\sphinxincludegraphics{{step5_enable_apis_and_services}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step6\_search\_calendar\_api.png}{\sphinxincludegraphics{{step6_search_calendar_api}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step7\_enable\_calendar\_api.png}{\sphinxincludegraphics{{step7_enable_calendar_api}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step8\_create\_credentials.png}{\sphinxincludegraphics{{step8_create_credentials}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step9\_credentials\_form.png}{\sphinxincludegraphics{{step9_credentials_form}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step10\_oauth\_client\_id.png}{\sphinxincludegraphics{{step10_oauth_client_id}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/google\_credentials/step11\_oauth\_consent\_screen.png}{\sphinxincludegraphics{{step11_oauth_consent_screen}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
-
 \begin{sphinxadmonition}{attention}{Attention:}
@@
 \subsection{Creating a calendar}
-\label{\detokenize{index:creating-a-calendar}}
+\label{\detokenize{index:creating-a-calendar}}\begin{enumerate}
+\item {} 
+Navigate to \sphinxurl{https://calendar.google.com} and sign in to the appropriate account.
+
+\item {} 
+Click on the settings icon next on the left side, and select \sphinxtitleref{New Calendar}.
+
+\end{enumerate}
+
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/google\_calendar/create\_calendar/step2\_new\_calendar.png}{\sphinxincludegraphics{{step2_new_calendar}.png}}\end{figure}
+\begin{enumerate}
+\setcounter{enumi}{2}
+\item {} 
+Enter a name for your calendar and click \sphinxtitleref{Create calendar}.
+
+\end{enumerate}
+
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/google\_calendar/create\_calendar/step3\_create.png}{\sphinxincludegraphics{{step3_create}.png}}\end{figure}
 
@@
 \label{\detokenize{index:adding-holiday-events}}
+\begin{sphinxadmonition}{note}{Note:}
+It is recommended to create full-day events for holidays and clinician requests.
+\end{sphinxadmonition}
+
+Create an event by clicking on any cell in the calendar. Make sure that
+the name of the event you create starts with “\sphinxstylestrong{{[}holiday{]}}”, so that the
+scheduler can recognize the event correctly. Click \sphinxtitleref{Save} once you are done.
+
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/google\_calendar/create\_holiday.png}{\sphinxincludegraphics{{create_holiday}.png}}\end{figure}
 
@@
 \label{\detokenize{index:adding-clinician-requests}}\label{\detokenize{index:id1}}
+Create an event by clicking on any cell in the calendar. Make sure that
+the name of the event you create has the format: “\sphinxstylestrong{{[}request{]} Name}”, so that the
+scheduler can recognize the event correctly. Click \sphinxtitleref{Save} once you are done.
+
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/google\_calendar/create\_request.png}{\sphinxincludegraphics{{create_request}.png}}\end{figure}
+
+\begin{sphinxadmonition}{warning}{Warning:}
+It is important that the name you use in the event name matches the name in
+the clinician configuration you create in {\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}.
+\end{sphinxadmonition}
 
@@
 \sphinxhref{\_static/images/authentication/open\_in\_browser.png}{\sphinxincludegraphics{{open_in_browser}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \centering
-\sphinxhref{\_static/images/authentication/access\_request.png}{\sphinxincludegraphics{{access_request}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
-\begin{quote}
-
-\sphinxstylestrong{(Optional)} If you opened the access request manually in step 1, you will
-receive an authentication code which needs to be pasted into the command prompt
-\end{quote}
+\capstart
+\sphinxhref{\_static/images/authentication/access\_request.png}{\sphinxincludegraphics{{access_request}.png}}\caption{\sphinxstylestrong{(Optional)} If you opened the access request manually in step 1, you will
+receive an authentication code which needs to be pasted into the command prompt}\label{\detokenize{index:id7}}\end{figure}
 
@@
 \sphinxhref{\_static/images/authentication/access\_request\_auth\_code.png}{\sphinxincludegraphics{{access_request_auth_code}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/configuration/add\_clinician/step1\_new\_clinician.png}{\sphinxincludegraphics{{step1_new_clinician}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/configuration/add\_clinician/step2\_add\_division.png}{\sphinxincludegraphics{{step2_add_division}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/configuration/add\_clinician/step3\_remove\_division.png}{\sphinxincludegraphics{{step3_remove_division}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/configuration/add\_clinician/step4\_add\_clinician.png}{\sphinxincludegraphics{{step4_add_clinician}.png}}\end{figure}
-
-\begin{DUlineblock}{0em}
-\item[] 
-\end{DUlineblock}
 
@@
 
-\textless{}image placeholder\textgreater{}
-
 Selecting the \sphinxtitleref{Yearly Export} option will generate an excel file with a single
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/export\_schedule/export\_yearly.png}{\sphinxincludegraphics{{export_yearly}.png}}\end{figure}
 
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/export\_schedule/export\_monthly.png}{\sphinxincludegraphics{{export_monthly}.png}}\end{figure}
 
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/publish\_schedule/publish\_schedule.png}{\sphinxincludegraphics{{publish_schedule}.png}}\end{figure}
 
@@
 
-\textless{}image placeholder\textgreater{}
+\begin{figure}[h!]
+\centering
+\sphinxhref{\_static/images/scheduling/publish\_schedule/clear\_schedule.png}{\sphinxincludegraphics{{clear_schedule}.png}}\end{figure}
 

### 2019-02-05 docs/source/conf.py
Place figures in Latex in the same order as they are in the code

@@ -123,7 +123,7 @@ latex_elements = {
     # Latex figure (float) alignment
     #
     # 'figure_align': 'htbp',
-    'figure_align': 'h!'
+    'figure_align': 'H'
 }
 
 # Grouping the document tree into LaTeX files. List of tuples

### 2019-02-19 docs/build/latex/ClinicianScheduler.tex
Place figures in Latex in the same order as they are in the code
Add sample output section to manual

@@
 \title{Clinician Scheduler Documentation}
-\date{Feb 05, 2019}
+\date{Feb 19, 2019}
 \release{1.0}
@@
 
+\item {} 
+\phantomsection\label{\detokenize{index:id28}}{\hyperref[\detokenize{index:sample-output}]{\sphinxcrossref{Sample Output}}}
+
 \end{itemize}
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
@@
 
+\chapter{Sample Output}
+\label{\detokenize{index:sample-output}}
+Yearly output format
+
+\begin{figure}[H]
+\centering
+\sphinxhref{\_static/images/scheduling/sample\_yearly.png}{\sphinxincludegraphics{{sample_yearly}.png}}\end{figure}
+
+Monthly output format
+
+\begin{figure}[H]
+\centering
+\sphinxhref{\_static/images/scheduling/sample\_monthly.png}{\sphinxincludegraphics{{sample_monthly}.png}}\end{figure}
+
 \renewcommand{\indexname}{Index}

### 2019-02-21 scripts/make_exe.sh
Update make script + README

@@ -22,7 +22,4 @@ fi
 
 # run pyinstaller
 echo "Building scheduler.exe..."
-pyinstaller ../src/main.py -F -n scheduler --distpath ../dist_32
-
-echo "Building config_manager.exe..."
-pyinstaller ../src/config_manager.py -F --distpath ../dist_32
+pyinstaller ../src/main.py -F -n scheduler --distpath ../dist --hidden-import=PyQt5.sip

### 2019-02-21 scripts/make_exe64.sh

@@ -1,28 +1,0 @@
-#!/bin/bash
-
-# check for python 3
-ret=$(python -c 'import sys; print("%i" % (sys.hexversion>0x03000000))')
-if [ $ret -eq 0 ]; then
-    echo "Need Python 3+ to continue."
-    exit 1
-fi
-
-# check for 64bit python
-ret=$(python -c "import platform; print(platform.architecture()[0])")
-if ! [ $ret = "64bit" ]; then
-    echo "Need 64bit version of Python to continue."
-    exit 1
-fi
-
-# check for PyInstaller
-if ! [ $(command -v pyinstaller) ]; then
-    echo "Need PyInstaller to continue."
-    exit 1
-fi
-
-# run pyinstaller
-echo "Building scheduler.exe..."
-pyinstaller ../src/main.py -F -n scheduler --distpath ../dist_64
-
-echo "Building config_manager.exe..."
-pyinstaller ../src/config_manager.py -F --distpath ../dist_64

### 2019-06-13 src/helpers/apihelper.py
Force gapi lib to use port 443 in local webserver
Fixes #16.

@@ -1,5 +1,6 @@
 import os
 import sys
+import argparse
 
 from googleapiclient.discovery import build
 from httplib2 import Http
@@ -21,7 +22,12 @@ class ApiHelper:
         if not creds or creds.invalid:
             flow = client.flow_from_clientsecrets(
                 self.get_client_secret(), ApiHelper.SCOPE)
-            creds = tools.run_flow(flow, store)
+
+            # hack: force gapi library to use port 443
+            flags = tools.argparser.parse_args()
+            flags.auth_host_port = [443]
+
+            creds = tools.run_flow(flow, store, flags=flags)
 
         # discover calendar API
         self._service = build(

### 2019-06-14 src/services/scheduler.py
Implement shuffling of clinicians
Add checkbox enabling user to shuffle list of clinicians in order to
generate a different schedule.

Closes #27.

@@ -236,9 +236,9 @@ class Scheduler:
         self.set_data(clin_data)
         self.set_timeoff(timeoff_data)
 
-    def generate(self, debug=False):
+    def generate(self, debug=False, shuffle=False):
         self.setup_solver()
-        self.setup_problem()
+        self.setup_problem(shuffle=shuffle)
         ret = self.problem.solve(self.solver) == pulp.LpStatusOptimal
 
         if debug:
@@ -397,7 +397,7 @@ class Scheduler:
         self.holiday_map = lw
         self.long_weekends = list(lw.values())
     
-    def setup_problem(self):
+    def setup_problem(self, shuffle=False):
         """
         Constructs an LP program based on the clinician, division data.
         """
@@ -408,6 +408,8 @@ class Scheduler:
 
         divisions = list(self.divisions.values())
         clinicians = list(self.clinicians.values())
+
+        if shuffle: random.shuffle(clinicians)
 
         self._build_clinician_variables(divisions, clinicians)
         self._build_coverage_constraints(divisions, clinicians)

### 2019-06-14 src/ui/mainwindow.py

@@ -213,6 +213,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         retrieveLongWeekends = self.retrieveLongWeekendsCheckBox.isChecked()
         calendarYear = self.calendarYearSpinBox.value()
         calendarId = self.gCalLineEdit.text()
+        shuffle = self.shuffleCheckBox.isChecked()
 
         requests = []
         holidays = []
@@ -246,8 +247,10 @@ class MainWindow(QMainWindow, Ui_MainWindow):
             )
 
         # init scheduler with all the given data
-        schedule = scheduler.Scheduler(
-            num_blocks=numBlocks, clin_data=self.configuration, timeoff_data=requests, long_weekends=holidays).generate(debug=True)
+        schedule = scheduler \
+            .Scheduler(
+                num_blocks=numBlocks, clin_data=self.configuration, timeoff_data=requests, long_weekends=holidays) \
+            .generate(debug=True, shuffle=shuffle)
         if schedule is None:
             QMessageBox.critical(self, "Could not generate schedule!",
                 "Could not generate a schedule based on the given constraints and configuration. Try adjusting min/max values in the configuration tab.")

### 2019-06-14 src/ui/ui_mainwindow.py

@@ -128,6 +128,9 @@ class Ui_MainWindow(object):
         self.scheduleActions.setObjectName("scheduleActions")
         self.horizontalLayout_3 = QtWidgets.QHBoxLayout(self.scheduleActions)
         self.horizontalLayout_3.setObjectName("horizontalLayout_3")
+        self.shuffleCheckBox = QtWidgets.QCheckBox(self.scheduleActions)
+        self.shuffleCheckBox.setObjectName("shuffleCheckBox")
+        self.horizontalLayout_3.addWidget(self.shuffleCheckBox)
         self.generateScheduleButton = QtWidgets.QPushButton(self.scheduleActions)
         self.generateScheduleButton.setObjectName("generateScheduleButton")
         self.horizontalLayout_3.addWidget(self.generateScheduleButton)
@@ -171,7 +174,7 @@ class Ui_MainWindow(object):
         MainWindow.setCentralWidget(self.centralwidget)
 
         self.retranslateUi(MainWindow)
-        self.tabWidget.setCurrentIndex(0)
+        self.tabWidget.setCurrentIndex(1)
         QtCore.QMetaObject.connectSlotsByName(MainWindow)
 
     def retranslateUi(self, MainWindow):
@@ -203,6 +206,7 @@ class Ui_MainWindow(object):
         self.numberOfBlocksLabel.setText(_translate("MainWindow", "Number of Blocks"))
         self.retrieveLongWeekendsLabel.setText(_translate("MainWindow", "Retrieve Long Weekends"))
         self.scheduleActions.setTitle(_translate("MainWindow", "Schedule"))
+        self.shuffleCheckBox.setText(_translate("MainWindow", "Shuffle?"))
         self.generateScheduleButton.setToolTip(_translate("MainWindow", "<html><head/><body><p>Generate a new schedule using the given configuration file and calendar information</p></body></html>"))
         self.generateScheduleButton.setText(_translate("MainWindow", "Generate"))
         self.exportScheduleButton.setToolTip(_translate("MainWindow", "<html><head/><body><p>Export generated schedule as an Excel spreadsheet</p></body></html>"))

### 2019-06-14 paper/scheduler.tex
Move docs/ -> paper/
Fix up subscript in scheduler paper
Normalize objectives of linear program
Add more details on how to normalize the objectives in writeup.

Closes #28.

--- 2019-01-02 docs/scheduler.tex
@@
 \usepackage{bbm}
+\usepackage{physics}
 
@@
 			\end{equation}
+		\item Weekends should be roughly equally distributed between clinicians.
+			\begin{equation}
+				\floor*{ \frac{W}{C} } \leq \sum_{w} Y_{c, w} \leq \ceil*{ \frac{W}{C} } \text{ for each } c
+			\end{equation}
 		\item Long weekends should be roughly equally distributed between clinicians.
@@
 			\end{equation}
-		Each clinician should work a given number of long weekends. Roughly, this is the number of long weekends divided by the number of clinicians
 		\item A clinician cannot work two consecutive blocks (across all divisions)
@@
 				\label{eqn_block_obj}
-				\min \sum_{c, b, d} (-1)^{ I(b \in \mc{S}_c) } \cdot X_{c, b, d}
+				\min_X Q_1(X) = \min_X \sum_{c, b, d} (-1)^{ I(b \in \mc{S}_c) } \cdot X_{c, b, d}
 			\end{equation}
@@
 				\label{eqn_weekend_obj}
-				\min \sum_{c, w} (-1)^{ I (w \in \mc{T}_c)} \cdot Y_{c, w}
+				\min_Y Q_2(Y) = \min_Y \sum_{c, w} (-1)^{ I (w \in \mc{T}_c)} \cdot Y_{c, w}
 			\end{equation}
@@
 				\label{eqn_adjacency_obj}
-				\max \sum_{c, b, d} Z_{c, b, d} = \min \left(- \sum_{c, b, d} Z_{c, b, d}\right)
+				\max_Z \sum_{c, b, d} Z_{c, b, d} = \min_Z \left(- \sum_{c, b, d} Z_{c, b, d}\right) = \min_Z Q_3(Z)
 			\end{equation}
@@
 
-We can optimize the objectives simultaneously by optimizing a weighted sum of the objective functions. We chose to weigh each objective equally.
+In order to optimize all the objectives simultaneously we can use the weighted sum method which guarantees our solutions to be Pareto optimal (??).
+It is necessary to normalize the objective functions, since their domains are not equivalent.
+In general, if the objective function is of the from
+\[
+ 	f(\vb{x}) = \sum_i a_ix_i
+\]
+Then we normalize it by dividing by the sum of absolute weights, thus giving
+\[
+	\bar{f}(\vb{x}) = \frac{f(\vb{x})}{\sum_i \abs{a_i}}
+\]
+In our case, this gives us the following objective functions:
+	\begin{align*}
+		\bar{Q}_1(X) = \frac{Q_1(X)}{\sum_{c, b, d} 1} = \frac{Q_1(X)}{C \cdot B \cdot D} \\
+		\bar{Q}_2(Y) = \frac{Q_2(Y)}{\sum_{c, w} 1} = \frac{Q_2(Y)}{C \cdot W} \\
+		\bar{Q}_3(Z) = \frac{Q_1(Z)}{\sum_{c, b, d} 1} = \frac{Q_3(Z)}{C \cdot B \cdot D}
+	\end{align*}
+We also chose to weigh the objectives equally, thus giving us the final objective:
+	\begin{equation}
+		\min_{X, Y, Z} \frac{\bar{Q}_1(X) + \bar{Q}_2(Y) + \bar{Q}_3(Z)}{3}
+	\end{equation}
 \end{document}

### 2019-06-14 src/services/scheduler.py

@@ -420,10 +420,14 @@ class Scheduler:
         self._build_adjacency_variables(divisions)
         appeasement_objs = self._build_appeasement_objectives(clinicians)
 
+        num_clin = len(clinicians)
+        num_div = len(divisions)
+
+        # make sure to normalize objectives, and weigh them equally
         self.problem.setObjective(
-            (1 / 3) * appeasement_objs[0]
-            + (1 / 3) * appeasement_objs[1]
-            + (1 / 3) * self._build_adjacency_objective(clinicians)
+              (1 / 3) * (1 / (num_clin * self.num_blocks * num_div)) * appeasement_objs[0]
+            + (1 / 3) * (1 / (num_clin * self.num_weekends)) * appeasement_objs[1]
+            + (1 / 3) * (1 / (num_clin * self.num_blocks * num_div)) * self._build_adjacency_objective(clinicians)
         )
 
     def _build_clinician_variables(self, divisions, clinicians):

### 2019-06-14 scripts/make_exe.sh
Bundle cbc.exe with scheduler executable
Closes #29.

@@ -22,4 +22,4 @@ fi
 
 # run pyinstaller
 echo "Building scheduler.exe..."
-pyinstaller ../src/main.py -F -n scheduler --distpath ../dist --hidden-import=PyQt5.sip
+pyinstaller --distpath ../dist scheduler.spec

### 2019-06-14 src/services/scheduler.py

@@ -275,8 +275,9 @@ class Scheduler:
         if getattr(sys, 'frozen', False):
             # running in a bundle
             cwd = os.getcwd()
-            exe = 'cbc-2.9.9-x86\\bin\\cbc.exe'
-            solverdir = os.path.join(cwd, exe)
+            exe = 'cbc\\bin\\cbc.exe'
+            if hasattr(sys, '_MEIPASS'): solverdir = os.path.join(sys._MEIPASS, exe)
+            else: solverdir = exe
             self.solver = pulp.COIN_CMD(path=solverdir)
         else:
             # running from source

### 2019-06-19 src/services/scheduler.py
Implement on-off-on-off-on constraint (i.e spread)
Prevent on-off-on-off-on type of assignments to improve spread of blocks
for a given clinician. Document in more detail in pdf.

@@ -416,6 +416,7 @@ class Scheduler:
         self._build_coverage_constraints(divisions, clinicians)
         self._build_minmax_constraints(divisions)
         self._build_consec_constraints(clinicians)
+        self._build_spread_constraints(clinicians)
         self._build_longweekend_constraints(clinicians)
         self._build_weekend_constraints(clinicians)
         self._build_adjacency_variables(divisions)
@@ -499,6 +500,20 @@ class Scheduler:
                     )]
                 )
                 self.problem.add(sum_ <= 1)
+
+    def _build_spread_constraints(self, clinicians):
+        # we need at least 5 consecutive blocks to implement this constraint
+        # on-off-on-off-on
+        if self.num_blocks <= 5: return
+
+        for clinician in clinicians:
+            for block_num in range(1, self.num_blocks - 3):
+                # constraint: X_i + X_{i+2} + X_{i+4} <= 2
+                sum_ = pulp.lpSum(
+                    [_.get_var() for _ in clinician.get_block_vars(
+                        lambda x, block_num=block_num: x.block_num in (block_num, block_num + 2, block_num + 4))]
+                )
+                self.problem.add(sum_ <= 2)
 
     def _build_longweekend_constraints(self, clinicians):
         if self.long_weekends:

### 2019-06-21 docs/build/latex/ClinicianScheduler.tex
Create homepage + links to manual & privacy policy
Also clean and simplify README for updated version of app.

@@
 \title{Clinician Scheduler Documentation}
-\date{Feb 19, 2019}
+\date{Jun 21, 2019}
 \release{1.0}
@@
 \sphinxtableofcontents
-\phantomsection\label{\detokenize{index::doc}}
+\phantomsection\label{\detokenize{manual::doc}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id8}}{\hyperref[\detokenize{index:setup}]{\sphinxcrossref{Setup}}}
+\phantomsection\label{\detokenize{manual:id8}}{\hyperref[\detokenize{manual:setup}]{\sphinxcrossref{Setup}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id9}}{\hyperref[\detokenize{index:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
+\phantomsection\label{\detokenize{manual:id9}}{\hyperref[\detokenize{manual:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id10}}{\hyperref[\detokenize{index:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
+\phantomsection\label{\detokenize{manual:id10}}{\hyperref[\detokenize{manual:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id11}}{\hyperref[\detokenize{index:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
+\phantomsection\label{\detokenize{manual:id11}}{\hyperref[\detokenize{manual:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id12}}{\hyperref[\detokenize{index:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
+\phantomsection\label{\detokenize{manual:id12}}{\hyperref[\detokenize{manual:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id13}}{\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
+\phantomsection\label{\detokenize{manual:id13}}{\hyperref[\detokenize{manual:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id14}}{\hyperref[\detokenize{index:usage}]{\sphinxcrossref{Usage}}}
+\phantomsection\label{\detokenize{manual:id14}}{\hyperref[\detokenize{manual:usage}]{\sphinxcrossref{Usage}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id15}}{\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{Authentication}}}
+\phantomsection\label{\detokenize{manual:id15}}{\hyperref[\detokenize{manual:authentication}]{\sphinxcrossref{Authentication}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id16}}{\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
+\phantomsection\label{\detokenize{manual:id16}}{\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id17}}{\hyperref[\detokenize{index:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
+\phantomsection\label{\detokenize{manual:id17}}{\hyperref[\detokenize{manual:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id18}}{\hyperref[\detokenize{index:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
+\phantomsection\label{\detokenize{manual:id18}}{\hyperref[\detokenize{manual:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id19}}{\hyperref[\detokenize{index:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+\phantomsection\label{\detokenize{manual:id19}}{\hyperref[\detokenize{manual:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id20}}{\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+\phantomsection\label{\detokenize{manual:id20}}{\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id21}}{\hyperref[\detokenize{index:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
+\phantomsection\label{\detokenize{manual:id21}}{\hyperref[\detokenize{manual:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id22}}{\hyperref[\detokenize{index:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
+\phantomsection\label{\detokenize{manual:id22}}{\hyperref[\detokenize{manual:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id23}}{\hyperref[\detokenize{index:scheduling}]{\sphinxcrossref{Scheduling}}}
+\phantomsection\label{\detokenize{manual:id23}}{\hyperref[\detokenize{manual:scheduling}]{\sphinxcrossref{Scheduling}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{index:id24}}{\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
+\phantomsection\label{\detokenize{manual:id24}}{\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id25}}{\hyperref[\detokenize{index:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
+\phantomsection\label{\detokenize{manual:id25}}{\hyperref[\detokenize{manual:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id26}}{\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
+\phantomsection\label{\detokenize{manual:id26}}{\hyperref[\detokenize{manual:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{index:id27}}{\hyperref[\detokenize{index:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
+\phantomsection\label{\detokenize{manual:id27}}{\hyperref[\detokenize{manual:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{index:id28}}{\hyperref[\detokenize{index:sample-output}]{\sphinxcrossref{Sample Output}}}
+\phantomsection\label{\detokenize{manual:id28}}{\hyperref[\detokenize{manual:sample-output}]{\sphinxcrossref{Sample Output}}}
 
@@
 \chapter{Setup}
-\label{\detokenize{index:setup}}
+\label{\detokenize{manual:setup}}
 
 \section{Generating Google API credentials}
-\label{\detokenize{index:generating-google-api-credentials}}
+\label{\detokenize{manual:generating-google-api-credentials}}
 \begin{sphinxadmonition}{note}{Note:}
@@
 \section{Google Calendar Configuration}
-\label{\detokenize{index:google-calendar-configuration}}
+\label{\detokenize{manual:google-calendar-configuration}}
 
 \subsection{Creating a calendar}
-\label{\detokenize{index:creating-a-calendar}}\begin{enumerate}
+\label{\detokenize{manual:creating-a-calendar}}\begin{enumerate}
 \item {} 
@@
 \subsection{Adding holiday events}
-\label{\detokenize{index:adding-holiday-events}}
+\label{\detokenize{manual:adding-holiday-events}}
 \begin{sphinxadmonition}{note}{Note:}
@@
 \subsection{Adding clinician requests}
-\label{\detokenize{index:adding-clinician-requests}}\label{\detokenize{index:id1}}
+\label{\detokenize{manual:adding-clinician-requests}}\label{\detokenize{manual:id1}}
 Create an event by clicking on any cell in the calendar. Make sure that
@@
 It is important that the name you use in the event name matches the name in
-the clinician configuration you create in {\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}.
+the clinician configuration you create in {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}.
 \end{sphinxadmonition}
@@
 \chapter{Usage}
-\label{\detokenize{index:usage}}
+\label{\detokenize{manual:usage}}
 
 \section{Authentication}
-\label{\detokenize{index:authentication}}\label{\detokenize{index:id2}}
+\label{\detokenize{manual:authentication}}\label{\detokenize{manual:id2}}
 When the scheduler first makes a connection to Google calendar, it needs
@@
 that you specified. You may encounter this when using any functionality that
-interacts with Google calendar, for example in {\hyperref[\detokenize{index:generating-a-schedule}]{\sphinxcrossref{\DUrole{std,std-ref}{Generating a schedule}}}}
-or in {\hyperref[\detokenize{index:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{\DUrole{std,std-ref}{Publishing a schedule to Google Calendar}}}}.
+interacts with Google calendar, for example in {\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{\DUrole{std,std-ref}{Generating a schedule}}}}
+or in {\hyperref[\detokenize{manual:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{\DUrole{std,std-ref}{Publishing a schedule to Google Calendar}}}}.
 \begin{enumerate}
@@
 \sphinxhref{\_static/images/authentication/access\_request.png}{\sphinxincludegraphics{{access_request}.png}}\caption{\sphinxstylestrong{(Optional)} If you opened the access request manually in step 1, you will
-receive an authentication code which needs to be pasted into the command prompt}\label{\detokenize{index:id7}}\end{figure}
+receive an authentication code which needs to be pasted into the command prompt}\label{\detokenize{manual:id7}}\end{figure}
 
@@
 \section{Clinician Configuration}
-\label{\detokenize{index:clinician-configuration}}\label{\detokenize{index:id3}}
+\label{\detokenize{manual:clinician-configuration}}\label{\detokenize{manual:id3}}
 Before we can generate a schedule, we need to create a configuration file
@@
 \subsection{Creating a new configuration file}
-\label{\detokenize{index:creating-a-new-configuration-file}}
+\label{\detokenize{manual:creating-a-new-configuration-file}}
 By default, you will get a blank configuration file when you launch
@@
 \subsection{Saving the configuration file}
-\label{\detokenize{index:saving-the-configuration-file}}
+\label{\detokenize{manual:saving-the-configuration-file}}
 When you are ready to save the configuration you created, click on
@@
 \subsection{Loading a configuration file}
-\label{\detokenize{index:loading-a-configuration-file}}
+\label{\detokenize{manual:loading-a-configuration-file}}
 If you would like to open a previously created configuration file, simply
@@
 \subsection{Adding a new clinician}
-\label{\detokenize{index:adding-a-new-clinician}}\label{\detokenize{index:id4}}
+\label{\detokenize{manual:adding-a-new-clinician}}\label{\detokenize{manual:id4}}
 1. From the configuration tab, click \sphinxtitleref{New Clinician}. You should see a
@@
 \subsection{Deleting an existing clinician}
-\label{\detokenize{index:deleting-an-existing-clinician}}
+\label{\detokenize{manual:deleting-an-existing-clinician}}
 To delete an existing clinician, simply select a row corresponding
@@
 \subsection{Editing an existing clinician}
-\label{\detokenize{index:editing-an-existing-clinician}}
+\label{\detokenize{manual:editing-an-existing-clinician}}
 To edit the information of a clinician, select a row corresponding
@@
 see a dialog window where you can change the information. For more
-details on how to enter data in the edit dialog, see {\hyperref[\detokenize{index:adding-a-new-clinician}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding a new clinician}}}}.
+details on how to enter data in the edit dialog, see {\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding a new clinician}}}}.
 
@@
 \section{Scheduling}
-\label{\detokenize{index:scheduling}}
+\label{\detokenize{manual:scheduling}}
 
 \subsection{Generating a schedule}
-\label{\detokenize{index:generating-a-schedule}}\label{\detokenize{index:id5}}
+\label{\detokenize{manual:generating-a-schedule}}\label{\detokenize{manual:id5}}
 Once you have created a configuration file, you can switch over to the
@@
 From the scheduling tab, click on \sphinxtitleref{Load} and select the configuration
-file that you generated in {\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
+file that you generated in {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
 
@@
 \sphinxstylestrong{(Optional)} In order to retrieve the time-off requests that were populated
-in Google calendar in {\hyperref[\detokenize{index:adding-clinician-requests}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding clinician requests}}}}, we need to specify the
+in Google calendar in {\hyperref[\detokenize{manual:adding-clinician-requests}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding clinician requests}}}}, we need to specify the
 calendar ID. Open the calendar in your browser, and navigate to the
@@
 
-\sphinxstylestrong{(Optional)} See {\hyperref[\detokenize{index:authentication}]{\sphinxcrossref{\DUrole{std,std-ref}{Authentication}}}} in case you encounter problems
+\sphinxstylestrong{(Optional)} See {\hyperref[\detokenize{manual:authentication}]{\sphinxcrossref{\DUrole{std,std-ref}{Authentication}}}} in case you encounter problems
 with Google calendar at this stage.
@@
 some constraints by changing the minimum and maximum number of blocks of
-clinicians in the configuration file. See {\hyperref[\detokenize{index:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
+clinicians in the configuration file. See {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
 for more information on changing the configuration file.
@@
 \subsection{Exporting a schedule}
-\label{\detokenize{index:exporting-a-schedule}}
+\label{\detokenize{manual:exporting-a-schedule}}
 If you are satisfied with the generated schedule, you can choose to export
@@
 \subsection{Publishing a schedule to Google Calendar}
-\label{\detokenize{index:publishing-a-schedule-to-google-calendar}}\label{\detokenize{index:id6}}
+\label{\detokenize{manual:publishing-a-schedule-to-google-calendar}}\label{\detokenize{manual:id6}}
 If you are satisfied with the generated schedule, you can publish it
@@
 \subsection{Clearing a published schedule}
-\label{\detokenize{index:clearing-a-published-schedule}}
+\label{\detokenize{manual:clearing-a-published-schedule}}
 If you want to remove all the published events created by the application,
@@
 \chapter{Sample Output}
-\label{\detokenize{index:sample-output}}
+\label{\detokenize{manual:sample-output}}
 Yearly output format

### 2019-06-21 docs/source/conf.py

@@ -130,7 +130,7 @@ latex_elements = {
 # (source start file, target name, title,
 #  author, documentclass [howto, manual, or own class]).
 latex_documents = [
-    (master_doc, 'ClinicianScheduler.tex', 'Clinician Scheduler Documentation',
+    ('manual', 'ClinicianScheduler.tex', 'Clinician Scheduler Documentation',
      'David Landsman', 'manual'),
 ]
 

### 2019-06-24 src/ui/ui_mainwindow.py
Work on UI to remove GCal and add excel requests
* Remove GCal input fields + buttons
* Add Excel Requests & Holidays load button
* Add output window (for logging purposes)

@@ -1,8 +1,8 @@
 # -*- coding: utf-8 -*-
 
-# Form implementation generated from reading ui file 'designer\mainwindow.ui'
+# Form implementation generated from reading ui file 'src\designer\mainwindow.ui'
 #
-# Created by: PyQt5 UI code generator 5.11.2
+# Created by: PyQt5 UI code generator 5.11.3
 #
 # WARNING! All changes made in this file will be lost!
 
@@ -62,6 +62,8 @@ class Ui_MainWindow(object):
         self.schedulerTab.setObjectName("schedulerTab")
         self.verticalLayout_2 = QtWidgets.QVBoxLayout(self.schedulerTab)
         self.verticalLayout_2.setObjectName("verticalLayout_2")
+        self.settingsLayout = QtWidgets.QHBoxLayout()
+        self.settingsLayout.setObjectName("settingsLayout")
         self.settingsGroupBox = QtWidgets.QGroupBox(self.schedulerTab)
         sizePolicy = QtWidgets.QSizePolicy(QtWidgets.QSizePolicy.Fixed, QtWidgets.QSizePolicy.Preferred)
         sizePolicy.setHorizontalStretch(0)
@@ -69,54 +71,49 @@ class Ui_MainWindow(object):
         sizePolicy.setHeightForWidth(self.settingsGroupBox.sizePolicy().hasHeightForWidth())
         self.settingsGroupBox.setSizePolicy(sizePolicy)
         self.settingsGroupBox.setObjectName("settingsGroupBox")
-        self.horizontalLayout_2 = QtWidgets.QHBoxLayout(self.settingsGroupBox)
-        self.horizontalLayout_2.setObjectName("horizontalLayout_2")
+        self.horizontalLayout_6 = QtWidgets.QHBoxLayout(self.settingsGroupBox)
+        self.horizontalLayout_6.setObjectName("horizontalLayout_6")
         self.settingsForm = QtWidgets.QFormLayout()
         self.settingsForm.setObjectName("settingsForm")
         self.configFileLabel = QtWidgets.QLabel(self.settingsGroupBox)
         self.configFileLabel.setObjectName("configFileLabel")
         self.settingsForm.setWidget(0, QtWidgets.QFormLayout.LabelRole, self.configFileLabel)
-        self.loadButton = QtWidgets.QPushButton(self.settingsGroupBox)
-        self.loadButton.setObjectName("loadButton")
[…]

### 2019-06-24 src/helpers/logger.py
Add logging functionality
Logger class allows to write lines with different levels of priority to the
output window

@@ -1,0 +1,16 @@
+from datetime import datetime
+
+LEVELCOLOURS = {
+    'INFO': 'black',
+    'WARNING': 'orange',
+    'ERROR': 'red'
+}
+
+class Logger:
+    def __init__(self, output):
+        self.output = output
+
+    def write_line(self, line, level='INFO'):
+        timestamp = datetime.now().time().strftime('%H:%M:%S')
+        colour = LEVELCOLOURS[level]
+        self.output.insertHtml('<p style="color:{2}">[{0}] {1}</p><br>'.format(timestamp, line, colour))

### 2019-06-24 src/ui/mainwindow.py

@@ -12,6 +12,7 @@ from constants import WEEK_HOURS, WEEKEND_HOURS
 from helpers.apihelper import ApiHelper
 from helpers.excelhelper import ExcelHelper
 from helpers.uihelper import UiHelper
+from helpers.logger import Logger
 from services import scheduler
 
 from .dialog import DialogWindow
@@ -27,11 +28,12 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         self._configPath = ''
 
         self.setupUi(self)
+
+        self._logger = Logger(self.outputTextEdit)
         self.tabWidget.currentChanged[int].connect(self.updateTabText)
         self.setupConfigurationTab()
         self.setupSchedulerTab()
 
-        self.handleCalendarIdChange(self.gCalLineEdit.text())
         self.tabWidget.setCurrentIndex(0)
 
         self.show()
@@ -76,15 +78,10 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         self.newConfig()
 
     def setupSchedulerTab(self):
-        # action setup
-        self.gCalLineEdit.textChanged[str].connect(self.handleCalendarIdChange)
-
-        self.loadButton.clicked.connect(self.openConfig)
+        self.loadConfigButton.clicked.connect(self.openConfig)
         self.generateScheduleButton.clicked.connect(self.generateSchedule)
         self.exportScheduleButton.clicked.connect(self.exportSchedule)
         self.exportMonthlyButton.clicked.connect(self.exportMonthlySchedule)
-        self.publishCalendarButton.clicked.connect(self.publishSchedule)
-        self.clearCalendarButton.clicked.connect(self.clearCalendar)
 
         # misc vars
         self.holidayMap = {}
@@ -134,6 +131,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
                     self.configPath = path
                     self.configuration = json.load(f)
                     UiHelper.syncTreeView(self.treeView, self.model, self.configuration)
+                    self._logger.write_line('Loaded configuration file: {}'.format(path))
 
             except Exception as ex:
                 QMessageBox.critical(self, "Unable to open file", str(ex))

### 2019-06-24 src/ui/ui_mainwindow.py

@@ -109,9 +109,10 @@ class Ui_MainWindow(object):
         self.outputGroupBox.setObjectName("outputGroupBox")
         self.horizontalLayout_9 = QtWidgets.QHBoxLayout(self.outputGroupBox)
         self.horizontalLayout_9.setObjectName("horizontalLayout_9")
-        self.outputTextBrowser = QtWidgets.QTextBrowser(self.outputGroupBox)
-        self.outputTextBrowser.setObjectName("outputTextBrowser")
-        self.horizontalLayout_9.addWidget(self.outputTextBrowser)
+        self.outputTextEdit = QtWidgets.QTextEdit(self.outputGroupBox)
+        self.outputTextEdit.setReadOnly(True)
+        self.outputTextEdit.setObjectName("outputTextEdit")
+        self.horizontalLayout_9.addWidget(self.outputTextEdit)
         self.settingsLayout.addWidget(self.outputGroupBox)
         self.verticalLayout_2.addLayout(self.settingsLayout)
         self.schedulerButtonLayout = QtWidgets.QHBoxLayout()

### 2019-06-25 src/helpers/apihelper.py
Remove GCal-related methods + vars

@@ -1,112 +1,0 @@
-import os
-import sys
-import argparse
-
-from googleapiclient.discovery import build
-from httplib2 import Http
-from oauth2client import client
-from oauth2client import file as oauth_file
-from oauth2client import tools
-
-
-class ApiHelper:
-    """
-    A helper class that wraps some of the API methods of Google calendar.
-    """
-    SCOPE = 'https://www.googleapis.com/auth/calendar'
-
-    def __init__(self, calendar_id):
-        # retrieve credentials / create new
-        store = oauth_file.Storage('token.json')
-        creds = store.get()
-        if not creds or creds.invalid:
-            flow = client.flow_from_clientsecrets(
-                self.get_client_secret(), ApiHelper.SCOPE)
-
-            # hack: force gapi library to use port 443
-            flags = tools.argparser.parse_args()
-            flags.auth_host_port = [443]
-
-            creds = tools.run_flow(flow, store, flags=flags)
-
-        # discover calendar API
-        self._service = build(
-            'calendar', 'v3', http=creds.authorize(Http()))
-        self.calendar_id = calendar_id
-        self.time_zone = self.get_timezone()
-
-    def get_client_secret(self):
-        if getattr(sys, 'frozen', False):
-            # running in a bundle
-            cwd = os.getcwd()
-            file = 'credentials.json'
-            return os.path.join(cwd, file)
-        else:
-            # running from source
-            return 'credentials.json'
-
-    def get_timezone(self):
-        """
-        Returns the timezone of the calendar given by `calendar_id`.
-        """
-        # pylint: disable=maybe-no-member
-        result = self._service.calendars().get(
-            calendarId=self.calendar_id).execute()
-        return result.get('timeZone')
-
-    def get_events(self, start, end, search_str='', max_res=1000):
-        """
-        Returns a list of at most `max_res` events with dates between 
-        start and end.
-        """
-        # pylint: disable=maybe-no-member
-        result = self._service.events().list(
-            calendarId=self.calendar_id,
-            timeMin=start,
-            timeMax=end,
-            maxResults=max_res,
-            singleEvents=True,
-            orderBy='startTime').execute()
-
-        events = result.get('items', [])
-        return list(filter(lambda x, s=search_str: s in x['summary'], events))
-
-    def create_event(self, start, end, attendees, summary):
-        """
-        Creates a new event starting at `start` and ending at `end` with
-        the given `attendees` and `summary`
-        """
-        event = {
-            'summary': summary,
[…]

### 2019-06-25 src/ui/mainwindow.py

@@ -9,7 +9,6 @@ from PyQt5.QtGui import *
 from PyQt5.QtWidgets import *
 
 from constants import WEEK_HOURS, WEEKEND_HOURS
-from helpers.apihelper import ApiHelper
 from helpers.excelhelper import ExcelHelper
 from helpers.uihelper import UiHelper
 from helpers.logger import Logger
@@ -85,20 +84,6 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
         # misc vars
         self.holidayMap = {}
-
-    def handleCalendarIdChange(self, text):
-        if len(text) == 0:
-            # first, uncheck "retrieve" checkboxes
-            self.retrieveTimeOffRequestsCheckBox.setChecked(False)
-            self.retrieveLongWeekendsCheckBox.setChecked(False)
-            # then disable them altogether
-            self.retrieveTimeOffRequestsCheckBox.setDisabled(True)
-            self.retrieveLongWeekendsCheckBox.setDisabled(True)
-
-        else:
-            # enable "retrieve" checkboxes
-            self.retrieveTimeOffRequestsCheckBox.setDisabled(False)
-            self.retrieveLongWeekendsCheckBox.setDisabled(False)
 
     def updateTabText(self, selectedTabIdx):
         # clear (configPath) from all other tabs
@@ -207,42 +192,11 @@ class MainWindow(QMainWindow, Ui_MainWindow):
             return
 
         numBlocks = self.numberOfBlocksSpinBox.value()
-        retrieveTimeOff = self.retrieveTimeOffRequestsCheckBox.isChecked()
-        retrieveLongWeekends = self.retrieveLongWeekendsCheckBox.isChecked()
         calendarYear = self.calendarYearSpinBox.value()
-        calendarId = self.gCalLineEdit.text()
         shuffle = self.shuffleCheckBox.isChecked()
 
         requests = []
         holidays = []
-
-        if retrieveTimeOff:    
-            if not calendarId:
-                QMessageBox.critical(self, "Missing Calendar ID", "Please supply a calendar ID!")
-                return
-
-            # read timeoff requests from gcal
-            startDate = datetime(calendarYear, 1, 1)
-            endDate = startDate + timedelta(weeks=52)
-            requests = ApiHelper(calendarId).get_events(
-                start=startDate.isoformat() + 'Z',
-                end=endDate.isoformat() + 'Z',
-                search_str='[request]'
-            )
-
-        if retrieveLongWeekends:
-            if not calendarId:
[…]

### 2019-06-25 src/helpers/logger.py
Add logging to scheduler actions
Also fix bug where schedule generation crashed the app if using a local
configuration file (not saved)

@@ -13,4 +13,4 @@ class Logger:
     def write_line(self, line, level='INFO'):
         timestamp = datetime.now().time().strftime('%H:%M:%S')
         colour = LEVELCOLOURS[level]
-        self.output.insertHtml('<p style="color:{2}">[{0}] {1}</p><br>'.format(timestamp, line, colour))
+        self.output.insertHtml('<span style="color:{2}">[{0}] {1}</span><br>'.format(timestamp, line, colour))

### 2019-06-25 src/services/scheduler.py

@@ -223,7 +223,7 @@ class Scheduler:
     schedule based on the supplied clinician and division data.
     """
 
-    def __init__(self, num_blocks, clin_data={}, timeoff_data=[], long_weekends=[]):
+    def __init__(self, logger, num_blocks, clin_data={}, timeoff_data=[], long_weekends=[]):
         self.num_blocks = num_blocks
         self.num_weekends = num_blocks * BLOCK_SIZE
 
@@ -231,6 +231,7 @@ class Scheduler:
         self.divisions = {}
         self.holiday_map = {}
         self.long_weekends = []
+        self._logger = logger
         
         self.set_long_weekends(long_weekends)
         self.set_data(clin_data)
@@ -240,22 +241,23 @@ class Scheduler:
         self.setup_solver()
         self.setup_problem(shuffle=shuffle)
         ret = self.problem.solve(self.solver) == pulp.LpStatusOptimal
-
-        if debug:
-            print('objective value = {}'.format(pulp.value(self.problem.objective)))
-            print('conflicts per doc:')
-            for clinician in self.clinicians.values():
-                print(clinician.name)
-                assigned_blocksoff = clinician.get_block_vars(
-                    lambda x, clinician=clinician: x.block_num in clinician.blocks_off and x.get_value() == 1.0)
-                assigned_weekendsoff = clinician.get_weekend_vars(
-                    lambda x, clinician=clinician: x.week_num in clinician.weekends_off and x.get_value() == 1.0)
-                print('\t{} out of {} blocks'.format(
-                    len(assigned_blocksoff), len(clinician.blocks_off)))
-                print('\t{} out of {} weekends'.format(
-                    len(assigned_weekendsoff), len(clinician.weekends_off)))
                     
         if ret:
+            if debug:
+                self._logger.write_line('Objective function value: {}'.format(pulp.value(self.problem.objective)))
+                conflicts_str = 'Schedule Conflicts:'
+                for clinician in self.clinicians.values():
+                    assigned_blocksoff = clinician.get_block_vars(
+                        lambda x, clinician=clinician: x.block_num in clinician.blocks_off and x.get_value() == 1.0)
+                    assigned_weekendsoff = clinician.get_weekend_vars(
+                        lambda x, clinician=clinician: x.week_num in clinician.weekends_off and x.get_value() == 1.0)
+                    
+                    conflicts_str += ' {0} ({1}/{2} blocks, {3}/{4} weekends)'.format(
[…]

### 2019-06-25 src/ui/mainwindow.py

@@ -133,6 +133,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
                     UiHelper.convertTreeToDict(self.model.rootItem, data)
                     json.dump(data, f)
                     self.configPath = fileName
+                    self._logger.write_line('Saved configuration file: {}'.format(fileName))
 
             except Exception as ex:
                 QMessageBox.critical(self, "Unable to save file", str(ex))
@@ -200,14 +201,14 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
         # init scheduler with all the given data
         schedule = scheduler \
-            .Scheduler(
-                num_blocks=numBlocks, clin_data=self.configuration, timeoff_data=requests, long_weekends=holidays) \
+            .Scheduler(logger=self._logger, num_blocks=numBlocks, clin_data=self.configuration, \
+                 timeoff_data=requests, long_weekends=holidays) \
             .generate(debug=True, shuffle=shuffle)
         if schedule is None:
-            QMessageBox.critical(self, "Could not generate schedule!",
-                "Could not generate a schedule based on the given constraints and configuration. Try adjusting min/max values in the configuration tab.")
+            self._logger.write_line('Could not generate schedule! Try adjusting min/max values in the configuration tab.', level='ERROR')
         
         else:
+            self._logger.write_line('Generated schedule!')
             divAssignments = schedule[0]
             weekendAssignments = schedule[1]
             self.holidayMap = schedule[2]

### 2019-06-25 src/helpers/logger.py
Make UI more compact by grouping Setup + Schedule
Fix bug where logger output would be overwritten if user had selected text.

@@ -1,4 +1,6 @@
+# pylint: disable=undefined-variable
 from datetime import datetime
+from PyQt5.QtGui import *
 
 LEVELCOLOURS = {
     'INFO': 'black',
@@ -13,4 +15,6 @@ class Logger:
     def write_line(self, line, level='INFO'):
         timestamp = datetime.now().time().strftime('%H:%M:%S')
         colour = LEVELCOLOURS[level]
+
+        self.output.moveCursor(QTextCursor.End)
         self.output.insertHtml('<span style="color:{2}">[{0}] {1}</span><br>'.format(timestamp, line, colour))

### 2019-06-25 src/main.py

@@ -11,6 +11,11 @@ def main():
     app = QApplication([])
     app.setApplicationName("Clinician Scheduler")
 
+    # increase global font size
+    global_font = app.font()
+    global_font.setPointSize(11)
+    app.setFont(global_font)
+
     window = MainWindow()
     app.exec_()
 

### 2019-06-25 src/ui/ui_mainwindow.py

@@ -62,49 +62,77 @@ class Ui_MainWindow(object):
         self.schedulerTab.setObjectName("schedulerTab")
         self.verticalLayout_2 = QtWidgets.QVBoxLayout(self.schedulerTab)
         self.verticalLayout_2.setObjectName("verticalLayout_2")
-        self.settingsLayout = QtWidgets.QHBoxLayout()
-        self.settingsLayout.setObjectName("settingsLayout")
-        self.settingsGroupBox = QtWidgets.QGroupBox(self.schedulerTab)
+        self.controlsLayout = QtWidgets.QHBoxLayout()
+        self.controlsLayout.setObjectName("controlsLayout")
+        self.setupScheduleGroupBox = QtWidgets.QVBoxLayout()
+        self.setupScheduleGroupBox.setObjectName("setupScheduleGroupBox")
+        self.setupGroupBox = QtWidgets.QGroupBox(self.schedulerTab)
         sizePolicy = QtWidgets.QSizePolicy(QtWidgets.QSizePolicy.Fixed, QtWidgets.QSizePolicy.Preferred)
         sizePolicy.setHorizontalStretch(0)
         sizePolicy.setVerticalStretch(0)
-        sizePolicy.setHeightForWidth(self.settingsGroupBox.sizePolicy().hasHeightForWidth())
-        self.settingsGroupBox.setSizePolicy(sizePolicy)
-        self.settingsGroupBox.setObjectName("settingsGroupBox")
-        self.horizontalLayout_6 = QtWidgets.QHBoxLayout(self.settingsGroupBox)
+        sizePolicy.setHeightForWidth(self.setupGroupBox.sizePolicy().hasHeightForWidth())
+        self.setupGroupBox.setSizePolicy(sizePolicy)
+        self.setupGroupBox.setObjectName("setupGroupBox")
+        self.horizontalLayout_6 = QtWidgets.QHBoxLayout(self.setupGroupBox)
         self.horizontalLayout_6.setObjectName("horizontalLayout_6")
-        self.settingsForm = QtWidgets.QFormLayout()
-        self.settingsForm.setObjectName("settingsForm")
-        self.configFileLabel = QtWidgets.QLabel(self.settingsGroupBox)
+        self.setupForm = QtWidgets.QFormLayout()
+        self.setupForm.setObjectName("setupForm")
+        self.configFileLabel = QtWidgets.QLabel(self.setupGroupBox)
         self.configFileLabel.setObjectName("configFileLabel")
[…]

### 2019-06-25 src/main.py
Implement Verbose Output checkbox functionality
Make font change in .ui file rather than in main.

@@ -11,11 +11,6 @@ def main():
     app = QApplication([])
     app.setApplicationName("Clinician Scheduler")
 
-    # increase global font size
-    global_font = app.font()
-    global_font.setPointSize(11)
-    app.setFont(global_font)
-
     window = MainWindow()
     app.exec_()
 

### 2019-06-25 src/services/scheduler.py

@@ -237,13 +237,13 @@ class Scheduler:
         self.set_data(clin_data)
         self.set_timeoff(timeoff_data)
 
-    def generate(self, debug=False, shuffle=False):
+    def generate(self, verbose=False, shuffle=False):
         self.setup_solver()
         self.setup_problem(shuffle=shuffle)
         ret = self.problem.solve(self.solver) == pulp.LpStatusOptimal
                     
         if ret:
-            if debug:
+            if verbose:
                 self._logger.write_line('Objective function value: {}'.format(pulp.value(self.problem.objective)))
                 conflicts_str = 'Schedule Conflicts:'
                 for clinician in self.clinicians.values():

### 2019-06-25 src/ui/mainwindow.py

@@ -195,6 +195,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         numBlocks = self.numberOfBlocksSpinBox.value()
         calendarYear = self.calendarYearSpinBox.value()
         shuffle = self.shuffleCheckBox.isChecked()
+        verbose = self.verboseCheckBox.isChecked()
 
         requests = []
         holidays = []
@@ -203,7 +204,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         schedule = scheduler \
             .Scheduler(logger=self._logger, num_blocks=numBlocks, clin_data=self.configuration, \
                  timeoff_data=requests, long_weekends=holidays) \
-            .generate(debug=True, shuffle=shuffle)
+            .generate(verbose=verbose, shuffle=shuffle)
         if schedule is None:
             self._logger.write_line('Could not generate schedule! Try adjusting min/max values in the configuration tab.', level='ERROR')
         

### 2019-06-25 src/ui/ui_mainwindow.py

@@ -12,11 +12,17 @@ class Ui_MainWindow(object):
     def setupUi(self, MainWindow):
         MainWindow.setObjectName("MainWindow")
         MainWindow.resize(800, 600)
+        font = QtGui.QFont()
+        font.setPointSize(11)
+        MainWindow.setFont(font)
         self.centralwidget = QtWidgets.QWidget(MainWindow)
         self.centralwidget.setObjectName("centralwidget")
         self.gridLayout = QtWidgets.QGridLayout(self.centralwidget)
         self.gridLayout.setObjectName("gridLayout")
         self.tabWidget = QtWidgets.QTabWidget(self.centralwidget)
+        font = QtGui.QFont()
+        font.setPointSize(10)
+        self.tabWidget.setFont(font)
         self.tabWidget.setObjectName("tabWidget")
         self.configurationTab = QtWidgets.QWidget()
         self.configurationTab.setObjectName("configurationTab")

### 2019-06-26 src/helpers/excelhelper.py
Work on method to load requests from spreadsheet
Also split requests & holidays button into two

@@ -4,7 +4,7 @@ import operator
 from collections import namedtuple
 from datetime import datetime, timedelta
 
-from openpyxl import Workbook
+from openpyxl import Workbook, load_workbook
 from openpyxl.styles import Border, Side, Alignment, colors
 from openpyxl.utils import get_column_letter
 
@@ -208,6 +208,32 @@ class ExcelHelper:
         # remove the automatically created first sheet
         wb.remove(wb.active)
         wb.save(filename)
+
+    @staticmethod
+    def loadRequests(filename):
+        Range = namedtuple('Range', ['start', 'end'])
+        request_dict = {}
+
+        wb = load_workbook(filename, read_only=True)
+        for sheet in wb.sheetnames:
+            ws = wb[sheet]
+            row = 1
+                        
+            while True:
+                start = ws['A{0}'.format(row)].value
+                end = ws['B{0}'.format(row)].value
+
+                if start is None or end is None: break 
+                    
+                if sheet in request_dict:
+                    request_dict[sheet].append(Range(start, end))
+                else:
+                    request_dict[sheet] = [Range(start, end)]
+                
+                row += 1
+
+        wb.close()
+        return request_dict
 
     @staticmethod
     def getColumn(table, col_idx):

### 2019-06-26 src/ui/mainwindow.py

@@ -25,6 +25,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
         self._configuration = {}
         self._configPath = ''
+        self._requests = {}
 
         self.setupUi(self)
 
@@ -78,6 +79,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
     def setupSchedulerTab(self):
         self.loadConfigButton.clicked.connect(self.openConfig)
+        self.loadRequestsButton.clicked.connect(self.openRequests)
         self.generateScheduleButton.clicked.connect(self.generateSchedule)
         self.exportScheduleButton.clicked.connect(self.exportSchedule)
         self.exportMonthlyButton.clicked.connect(self.exportMonthlySchedule)
@@ -120,6 +122,19 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
             except Exception as ex:
                 QMessageBox.critical(self, "Unable to open file", str(ex))
+
+    def openRequests(self):
+        path, _ = QFileDialog.getOpenFileName(
+            self, "Open file", "", "Excel files (*.xlsx *.xlsm)"
+        )
+
+        if path != '':
+            try:
+                self._requests = ExcelHelper.loadRequests(path)
+                self._logger.write_line('Loaded requests file: {}'.format(path))
+
+            except Exception as ex:
+                QMessageBox.critical(self, "Unable to load requests", str(ex))
 
     def saveConfig(self):
         fileName, _ = QFileDialog.getSaveFileName(

### 2019-06-26 src/ui/ui_mainwindow.py

@@ -13,16 +13,13 @@ class Ui_MainWindow(object):
         MainWindow.setObjectName("MainWindow")
         MainWindow.resize(800, 600)
         font = QtGui.QFont()
-        font.setPointSize(11)
+        font.setPointSize(10)
         MainWindow.setFont(font)
         self.centralwidget = QtWidgets.QWidget(MainWindow)
         self.centralwidget.setObjectName("centralwidget")
         self.gridLayout = QtWidgets.QGridLayout(self.centralwidget)
         self.gridLayout.setObjectName("gridLayout")
         self.tabWidget = QtWidgets.QTabWidget(self.centralwidget)
-        font = QtGui.QFont()
-        font.setPointSize(10)
-        self.tabWidget.setFont(font)
         self.tabWidget.setObjectName("tabWidget")
         self.configurationTab = QtWidgets.QWidget()
         self.configurationTab.setObjectName("configurationTab")
@@ -89,28 +86,34 @@ class Ui_MainWindow(object):
         self.loadConfigButton = QtWidgets.QPushButton(self.setupGroupBox)
         self.loadConfigButton.setObjectName("loadConfigButton")
         self.setupForm.setWidget(0, QtWidgets.QFormLayout.FieldRole, self.loadConfigButton)
-        self.requestsHolidaysLabel = QtWidgets.QLabel(self.setupGroupBox)
-        self.requestsHolidaysLabel.setObjectName("requestsHolidaysLabel")
-        self.setupForm.setWidget(1, QtWidgets.QFormLayout.LabelRole, self.requestsHolidaysLabel)
-        self.loadRequestsHolidaysButton = QtWidgets.QPushButton(self.setupGroupBox)
-        self.loadRequestsHolidaysButton.setObjectName("loadRequestsHolidaysButton")
-        self.setupForm.setWidget(1, QtWidgets.QFormLayout.FieldRole, self.loadRequestsHolidaysButton)
+        self.requestsLabel = QtWidgets.QLabel(self.setupGroupBox)
+        self.requestsLabel.setObjectName("requestsLabel")
+        self.setupForm.setWidget(1, QtWidgets.QFormLayout.LabelRole, self.requestsLabel)
+        self.loadRequestsButton = QtWidgets.QPushButton(self.setupGroupBox)
+        self.loadRequestsButton.setObjectName("loadRequestsButton")
[…]

### 2019-06-27 paper/paper.intro.tex
Add paper parts and files

In order for various hospital departments to function efficiently it is necessary for clinicians to have carefully allocated on-call schedules that simultaneously ensure sufficient resources are provided to their patients while not overworking clinicians to prevent \textit{errors} (?). When generating these schedules, there are many variables and constraints that need to be taken into account, such as preventing multiple consecutive assignments for a given clinician or ensuring that assignments are spread out evenly throughout a certain period of time. It is very common for departments to opt for a manual method of generating these schedules [??]. Unfortunately, such methods are unreliable and can lead to unfavourable assignments, since a manual method may miss certain constraints or forget to account for a certain variable during the process. \\

Automated methods for scheduling in hospital settings have been studied extensively, most notably in the context of the Nurse Scheduling Problem (NSP) in Operations Research. In an NSP, the goal is to assign nurses to shifts, while satisfying various constraints pertaining to both hospital regulations and nurse preferences. The constraints can be classified as soft or hard constraints, based on the importance of satisfying them. Typically, it is necessary for a solution to satisfy all of the hard constraints, while the soft constraints are considered optional. The NSP, in its general case, is known to be NP-complete, meaning that although it is possible to efficiently check whether a given schedule satisfies all the constraints using an algorithm, it is not yet known whether an efficient algorithm for finding a satisfying schedule exists, and it is in fact equivalent to the hardest computational problems in Computer Science, Operations Research, and \textit{etc...} [??]. In order to tackle NP-complete problems such as NSP in real-life, algorithms either try to use various heuristics to approximate a potential solution rather than finding an exact solution or they focus on relatively small problem sizes, ensuring that the runtime of the algorithm is still practical. \textit{Some examples of heuristic approaches...} . \textit{Some examples of exact approaches (LP, etc.)} . A comprehensive review of various approaches is presented by Burke et al. [??]. \\

In this paper we present the application of two methods, Network Flows and Integer Linear Programming (ILP), to a variation of the Nurse Scheduling Problem in order to generate a yearly schedule for clinicians working simultaneously at two different divisions of St. Michael's Hospital in Toronto, ON. The paper is organized as follows. Section [??] contains a detailed description of the scenario and a mathematical formalization of the problem. Section [??] describes first the Network Flow approach followed by the ILP approach to tackle the problem. Section [??] presents the results of both approaches when applied to the clinician data from St. Michael's Hospital. Section [??] presents results from simulations with a variety of problem sizes. We conclude the paper with a discussion of the real-world and simulation results in Section [??].

### 2019-06-27 paper/paper.tex
Add paper parts and files

\documentclass[]{article}

\begin{document}
	\begin{abstract}
		\input{paper.abstract.tex}
	\end{abstract}

	\section{Introduction} 
		\input{paper.intro.tex}
	\section{Problem}
		\input{paper.problem.tex}
	\section{Methods}
		\input{paper.methods.tex}
	\section{Results}
		\input{paper.results.tex}
	\section{Simulations}
		\input{paper.simulations.tex}
	\section{Discussion}
		\input{paper.discussion.tex}
\end{document}

### 2019-07-05 src/helpers/excelhelper.py
Improve error/exception presentation
Also fix bug where monthly export can crash app if no data in schedule table

@@ -215,7 +215,8 @@ class ExcelHelper:
             ExcelHelper.expandColumns(ws)
             
         # remove the automatically created first sheet
-        wb.remove(wb.active)
+        if len(wb.sheetnames) > 1:
+            wb.remove(wb.active)
         wb.save(filename)
 
     @staticmethod

### 2019-07-05 src/main.py

@@ -18,4 +18,4 @@ if __name__ == '__main__':
     try:
         main()
     except Exception as ex:
-        QMessageBox.critical(app, "Unknown error occurred!", str(ex))
+        QMessageBox.critical(app, "", "Unknown error occurred!\nDetails: {}".format(str(ex)))

### 2019-07-05 src/ui/mainwindow.py

@@ -125,7 +125,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
                     self._logger.write_line('Loaded configuration file: {}'.format(path))
 
             except Exception as ex:
-                QMessageBox.critical(self, "Unable to open file", str(ex))
+                QMessageBox.critical(self, "", "Unable to open file!\nDetails: {}".format(str(ex)))
                 self._logger.write_line('Unable to open configuration file. {}'.format(str(ex)), level='ERROR')
 
     def openRequests(self):
@@ -139,7 +139,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
                 self._logger.write_line('Loaded requests file: {}'.format(path))
 
             except Exception as ex:
-                QMessageBox.critical(self, "Unable to load requests", str(ex))
+                QMessageBox.critical(self, "", "Unable to load requests!\nDetails: {}".format(str(ex)))
                 self._logger.write_line('Unable to load requests. {}'.format(str(ex)), level='ERROR')
 
     def openHolidays(self):
@@ -153,7 +153,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
                 self._logger.write_line('Loaded holidays file: {}'.format(path))
 
             except Exception as ex:
-                QMessageBox.critical(self, "Unable to load holidays", str(ex))
+                QMessageBox.critical(self, "", "Unable to load holidays!\nDetails: {}".format(str(ex)))
                 self._logger.write_line('Unable to load holidays. {}'.format(str(ex)), level='ERROR')
 
     def saveConfig(self):
@@ -171,7 +171,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
                     self._logger.write_line('Saved configuration file: {}'.format(fileName))
 
             except Exception as ex:
-                QMessageBox.critical(self, "Unable to save file", str(ex))
+                QMessageBox.critical(self, "", "Unable to save file!\nDetails: {}".format(str(ex)))
 
     def newConfig(self):
         # clear data
@@ -224,7 +224,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
         numClinicians = len(self.configuration.keys())
         if numClinicians <= 0:
-            QMessageBox.critical(self, "No Clinicians Configured", "Please load a configuration file with at least one clinician!")
[…]

### 2019-07-05 paper/paper.intro.tex
Rewrite portions of introduction
Add bits of problem description

@@
-In order for various hospital departments to function efficiently it is necessary for clinicians to have carefully allocated on-call schedules that simultaneously ensure sufficient resources are provided to their patients while not overworking clinicians to prevent \textit{errors} (?).
-When generating these schedules, there are many variables and constraints that need to be taken into account, such as preventing multiple consecutive assignments for a given clinician or ensuring that assignments are spread out evenly throughout a certain period of time.
+Hospital departments provide crucial services to a great deal of people all over the world.
+The demand created by patients and other consumers of a hospital vastly exceeds the available supply.
+As such, it is crucial that a hospital department allocates its resources, such as the availability of clinicians, optimally, to provide the best possible service for its patients.
+In particular, in order for various hospital departments to function efficiently it is necessary for clinicians to have carefully allocated on-call schedules that simultaneously ensure sufficient resources are provided to their patients while not overworking clinicians to prevent costly mistakes.
+It is common practice for on-call schedules to be created manually, although this method is prone to errors.
+Namely, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method is unlikely to provide an optimal schedule.
+Moreover, such methods are likely to overlook certain constraints that must be maintained to have an operational department.
+For these reasons, it is important to develop automated methods that are able to find optimal schedules that satisfy the given constraints.
+\\
+
+%When generating these schedules, there are many variables and constraints that need to be taken into account, such as preventing multiple consecutive assignments for a given clinician or ensuring that assignments are spread out evenly throughout a certain period of time.
 It is very common for departments to opt for a manual method of generating these schedules [??].
@@
 
-Automated methods for scheduling in hospital settings have been studied extensively, most notably in the context of the Nurse Scheduling Problem (NSP) in Operations Research.
+Automated methods for tackling optimization problems related to scheduling have been studied extensively and applied in many industries, including transportation [??], manufacturing [??], [...].
+Of special interest to our work are the approaches to tackle the nurse scheduling problem.
+In this problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences.
+A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??].
+An extensive literature review of these and other methods is presented by [??].
+We will briefly summarize the main ideas in some of these approaches.
+\\
+
+[...] Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation.
+While this allows for finding precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians, it makes it very difficult to adapt a particular approach to other departments in the same hospital or other hospitals.
+Moreover, the large number of variables and constraints also leads to computational complexity issues, especially when using exact methods for finding the solution.
+\\
+
+[...]
+
+%Automated methods for scheduling in hospital settings have been studied extensively, most notably in the context of the Nurse Scheduling Problem (NSP) in Operations Research.
 In an NSP, the goal is to assign nurses to shifts, while satisfying various constraints pertaining to both hospital regulations and nurse preferences.
@@
 
-In this paper we present the application of two methods, Network Flows and Integer Linear Programming (ILP), to a variation of the Nurse Scheduling Problem in order to generate a yearly schedule for clinicians working simultaneously at two different divisions of St.
+%In this paper we present the application of two methods, Network Flows and Integer Linear Programming (ILP), to a variation of the Nurse Scheduling Problem in order to generate a yearly schedule for clinicians working simultaneously at two different divisions of St.
 Michael's Hospital in Toronto, ON.

### 2019-07-05 paper/paper.problem.tex
Rewrite portions of introduction
Add bits of problem description

Clinicians at the Infectious Diseases (ID) and HIV departments of St. Michael's Hospital are scheduled throughout the year to receive patients during on-call hours. Each clinician is typically scheduled for a full week or a full weekend of on-call service. To prevent under- and over-working of clinicians, they each have a minimum and maximum number of weeks that they are required to work. Moreover, during holidays weekends, the work-load for on-call service increases drastically, and it is important to distribute these weekends equally. At the same time, clinicians often request to not be put on service during certain days or weeks, and the generated schedule should attempt to respect those requests as best as possible. Other considerations also need to be taken into account, such as making sure someone is available for on-call service at any time of the year, and preventing multiple back-to-back assignments for a single clinician. It is clear that ensuring all of these conditions are met manually, especially with an increasing number of clinicians, is a difficult task and can lead to mismanagement of the schedule. Therefore, in this paper we attempt to develop an efficient algorithm that can generate a satisfying schedule, while optimizing clinician and patient contentment. \\

In order to develop an algorithm, we need to formalize the variables and constraints of the problem mathematically. Table [??] presents the sets and indices that are used in the definition of the constraints. [...] \\

[...]

### 2019-07-05 paper/paper.intro.tex
Expand intro to describe our goals

@@
 In this problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences.
-A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??].
+A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??].
 An extensive literature review of these and other methods is presented by [??].
-We will briefly summarize the main ideas in some of these approaches.
+We will briefly summarize the main ideas of some of these approaches.
 \\
@@
 Moreover, the large number of variables and constraints also leads to computational complexity issues, especially when using exact methods for finding the solution.
+In this paper, we will tackle a version of the nurse scheduling problem arising from a case study of two on-call divisions of St.
+Michael's Hospital in Toronto, ON.
+Our goal is to present a simple formulation for the problem that has been developed and tested at the hospital after switching from a manual approach to scheduling, as well as analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach.
+We hope that the formulation we present is adaptable, at least as a basic framework, for solving similar problems in other departments.
 \\

### 2019-07-08 docs/build/latex/ClinicianScheduler.tex
Remove GCal references in manual + rearrange it

@@
 \title{Clinician Scheduler Documentation}
-\date{Jun 21, 2019}
+\date{Jul 08, 2019}
 \release{1.0}
@@
 \item {} 
-\phantomsection\label{\detokenize{manual:id8}}{\hyperref[\detokenize{manual:setup}]{\sphinxcrossref{Setup}}}
+\phantomsection\label{\detokenize{manual:id5}}{\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{manual:id9}}{\hyperref[\detokenize{manual:generating-google-api-credentials}]{\sphinxcrossref{Generating Google API credentials}}}
+\phantomsection\label{\detokenize{manual:id6}}{\hyperref[\detokenize{manual:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id10}}{\hyperref[\detokenize{manual:google-calendar-configuration}]{\sphinxcrossref{Google Calendar Configuration}}}
+\phantomsection\label{\detokenize{manual:id7}}{\hyperref[\detokenize{manual:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
+
+\item {} 
+\phantomsection\label{\detokenize{manual:id8}}{\hyperref[\detokenize{manual:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+
+\item {} 
+\phantomsection\label{\detokenize{manual:id9}}{\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+
+\item {} 
+\phantomsection\label{\detokenize{manual:id10}}{\hyperref[\detokenize{manual:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
+
+\item {} 
+\phantomsection\label{\detokenize{manual:id11}}{\hyperref[\detokenize{manual:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{manual:id11}}{\hyperref[\detokenize{manual:creating-a-calendar}]{\sphinxcrossref{Creating a calendar}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id12}}{\hyperref[\detokenize{manual:adding-holiday-events}]{\sphinxcrossref{Adding holiday events}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id13}}{\hyperref[\detokenize{manual:adding-clinician-requests}]{\sphinxcrossref{Adding clinician requests}}}
-
-\end{itemize}
+\phantomsection\label{\detokenize{manual:id12}}{\hyperref[\detokenize{manual:scheduling}]{\sphinxcrossref{Scheduling}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{manual:id14}}{\hyperref[\detokenize{manual:usage}]{\sphinxcrossref{Usage}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id15}}{\hyperref[\detokenize{manual:authentication}]{\sphinxcrossref{Authentication}}}
+\phantomsection\label{\detokenize{manual:id13}}{\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id16}}{\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id17}}{\hyperref[\detokenize{manual:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
+\phantomsection\label{\detokenize{manual:id14}}{\hyperref[\detokenize{manual:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id18}}{\hyperref[\detokenize{manual:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
+\phantomsection\label{\detokenize{manual:id15}}{\hyperref[\detokenize{manual:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id19}}{\hyperref[\detokenize{manual:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id20}}{\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id21}}{\hyperref[\detokenize{manual:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id22}}{\hyperref[\detokenize{manual:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
+\phantomsection\label{\detokenize{manual:id16}}{\hyperref[\detokenize{manual:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{manual:id23}}{\hyperref[\detokenize{manual:scheduling}]{\sphinxcrossref{Scheduling}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id24}}{\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id25}}{\hyperref[\detokenize{manual:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id26}}{\hyperref[\detokenize{manual:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id27}}{\hyperref[\detokenize{manual:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
-
-\end{itemize}
-
-\end{itemize}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id28}}{\hyperref[\detokenize{manual:sample-output}]{\sphinxcrossref{Sample Output}}}
+\phantomsection\label{\detokenize{manual:id17}}{\hyperref[\detokenize{manual:sample-output}]{\sphinxcrossref{Sample Output}}}
 
@@
 
-The clinician scheduler allows you to automatically generate and publish
-schedules that satisfy common constraints in an on-call system, while taking
-into account the preferences of clinicians.
-
-\chapter{Setup}
-\label{\detokenize{manual:setup}}
-
-\section{Generating Google API credentials}
-\label{\detokenize{manual:generating-google-api-credentials}}
-\begin{sphinxadmonition}{note}{Note:}
-This process should only be done the very first time you are running
-the program. If you have previously generated a credential file using
-this process you should be able to re-use it. Just make sure that the
-credential file is placed in the same folder as the executable (\sphinxcode{\sphinxupquote{scheduler.exe}}).
-\end{sphinxadmonition}
-
-The application uses Google calendar for retrieving clinician
-time-off requests and long weekend information, as well as uploading the
-generated schedule to the calendar. These operations require the use
-of Google API credentials, which can be generated as follows.
-\begin{enumerate}
-\item {} 
-Sign into \sphinxurl{https://console.developers.google.com}.
-
-\item {} 
-Click on the project selector at the top left of the page.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step2\_select\_a\_project.png}{\sphinxincludegraphics{{step2_select_a_project}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{2}
-\item {} 
-Click on \sphinxtitleref{New Project}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step3\_new\_project.png}{\sphinxincludegraphics{{step3_new_project}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{3}
-\item {} 
-Enter “Clinician Scheduler” as the \sphinxtitleref{Project Name} and click \sphinxtitleref{Create}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step4\_create\_project.png}{\sphinxincludegraphics{{step4_create_project}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{4}
-\item {} 
-Now you should see the dashboard for the Clinician Scheduler project.
-You will need to enable the Google calendar API. Click on \sphinxtitleref{Enable APIs and Services}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step5\_enable\_apis\_and\_services.png}{\sphinxincludegraphics{{step5_enable_apis_and_services}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{5}
-\item {} 
-Search for “Google calendar API” using the search bar, and select it.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step6\_search\_calendar\_api.png}{\sphinxincludegraphics{{step6_search_calendar_api}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{6}
-\item {} 
-Click \sphinxtitleref{Enable}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step7\_enable\_calendar\_api.png}{\sphinxincludegraphics{{step7_enable_calendar_api}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{7}
-\item {} 
-Now you should see the overview page for the Google calendar API.
-To generate the credentials, click \sphinxtitleref{Create credentials}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step8\_create\_credentials.png}{\sphinxincludegraphics{{step8_create_credentials}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{8}
-\item {} 
-On the credentials form, choose “Google Calendar API” for \sphinxtitleref{Which API are you using?},
-then “Other UI (e.g. Windows, CLI tool)” for \sphinxtitleref{Where will you be calling the API from?}
-and “User data” for \sphinxtitleref{What data will you be accessing?}. Then click on \sphinxtitleref{What credentials do I need?}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step9\_credentials\_form.png}{\sphinxincludegraphics{{step9_credentials_form}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{9}
-\item {} 
-Enter “Client” for \sphinxtitleref{Name} and click on \sphinxtitleref{Create OAuth client ID}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step10\_oauth\_client\_id.png}{\sphinxincludegraphics{{step10_oauth_client_id}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{10}
-\item {} 
-Choose the email address associated to your account for \sphinxtitleref{Email address}
-and enter “Clinician Scheduler” for \sphinxtitleref{Product name shown to users}, then
-click \sphinxtitleref{Continue}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step11\_oauth\_consent\_screen.png}{\sphinxincludegraphics{{step11_oauth_consent_screen}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{11}
-\item {} 
-Your credentials are now generated! Make sure to download and save
-them in the same location that you unzipped the application, so that
-the credential file and the executable file (\sphinxcode{\sphinxupquote{scheduler.exe}}) are in the same folder.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_credentials/step12\_download\_credentials.png}{\sphinxincludegraphics{{step12_download_credentials}.png}}\end{figure}
-
-\begin{sphinxadmonition}{attention}{Attention:}
-Make sure the credential file is saved as \sphinxcode{\sphinxupquote{credentials.json}} (rename it, if necessary),
-or otherwise the application will not be able to recognize it!
-\end{sphinxadmonition}
-
-\section{Google Calendar Configuration}
-\label{\detokenize{manual:google-calendar-configuration}}
-
-\subsection{Creating a calendar}
-\label{\detokenize{manual:creating-a-calendar}}\begin{enumerate}
-\item {} 
-Navigate to \sphinxurl{https://calendar.google.com} and sign in to the appropriate account.
-
-\item {} 
-Click on the settings icon next on the left side, and select \sphinxtitleref{New Calendar}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_calendar/create\_calendar/step2\_new\_calendar.png}{\sphinxincludegraphics{{step2_new_calendar}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{2}
-\item {} 
-Enter a name for your calendar and click \sphinxtitleref{Create calendar}.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_calendar/create\_calendar/step3\_create.png}{\sphinxincludegraphics{{step3_create}.png}}\end{figure}
-
-\subsection{Adding holiday events}
-\label{\detokenize{manual:adding-holiday-events}}
-\begin{sphinxadmonition}{note}{Note:}
-It is recommended to create full-day events for holidays and clinician requests.
-\end{sphinxadmonition}
-
-Create an event by clicking on any cell in the calendar. Make sure that
-the name of the event you create starts with “\sphinxstylestrong{{[}holiday{]}}”, so that the
-scheduler can recognize the event correctly. Click \sphinxtitleref{Save} once you are done.
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_calendar/create\_holiday.png}{\sphinxincludegraphics{{create_holiday}.png}}\end{figure}
-
-\subsection{Adding clinician requests}
-\label{\detokenize{manual:adding-clinician-requests}}\label{\detokenize{manual:id1}}
-Create an event by clicking on any cell in the calendar. Make sure that
-the name of the event you create has the format: “\sphinxstylestrong{{[}request{]} Name}”, so that the
-scheduler can recognize the event correctly. Click \sphinxtitleref{Save} once you are done.
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/google\_calendar/create\_request.png}{\sphinxincludegraphics{{create_request}.png}}\end{figure}
-
-\begin{sphinxadmonition}{warning}{Warning:}
-It is important that the name you use in the event name matches the name in
-the clinician configuration you create in {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}.
-\end{sphinxadmonition}
-
-\chapter{Usage}
-\label{\detokenize{manual:usage}}
-
-\section{Authentication}
-\label{\detokenize{manual:authentication}}\label{\detokenize{manual:id2}}
-When the scheduler first makes a connection to Google calendar, it needs
-to create an authentication token, which will be used to simplify future
-connections.
-
-This requires you to allow the scheduler application to access a calendar
-that you specified. You may encounter this when using any functionality that
-interacts with Google calendar, for example in {\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{\DUrole{std,std-ref}{Generating a schedule}}}}
-or in {\hyperref[\detokenize{manual:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{\DUrole{std,std-ref}{Publishing a schedule to Google Calendar}}}}.
-\begin{enumerate}
-\item {} 
-The scheduler should have automatically opened the access request page
-in a browser window, in which case you can skip to step 2. \sphinxstylestrong{Otherwise},
-locate the command prompt for the scheduler. You should see instructions
-to open the access request printed on the command prompt. Follow these instructions
-in order to open the access request page manually.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/authentication/open\_in\_browser.png}{\sphinxincludegraphics{{open_in_browser}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{1}
-\item {} 
-Follow the instructions on the access request page in order to allow
-the scheduler application to read/write from/to the calendar you specified.
-
-\end{enumerate}
-
-\begin{figure}[H]
-\centering
-\capstart
-\sphinxhref{\_static/images/authentication/access\_request.png}{\sphinxincludegraphics{{access_request}.png}}\caption{\sphinxstylestrong{(Optional)} If you opened the access request manually in step 1, you will
-receive an authentication code which needs to be pasted into the command prompt}\label{\detokenize{manual:id7}}\end{figure}
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/authentication/access\_request\_auth\_code.png}{\sphinxincludegraphics{{access_request_auth_code}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{2}
-\item {} 
-Once you have completed the authentication process, the scheduler will
-automatically resume its functionality.
-
-\end{enumerate}
-
-\section{Clinician Configuration}
-\label{\detokenize{manual:clinician-configuration}}\label{\detokenize{manual:id3}}
+\chapter{Clinician Configuration}
+\label{\detokenize{manual:clinician-configuration}}\label{\detokenize{manual:id1}}
 Before we can generate a schedule, we need to create a configuration file
@@
 
-\subsection{Creating a new configuration file}
+\section{Creating a new configuration file}
 \label{\detokenize{manual:creating-a-new-configuration-file}}
@@
 
-\subsection{Saving the configuration file}
+\section{Saving the configuration file}
 \label{\detokenize{manual:saving-the-configuration-file}}
@@
 
-\subsection{Loading a configuration file}
+\section{Loading a configuration file}
 \label{\detokenize{manual:loading-a-configuration-file}}
@@
 
-\subsection{Adding a new clinician}
-\label{\detokenize{manual:adding-a-new-clinician}}\label{\detokenize{manual:id4}}
+\section{Adding a new clinician}
+\label{\detokenize{manual:adding-a-new-clinician}}\label{\detokenize{manual:id2}}
 1. From the configuration tab, click \sphinxtitleref{New Clinician}. You should see a
@@
 
-\subsection{Deleting an existing clinician}
+\section{Deleting an existing clinician}
 \label{\detokenize{manual:deleting-an-existing-clinician}}
@@
 
-\subsection{Editing an existing clinician}
+\section{Editing an existing clinician}
 \label{\detokenize{manual:editing-an-existing-clinician}}
@@
 
-\section{Scheduling}
+\subsection{Scheduling}
 \label{\detokenize{manual:scheduling}}
 
-\subsection{Generating a schedule}
-\label{\detokenize{manual:generating-a-schedule}}\label{\detokenize{manual:id5}}
+\section{Generating a schedule}
+\label{\detokenize{manual:generating-a-schedule}}\label{\detokenize{manual:id3}}
 Once you have created a configuration file, you can switch over to the
@@
 \sphinxstylestrong{(Optional)} In order to retrieve the time-off requests that were populated
-in Google calendar in {\hyperref[\detokenize{manual:adding-clinician-requests}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding clinician requests}}}}, we need to specify the
+in Google calendar in \DUrole{xref,std,std-ref}{adding-clinician-requests}, we need to specify the
 calendar ID. Open the calendar in your browser, and navigate to the
@@
 
-\sphinxstylestrong{(Optional)} See {\hyperref[\detokenize{manual:authentication}]{\sphinxcrossref{\DUrole{std,std-ref}{Authentication}}}} in case you encounter problems
+\sphinxstylestrong{(Optional)} See \DUrole{xref,std,std-ref}{authentication} in case you encounter problems
 with Google calendar at this stage.
@@
 
-\subsection{Exporting a schedule}
+\section{Exporting a schedule}
 \label{\detokenize{manual:exporting-a-schedule}}
@@
 
-\subsection{Publishing a schedule to Google Calendar}
-\label{\detokenize{manual:publishing-a-schedule-to-google-calendar}}\label{\detokenize{manual:id6}}
+\section{Publishing a schedule to Google Calendar}
+\label{\detokenize{manual:publishing-a-schedule-to-google-calendar}}\label{\detokenize{manual:id4}}
 If you are satisfied with the generated schedule, you can publish it
@@
 
-\subsection{Clearing a published schedule}
+\section{Clearing a published schedule}
 \label{\detokenize{manual:clearing-a-published-schedule}}

### 2019-07-09 paper/paper.intro.tex
Update paper.intro.tex
Intro edited and some comments inserted. Reads very well and I like the overall organization/structure.

--- 2019-07-05 paper/paper.intro.tex
@@
-Hospital departments provide crucial services to a great deal of people all over the world.
-The demand created by patients and other consumers of a hospital vastly exceeds the available supply.
-As such, it is crucial that a hospital department allocates its resources, such as the availability of clinicians, optimally, to provide the best possible service for its patients.
-In particular, in order for various hospital departments to function efficiently it is necessary for clinicians to have carefully allocated on-call schedules that simultaneously ensure sufficient resources are provided to their patients while not overworking clinicians to prevent costly mistakes.
-It is common practice for on-call schedules to be created manually, although this method is prone to errors.
-Namely, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method is unlikely to provide an optimal schedule.
-Moreover, such methods are likely to overlook certain constraints that must be maintained to have an operational department.
-For these reasons, it is important to develop automated methods that are able to find optimal schedules that satisfy the given constraints.
+On-call schedules for a fixed number of health-care providers are central to the efficient running of hospitals.
+Hospital departments provide services where patient needs, and thus the system's demands, often exceed the available supply.
+For example, it is important that a hospital department allocates its resources, such as the availability of a finite number of clinicians, optimally, to ensure the best possible service for its patients.
+Carefully allocated on-call schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes [ref].
+It is common practice for on-call schedules to be created manually.
+Yet manually-created schedules are prone to errors and potential for biases [ref].
+First, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method may not provide an optimal schedule.
+Second, such methods are likely to overlook certain constraints that must be maintained to have an operational department, such as XXXX.
+Third, manual scheduling is often time-consuming for the person developing the schedule.
+For these reasons, it is important to develop automated methods that can generate optimal schedules that satisfy the given constraints of the hospital department.
 \\
@@
 
-Automated methods for tackling optimization problems related to scheduling have been studied extensively and applied in many industries, including transportation [??], manufacturing [??], [...].
-Of special interest to our work are the approaches to tackle the nurse scheduling problem.
-In this problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences.
-A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??].
-An extensive literature review of these and other methods is presented by [??].
-We will briefly summarize the main ideas in some of these approaches.
-\\
+Automated methods to optimize schedules have been studied and applied in many industries, including transportation [??], manufacturing [??], [...].
+Of special interest to a clinician on-call scheduling problem are the approaches to schedule nurses, who often work in shifts.
+In the nurse scheduling problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences.
+A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??].
+
+%An extensive literature review of these and other methods is presented by [??].
+We will briefly summarize the main ideas of some of these approaches.
+\\ SM - don't need to introduce that you will do this for this type of paper I think - would do in a thesis chapter though.
 
 [...] Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation.
-While this allows for finding precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians, it makes it very difficult to adapt a particular approach to other departments in the same hospital or other hospitals.
-Moreover, the large number of variables and constraints also leads to computational complexity issues, especially when using exact methods for finding the solution.
-\\
+While these department-specific approaches allow end-users to find precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians in that department, they are difficult to readily adapt to other departments in the same hospital or other hospitals.
+% explain why hard to adapt?...
+what makes their generalizablility/adaptability limited?
+Moreover, the large number of variables and constraints also leads to computational complexity issues [ref], especially when using exact methods for finding the solution.
+In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease consults; and HIV consults service) at St.
+Michael's Hospital in Toronto, Canada.
+Our goal is to (1) present a simple formulation for the problem developed and tested at the hospital after switching from a manual approach to scheduling; and (2) analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach; and (3) describe the adaptability of the formulation as a basic framework for solving similar problems in other departments.
+\\ %make #3 an objective
+We begin by describing the problem, then...% list the main contents/elements of the paper.
 

### 2019-07-09 docs/build/latex/ClinicianScheduler.tex
Update Latex version of manual

@@
 \title{Clinician Scheduler Documentation}
-\date{Jul 08, 2019}
-\release{1.0}
+\date{Jul 09, 2019}
+\release{2.0}
 \author{David Landsman}
@@
 \item {} 
-\phantomsection\label{\detokenize{manual:id5}}{\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
+\phantomsection\label{\detokenize{manual:id4}}{\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
 \begin{itemize}
 \item {} 
-\phantomsection\label{\detokenize{manual:id6}}{\hyperref[\detokenize{manual:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
+\phantomsection\label{\detokenize{manual:id5}}{\hyperref[\detokenize{manual:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id7}}{\hyperref[\detokenize{manual:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
+\phantomsection\label{\detokenize{manual:id6}}{\hyperref[\detokenize{manual:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id8}}{\hyperref[\detokenize{manual:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
+\phantomsection\label{\detokenize{manual:id7}}{\hyperref[\detokenize{manual:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id9}}{\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
+\phantomsection\label{\detokenize{manual:id8}}{\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id10}}{\hyperref[\detokenize{manual:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
+\phantomsection\label{\detokenize{manual:id9}}{\hyperref[\detokenize{manual:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id11}}{\hyperref[\detokenize{manual:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id12}}{\hyperref[\detokenize{manual:scheduling}]{\sphinxcrossref{Scheduling}}}
+\phantomsection\label{\detokenize{manual:id10}}{\hyperref[\detokenize{manual:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
 
@@
 \item {} 
-\phantomsection\label{\detokenize{manual:id13}}{\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
+\phantomsection\label{\detokenize{manual:id11}}{\hyperref[\detokenize{manual:requests-holiday-weekends}]{\sphinxcrossref{Requests \& Holiday Weekends}}}
+\begin{itemize}
+\item {} 
+\phantomsection\label{\detokenize{manual:id12}}{\hyperref[\detokenize{manual:clinician-requests-excel-file}]{\sphinxcrossref{Clinician Requests Excel File}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id14}}{\hyperref[\detokenize{manual:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
+\phantomsection\label{\detokenize{manual:id13}}{\hyperref[\detokenize{manual:holiday-weekends-excel-file}]{\sphinxcrossref{Holiday Weekends Excel File}}}
+
+\end{itemize}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id15}}{\hyperref[\detokenize{manual:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}
+\phantomsection\label{\detokenize{manual:id14}}{\hyperref[\detokenize{manual:scheduling}]{\sphinxcrossref{Scheduling}}}
+\begin{itemize}
+\item {} 
+\phantomsection\label{\detokenize{manual:id15}}{\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
 
 \item {} 
-\phantomsection\label{\detokenize{manual:id16}}{\hyperref[\detokenize{manual:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}
+\phantomsection\label{\detokenize{manual:id16}}{\hyperref[\detokenize{manual:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
 
@@
 \phantomsection\label{\detokenize{manual:id17}}{\hyperref[\detokenize{manual:sample-output}]{\sphinxcrossref{Sample Output}}}
+\begin{itemize}
+\item {} 
+\phantomsection\label{\detokenize{manual:id18}}{\hyperref[\detokenize{manual:yearly-excel-format}]{\sphinxcrossref{Yearly Excel Format}}}
+
+\item {} 
+\phantomsection\label{\detokenize{manual:id19}}{\hyperref[\detokenize{manual:monthly-excel-format}]{\sphinxcrossref{Monthly Excel Format}}}
+
+\end{itemize}
 
@@
 \label{\detokenize{manual:clinician-configuration}}\label{\detokenize{manual:id1}}
-Before we can generate a schedule, we need to create a configuration file
-that specifies which clinicians are available, and how many weeks each
-clinician should fulfill.
+The clinician configuration specifies which clinicians are available,
+which divisions they are covering, and how many weeks they should cover
+in each of their divisions.
 
@@
 Unsaved changes to a configuration file will be discarded upon clicking
-\sphinxtitleref{New Config}.
+on \sphinxtitleref{New Config}.
 \end{sphinxadmonition}
@@
 If you would like to open a previously created configuration file, simply
-click on \sphinxtitleref{Open Config}, navigate to the location of the configuration file,
+click on \sphinxtitleref{Open Config}, navigate to the location of the file,
 and select it.
@@
 1. From the configuration tab, click \sphinxtitleref{New Clinician}. You should see a
-form for supplying details.
+form for supplying clinician details.
 
@@
 
-\subsection{Scheduling}
+\chapter{Requests \& Holiday Weekends}
+\label{\detokenize{manual:requests-holiday-weekends}}\label{\detokenize{manual:requests-and-holidays}}
+Clinician requests for time-off and the dates of holiday weekends are
+supplied using a pair of excel files.
+
+\section{Clinician Requests Excel File}
+\label{\detokenize{manual:clinician-requests-excel-file}}
+The requests file should have a single sheet per clinician that is in the
+configuration file you created in {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}. Inside
+each sheet, each request is entered on a separate row, containing the start
+and end date (inclusive) of the request.
+
+\begin{sphinxadmonition}{warning}{Warning:}
+Make sure each value in the Excel file is formatted as a date!
+\end{sphinxadmonition}
+
+\begin{figure}[H]
+\centering
+\sphinxhref{\_static/images/requests\_holidays/excel\_requests.png}{\sphinxincludegraphics{{excel_requests}.png}}\end{figure}
+
+\begin{sphinxadmonition}{note}{Note:}
+To enter a request for a single day, use the same value for both start
+and end date.
+\end{sphinxadmonition}
+
+\section{Holiday Weekends Excel File}
+\label{\detokenize{manual:holiday-weekends-excel-file}}
+The holiday weekends file should have a single sheet with the dates of the
+holiday weekends, one in each row.
+
+\begin{sphinxadmonition}{warning}{Warning:}
+Make sure each value in the Excel file is formatted as a date!
+\end{sphinxadmonition}
+
+\begin{figure}[H]
+\centering
+\sphinxhref{\_static/images/requests\_holidays/excel\_holidays.png}{\sphinxincludegraphics{{excel_holidays}.png}}\end{figure}
+
+\chapter{Scheduling}
 \label{\detokenize{manual:scheduling}}
@@
 \item {} 
-From the scheduling tab, click on \sphinxtitleref{Load} and select the configuration
-file that you generated in {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
+Load the configuration, requests, and holidays files that you created
+earlier in {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}} and {\hyperref[\detokenize{manual:requests-and-holidays}]{\sphinxcrossref{\DUrole{std,std-ref}{Requests \& Holiday Weekends}}}}.
+
+\end{enumerate}
+
+\begin{sphinxadmonition}{note}{Note:}
+You can be sure that everything loaded correctly by checking the output
+on the right.
+\end{sphinxadmonition}
+
+\begin{figure}[H]
+\centering
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step1\_load.png}{\sphinxincludegraphics{{step1_load}.png}}\end{figure}
+\begin{enumerate}
+\setcounter{enumi}{1}
+\item {} 
+Set the starting calendar year for the schedule, as well as the number
+of 2-week blocks that you want the schedule to cover. By default, the calendar
+year is the upcoming year, and the number of blocks is 26, to cover a full
+year.
 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step1\_load\_config.png}{\sphinxincludegraphics{{step1_load_config}.png}}\end{figure}
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step2\_year\_blocks.png}{\sphinxincludegraphics{{step2_year_blocks}.png}}\end{figure}
 \begin{enumerate}
-\setcounter{enumi}{1}
+\setcounter{enumi}{2}
 \item {} 
-Enter the calendar year for the schedule that you want to generate.
+\sphinxstylestrong{(Optional)} Check the \sphinxtitleref{Shuffle?} checkbox if you would like the scheduler
+to generate a slightly different schedule each time you click on \sphinxtitleref{Generate}.
 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step2\_calendar\_year.png}{\sphinxincludegraphics{{step2_calendar_year}.png}}\end{figure}
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step3\_optional\_shuffle.png}{\sphinxincludegraphics{{step3_optional_shuffle}.png}}\end{figure}
 \begin{enumerate}
-\setcounter{enumi}{2}
+\setcounter{enumi}{3}
 \item {} 
-\sphinxstylestrong{(Optional)} In order to retrieve the time-off requests that were populated
-in Google calendar in \DUrole{xref,std,std-ref}{adding-clinician-requests}, we need to specify the
-calendar ID. Open the calendar in your browser, and navigate to the
-\sphinxtitleref{Settings and sharing} page.
-On the setting page, navigate to the \sphinxtitleref{Integrate calendar} section, and copy
-the value under \sphinxtitleref{Calendar ID} to your clipboard.
-Paste this value into the \sphinxtitleref{Google Calendar ID} textbox on the \sphinxtitleref{Scheduler}
-tab.
+\sphinxstylestrong{(Optional)} Check the \sphinxtitleref{Verbose Output} checkbox if you would like
+to see detailed output on the right after the scheduler finishes generating
+the schedule.
 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step3\_calendar\_id.png}{\sphinxincludegraphics{{step3_calendar_id}.png}}\end{figure}
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step4\_optional\_verbose.png}{\sphinxincludegraphics{{step4_optional_verbose}.png}}\end{figure}
 \begin{enumerate}
-\setcounter{enumi}{3}
+\setcounter{enumi}{4}
 \item {} 
-\sphinxstylestrong{(Optional)} If you supplied a calendar ID in step 3, you can configure
-the options \sphinxtitleref{Retrieve Time-off Requests} and \sphinxtitleref{Retrieve Long Weekends} as necessary.
-Enabling \sphinxtitleref{Retrieve Time-off Requests} will read the time-off calendar events
-from the calendar specified in \sphinxtitleref{Calendar ID}, while enabling
-\sphinxtitleref{Retrieve Long Weekends} will read the holiday events from that calendar.
+Click on \sphinxtitleref{Generate} to generate a schedule.
 
@@
 
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step4\_retrieve\_events.png}{\sphinxincludegraphics{{step4_retrieve_events}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{4}
-\item {} 
-\sphinxstylestrong{(Optional)} In the case that you only need to generate a schedule for
-a subset of the calendar year, you can select the amount of blocks you need
-to generate by specifying a value in \sphinxtitleref{Number of Blocks}. By default, the
-scheduler will generate a full schedule, corresponding to 26 blocks.
-
-\end{enumerate}
+\begin{sphinxadmonition}{warning}{Warning:}
+Depending on the amount of clinicians and requests provided, it may take
+some time to find an optimal schedule.
+\end{sphinxadmonition}
 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step5\_number\_of\_blocks.png}{\sphinxincludegraphics{{step5_number_of_blocks}.png}}\end{figure}
-\begin{enumerate}
-\setcounter{enumi}{5}
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step5\_generate.png}{\sphinxincludegraphics{{step5_generate}.png}}\end{figure}
+
+\begin{sphinxadmonition}{warning}{Warning:}
+It is possible that the scheduler will not be able to come up with a
+schedule that satisfies your constraints. This can be a result of either
+of the following reasons:
+\begin{itemize}
 \item {} 
-Click on \sphinxtitleref{Generate}, and after a few moments you should see a preview of
-the generated schedule in the table.
+There are not enough clinicians to distribute evenly throughout the year
 
-\sphinxstylestrong{(Optional)} See \DUrole{xref,std,std-ref}{authentication} in case you encounter problems
-with Google calendar at this stage.
+\item {} 
+The minimum and maximum number of blocks of clinicians are too restrictive
 
-\end{enumerate}
+\end{itemize}
 
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step6\_generate\_schedule.png}{\sphinxincludegraphics{{step6_generate_schedule}.png}}\end{figure}
+Try changing the configuration file by adding new clinicians, or changing
+the min/max blocks of different clinicians to allow for more flexibility.
 
-\begin{sphinxadmonition}{note}{Note:}
-It is possible that the scheduler will not be able to come up with a
-schedule that satisfies your constraints. You can try adjusting
-some constraints by changing the minimum and maximum number of blocks of
-clinicians in the configuration file. See {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
-for more information on changing the configuration file.
+See {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}} for more information on modifying the configuration file.
 \end{sphinxadmonition}
@@
 If you are satisfied with the generated schedule, you can choose to export
-it as an Excel file. There are two format options: \sphinxtitleref{Yearly Export} and
+it as an Excel file. There are two Excel format options: \sphinxtitleref{Yearly Export} and
 \sphinxtitleref{Monthly Export}.
@@
 
-\section{Publishing a schedule to Google Calendar}
-\label{\detokenize{manual:publishing-a-schedule-to-google-calendar}}\label{\detokenize{manual:id4}}
-If you are satisfied with the generated schedule, you can publish it
-to Google calendar, by clicking on \sphinxtitleref{Publish}.
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/scheduling/publish\_schedule/publish\_schedule.png}{\sphinxincludegraphics{{publish_schedule}.png}}\end{figure}
-
-\begin{sphinxadmonition}{warning}{Warning:}
-This process might take some time, especially if the number of events
-to be published is quite large.
-\end{sphinxadmonition}
-
-\section{Clearing a published schedule}
-\label{\detokenize{manual:clearing-a-published-schedule}}
-If you want to remove all the published events created by the application,
-simply click on \sphinxtitleref{Clear}.
-
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/scheduling/publish\_schedule/clear\_schedule.png}{\sphinxincludegraphics{{clear_schedule}.png}}\end{figure}
-
-\begin{sphinxadmonition}{warning}{Warning:}
-This process might take some time, especially if the number of published
-events is quite large.
-\end{sphinxadmonition}
-
 \chapter{Sample Output}
 \label{\detokenize{manual:sample-output}}
-Yearly output format
 
+\section{Yearly Excel Format}
+\label{\detokenize{manual:yearly-excel-format}}
 \begin{figure}[H]
@@
 
-Monthly output format
-
+\section{Monthly Excel Format}
+\label{\detokenize{manual:monthly-excel-format}}
 \begin{figure}[H]

### 2019-07-09 docs/source/conf.py

@@ -26,7 +26,7 @@ author = 'David Landsman'
 # The short X.Y version
 version = ''
 # The full version, including alpha/beta/rc tags
-release = '1.0'
+release = '2.0'
 
 
 # -- General configuration ---------------------------------------------------

### 2019-07-09 src/helpers/excelhelper.py
Add better error messages for Excel loading

@@ -17,7 +17,7 @@ class ExcelFormatException(Exception):
         self.message = message
 
     def __str__(self):
-        return self.message + ' Filename: {}'.format(self.file)
+        return self.message + '\nFilename: {}'.format(self.file)
 
 
 class ExcelHelper:
@@ -236,6 +236,20 @@ class ExcelHelper:
                 if start is None and end is None: break 
                 elif start is None or end is None:
                     raise ExcelFormatException(file=filename, message='Incorrect format for requests file!')
+
+                if type(start) is str:
+                    try:
+                        start = datetime.strptime(start, '%d-%b-%Y')
+                    except ValueError:
+                        wb.close()
+                        raise ExcelFormatException(file=filename, message='Incorrect date format! Make sure all dates are of the same form as 01-Jan-2019')
+
+                if type(end) is str:
+                    try:
+                        end = datetime.strptime(end, '%d-%b-%Y')
+                    except ValueError:
+                        wb.close()
+                        raise ExcelFormatException(file=filename, message='Incorrect date format! Make sure all dates are of the same form as 01-Jan-2019')
                     
                 if sheet in request_dict:
                     request_dict[sheet].append(Range(start, end))
@@ -258,6 +272,14 @@ class ExcelHelper:
             date = ws['A{0}'.format(row)].value
 
             if date is None: break
+
+            if type(date) is str:
+                try:
+                    date = datetime.strptime(date, '%d-%b-%Y')
+                except ValueError:
+                    wb.close()
+                    raise ExcelFormatException(file=filename, message='Incorrect date format! Make sure all dates are of the same form as 01-Jan-2019')
+
             holidays.append(date)
             row += 1
 

### 2019-07-10 paper/scheduler.tex
Implement on-off-on-off-on constraint (i.e spread)
Prevent on-off-on-off-on type of assignments to improve spread of blocks
for a given clinician. Document in more detail in pdf.
Implement a constraint to spread out weekends
Instead of preventing consecutive weekend assignments (X_i + X_{i+1} <= 1),
restrict to a single possible weekend assignment in the span of 4 weekends
(X_i + X_{i+1} + X_{i+2} + X_{i+3} <= 1). This constraint provides a better
spread of weekend assignments.

@@
 			\end{equation}
-		\item A clinician cannot work two consecutive weekends
+		\item Prevent on-off-on-off-on block assignments (across all divisions) to improve the spread of assignments for any given clinician
 			\begin{equation}
-				Y_{c, w} + Y_{c, w+1} \leq 1 \text{ for each $c$ and each $w \leq W - 1$}
+				\sum_d \left(X_{c, b, d} + X_{c, b + 2, d} + X_{c, b + 4, d}\right) \leq 2 \text{ for each $c$ and each $b \leq B - 4$}
+			\end{equation}
+		\item Improve spread of weekend assignments by preventing more than 1 assignment for a given clinician in a span of 4 weekends
+			\begin{equation}
+				Y_{c, w} + Y_{c, w+1} + Y_{c, w+2} + Y_{c, w+3} \leq 1 \text{ for each $c$ and each $w \leq W - 3$}
 			\end{equation}

### 2019-07-10 src/services/scheduler.py

@@ -408,8 +408,9 @@ class Scheduler:
         self._build_clinician_variables(divisions, clinicians)
         self._build_coverage_constraints(divisions, clinicians)
         self._build_minmax_constraints(divisions)
-        self._build_consec_constraints(clinicians)
-        self._build_spread_constraints(clinicians)
+        self._build_consec_blocks_constraints(clinicians)
+        self._build_spread_blocks_constraints(clinicians)
+        self._build_spread_weekends_constraints(clinicians)
         self._build_longweekend_constraints(clinicians)
         self._build_weekend_constraints(clinicians)
         self._build_adjacency_variables(divisions)
@@ -473,7 +474,7 @@ class Scheduler:
                 self.problem.add(sum_ <= max_)
                 self.problem.add(sum_ >= min_)
 
-    def _build_consec_constraints(self, clinicians):
+    def _build_consec_blocks_constraints(self, clinicians):
         for clinician in clinicians:
             for block_num in range(1, self.num_blocks):
                 # if a clinician works a given block, they should not work any
@@ -484,20 +485,10 @@ class Scheduler:
                 )
                 self.problem.add(sum_ <= 1)
 
-            # at most 1 consecutive weekend of work
-            for week_num in range(1, self.num_weekends):
-                sum_ = pulp.lpSum(
-                    [_.get_var() for _ in clinician.get_weekend_vars(
-                        lambda x, week_num=week_num: x.week_num in (
-                            week_num, week_num + 1)
-                    )]
-                )
-                self.problem.add(sum_ <= 1)
-
-    def _build_spread_constraints(self, clinicians):
+    def _build_spread_blocks_constraints(self, clinicians):
+        # on-off-on-off-on constraint for block assignment
         # we need at least 5 consecutive blocks to implement this constraint
-        # on-off-on-off-on
-        if self.num_blocks <= 5: return
+        if self.num_blocks < 5: return
 
         for clinician in clinicians:
             for block_num in range(1, self.num_blocks - 3):
@@ -507,6 +498,21 @@ class Scheduler:
                         lambda x, block_num=block_num: x.block_num in (block_num, block_num + 2, block_num + 4))]
                 )
                 self.problem.add(sum_ <= 2)
+
+    def _build_spread_weekends_constraints(self, clinicians):
+        # spreading out weekend assignments
+        # needs at least 4 weekends
+        if self.num_weekends < 4: return
+
+        for clinician in clinicians:
[…]

### 2019-07-11 src/services/scheduler.py
Fix normalization of LP objectives
Previous normalization factor did not account for certain clinicians not
working in all divisions.

@@ -414,18 +414,18 @@ class Scheduler:
         self._build_longweekend_constraints(clinicians)
         self._build_weekend_constraints(clinicians)
         self._build_adjacency_variables(divisions)
-        appeasement_objs = self._build_appeasement_objectives(clinicians)
 
-        num_clin = len(clinicians)
-        num_div = len(divisions)
+        block_conflicts_obj = self._build_block_objective(clinicians)
+        weekend_conflicts_obj = self._build_weekend_objective(clinicians)
+        adjacency_obj = self._build_adjacency_objective(clinicians)
 
         # make sure to normalize objectives, and weigh them equally
         self.problem.setObjective(
-              (1 / 3) * (1 / (num_clin * self.num_blocks * num_div)) * appeasement_objs[0]
-            + (1 / 3) * (1 / (num_clin * self.num_weekends)) * appeasement_objs[1]
-            + (1 / 3) * (1 / (num_clin * self.num_blocks * num_div)) * self._build_adjacency_objective(clinicians)
+              (1 / 3) * (1 / len(block_conflicts_obj)) * block_conflicts_obj
+            + (1 / 3) * (1 / len(weekend_conflicts_obj)) * weekend_conflicts_obj
+            + (1 / 3) * (1 / len(adjacency_obj)) * adjacency_obj
         )
-
+        
     def _build_clinician_variables(self, divisions, clinicians):
         # create clinician BlockVariables
         for div in divisions:
@@ -593,7 +593,7 @@ class Scheduler:
             )
         return pulp.lpSum(adjacency_vars)
 
-    def _build_appeasement_objectives(self, clinicians):
+    def _build_weekend_objective(self, clinicians):
         wa_variables = []
         for clinician in clinicians:
             wa_variables.extend(
@@ -608,6 +608,9 @@ class Scheduler:
                  ]
             )
 
+        return pulp.lpSum(wa_variables)
+
+    def _build_block_objective(self, clinicians):
         ba_variables = []
         for clinician in clinicians:
             ba_variables.extend(
@@ -622,10 +625,7 @@ class Scheduler:
                  ]
             )
 
-        return (
-            pulp.lpSum(wa_variables),
-            pulp.lpSum(ba_variables)
-        )
+        return pulp.lpSum(ba_variables)
 
     def assign_schedule(self):
         """

### 2019-07-11 paper/paper.intro.tex
Work on problem section

@@
 Moreover, the large number of variables and constraints also leads to computational complexity issues [ref], especially when using exact methods for finding the solution.
-In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease consults; and HIV consults service) at St.
+In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease (ID) consults; and HIV consults service) at St.
 Michael's Hospital in Toronto, Canada.

### 2019-07-11 paper/paper.problem.tex
Work on problem section

@@
-Clinicians at the Infectious Diseases (ID) and HIV departments of St.
+At St.
+Michael's Hospital, the (???) division offers general ID and HIV consultation throughout the whole year, during regular work weeks as well as weekends and holidays.
+The clinicians in the division typically receive a schedule in advance, outlining their on-call service dates for the full year.
+In the yearly schedule each clinician is assigned to blocks of regular work weeks, as well as weekends.
+Each block corresponds to two consecutive work weeks.
+Apart from long (holiday) weekends, a work week starts on Monday at 8 A.M.
+and ends on Friday at 5 P.M.
+Conversely, weekend service starts on Friday at 5 P.M., and goes on until the start of the next work week on Monday at 8 A.M.
+During the weekend, all clinicians in the department provide both ID and HIV consultation, while during the week some clinicians only provide one of the two services.
+\\
+
+In order to provide quality service and ensure patients get the best care they can, it is important to prevent under- and over-working of clinicians.
+Several constraints are placed on the assignments in hopes of preventing such issues.
+Firstly, each clinician has limits on the number of blocks they can and must work during the year, depending on the type of consultation.
+For instance, a clinician might have to provide 3-5 blocks of general ID consultation as well as 2-3 blocks of HIV consultation throughout the year.
+These limits may change from year to year as the number of clinicians in the department changes.
+Moreover, the schedule does not assign a clinician to work for two blocks or two weekends in a row.
+It is especially important to prevent over-working during weekends, as the demand for the on-call service increases drastically.
+Hence, the schedule attempts to distribute both regular and holiday weekends equally among all clinicians.
+\\
+
+Apart from maintaining a balanced work load among clinicians, the schedule also tries to accommodate their preferences.
+Clinicians provide their requests for time off ahead of schedule generation so that they can be integrated into the schedule.
+They are free to specify days, weeks or weekends off, with the understanding that any blocks overlapping with their request will be assigned to a different clinician, if possible.
+For example, if a clinician only requests Monday and Tuesday off, the schedule will generally avoid assigning the entire block to that clinician.
+Clinicians also prefer to have their weekend and block assignments close together, so the schedule tries to take this into account when distributing assignments.
+
+%Clinicians offering ID and HIV services at St.
 Michael's Hospital are scheduled throughout the year to receive patients during on-call hours.
@@
 
-In order to develop an algorithm, we need to formalize the variables and constraints of the problem mathematically.
+%In order to develop an algorithm, we need to formalize the variables and constraints of the problem mathematically.
 Table [??] presents the sets and indices that are used in the definition of the constraints.

### 2019-07-11 paper/paper.problem.tex
Add math formulation subsection to problem

@@
+\subsection{Overview}
 At St.
@@
 Clinicians also prefer to have their weekend and block assignments close together, so the schedule tries to take this into account when distributing assignments.
-
-%Clinicians offering ID and HIV services at St.
-Michael's Hospital are scheduled throughout the year to receive patients during on-call hours.
-Each clinician is typically scheduled for a full week or a full weekend of on-call service.
-To prevent under- and over-working of clinicians, they each have a minimum and maximum number of weeks that they are required to work.
-Moreover, during holidays weekends, the work-load for on-call service increases drastically, and it is important to distribute these weekends equally.
-At the same time, clinicians often request to not be put on service during certain days or weeks, and the generated schedule should attempt to respect those requests as best as possible.
-Other considerations also need to be taken into account, such as making sure someone is available for on-call service at any time of the year, and preventing multiple back-to-back assignments for a single clinician.
-It is clear that ensuring all of these conditions are met manually, especially with an increasing number of clinicians, is a difficult task and can lead to mismanagement of the schedule.
-Therefore, in this paper we attempt to develop an efficient algorithm that can generate a satisfying schedule, while optimizing clinician and patient contentment.
 \\
 
-%In order to develop an algorithm, we need to formalize the variables and constraints of the problem mathematically.
-Table [??] presents the sets and indices that are used in the definition of the constraints.
+[...]
+
+\subsection{Mathematical Formulation}
 [...] \\
 
-[...]
+Table \ref{tbl:sets-indices} presents the sets and indices that are used in the definition of the constraints. Table \ref{tbl:variables-constants} presents the constants and variables in the problem.
+
+\begin{table}[h]
+	\centering
+	\begin{tabular}{ c c l }
+		\hline
+		set                                  & index & description                       \\ \hline
+		$\mathcal{D} = \{1, \ldots, D \}$    & $d$   & services/divisions                \\
+		$\mathcal{C} = \{1, \ldots, C \}$    & $c$   & clinicians                        \\
+		$\mathcal{B} = \{1, \ldots, B \}$    & $b$   & blocks                            \\
+		$\mathcal{W} = \{1, \ldots, W \}$    & $w$   & weekends                          \\
+		$\mathcal{L} \subset \mathcal{W}$    &       & long weekends                     \\
+		$\mathcal{BR}_c \subset \mathcal{B}$ &       & block requests of clinician $c$   \\
+		$\mathcal{WR}_c \subset \mathcal{W}$ &       & weekend requests of clinician $c$
+	\end{tabular}
+	\caption{Description of sets and indices in the problem}
+	\label{tbl:sets-indices}
+\end{table}
+
+\begin{table}[h]
+	\centering
+	\begin{tabular}{ c l }
+		\hline
+		name                       & description                                                          \\ \hline
+		$X_{c, b, d} \in \{0, 1\}$ & assignment of clinician $c$ for division $d$ on block $b$            \\
+		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                           \\
+		$m_{c, d}$                 & minimum number of blocks clinician $c$ should cover for division $d$ \\
+		$M_{c, d}$                 & maximum number of blocks clinician $c$ should cover for division $d$
+	\end{tabular}
+	\caption{Description of variables and constants in the problem}
+	\label{tbl:variables-constants}
+\end{table}

### 2019-07-15 paper/paper.problem.tex
Add summary of constraints + more details
Move mathematical formulation to methods section, for now
Move \usepackage lines to separate file

@@
-\subsection{Overview}
 At St.
@@
 Clinicians also prefer to have their weekend and block assignments close together, so the schedule tries to take this into account when distributing assignments.
-\\
-
-[...]
-
-\subsection{Mathematical Formulation}
-[...] \\
-
-Table \ref{tbl:sets-indices} presents the sets and indices that are used in the definition of the constraints. Table \ref{tbl:variables-constants} presents the constants and variables in the problem.
+A summary of the outlined constraints is given in Table \ref{tbl:constraint-summary}.
 
@@
 	\centering
-	\begin{tabular}{ c c l }
+	\begin{tabular}{ l l }
 		\hline
-		set                                  & index & description                       \\ \hline
-		$\mathcal{D} = \{1, \ldots, D \}$    & $d$   & services/divisions                \\
-		$\mathcal{C} = \{1, \ldots, C \}$    & $c$   & clinicians                        \\
-		$\mathcal{B} = \{1, \ldots, B \}$    & $b$   & blocks                            \\
-		$\mathcal{W} = \{1, \ldots, W \}$    & $w$   & weekends                          \\
-		$\mathcal{L} \subset \mathcal{W}$    &       & long weekends                     \\
-		$\mathcal{BR}_c \subset \mathcal{B}$ &       & block requests of clinician $c$   \\
-		$\mathcal{WR}_c \subset \mathcal{W}$ &       & weekend requests of clinician $c$
+		\textbf{Constraint Name} & \textbf{Description}                                                                                                           \\ \hline
+		Block Coverage           & \makecell[l]{each division needs to have exactly \\ one clinician that covers any given block}                                 \\ \hline
+		Weekend Coverage         & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              \\ \hline
+		Min/Max                  & \makecell[l]{in a given division, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} \\ \hline
+		No Consecutive Blocks    & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all divisions}                                    \\ \hline
+		No Consecutive Weekends  & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        \\ \hline
+		Equal Weekends           & \makecell[l]{weekends should be equally distributed \\ between clinicians}                                                     \\ \hline
+		Equal Holidays           & \makecell[l]{long weekends should be equally distributed \\ between clinicians}                                                \\ \hline
+		Block Requests           & \makecell[l]{each clinician can request to be off service \\ during certain blocks throughout the year}                        \\ \hline
+		Weekend Requests         & \makecell[l]{each clinician can request to be off service \\ during certain weekends throughout the year}                      \\ \hline
+		Block-Weekend Adjacency  & \makecell[l]{the block and weekend assignments of a given \\ clinician should be adjacent}
 	\end{tabular}
-	\caption{Description of sets and indices in the problem}
-	\label{tbl:sets-indices}
+	\caption{Summary of the constraints for the clinician scheduling problem}
+	\label{tbl:constraint-summary}
 \end{table}
 
-\begin{table}[h]
-	\centering
-	\begin{tabular}{ c l }
-		\hline
-		name                       & description                                                          \\ \hline
-		$X_{c, b, d} \in \{0, 1\}$ & assignment of clinician $c$ for division $d$ on block $b$            \\
-		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                           \\
-		$m_{c, d}$                 & minimum number of blocks clinician $c$ should cover for division $d$ \\
-		$M_{c, d}$                 & maximum number of blocks clinician $c$ should cover for division $d$
-	\end{tabular}
-	\caption{Description of variables and constants in the problem}
-	\label{tbl:variables-constants}
-\end{table}
+In most scheduling problems, the constraints can be divided into hard and soft constraints.
+Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions.
+Typically, soft constraints are encoded as an objective that needs to be maximized or minimized, rather than a constraint.
+In the case of the clinician scheduling problem, we chose Block Requests, Weekend Requests and Block-Weekend Adjacency as our soft constraints, while the rest of the constraints are hard.
+Though it is important to take clinician requests into account when constructing the schedule, it is crucial to that the work-load of the schedule is balanced among all clinicians, and the needs of the patients are fulfilled.
+\\
+
+It is important to note that even though the yearly schedule is generally accurate and adhered to, clinicians are able to exchange certain weeks or days throughout the year.
+The approach to solving the clinician scheduling problem does not attempt to take these exchanges into account, and only focuses on the full year time horizon.

### 2019-07-15 paper/paper.methods.tex
Add summary of constraints + more details
Move mathematical formulation to methods section, for now
Move \usepackage lines to separate file

Table \ref{tbl:sets-indices} presents the sets and indices that are used in the definition of the constraints. Table \ref{tbl:variables-constants} presents the constants and variables in the problem.

\begin{table}[h]
	\centering
	\begin{tabular}{ c c l }
		\hline
		\textbf{Set}                         & \textbf{Index} & \textbf{Description}              \\ \hline
		$\mathcal{D} = \{1, \ldots, D \}$    & $d$            & services/divisions                \\
		$\mathcal{C} = \{1, \ldots, C \}$    & $c$            & clinicians                        \\
		$\mathcal{B} = \{1, \ldots, B \}$    & $b$            & blocks                            \\
		$\mathcal{W} = \{1, \ldots, W \}$    & $w$            & weekends                          \\
		$\mathcal{L} \subset \mathcal{W}$    &                & long weekends                     \\
		$\mathcal{BR}_c \subset \mathcal{B}$ &                & block requests of clinician $c$   \\
		$\mathcal{WR}_c \subset \mathcal{W}$ &                & weekend requests of clinician $c$
	\end{tabular}
	\caption{Description of sets and indices in the problem}
	\label{tbl:sets-indices}
\end{table}

\begin{table}[h]
	\centering
	\begin{tabular}{ c l }
		\hline
		\textbf{Name}              & \textbf{Description}                                                 \\ \hline
		$X_{c, b, d} \in \{0, 1\}$ & assignment of clinician $c$ for division $d$ on block $b$            \\
		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                           \\
		$m_{c, d}$                 & minimum number of blocks clinician $c$ should cover for division $d$ \\
		$M_{c, d}$                 & maximum number of blocks clinician $c$ should cover for division $d$
	\end{tabular}
	\caption{Description of variables and constants in the problem}
	\label{tbl:variables-constants}
\end{table}

### 2019-07-16 paper/paper.methods.tex
Add more detailed descriptions to methods section
Start working on formalizing constraints

@@
-Table \ref{tbl:sets-indices} presents the sets and indices that are used in the definition of the constraints. Table \ref{tbl:variables-constants} presents the constants and variables in the problem.
+In this section, we present an application of 0-1 Linear Programming to solve the clinician scheduling problem presented in Section \ref{???}.
+First, we describe the sets, indices and variables present in the formulation of the program.
+Later we convert the constraints given in Table \ref{tbl:constraint-summary} into mathematical terms, and outline the objective function of the linear program.
+
+\subsection{Sets and Indices}
+We denote the set of all services/divisions that clinicians in the department can provide as $\mathcal{D}$.
+It is important to note that in the following formulation, it is assumed that all clinicians are able to provide all services in this set, although this may not be true in practice.
+The set of all clinicians in the department is denoted as $\mathcal{C}$.
+The sets of blocks and weekends that clinicians will be assigned to are denoted as $\mathcal{B}$ and $\mathcal{W}$ respectively.
+Note that the size of a block is not constrained, and can be adapted to the needs of the given department.
+A subset of weekends are denoted $\mathcal{L}$, corresponding to the long/holiday weekends throughout the year.
+Lastly, we denote with $\mathcal{BR}_c$ and $\mathcal{WR}_c$ the block and weekend requests of clinicians as subsets of all blocks and weekends, respectively.
+For instance, if clinician $c$'s requests intersect with blocks 1 and 2, and weekend 1, then $\mathcal{BR}_c = \{1, 2\}$ and $\mathcal{WR}_c = \{1\}$.
+Table \ref{tbl:sets-indices} presents a summary of the sets and indices described.
 
@@
 
+\subsection{Variables}
+Since each clinician may be assigned to work for any service, during any block of the year, we denote such an assignment as $X_{c, b, d}$.
+A value of 1 indicates that the given clinician $c$ is assigned to cover division $d$ during block $b$.
+Similarly, we define weekend assignments as $Y_{c, w}$, although these are not indexed by service, as clinicians are expected to provide all services during the weekends.
+The optimization of the soft constraint Block-Weekend Adjacency is done by maximizing the product $X_{c, b, d} \cdot Y_{c, w}$ for adjacent blocks and weekends.
+In order to formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, d}$, with additional constraints on its range.
+Further details regarding this variable are described in Sections \ref{???} and \ref{???}.
+Lastly, we introduce a set of constants $m_{c, d}$ and $M_{c, d}$ to constrain the number of blocks each clinician is allowed to work during the year.
+Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.
+
 \begin{table}[h]
@@
 		\textbf{Name}              & \textbf{Description}                                                 \\ \hline
-		$X_{c, b, d} \in \{0, 1\}$ & assignment of clinician $c$ for division $d$ on block $b$            \\
+		$X_{c, b, d} \in \{0, 1\}$ & assignment of clinician $c$ for service/division $d$ on block $b$    \\
 		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                           \\
+		$Z_{c, b, d} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend adjacency               \\
 		$m_{c, d}$                 & minimum number of blocks clinician $c$ should cover for division $d$ \\
@@
 \end{table}
+
+\subsection{Constraints}
+We now present the hard constraints in Table \ref{tbl:constraint-summary} in light of the mathematical setup given above.
+
+	\begin{align}
+		&\sum_{c=1}^{C} X_{c, b, d} = 1 &&\forall\; b\in \mathcal{B}, d \in \mathcal{D} \tag{Block Coverage} \label{eqn:constr-block-cov} \\
+		&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall\; w\in \mathcal{W} \tag{Weekend Coverage} \label{eqn:constr-weekend-cov} \\
+		&m_{c, d} \leq \sum_{b=1}^{B} X_{c, b, d} \leq M_{c, d} &&\forall\; c\in\mathcal{C}, d\in\mathcal{D} \tag{Min/Max} \label{eqn:constr-min-max} \\
+		&X_{c, b, d} + X_{c, b + 1, d} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, d\in\mathcal{D} \tag{No Consecutive Blocks} \label{eqn:constr-no-consec-blocks} \\
+		&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{No Consecutive Weekends} \label{eqn:constr-no-consec-weekends}
+%		\sum_{c=1}^{C} X_{c, b, d} = 1 \qquad\forall\; b\in \mathcal{B}, d \in \mathcal{D} \tag{Block Coverage} \label{eqn:constr-block-cov} \\		
+	\end{align}

### 2019-07-17 paper/paper.methods.tex
Add all hard constraints to methods

@@
 \begin{align}
-		&\sum_{c=1}^{C} X_{c, b, d} = 1 &&\forall\; b\in \mathcal{B}, d \in \mathcal{D} \tag{Block Coverage} \label{eqn:constr-block-cov} \\
-		&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall\; w\in \mathcal{W} \tag{Weekend Coverage} \label{eqn:constr-weekend-cov} \\
-		&m_{c, d} \leq \sum_{b=1}^{B} X_{c, b, d} \leq M_{c, d} &&\forall\; c\in\mathcal{C}, d\in\mathcal{D} \tag{Min/Max} \label{eqn:constr-min-max} \\
+	&\sum_{c=1}^{C} X_{c, b, d} = 1 &&\forall b\in \mathcal{B}, d \in \mathcal{D} \tag{Block Coverage} \label{eqn:constr-block-cov} \\
+	&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{Weekend Coverage} \label{eqn:constr-weekend-cov} \\
+	&m_{c, d} \leq \sum_{b=1}^{B} X_{c, b, d} \leq M_{c, d} &&\forall\ c\in\mathcal{C}, d\in\mathcal{D} \tag{Min/Max} \label{eqn:constr-min-max} \\
 	&X_{c, b, d} + X_{c, b + 1, d} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, d\in\mathcal{D} \tag{No Consecutive Blocks} \label{eqn:constr-no-consec-blocks} \\
-		&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{No Consecutive Weekends} \label{eqn:constr-no-consec-weekends}
-%		\sum_{c=1}^{C} X_{c, b, d} = 1 \qquad\forall\; b\in \mathcal{B}, d \in \mathcal{D} \tag{Block Coverage} \label{eqn:constr-block-cov} \\		
+	&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{No Consecutive Weekends} \label{eqn:constr-no-consec-weekends} \\
+	&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}} &&\forall c\in\mathcal{C} \tag{Equal Weekends} \label{eqn:constr-equal-weekends} \\
+	&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq \ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C} \tag{Equal Holidays} \label{eqn:constr-equal-holidays}
 \end{align}

### 2019-07-17 paper/paper.methods.tex
Add rough description of objective functions

@@
 \end{align}
+
+\subsection{Objectives}
+As mentioned briefly in Section \ref{???}, the soft constraints of the clinician scheduling problem are: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks.
+We convert these constraints into linear objective functions of the binary variables defined in Section \ref{???}.
+The first two objectives have a straight-forward formulation as linear functions:
+\begin{align}
+	&\max_X Q_1(X) = \max_X \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, d} \\
+	&\max_Y Q_2(Y) = \max_Y \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w}
+\end{align}
+where $\ind(\cdot)$ is the indicator function that has value 1 when the condition is met and 0 otherwise.
+In these two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments.
+The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, d}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$.
+For instance, if clinician $c$ wants to work during weeks 1 and 2 (that is, block $b = 1$), as well as weekend 1, we would like the product $X_{c, b=1, d} \cdot Y_{c, w=1}$ to be 1, since that indicates both variables are assigned.
+If at least one of the two variables is not assigned, the product will be 0.
+This leads to the obvious maximization objective
+\begin{align}
+	&\max_{X, Y} Q_3(X, Y) = \max_{X, Y} \sum_{min}^{max} ...
+\end{align}
+However, as it is, $Q_3$ is not a linear function of its variables and cannot be solved in a linear programming framework.
+The general approach used to convert such functions into linear objectives is by introducing a helper variable [ref \ref{???}].
+In our case, introducing a variable $Z_{c, b, d}$ for every product $X_{c, b, d} \cdot Y_{c, w}$ with $w = ...$, and constraining $Z$ such that
+\begin{align}
+	&Z_{c, b, d} \leq X_{c, b, d} \\
+	&Z_{c, b, d} \leq Y_{c, w=...} &&\forall d\in\mathcal{D}
+\end{align}
+allows us to rewrite $Q_3$ as linear function of $Z$,
+\begin{align*}
+	&\max_{Z} Q_3(Z) = \max_{Z} \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} Z_{c, b, d}
+\end{align*}
+Indeed, whenever $X_{c, b, d} \cdot Y_{c, w} = 1$, $Z_{c, b, d}$ can attain a maximum value of 1, and whenever $X_{c, b, d} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, d}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, d}$ will be constrained to attain a maximum value of 0.
+\\
+
+[...]

### 2019-07-18 paper/paper.methods.tex
Add details of weighted, normalized optimization

@@
 \begin{align}
-	&\max_X Q_1(X) = \max_X \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, d} \\
-	&\max_Y Q_2(Y) = \max_Y \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w}
+	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, d} \tag{Block Requests} \label{eqn:obj-block-requests}\\
+	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{Weekend Requests} \label{eqn:obj-weekend-requests}
 \end{align}
@@
 \begin{align}
-	&\max_{X, Y} Q_3(X, Y) = \max_{X, Y} \sum_{min}^{max} ...
+	&Q_3(X, Y) = \sum_{min}^{max} ... \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
 \end{align}
-However, as it is, $Q_3$ is not a linear function of its variables and cannot be solved in a linear programming framework.
-The general approach used to convert such functions into linear objectives is by introducing a helper variable [ref \ref{???}].
+However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework.
+The general approach used to convert such functions into linear objectives is by introducing a helper variable and additional constraints [ref \ref{???}].
 In our case, introducing a variable $Z_{c, b, d}$ for every product $X_{c, b, d} \cdot Y_{c, w}$ with $w = ...$, and constraining $Z$ such that
@@
 allows us to rewrite $Q_3$ as linear function of $Z$,
-\begin{align*}
-	&\max_{Z} Q_3(Z) = \max_{Z} \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} Z_{c, b, d}
-\end{align*}
-Indeed, whenever $X_{c, b, d} \cdot Y_{c, w} = 1$, $Z_{c, b, d}$ can attain a maximum value of 1, and whenever $X_{c, b, d} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, d}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, d}$ will be constrained to attain a maximum value of 0.
+\begin{align}
+	&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} Z_{c, b, d}
+\end{align}
+Indeed, whenever $X_{c, b, d} \cdot Y_{c, w} = 1$, $Z_{c, b, d}$ can attain a maximum value of 1, and whenever $X_{c, b, d} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, d}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, d}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective.
 \\
 
+As we can see, these objectives turn our clinician scheduling problem into a multiple objective optimization problem.
+The most common approach to solving such problems is by optimizing a weighted sum of the normalized objective functions, as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}].
+This is the approach we decided to use in our problem, to ensure all three objectives are considered when finding a solution.
+Under the assumption that each clinician in $\mathcal{C}$ provides all types of services in $\mathcal{D}$, the normalized objectives can be written as follows,
+\begin{align}
+	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot D} \tag{Block Requests} \label{eqn:norm-obj-block-requests}\\
+	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests} \label{eqn:norm-obj-weekend-requests} \\
+	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot D} \tag{Block-Weekend Adjacency} \label{eqn:norm-obj-block-weekend-adj}
+\end{align}
+where we divided each of the original objective functions by the sum of the absolute value of its coefficients [ref \ref{???}]. The final weighted objective is given by
+\begin{equation}
+	\max_{X, Y, Z} \alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
+\end{equation}
+with $0 \leq \alpha, \beta \leq 1$. \\
+
 [...]

### 2019-07-18 paper/paper.methods.tex
Clarify adjacency mapping from block to weekend

@@
 \begin{align}
-	&Q_3(X, Y) = \sum_{min}^{max} ... \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
+	&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} X_{c, b, d}\cdot Y_{c, w=\varphi(b)} \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
 \end{align}
+where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by whichever definition of adjacency being considered. For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\
+
 However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework.
 The general approach used to convert such functions into linear objectives is by introducing a helper variable and additional constraints [ref \ref{???}].
-In our case, introducing a variable $Z_{c, b, d}$ for every product $X_{c, b, d} \cdot Y_{c, w}$ with $w = ...$, and constraining $Z$ such that
+In our case, introducing a variable $Z_{c, b, d}$ for every product $X_{c, b, d} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that
 \begin{align}
 	&Z_{c, b, d} \leq X_{c, b, d} \\
-	&Z_{c, b, d} \leq Y_{c, w=...} &&\forall d\in\mathcal{D}
+	&Z_{c, b, d} \leq Y_{c, w=\varphi(b)} &&\forall d\in\mathcal{D}
 \end{align}

### 2019-07-18 paper/paper.methods.tex
Fix typos and clarify example

@@
 The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, d}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$.
-For instance, if clinician $c$ wants to work during weeks 1 and 2 (that is, block $b = 1$), as well as weekend 1, we would like the product $X_{c, b=1, d} \cdot Y_{c, w=1}$ to be 1, since that indicates both variables are assigned.
+For instance, clinicians might want to be assigned during a weekend that falls within an assigned block.
+So if clinician $c$ is assigned to work during block 3, corresponding to weeks 5 and 6 assuming 2-week blocks, they might also want to be assigned to work during weekend 5.
+In that case, we would like $X_{c, b=3, d} \cdot Y_{c, w=5}$ to be 1, since that indicates both variables are assigned.
 If at least one of the two variables is not assigned, the product will be 0.
@@
 \end{align}
-where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by whichever definition of adjacency being considered. For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\
+where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by some appropriate definition of adjacency. For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\
 
@@
 \end{align}
-allows us to rewrite $Q_3$ as linear function of $Z$,
+allows us to rewrite $Q_3$ as a linear function of $Z$,
 \begin{align}
@@
 \end{align}
-where we divided each of the original objective functions by the sum of the absolute value of its coefficients [ref \ref{???}]. The final weighted objective is given by
+where we divided each of the original objective functions by the sum of the absolute values of its coefficients [ref \ref{???}]. The final weighted objective is given by
 \begin{equation}
-	\max_{X, Y, Z} \alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
+	\alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
 \end{equation}

### 2019-07-19 paper/paper.problem.tex
SM edits to problem

@@
 At St.
-Michael's Hospital, the (???) division offers general ID and HIV consultation throughout the whole year, during regular work weeks as well as weekends and holidays.
+Michael's Hospital, the division of infectious diseases (ID) offers general ID and HIV consultation on inpatients as two parallel services.
+Each service provides clinical care throughout the year, during regular work weeks as well as weekends and holidays.
 The clinicians in the division typically receive a schedule in advance, outlining their on-call service dates for the full year.
-In the yearly schedule each clinician is assigned to blocks of regular work weeks, as well as weekends.
+In the yearly schedule each clinician is assigned to blocks of regular work weeks and weekends.
 Each block corresponds to two consecutive work weeks.
@@
 and ends on Friday at 5 P.M.
-Conversely, weekend service starts on Friday at 5 P.M., and goes on until the start of the next work week on Monday at 8 A.M.
-During the weekend, all clinicians in the department provide both ID and HIV consultation, while during the week some clinicians only provide one of the two services.
+Accordingly, weekend service starts on Friday at 5 P.M., and ends on Monday at 8 A.M.
+During the weekend, ID and HIV consultation services are combined and provided by one clinician.
+During the regular work week the ID and HIV services are led by one clinician each.
 \\
 
-In order to provide quality service and ensure patients get the best care they can, it is important to prevent under- and over-working of clinicians.
-Several constraints are placed on the assignments in hopes of preventing such issues.
-Firstly, each clinician has limits on the number of blocks they can and must work during the year, depending on the type of consultation.
+Several constraints are placed on the on-call assignments.
+First, each clinician has limits on the number of blocks they can and must work during the year, depending on the type of consultation.
 For instance, a clinician might have to provide 3-5 blocks of general ID consultation as well as 2-3 blocks of HIV consultation throughout the year.
 These limits may change from year to year as the number of clinicians in the department changes.
-Moreover, the schedule does not assign a clinician to work for two blocks or two weekends in a row.
-It is especially important to prevent over-working during weekends, as the demand for the on-call service increases drastically.
-Hence, the schedule attempts to distribute both regular and holiday weekends equally among all clinicians.
+Moreover, the schedule should not assign a clinician to work for two blocks or two weekends in a row.
+In addition, the schedule should distribute both regular and holiday weekends equally among all clinicians.
 \\
 
-Apart from maintaining a balanced work load among clinicians, the schedule also tries to accommodate their preferences.
-Clinicians provide their requests for time off ahead of schedule generation so that they can be integrated into the schedule.
-They are free to specify days, weeks or weekends off, with the understanding that any blocks overlapping with their request will be assigned to a different clinician, if possible.
-For example, if a clinician only requests Monday and Tuesday off, the schedule will generally avoid assigning the entire block to that clinician.
-Clinicians also prefer to have their weekend and block assignments close together, so the schedule tries to take this into account when distributing assignments.
+In addition to maintaining a balanced work load among clinicians, the schedule should also accommodate their preferences.
+Clinicians provide their requests for time off ahead of schedule generation so that the requests may be integrated into the schedule.
+Individuals may specify days, weeks or weekends off, with the understanding that any blocks overlapping with their request will be assigned to a different clinician where possible.
+For example, if a clinician only requests a given Monday and Tuesday off, the schedule will generally avoid assigning the entire block to that clinician.
+Clinicians also prefer to have their weekend and block assignments close together, so the schedule should account for this when distributing assignments.
 A summary of the outlined constraints is given in Table \ref{tbl:constraint-summary}.
@@
 In most scheduling problems, the constraints can be divided into hard and soft constraints.
-Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions.
+Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions [ref].
 Typically, soft constraints are encoded as an objective that needs to be maximized or minimized, rather than a constraint.
@@
 
-It is important to note that even though the yearly schedule is generally accurate and adhered to, clinicians are able to exchange certain weeks or days throughout the year.
-The approach to solving the clinician scheduling problem does not attempt to take these exchanges into account, and only focuses on the full year time horizon.
+It is important to note that - irrespective of how the schedule is generated - clinicians may exchange certain weeks or days throughout the year after the schedule is implemented.
+The approach to solving the clinician scheduling problem does not account for these future exchanges, and only focuses on the full year time horizon.

### 2019-07-19 paper/paper.methods.tex
Minor fixes to problem, methods sections

@@
 Further details regarding this variable are described in Sections \ref{???} and \ref{???}.
-Lastly, we introduce a set of constants $m_{c, d}$ and $M_{c, d}$ to constrain the number of blocks each clinician is allowed to work during the year.
+Lastly, we introduce a set of constants $m_{c, d}$ and $M_{c, d}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year.
 Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.
@@
 We convert these constraints into linear objective functions of the binary variables defined in Section \ref{???}.
-The first two objectives have a straight-forward formulation as linear functions:
+The first two objectives have a straightforward formulation as linear functions:
 \begin{align}

### 2019-07-19 paper/paper.problem.tex
Minor fixes to problem, methods sections

@@
 Moreover, the schedule should not assign a clinician to work for two blocks or two weekends in a row.
-In addition, the schedule should distribute both regular and holiday weekends equally among all clinicians.
+The schedule should also distribute both regular and holiday weekends equally among all clinicians.
 \\
@@
 	\centering
-	\begin{tabular}{ l l }
+	\begin{tabular}{ l l l }
 		\hline
-		\textbf{Constraint Name} & \textbf{Description}                                                                                                           \\ \hline
-		Block Coverage           & \makecell[l]{each division needs to have exactly \\ one clinician that covers any given block}                                 \\ \hline
-		Weekend Coverage         & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              \\ \hline
-		Min/Max                  & \makecell[l]{in a given division, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} \\ \hline
-		No Consecutive Blocks    & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all divisions}                                    \\ \hline
-		No Consecutive Weekends  & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        \\ \hline
-		Equal Weekends           & \makecell[l]{weekends should be equally distributed \\ between clinicians}                                                     \\ \hline
-		Equal Holidays           & \makecell[l]{long weekends should be equally distributed \\ between clinicians}                                                \\ \hline
-		Block Requests           & \makecell[l]{each clinician can request to be off service \\ during certain blocks throughout the year}                        \\ \hline
-		Weekend Requests         & \makecell[l]{each clinician can request to be off service \\ during certain weekends throughout the year}                      \\ \hline
-		Block-Weekend Adjacency  & \makecell[l]{the block and weekend assignments of a given \\ clinician should be adjacent}
+		\textbf{Constraint Name} & \textbf{Description}                                                                                                           & \textbf{Type} \\ \hline
+		Block Coverage           & \makecell[l]{each division needs to have exactly \\ one clinician that covers any given block}                                 & Hard          \\ \hline
+		Weekend Coverage         & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              & Hard          \\ \hline
+		Min/Max                  & \makecell[l]{in a given division, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} & Hard          \\ \hline
+		No Consecutive Blocks    & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all divisions}                                    & Hard          \\ \hline
+		No Consecutive Weekends  & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        & Hard          \\ \hline
+		Equal Weekends           & \makecell[l]{weekends should be equally distributed \\ between clinicians}                                                     & Hard          \\ \hline
+		Equal Holidays           & \makecell[l]{long weekends should be equally distributed \\ between clinicians}                                                & Hard          \\ \hline
+		Block Requests           & \makecell[l]{each clinician can request to be off service \\ during certain blocks throughout the year}                        & Soft          \\ \hline
+		Weekend Requests         & \makecell[l]{each clinician can request to be off service \\ during certain weekends throughout the year}                      & Soft          \\ \hline
+		Block-Weekend Adjacency  & \makecell[l]{the block and weekend assignments of a given \\ clinician should be adjacent}                                     & Soft
 	\end{tabular}
@@
 Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions [ref].
-Typically, soft constraints are encoded as an objective that needs to be maximized or minimized, rather than a constraint.
-In the case of the clinician scheduling problem, we chose Block Requests, Weekend Requests and Block-Weekend Adjacency as our soft constraints, while the rest of the constraints are hard.
-Though it is important to take clinician requests into account when constructing the schedule, it is crucial to that the work-load of the schedule is balanced among all clinicians, and the needs of the patients are fulfilled.
+Typically, soft constraints are encoded as objective functions rather than actual constraints.
+These objective functions are maximized or minimized when finding a solution.
+In the case of our clinician scheduling problem, we chose Block Requests, Weekend Requests and Block-Weekend Adjacency as our soft constraints, while the rest of the constraints are hard.
+Though it is important to take clinician requests into account when constructing the schedule, it is crucial that the work-load of the schedule is balanced among all clinicians, and the needs of the patients are fulfilled.
 \\

### 2019-07-24 paper/paper.methods.tex
SM edits/comments Methods

--- 2019-07-18 paper/paper.methods.tex
@@
 First, we describe the sets, indices and variables present in the formulation of the program.
-Later we convert the constraints given in Table \ref{tbl:constraint-summary} into mathematical terms, and outline the objective function of the linear program.
+We then convert the constraints given in Table \ref{tbl:constraint-summary} into mathematical terms, and outline the objective function of the linear program.
 
 \subsection{Sets and Indices}
-We denote the set of all services/divisions that clinicians in the department can provide as $\mathcal{D}$.
-It is important to note that in the following formulation, it is assumed that all clinicians are able to provide all services in this set, although this may not be true in practice.
+We denote the set of all services (or sometimes referred to as divisions) %or I think services is good (vs.
+services/divisions) as you define two types of services (ID vs.
+HIV) in the problem statement
+that clinicians in a single department can provide as $\mathcal{D}$.
+The following formulation assumes that all clinicians are able to provide all services.
 The set of all clinicians in the department is denoted as $\mathcal{C}$.
 The sets of blocks and weekends that clinicians will be assigned to are denoted as $\mathcal{B}$ and $\mathcal{W}$ respectively.
-Note that the size of a block is not constrained, and can be adapted to the needs of the given department.
-A subset of weekends are denoted $\mathcal{L}$, corresponding to the long/holiday weekends throughout the year.
+The size of a block is not constrained and can be adapted %how?
+to the needs of the given department.
+A subset of weekends are denoted $\mathcal{L}$, corresponding to established long/holiday weekends such as the Canadian Civic Day weekend in August.
 Lastly, we denote with $\mathcal{BR}_c$ and $\mathcal{WR}_c$ the block and weekend requests of clinicians as subsets of all blocks and weekends, respectively.
@@
 \subsection{Variables}
-Since each clinician may be assigned to work for any service, during any block of the year, we denote such an assignment as $X_{c, b, d}$.
+Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as $X_{c, b, d}$.
 A value of 1 indicates that the given clinician $c$ is assigned to cover division $d$ during block $b$.
 Similarly, we define weekend assignments as $Y_{c, w}$, although these are not indexed by service, as clinicians are expected to provide all services during the weekends.
-The optimization of the soft constraint Block-Weekend Adjacency is done by maximizing the product $X_{c, b, d} \cdot Y_{c, w}$ for adjacent blocks and weekends.
-In order to formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, d}$, with additional constraints on its range.
+To optimize the soft constraint Block-Weekend Adjacency, we maximize the product $X_{c, b, d} \cdot Y_{c, w}$ for adjacent blocks and weekends.
+To formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, d}$, with additional constraints on its range.
 Further details regarding this variable are described in Sections \ref{???} and \ref{???}.
-Lastly, we introduce a set of constants $m_{c, d}$ and $M_{c, d}$ to constrain the number of blocks each clinician is allowed to work during the year.
+We then introduce a set of constants $m_{c, d}$ and $M_{c, d}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year.
 Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.
@@
 \subsection{Constraints}
-We now present the hard constraints in Table \ref{tbl:constraint-summary} in light of the mathematical setup given above.
+We now present the hard constraints in Table \ref{tbl:constraint-summary} using the variables defined above.
 
@@
 \subsection{Objectives}
-As mentioned briefly in Section \ref{???}, the soft constraints of the clinician scheduling problem are: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks.
-We convert these constraints into linear objective functions of the binary variables defined in Section \ref{???}.
-The first two objectives have a straight-forward formulation as linear functions:
+As described in Section \ref{???}, the soft constraints of the clinician scheduling problem include: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks.
+We convert these soft constraints into linear objective functions of the binary variables defined in Section \ref{???}.
+The first two objectives can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to.
 \begin{align}
@@
 where $\ind(\cdot)$ is the indicator function that has value 1 when the condition is met and 0 otherwise.
-In these two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments.
+In the above two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments.
 The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, d}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$.
@@
 If at least one of the two variables is not assigned, the product will be 0.
-This leads to the obvious maximization objective
+This leads to the maximization objective  %avoid editorial adjectives
 \begin{align}
@@
 where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by some appropriate definition of adjacency. For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\
+% I really liked the example explanation!
 
 However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework.
-The general approach used to convert such functions into linear objectives is by introducing a helper variable and additional constraints [ref \ref{???}].
+An approach used to convert such functions into linear objectives involves introducing a helper variable and additional constraints [ref \ref{???}].
 In our case, introducing a variable $Z_{c, b, d}$ for every product $X_{c, b, d} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that
@@
 
-As we can see, these objectives turn our clinician scheduling problem into a multiple objective optimization problem.
-The most common approach to solving such problems is by optimizing a weighted sum of the normalized objective functions, as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}].
+Thus, our clinician scheduling problem is a multiple objective optimization problem.
+The most common approach to solving multiple objective optimization problems is by optimizing a weighted sum of the normalized objective functions, as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}].
 This is the approach we decided to use in our problem, to ensure all three objectives are considered when finding a solution.
@@
 \end{align}
-where we divided each of the original objective functions by the sum of the absolute values of its coefficients [ref \ref{???}]. The final weighted objective is given by
+where we divide each of the original objective functions by the sum of the absolute values of its coefficients [ref \ref{???}]. The final weighted objective is given by
 \begin{equation}

### 2019-07-25 paper/paper.methods.tex
Minor fix + update pdf

@@
 We convert these soft constraints into linear objective functions of the binary variables defined in Section \ref{???}.
-The first two objectives can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to.
+Objectives (1) and (2) can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to.
 \begin{align}

### 2019-07-25 src/ui/ui_mainwindow.py
Resolve conflicts from master

@@ -1,6 +1,6 @@
 # -*- coding: utf-8 -*-
 
-# Form implementation generated from reading ui file 'src\designer\mainwindow.ui'
+# Form implementation generated from reading ui file 'designer\mainwindow.ui'
 #
 # Created by: PyQt5 UI code generator 5.11.3
 #
@@ -163,6 +163,68 @@ class Ui_MainWindow(object):
         self.scheduleTable.verticalHeader().setVisible(False)
         self.verticalLayout_2.addWidget(self.scheduleTable)
         self.tabWidget.addTab(self.schedulerTab, "")
+        self.settingsTab = QtWidgets.QWidget()
+        self.settingsTab.setObjectName("settingsTab")
+        self.verticalLayout_3 = QtWidgets.QVBoxLayout(self.settingsTab)
+        self.verticalLayout_3.setObjectName("verticalLayout_3")
+        self.constraintsGroupBox = QtWidgets.QGroupBox(self.settingsTab)
+        self.constraintsGroupBox.setObjectName("constraintsGroupBox")
+        self.formLayout = QtWidgets.QFormLayout(self.constraintsGroupBox)
+        self.formLayout.setObjectName("formLayout")
+        self.constraintsForm = QtWidgets.QFormLayout()
+        self.constraintsForm.setObjectName("constraintsForm")
+        self.blockCoverageLabel = QtWidgets.QLabel(self.constraintsGroupBox)
+        self.blockCoverageLabel.setObjectName("blockCoverageLabel")
+        self.constraintsForm.setWidget(0, QtWidgets.QFormLayout.LabelRole, self.blockCoverageLabel)
+        self.blockCoverageCheckBox = QtWidgets.QCheckBox(self.constraintsGroupBox)
+        self.blockCoverageCheckBox.setChecked(True)
+        self.blockCoverageCheckBox.setObjectName("blockCoverageCheckBox")
+        self.constraintsForm.setWidget(0, QtWidgets.QFormLayout.FieldRole, self.blockCoverageCheckBox)
+        self.weekendCoverageLabel = QtWidgets.QLabel(self.constraintsGroupBox)
+        self.weekendCoverageLabel.setObjectName("weekendCoverageLabel")
+        self.constraintsForm.setWidget(1, QtWidgets.QFormLayout.LabelRole, self.weekendCoverageLabel)
[…]

### 2019-07-26 src/ui/dialog.py
Fix app crash if empty row in division table

@@ -155,9 +155,12 @@ class DialogWindow(QDialog, Ui_Dialog):
         for i in range(self.divisionTable.rowCount()):
             division = dict()
 
-            divName = self.divisionTable.item(i, 0).text()
-            divMin = self.divisionTable.item(i, 1).text()
-            divMax = self.divisionTable.item(i, 2).text()
+            try:
+                divName = self.divisionTable.item(i, 0).text()
+                divMin = self.divisionTable.item(i, 1).text()
+                divMax = self.divisionTable.item(i, 2).text()
+            except AttributeError:
+                break
 
             division['min'] = int(divMin)
             division['max'] = int(divMax)

### 2019-07-29 paper/paper.results.tex
Add very rough idea of results section w/table

\subsection{Software}
We developed a software package with a user interface that implements the above LP problem and allows configuration of clinicians at [ref \ref{???}], to be used by the ID division at St. Michael's Hospital. The software was used to generate the results in the following sections, using real data as well as simulated data as input.

\subsection{Infectious Diseases Division}
We used clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem. Tables \ref{tbl:2018-schedule-comparison}, \ref{??}, \ref{??} present the optimal schedule generated using the software as well as the manually created schedule, color-coded to distinguish between the different clinicians assigned.

\include{tbl/2018_schedule_comparison}

In order to evaluate the generated schedule and compare it with the manually created schedule, we outline the adherence of each schedule to the constraints presented in section \ref{???}. As we can see, the generated schedule was able to satisfy all mandatory constraints, however for the years of [???] it could not find an optimal schedule while ensuring that the block assignments are spread out ([???] constraint). On the other hand, we can see that the manually created schedule was not able to satisfy all mandatory constraints. In particular, we see that it contains consecutive block assignments for the years of [???], [...]. Evaluating the objectives, we see that [...\textit{request conflicts}]. Moreover, the manual schedule does not attempt to align weekend assignments with block assignments, unlike the optimal schedule found by the software.

\subsection{Simulations}

### 2019-07-29 paper/tbl/2018_schedule_comparison.tex
Add very rough idea of results section w/table

% Table generated by Excel2LaTeX from sheet '2018'
\begin{table}[h]
	\tiny
 	\centering
    \begin{tabular}{c||ccc||ccc}
    	\multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} & \multicolumn{3}{c||}{LP Solution}                                                                                                                                       & \multicolumn{3}{c}{Historical Data}                                                                                                                                     \\
    	                                                  &                  HIV                  &                                 ID                                 &                              Weekend                               &                  HIV                  &                                 ID                                 &                              Weekend                               \\ \midrule\midrule
    	                        1                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                        2                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                        3                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                        4                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                        5                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                        6                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                        7                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .608,  .761,  .902}B                \\
    	                        8                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                        9                         & \cellcolor[rgb]{ .788,  .788,  .788}D &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .608,  .761,  .902}B &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       10                         & \cellcolor[rgb]{ .788,  .788,  .788}D &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       11                         & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       12                         & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       13                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       14                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       15                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       16                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       17                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       18                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       19                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .663,  .816,  .557}A &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       20                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       21                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       22                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       23                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       24                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       25                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .663,  .816,  .557}A                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       26                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       27                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       28                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       29                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       30                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       31                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       32                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       33                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       34                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       35                         & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       36                         & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       37                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       38                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .788,  .788,  .788}D &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       39                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       40                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       41                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       42                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .788,  .788,  .788}D &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       43                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       44                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       45                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .788,  .788,  .788}D &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       46                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       47                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       48                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       49                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       50                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       51                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E
    \end{tabular}%
	\caption{Comparison of schedules for 2018}
	\label{tbl:2018-schedule-comparison}%
\end{table}%

### 2019-07-29 paper/paper.results.tex
Add table for 2017 and comparison table

@@
 We used clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem.
-Tables \ref{tbl:2018-schedule-comparison}, \ref{??}, \ref{??} present the optimal schedule generated using the software as well as the manually created schedule, color-coded to distinguish between the different clinicians assigned.
+Tables \ref{tbl:2017-schedule-comparison}, \ref{tbl:2018-schedule-comparison}, \ref{??}, \ref{??} present the optimal schedule generated using the software as well as the manually created schedule, color-coded to distinguish between the different clinicians assigned.
 
+\include{tbl/2017_schedule_comparison}
 \include{tbl/2018_schedule_comparison}
@@
 In order to evaluate the generated schedule and compare it with the manually created schedule, we outline the adherence of each schedule to the constraints presented in section \ref{???}.
-As we can see, the generated schedule was able to satisfy all mandatory constraints, however for the years of [???] it could not find an optimal schedule while ensuring that the block assignments are spread out ([???] constraint).
+As we can see from table \ref{tbl:constraints-comparison}, the generated schedule was able to satisfy all mandatory constraints, however for the years of [???] it could not find an optimal schedule while ensuring that the block assignments are spread out ([???] constraint).
 On the other hand, we can see that the manually created schedule was not able to satisfy all mandatory constraints.
@@
 
+\include{tbl/constraints_comparison}
+
 \subsection{Simulations}

### 2019-07-29 paper/tbl/2017_schedule_comparison.tex
Add table for 2017 and comparison table

% Table generated by Excel2LaTeX from sheet '2017'
\begin{table}[h]
	\tiny
	\centering
    \begin{tabular}{c||ccc||ccc}
    	\multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} & \multicolumn{3}{c||}{LP Solution}                                                                                                                                                              & \multicolumn{3}{c}{Historical Data}                                                                                                                                                            \\
    	                                                  &                             HIV                             &                                 ID                                 &                              Weekend                               &                             HIV                             &                                 ID                                 &                              Weekend                               \\ \midrule\midrule
    	                        1                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .788,  .788,  .788}D                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                        2                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                        3                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                        4                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .663,  .816,  .557}A                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                        5                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                        6                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                        7                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                        8                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                        9                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                       10                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       11                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       12                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       13                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       14                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       15                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       16                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &                \cellcolor[rgb]{ .957,  .69,  .518}E                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                       17                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &              \cellcolor[rgb]{ 1,  .851,  .4}C               &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       18                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       19                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       20                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       21                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       22                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &            \cellcolor[rgb]{ .788,  .788,  .788}D            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       23                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       24                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       25                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .663,  .816,  .557}A                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       26                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       27                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       28                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       29                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       30                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       31                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .663,  .816,  .557}A                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .608,  .761,  .902}B                \\
    	                       32                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       33                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       34                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       35                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                       36                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       37                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &                \cellcolor[rgb]{ .957,  .69,  .518}E                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       38                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                       39                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       40                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       41                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       42                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       43                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       44                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       45                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       46                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .608,  .761,  .902}B            & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       47                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       48                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       49                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       50                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       51                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E
    \end{tabular}%
	\caption{Comparison of schedules for 2017}
	\label{tbl:2017-schedule-comparison}%
\end{table}%

### 2019-07-30 paper/paper.results.tex
Start simulations section

@@
 \subsection{Simulations}
+In this section, we want to analyze properties of the LP problem on simulated data.
+Our goal is to see whether the aforementioned formulation of the problem is going to suffer from run-time issues, when presented with a large set of constraints arising from a hypothetical department with many clinicians, or clinicians with many time-off requests throughout the year.
+We also want to analyze the behaviour of the LP problem when applied to a longer time horizon than just a single year.

### 2019-07-30 paper/paper.results.tex
Add a few plots w/descriptions to results

--- 2019-07-29 paper/paper.results.tex
@@
 \subsection{Simulations}
+In this section, we want to analyze properties of the LP problem on simulated data.
+Our goal is to see whether the aforementioned formulation of the problem is going to suffer from runtime issues, when presented with a large set of constraints arising from a hypothetical department with many clinicians, or clinicians with many time-off requests throughout the year.
+We also want to analyze the effect of a longer time-horizon or a large number of divisions in a department, on the run-time of the algorithm.
+\\
+
+We began by evaluating the runtime in a single-division department, with an increasing number of clinicians, over a 1-year time-span.
+Figure \ref{fig:avg-runtime-clinicians} shows the average runtime in seconds of 10 runs, for different numbers of clinicians, $C = \{5, 10, 15, 25, 50, 100\}$.
+From the graph, we can see that the LP formulation performs extremely well on a typical hospital department with no more than 30 clinicians [ref \ref{??}], and can even handle busier departments with upwards of 100 clinicians.
+\\
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=.6]{fig/avg_runtime_clinicians}
+	\caption{Plot of average runtime for LP with an increasing number of clinicians}
+	\label{fig:avg-runtime-clinicians}
+\end{figure}
+
+Next we evaluated the effect of increasing the number of departments.
+Figure \ref{fig:avg-runtime-divisions} shows the average runtime for a department with a 10 clinicians, and a range of divisions, $D = \{1, 2, 3\}$.
+We can see that increasing the number of divisions even by 2 greatly affects the runtime.
+The LP solver had significant trouble when we increased the number of divisions further, making it impractical to quickly generate schedules for departments with many divisions.
+\\
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=.6]{fig/avg_runtime_divisions}
+	\caption{Plot of average runtime for LP with an increasing number of divisions}
+	\label{fig:avg-runtime-divisions}
+\end{figure}
+
+We plotted the effect of an increasing number of requests per clinician on the average runtime of the LP solver in figure \ref{fig:avg-runtime-requests}.
+We ran the program on a single-division department with 10 clinicians, and $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{1, 2, 3, 5, 10\}$ non-overlapping requests per clinician.
+The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can handle and accommodate much flexibility in clinician requests.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=.6]{fig/avg_runtime_requests}
+	\caption{Plot of average runtime for LP with an increasing number of requests per clinician}
+	\label{fig:avg-runtime-requests}
+\end{figure}

### 2019-07-30 paper/paper.methods.tex
Add section labels

@@
-In this section, we present an application of 0-1 Linear Programming to solve the clinician scheduling problem presented in Section \ref{???}.
+In this section, we present an application of 0-1 Linear Programming to solve the clinician scheduling problem presented in Section \ref{sec:problem}.
 First, we describe the sets, indices and variables present in the formulation of the program.
@@
 
-\subsection{Sets and Indices}
+\subsection{Sets and Indices} \label{sec:meth-sets-indices}
 We denote the set of all services (or sometimes referred to as divisions) %or I think services is good (vs.
@@
 
-\subsection{Variables}
+\subsection{Variables} \label{sec:meth-variables}
 Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as $X_{c, b, d}$.
@@
 To formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, d}$, with additional constraints on its range.
-Further details regarding this variable are described in Sections \ref{???} and \ref{???}.
+Further details regarding this variable are described in Section \ref{sec:meth-objectives}.
 We then introduce a set of constants $m_{c, d}$ and $M_{c, d}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year.
@@
 
-\subsection{Constraints}
+\subsection{Constraints} \label{sec:meth-constraints}
 We now present the hard constraints in Table \ref{tbl:constraint-summary} using the variables defined above.
@@
 
-\subsection{Objectives}
-As described in Section \ref{???}, the soft constraints of the clinician scheduling problem include: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks.
-We convert these soft constraints into linear objective functions of the binary variables defined in Section \ref{???}.
+\subsection{Objectives} \label{sec:meth-objectives}
+As described in Section \ref{sec:problem}, the soft constraints of the clinician scheduling problem include: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks.
+We convert these soft constraints into linear objective functions of the binary variables defined in Section \ref{sec:meth-variables}.
 Objectives (1) and (2) can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to.

### 2019-07-30 paper/paper.results.tex
Add section labels

@@
 
-In order to evaluate the generated schedule and compare it with the manually created schedule, we outline the adherence of each schedule to the constraints presented in section \ref{???}.
+In order to evaluate the generated schedule and compare it with the manually created schedule, we outline the adherence of each schedule to the constraints presented in table \ref{tbl:constraint-summary}.
 As we can see from table \ref{tbl:constraints-comparison}, the generated schedule was able to satisfy all mandatory constraints, however for the years of [???] it could not find an optimal schedule while ensuring that the block assignments are spread out ([???] constraint).
@@
 The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can handle and accommodate much flexibility in clinician requests.
+\\
 
@@
 \end{figure}
+
+[...]

### 2019-07-30 paper/paper.results.tex
Minor tweaks in results section

@@
 	\centering
-	\includegraphics[scale=.6]{fig/avg_runtime_clinicians}
+	\includegraphics[scale=.7]{fig/avg_runtime_clinicians}
 	\caption{Plot of average runtime for LP with an increasing number of clinicians}
@@
 	\centering
-	\includegraphics[scale=.6]{fig/avg_runtime_divisions}
+	\includegraphics[scale=.7]{fig/avg_runtime_divisions}
 	\caption{Plot of average runtime for LP with an increasing number of divisions}
@@
 We ran the program on a single-division department with 10 clinicians, and $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{1, 2, 3, 5, 10\}$ non-overlapping requests per clinician.
-The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can handle and accommodate much flexibility in clinician requests.
+The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can accommodate a lot of flexibility in clinician requests.
 \\
@@
 	\centering
-	\includegraphics[scale=.6]{fig/avg_runtime_requests}
+	\includegraphics[scale=.7]{fig/avg_runtime_requests}
 	\caption{Plot of average runtime for LP with an increasing number of requests per clinician}

### 2019-07-30 paper/paper.results.tex
Fix past->present tense in results for consistency

@@
 \subsection{Software}
-We developed a software package with a user interface that implements the above LP problem and allows configuration of clinicians at [ref \ref{???}], to be used by the ID division at St.
+We develop a software package with a user interface that implements the above LP problem and allows configuration of clinicians at [ref \ref{???}], to be used by the ID division at St.
 Michael's Hospital.
@@
 \subsection{Infectious Diseases Division}
-We used clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem.
+We use clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem.
 Tables \ref{tbl:2017-schedule-comparison}, \ref{tbl:2018-schedule-comparison}, \ref{??}, \ref{??} present the optimal schedule generated using the software as well as the manually created schedule, color-coded to distinguish between the different clinicians assigned.
@@
 
-We began by evaluating the runtime in a single-division department, with an increasing number of clinicians, over a 1-year time-span.
+We begin by evaluating the runtime in a single-division department, with an increasing number of clinicians, over a 1-year time-span.
 Figure \ref{fig:avg-runtime-clinicians} shows the average runtime in seconds of 10 runs, for different numbers of clinicians, $C = \{5, 10, 15, 25, 50, 100\}$.
@@
 
-Next we evaluated the effect of increasing the number of departments.
+Next we evaluate the effect of increasing the number of departments.
 Figure \ref{fig:avg-runtime-divisions} shows the average runtime for a department with a 10 clinicians, and a range of divisions, $D = \{1, 2, 3\}$.
@@
 
-We plotted the effect of an increasing number of requests per clinician on the average runtime of the LP solver in figure \ref{fig:avg-runtime-requests}.
-We ran the program on a single-division department with 10 clinicians, and $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{1, 2, 3, 5, 10\}$ non-overlapping requests per clinician.
+We plot the effect of an increasing number of requests per clinician on the average runtime of the LP solver in figure \ref{fig:avg-runtime-requests}.
+We run the program on a single-division department with 10 clinicians, and $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{1, 2, 3, 5, 10\}$ non-overlapping requests per clinician.
 The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can accommodate a lot of flexibility in clinician requests.

### 2019-08-12 paper/paper.results.tex
Add avg runtime for increasing # blocks plot

@@
 
-[...]
+We evaluate the effect of a longer time-horizon on the runtime of the algorithm.
+Figure \ref{fig:avg-runtime-blocks} presents the average runtime over 10 runs for an increasing number of 2-week blocks $B = \{26, 52, 78, 104\}$ to simulate a department that generates schedules for a few years in advance.
+We can see that there is an increase in the runtime, however the solver is still able to find an optimal schedule for a 4-year time horizon in a very reasonable amount of time.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=.7]{fig/avg_runtime_blocks}
+	\caption{Plot of average runtime for LP with an increasing number of 2-week blocks}
+	\label{fig:avg-runtime-blocks}
+\end{figure}

### 2019-08-12 paper/paper.results.tex
Clean up table & wording in results section

@@
 In order to evaluate the generated schedule and compare it with the manually created schedule, we outline the adherence of each schedule to the constraints presented in table \ref{tbl:constraint-summary}.
-As we can see from table \ref{tbl:constraints-comparison}, the generated schedule was able to satisfy all mandatory constraints, however for the years of [???] it could not find an optimal schedule while ensuring that the block assignments are spread out ([???] constraint).
+As we can see from table \ref{tbl:constraints-comparison}, the optimal schedule generated using the software was able to satisfy all mandatory constraints.
 On the other hand, we can see that the manually created schedule was not able to satisfy all mandatory constraints.
-In particular, we see that it contains consecutive block assignments for the years of [???], [...].
+In particular, we see that it assigned clinicians to multiple consecutive blocks for all four years.
+Moreover, the manual schedule did not have an equal distribution of weekends and holidays for all four years of data.
 Evaluating the objectives, we see that [...\textit{request conflicts}].
-Moreover, the manual schedule does not attempt to align weekend assignments with block assignments, unlike the optimal schedule found by the software.
+Moreover, the manual schedule does not attempt to align weekend assignments adjacent to block assignments.
 

### 2019-08-13 paper/paper.results.tex
Fix minor typos in results

@@
 Next we evaluate the effect of increasing the number of departments.
-Figure \ref{fig:avg-runtime-divisions} shows the average runtime for a department with a 10 clinicians, and a range of divisions, $D = \{1, 2, 3\}$.
+Figure \ref{fig:avg-runtime-divisions} shows the average runtime for a department with 10 clinicians, and a range of divisions, $D = \{1, 2, 3\}$.
 We can see that increasing the number of divisions even by 2 greatly affects the runtime.
@@
 Figure \ref{fig:avg-runtime-blocks} presents the average runtime over 10 runs for an increasing number of 2-week blocks $B = \{26, 52, 78, 104\}$ to simulate a department that generates schedules for a few years in advance.
-We can see that there is an increase in the runtime, however the solver is still able to find an optimal schedule for a 4-year time horizon in a very reasonable amount of time.
+We can see that there is an increasing trend in the runtime, however the solver is still able to find an optimal schedule for a 4-year time horizon in a very reasonable amount of time.
 

### 2019-08-13 paper/paper.results.tex
Complete LP vs historical comparison

@@
 Moreover, the manual schedule did not have an equal distribution of weekends and holidays for all four years of data.
-Evaluating the objectives, we see that [...\textit{request conflicts}].
-Moreover, the manual schedule does not attempt to align weekend assignments adjacent to block assignments.
+Evaluating the objectives, we see that LP outperforms the manually created schedule in all four years, by accommodating almost all time-off requests and ensuring that weekends are always assigned close to blocks.
 

### 2019-08-19 src/services/scheduler.py
Add CBC logging when debugging

@@ -244,13 +244,11 @@ class Scheduler:
         self.setup_solver()
         self.setup_problem(shuffle=shuffle)
 
-        started_solving = time.clock()
         ret = self.problem.solve(self.solver) == pulp.LpStatusOptimal
-        finished_solving = time.clock()
                     
         if ret:
             if verbose:
-                self._logger.write_line('Solved in: {} s'.format(finished_solving - started_solving))
+                self._logger.write_line('Solved in: {} seconds'.format(self.problem.solutionTime))
                 self._logger.write_line('Objective function value: {}'.format(pulp.value(self.problem.objective)))
                 conflicts_str = 'Schedule Conflicts:'
                 for clinician in self.clinicians.values():
@@ -290,7 +288,7 @@ class Scheduler:
             self.solver = pulp.COIN_CMD(path=solverdir)
         else:
             # running from source
-            self.solver = pulp.LpSolverDefault
+            self.solver = pulp.PULP_CBC_CMD(msg=1)
 
     def set_data(self, data):
         """

### 2019-08-19 paper/paper.results.tex
Replace divisions graph with table
Move divisions results to end of results section, rewrite its analysis portion

@@
 
-We begin by evaluating the runtime in a single-division department, with an increasing number of clinicians, over a 1-year time-span.
+%We begin by evaluating the runtime in a single-division department, with an increasing number of clinicians, over a 1-year time-span.
 Figure \ref{fig:avg-runtime-clinicians} shows the average runtime in seconds of 10 runs, for different numbers of clinicians, $C = \{5, 10, 15, 25, 50, 100\}$.
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=.7]{fig/avg_runtime_clinicians}
-	\caption{Plot of average runtime for LP with an increasing number of clinicians}
-	\label{fig:avg-runtime-clinicians}
-\end{figure}
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[scale=.7]{fig/avg_runtime_clinicians}
+%	\caption{Plot of average runtime for LP with an increasing number of clinicians}
+%	\label{fig:avg-runtime-clinicians}
+%\end{figure}
 
-Next we evaluate the effect of increasing the number of departments.
+%Next we evaluate the effect of increasing the number of departments.
 Figure \ref{fig:avg-runtime-divisions} shows the average runtime for a department with 10 clinicians, and a range of divisions, $D = \{1, 2, 3\}$.
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=.7]{fig/avg_runtime_divisions}
-	\caption{Plot of average runtime for LP with an increasing number of divisions}
-	\label{fig:avg-runtime-divisions}
-\end{figure}
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[scale=.7]{fig/avg_runtime_divisions}
+%	\caption{Plot of average runtime for LP with an increasing number of divisions}
+%	\label{fig:avg-runtime-divisions}
+%\end{figure}
 
@@
 We can see that there is an increasing trend in the runtime, however the solver is still able to find an optimal schedule for a 4-year time horizon in a very reasonable amount of time.
+\\
 
@@
 \end{figure}
+
+Lastly, we analyze the effect of increasing the number of divisions on the runtime of the program.
+Table \ref{tbl:runtime-divisions-comparison} presents the runtime for $D = \{1, 2, 3\}$ divisions and $C = \{10, 20, 30, 50\}$ clinicians in total across all divisions.
+For 2 divisions, a roster of more than 20 clinicians becomes impractical, if all constraints must be satisfied, but the runtime can be improved if we allow consecutive blocks to be assigned.
+For 3 divisions, the problem becomes impractical faster, with any more than 10 clinicians leading to unreasonable runtime for finding an optimal solution, and the problem cannot be improved by relaxing the consecutive blocks constraint.
+
+\include{tbl/runtime_divisions}

### 2019-08-19 paper/paper.results.tex
Remove "Plot of.." in figure captions

@@
 	\includegraphics[scale=.7]{fig/avg_runtime_requests}
-	\caption{Plot of average runtime for LP with an increasing number of requests per clinician}
+	\caption{Average runtime for LP with an increasing number of requests per clinician}
 	\label{fig:avg-runtime-requests}
@@
 	\includegraphics[scale=.7]{fig/avg_runtime_blocks}
-	\caption{Plot of average runtime for LP with an increasing number of 2-week blocks}
+	\caption{Average runtime for LP with an increasing number of 2-week blocks}
 	\label{fig:avg-runtime-blocks}

### 2019-08-19 paper/tbl/runtime_divisions.tex
Change unsatisfied to satisfied requests in table
Add missing tex file for runtime comparison

\begin{table}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
		\hline
		&       \multicolumn{4}{c|}{Number of Divisions}        \\ \hline
		Number of Clinicians &  1   &   2    & 2 (Consecutive Blocks Allowed) &  3   \\ \hline
		10          & 0.22 &  0.51  &              0.26              & 1.49 \\ \hline
		20          & 0.42 & 139.42 &              0.50              &     \\ \hline
		30          & 0.63 &   -    &              0.90              &     \\ \hline
		50          & 1.06 &   -    &              1.30              &     \\ \hline
	\end{tabular}
	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians. ``-'' indicates that a solution was not found after 24 hours.}
	\label{tbl:runtime-divisions-comparison}%
\end{table}

### 2019-08-19 paper/paper.results.tex
Update division runtime table and rephrase

@@
 Table \ref{tbl:runtime-divisions-comparison} presents the runtime for $D = \{1, 2, 3\}$ divisions and $C = \{10, 20, 30, 50\}$ clinicians in total across all divisions.
-For 2 divisions, a roster of more than 20 clinicians becomes impractical, if all constraints must be satisfied, but the runtime can be improved if we allow consecutive blocks to be assigned.
-For 3 divisions, the problem becomes impractical faster, with any more than 10 clinicians leading to unreasonable runtime for finding an optimal solution, and the problem cannot be improved by relaxing the consecutive blocks constraint.
+For 2 divisions, a roster of more than 20 clinicians becomes impractical to solve, as searching requires more than 24 hours.
+For 3 divisions we can see issues with runtime occur earlier, where configurations of more than 10 clinicians are unable to be solved in 24 hours.
+When relaxing the consecutive blocks constraint, we see a great improvement in runtime for 2 and 3 divisions, and even a schedule for 50 clinicians or more can be generated.
 

### 2019-08-19 paper/tbl/runtime_divisions.tex
Update division runtime table and rephrase

@@
 	\centering
-	\begin{tabular}{|c|c|c|c|c|}
+	\begin{tabular}{|c|c|c|c|c|c|}
 		\hline
-		&       \multicolumn{4}{c|}{Number of Divisions}        \\ \hline
-		Number of Clinicians &  1   &   2    & 2 (Consecutive Blocks Allowed) &  3   \\ \hline
-		10          & 0.22 &  0.51  &              0.26              & 1.49 \\ \hline
-		20          & 0.42 & 139.42 &              0.50              &     \\ \hline
-		30          & 0.63 &   -    &              0.90              &     \\ \hline
-		50          & 1.06 &   -    &              1.30              &     \\ \hline
+		                                      & \multicolumn{5}{c|}{Number of Divisions} \\ \hline
+		\makecell[l]{Number of \\ Clinicians} &  1   &   2   & 2 (CBA) &  3   & 3 (CBA)  \\ \hline
+		                 10                   & 0.18 & 0.94  &  0.24   & 2.02 &   0.30   \\ \hline
+		                 20                   & 0.33 & 13.84 &  0.42   &      &   1.17   \\ \hline
+		                 30                   & 0.51 &   -   &  0.68   &      &   0.93   \\ \hline
+		                 50                   & 0.83 &   -   &  1.12   &      &   1.55   \\ \hline
 	\end{tabular}
-	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians. ``-'' indicates that a solution was not found after 24 hours.}
+	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians.
+``-'' indicates that a solution was not found after 24 hours.
+``CBA'' indicates that \underline{c}onsecutive \underline{b}locks were \underline{a}llowed for these schedules.}
 	\label{tbl:runtime-divisions-comparison}%

### 2019-08-19 paper/paper.results.tex
Remove 2017 table + use adjustbox to scale table

@@
 We use clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem.
-Tables \ref{tbl:2017-schedule-comparison}, \ref{tbl:2018-schedule-comparison}, \ref{??}, \ref{??} present the optimal schedule generated using the software as well as the manually created schedule, color-coded to distinguish between the different clinicians assigned.
+Table \ref{tbl:2018-schedule-comparison} presents a the optimal schedule generated using the software as well as the manually created schedule for data from 2018, color-coded to distinguish between the different clinicians assigned.
 
-\include{tbl/2017_schedule_comparison}
 \include{tbl/2018_schedule_comparison}

### 2019-08-19 src/services/scheduler.py
Revert back to consecutive weekends constraint

@@ -479,6 +479,17 @@ class Scheduler:
                 )
                 self.problem.add(sum_ <= 1)
 
+    def _constraint_consec_weekends(self, divisions=None, clinicians=None):
+        for clinician in clinicians:
+            for week_num in range(1, self.num_weekends):
+                    sum_ = pulp.lpSum(
+                        [_.get_var() for _ in clinician.get_weekend_vars(
+                            lambda x, week_num=week_num: x.week_num in (
+                                week_num, week_num + 1)
+                        )]
+                    )
+                    self.problem.add(sum_ <= 1)
+
     def _constraint_spread_blocks(self, divisions=None, clinicians=None):
         # on-off-on-off-on constraint for block assignment
         # we need at least 5 consecutive blocks to implement this constraint
@@ -644,8 +655,9 @@ class Scheduler:
         'coverage': _constraint_coverage,
         'minMaxBlocks': _constraint_minmax_blocks,
         'preventConsecutiveBlocks': _constraint_consec_blocks,
-        'spreadBlocks': _constraint_spread_blocks,
-        'spreadWeekends': _constraint_spread_weekends,
+        'preventConsecutiveWeekends': _constraint_consec_weekends,
+        # 'spreadBlocks': _constraint_spread_blocks,
+        # 'spreadWeekends': _constraint_spread_weekends,
         'balancedLongWeekends': _constraint_balance_longweekends,
         'balancedWeekends': _constraint_balance_weekends
     }

### 2019-08-19 src/ui/ui_mainwindow.py

@@ -1,6 +1,6 @@
 # -*- coding: utf-8 -*-
 
-# Form implementation generated from reading ui file 'designer\mainwindow.ui'
+# Form implementation generated from reading ui file 'src\designer\mainwindow.ui'
 #
 # Created by: PyQt5 UI code generator 5.11.3
 #
@@ -208,20 +208,13 @@ class Ui_MainWindow(object):
         self.preventConsecutiveBlocksCheckBox.setChecked(True)
         self.preventConsecutiveBlocksCheckBox.setObjectName("preventConsecutiveBlocksCheckBox")
         self.constraintsForm.setWidget(4, QtWidgets.QFormLayout.FieldRole, self.preventConsecutiveBlocksCheckBox)
-        self.spreadBlocksLabel = QtWidgets.QLabel(self.constraintsGroupBox)
-        self.spreadBlocksLabel.setObjectName("spreadBlocksLabel")
-        self.constraintsForm.setWidget(5, QtWidgets.QFormLayout.LabelRole, self.spreadBlocksLabel)
-        self.spreadBlocksCheckBox = QtWidgets.QCheckBox(self.constraintsGroupBox)
-        self.spreadBlocksCheckBox.setChecked(True)
-        self.spreadBlocksCheckBox.setObjectName("spreadBlocksCheckBox")
-        self.constraintsForm.setWidget(5, QtWidgets.QFormLayout.FieldRole, self.spreadBlocksCheckBox)
-        self.spreadWeekendsLabel = QtWidgets.QLabel(self.constraintsGroupBox)
-        self.spreadWeekendsLabel.setObjectName("spreadWeekendsLabel")
-        self.constraintsForm.setWidget(6, QtWidgets.QFormLayout.LabelRole, self.spreadWeekendsLabel)
-        self.spreadWeekendsCheckBox = QtWidgets.QCheckBox(self.constraintsGroupBox)
-        self.spreadWeekendsCheckBox.setChecked(True)
-        self.spreadWeekendsCheckBox.setObjectName("spreadWeekendsCheckBox")
-        self.constraintsForm.setWidget(6, QtWidgets.QFormLayout.FieldRole, self.spreadWeekendsCheckBox)
+        self.preventConsecutiveWeekendsLabel = QtWidgets.QLabel(self.constraintsGroupBox)
+        self.preventConsecutiveWeekendsLabel.setObjectName("preventConsecutiveWeekendsLabel")
[…]

### 2019-08-19 paper/paper.discussion.tex
Fix runtime table + start discussion section

In this paper our goal was to develop a rather simple, yet flexible, integer linear programming formulation to generate schedules for clinical consultation (?) departments at hospitals. The difficulty in applying a linear programming approach to this task lies in the fact that ILP is an NP-hard problem, and as such the time to find an optimal solution grows exponentially as we increase the size of the problem. As a result, the vast majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time. However, even with the limitations of time complexity we are able to use the tool to generate schedules for a wide variety of simulated departments, supporting upwards of 50 total clinicians. Moreover, our formulation is able to accommodate the requests of clinicians without negatively affecting the runtime. Comparing the generated schedule to a manually created schedule in a real clinical department at a hospital, we see adherence to all required constraints as well as improved fulfillment of clinician requests, which are sometimes overlooked when creating the schedule by hand. The scenarios when the ILP formulation starts to have trouble are ones with multiple services offered within a single clinical department. Once there are more than 10 clinicians in total, providing services in 2 or more divisions, the constraint space becomes very complex, making it difficult for the tool to find an optimal solution in a reasonable amount of time. For these cases, we recommend relaxing the `No Consecutive Blocks` constraint, to simplify the problem and find a solution much faster.

### 2019-08-19 paper/paper.discussion.tex
Work on discussion: split into multiple paragraphs

@@
 As a result, the vast majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time.
-However, even with the limitations of time complexity we are able to use the tool to generate schedules for a wide variety of simulated departments, supporting upwards of 50 total clinicians.
+\\
+
+Our formulation includes hard constraints to ensure the schedule satisfies hospital and logistics requirements.
+It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function.
+We are able to use the tool to generate schedules for a wide variety of simulated departments, supporting upwards of 50 total clinicians.
 Moreover, our formulation is able to accommodate the requests of clinicians without negatively affecting the runtime.
-Comparing the generated schedule to a manually created schedule in a real clinical department at a hospital, we see adherence to all required constraints as well as improved fulfillment of clinician requests, which are sometimes overlooked when creating the schedule by hand.
 The scenarios when the ILP formulation starts to have trouble are ones with multiple services offered within a single clinical department.
 Once there are more than 10 clinicians in total, providing services in 2 or more divisions, the constraint space becomes very complex, making it difficult for the tool to find an optimal solution in a reasonable amount of time.
-For these cases, we recommend relaxing the `No Consecutive Blocks` constraint, to simplify the problem and find a solution much faster.
+For these cases, we recommend allowing consecutive blocks, to simplify the problem and find a solution much faster.
+\\
+
+Our tool is also employed at the Infectious Diseases Department at St.
+Michael's Hospital.
+It provides a simple and easy-to-use user interface that can be used by any hospital staff to generate schedules well in advance.
+We compare the manually created schedule at St.
+Michael's Hospital to the optimal schedule generated with the help of our tool.
+We see adherence to all required hard constraints in the generated schedule, while some of them were overlooked in the manual schedule.
+Moreover, we see a much better fulfillment of clinician requests than in the manual schedule.
+This reinforces the idea that it is vital to employ automated tools when generating schedules in hospital departments, to reduce human error, balance the work-load of clinicians and improve the service provided to patients.

### 2019-08-19 paper/paper.results.tex
Work on discussion: split into multiple paragraphs

@@
 Table \ref{tbl:runtime-divisions-comparison} presents the runtime for $D = \{1, 2, 3\}$ divisions and $C = \{10, 20, 30, 50\}$ clinicians in total across all divisions.
-For 2 divisions, a roster of more than 20 clinicians becomes impractical to solve, as searching requires more than 24 hours.
-For 3 divisions we can see issues with runtime occur earlier, where configurations of more than 10 clinicians are unable to be solved in 24 hours.
-When relaxing the consecutive blocks constraint, we see a great improvement in runtime for 2 and 3 divisions, and even a schedule for 50 clinicians or more can be generated.
+For 2 or more divisions, a roster of more than 10 clinicians becomes impractical to schedule, as searching requires more than 24 hours.
+However, when relaxing the `No Consecutive Blocks' constraint, we see a great improvement in runtime for 2 and 3 divisions, and we are able to generate a schedule with upwards of 50 clinicians.
 

### 2019-08-19 paper/tbl/runtime_divisions.tex
Fix runtime table + start discussion section
Work on discussion: split into multiple paragraphs

--- 2019-08-19 paper/tbl/runtime_divisions.tex
@@
-\begin{table}
+\begin{table}[htbp]
 	\centering
-	\begin{tabular}{|c|c|c|c|c|}
+	\begin{tabular}{|c|c||c|c||c|c|}
 		\hline
-		&       \multicolumn{4}{c|}{Number of Divisions}        \\ \hline
-		Number of Clinicians &  1   &   2    & 2 (Consecutive Blocks Allowed) &  3   \\ \hline
-		10          & 0.22 &  0.51  &              0.26              & 1.49 \\ \hline
-		20          & 0.42 & 139.42 &              0.50              &     \\ \hline
-		30          & 0.63 &   -    &              0.90              &     \\ \hline
-		50          & 1.06 &   -    &              1.30              &     \\ \hline
+		                                      & \multicolumn{5}{c|}{Number of Divisions} \\ \hline
+		\makecell[l]{Number of \\ Clinicians} &  1   &  2   & 2 (CBA) &  3   &  3 (CBA)  \\ \hline
+		                 10                   & 0.19 & 0.95 &  0.21   & 1.90 &   0.31    \\ \hline
+		                 20                   & 0.31 &  -   &  0.41   &  -   &   0.56    \\ \hline
+		                 30                   & 0.57 &  -   &  0.68   &  -   &   0.90    \\ \hline
+		                 50                   & 0.78 &  -   &  1.09   &  -   &   1.66    \\ \hline
 	\end{tabular}
-	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians. ``-'' indicates that a solution was not found after 24 hours.}
+	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians.
+`-' indicates that a solution was not found after 24 hours.
+`CBA' indicates that \underline{c}onsecutive \underline{b}locks were \underline{a}llowed when generating these schedules.}
 	\label{tbl:runtime-divisions-comparison}%

### 2019-08-19 paper/paper.discussion.tex
Add future work paragraph to discussion

@@
 This reinforces the idea that it is vital to employ automated tools when generating schedules in hospital departments, to reduce human error, balance the work-load of clinicians and improve the service provided to patients.
+\\
+
+In future work, we can look first and foremost into improving the runtime of the tool by modifying constraints and objectives, so that it can be used in larger departments offering multiple services for their patients.
+On top of that, we can introduce new objectives that incorporate additional clinician preferences and allow for better distribution of work-load.
+However, since our ultimate goal is to maintain a ILP formulation, care must be taken not to overly complicate the model and risk running into additional time complexity issues.

### 2019-08-20 src/services/scheduler.py
Add CBC printing to compiled version

@@ -285,7 +285,7 @@ class Scheduler:
             exe = 'cbc\\bin\\cbc.exe'
             if hasattr(sys, '_MEIPASS'): solverdir = os.path.join(sys._MEIPASS, exe)
             else: solverdir = exe
-            self.solver = pulp.COIN_CMD(path=solverdir)
+            self.solver = pulp.COIN_CMD(path=solverdir, msg=1)
         else:
             # running from source
             self.solver = pulp.PULP_CBC_CMD(msg=1)

### 2019-08-21 paper/paper.results.tex
SM comments/edits to results
Very good job. I like the variables you chose to examine with the simulated data. I cannot think of others at this time.

@@
 \subsection{Software}
-We develop a software package with a user interface that implements the above LP problem and allows configuration of clinicians at [ref \ref{???}], to be used by the ID division at St.
+We developed a software package with a user interface that implements the above LP problem and allows configuration of clinicians at [ref \ref{???}], to be used by the ID division at St.
 Michael's Hospital.
@@
 
-\subsection{Infectious Diseases Division}
+\subsection{Infectious Diseases Division}  %main comment  = I think use past tense in writing
 We use clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem.
-Table \ref{tbl:2018-schedule-comparison} presents a the optimal schedule generated using the software as well as the manually created schedule for data from 2018, color-coded to distinguish between the different clinicians assigned.
+Table \ref{tbl:2018-schedule-comparison} presents a the optimal schedule generated using the software as well as the manually-created schedule for data from 2018.
+The schedule is color-coded to distinguish between the different clinicians assigned.
 
 \include{tbl/2018_schedule_comparison}
-
-In order to evaluate the generated schedule and compare it with the manually created schedule, we outline the adherence of each schedule to the constraints presented in table \ref{tbl:constraint-summary}.
-As we can see from table \ref{tbl:constraints-comparison}, the optimal schedule generated using the software was able to satisfy all mandatory constraints.
-On the other hand, we can see that the manually created schedule was not able to satisfy all mandatory constraints.
-In particular, we see that it assigned clinicians to multiple consecutive blocks for all four years.
+First, we evaluated the software-generated schedule by comparing it with the manually-created schedule.
+Specifically, we examined the adherence of each schedule to the constraints presented in table \ref{tbl:constraint-summary}.
+As shown in table \ref{tbl:constraints-comparison}, the optimal schedule generated using the software satisfied all mandatory constraints.
+In contrast, the manually-created schedule was not able to satisfy all mandatory constraints.
+In particular, we see that manual schedule assigned clinicians to multiple consecutive blocks for all four years.
 Moreover, the manual schedule did not have an equal distribution of weekends and holidays for all four years of data.
@@
 \subsection{Simulations}
-In this section, we want to analyze properties of the LP problem on simulated data.
-Our goal is to see whether the aforementioned formulation of the problem is going to suffer from runtime issues, when presented with a large set of constraints arising from a hypothetical department with many clinicians, or clinicians with many time-off requests throughout the year.
-We also want to analyze the effect of a longer time-horizon or a large number of divisions in a department, on the run-time of the algorithm.
+We then examined the properties of the LP problem using simulated data.
+Specifically, we examined the influence of expanding the following constraints on the runtime of the algorithm: larger number of  clinicians; clinicians with many time-off requests throughout the year.
+We then examined the effect of a longer time-horizon or a large number of services within a single department serviced by the same clinicians, on the runtime.
 \\
@@
 
-We plot the effect of an increasing number of requests per clinician on the average runtime of the LP solver in figure \ref{fig:avg-runtime-requests}.
-We run the program on a single-division department with 10 clinicians, and $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{1, 2, 3, 5, 10\}$ non-overlapping requests per clinician.
+The effect of an increasing number of requests per clinician on the average runtime of the LP solver is shown in figure \ref{fig:avg-runtime-requests}.
+We executed the program on a single-division department with 10 clinicians, and $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{1, 2, 3, 5, 10\}$ non-overlapping requests per clinician.
 The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can accommodate a lot of flexibility in clinician requests.
@@
 
-We evaluate the effect of a longer time-horizon on the runtime of the algorithm.
 Figure \ref{fig:avg-runtime-blocks} presents the average runtime over 10 runs for an increasing number of 2-week blocks $B = \{26, 52, 78, 104\}$ to simulate a department that generates schedules for a few years in advance.
-We can see that there is an increasing trend in the runtime, however the solver is still able to find an optimal schedule for a 4-year time horizon in a very reasonable amount of time.
-\\
+Although here was an increasing trend in the runtime, the solver was able to find an optimal schedule for a 4-year time horizon in a very reasonable amount of time.
+\\ % 'very reasonable amount of time" is vague.
+what is reasonable?
+better to give a range in value.
 
@@
 
-Lastly, we analyze the effect of increasing the number of divisions on the runtime of the program.
-Table \ref{tbl:runtime-divisions-comparison} presents the runtime for $D = \{1, 2, 3\}$ divisions and $C = \{10, 20, 30, 50\}$ clinicians in total across all divisions.
+The effect of increasing the number of divisions %do you mean services?
+on the runtime of the program is shown in Table \ref{tbl:runtime-divisions-comparison} which presents the runtime for $D = \{1, 2, 3\}$ divisions and $C = \{10, 20, 30, 50\}$ clinicians in total across all divisions.
 For 2 or more divisions, a roster of more than 10 clinicians becomes impractical to schedule, as searching requires more than 24 hours.
 However, when relaxing the `No Consecutive Blocks' constraint, we see a great improvement in runtime for 2 and 3 divisions, and we are able to generate a schedule with upwards of 50 clinicians.
+%what does division mean?
+do you mean services?
 

### 2019-08-21 paper/paper.discussion.tex
Rename divisions -> services throughout
Also additional changes to results section following comments

@@
 The scenarios when the ILP formulation starts to have trouble are ones with multiple services offered within a single clinical department.
-Once there are more than 10 clinicians in total, providing services in 2 or more divisions, the constraint space becomes very complex, making it difficult for the tool to find an optimal solution in a reasonable amount of time.
+Once there are more than 10 clinicians in total, providing 2 or more services, the search space becomes very complex, making it difficult for the tool to find an optimal solution in a reasonable amount of time.
 For these cases, we recommend allowing consecutive blocks, to simplify the problem and find a solution much faster.

### 2019-08-21 paper/paper.problem.tex
Rename divisions -> services throughout
Also additional changes to results section following comments

@@
 		\textbf{Constraint Name} & \textbf{Description}                                                                                                           & \textbf{Type} \\ \hline
-		Block Coverage           & \makecell[l]{each division needs to have exactly \\ one clinician that covers any given block}                                 & Hard          \\ \hline
+		Block Coverage           & \makecell[l]{each service needs to have exactly \\ one clinician that covers any given block}                                 & Hard          \\ \hline
 		Weekend Coverage         & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              & Hard          \\ \hline
-		Min/Max                  & \makecell[l]{in a given division, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} & Hard          \\ \hline
-		No Consecutive Blocks    & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all divisions}                                    & Hard          \\ \hline
+		Min/Max                  & \makecell[l]{for a given service, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} & Hard          \\ \hline
+		No Consecutive Blocks    & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all services}                                    & Hard          \\ \hline
 		No Consecutive Weekends  & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        & Hard          \\ \hline

### 2019-08-21 paper/paper.results.tex
Rename divisions -> services throughout
Also additional changes to results section following comments

@@
 \subsection{Infectious Diseases Division}  %main comment  = I think use past tense in writing
-We use clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem.
-Table \ref{tbl:2018-schedule-comparison} presents a the optimal schedule generated using the software as well as the manually-created schedule for data from 2018.
+We used clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem.
+Table \ref{tbl:2018-schedule-comparison} presents the optimal schedule generated using the software as well as the manually-created schedule for data from 2018.
 The schedule is color-coded to distinguish between the different clinicians assigned.
+\\
 
 \include{tbl/2018_schedule_comparison}
+
 First, we evaluated the software-generated schedule by comparing it with the manually-created schedule.
@@
 In contrast, the manually-created schedule was not able to satisfy all mandatory constraints.
-In particular, we see that manual schedule assigned clinicians to multiple consecutive blocks for all four years.
+In particular, we see that the manual schedule assigned clinicians to multiple consecutive blocks in all four years.
 Moreover, the manual schedule did not have an equal distribution of weekends and holidays for all four years of data.
@@
 We then examined the properties of the LP problem using simulated data.
-Specifically, we examined the influence of expanding the following constraints on the runtime of the algorithm: larger number of  clinicians; clinicians with many time-off requests throughout the year.
-We then examined the effect of a longer time-horizon or a large number of services within a single department serviced by the same clinicians, on the runtime.
+Specifically, we examined the influence of expanding the following parameters on the runtime of the algorithm: number of clinicians; number of services offered in the division; amount of time-off requests per clinician throughout the year.
+We also examined the effect of a longer time-horizon on the runtime.
 \\
@@
 Figure \ref{fig:avg-runtime-blocks} presents the average runtime over 10 runs for an increasing number of 2-week blocks $B = \{26, 52, 78, 104\}$ to simulate a department that generates schedules for a few years in advance.
-Although here was an increasing trend in the runtime, the solver was able to find an optimal schedule for a 4-year time horizon in a very reasonable amount of time.
+Although here was an increasing trend in the runtime, the solver was able to find an optimal schedule for all time horizons in under XX amount of time.
 \\ % 'very reasonable amount of time" is vague.
@@
 
-The effect of increasing the number of divisions %do you mean services?
-on the runtime of the program is shown in Table \ref{tbl:runtime-divisions-comparison} which presents the runtime for $D = \{1, 2, 3\}$ divisions and $C = \{10, 20, 30, 50\}$ clinicians in total across all divisions.
-For 2 or more divisions, a roster of more than 10 clinicians becomes impractical to schedule, as searching requires more than 24 hours.
-However, when relaxing the `No Consecutive Blocks' constraint, we see a great improvement in runtime for 2 and 3 divisions, and we are able to generate a schedule with upwards of 50 clinicians.
+The effect of increasing the number of services %divisions %do you mean services?
+on the runtime of the program is shown in Table \ref{tbl:runtime-services-clinicians-comparison}.
+We executed the algorithm for $S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in total across all services.
+For 2 or more services, a roster of more than XX clinicians becomes impractical to schedule, as searching requires more than 24 hours.
+However, when relaxing the `No Consecutive Blocks' constraint, we see a great improvement in runtime for 2 and 3 services, and we are able to generate a schedule with upwards of 50 clinicians in under XX amount of time.
 %what does division mean?
@@
 
-\include{tbl/runtime_divisions}
+\include{tbl/runtime_services_clinicians}

### 2019-08-21 paper/paper.methods.tex
Rename divisions -> services throughout
Also additional changes to results section following comments

@@
 \subsection{Sets and Indices} \label{sec:meth-sets-indices}
-We denote the set of all services (or sometimes referred to as divisions) %or I think services is good (vs.
-services/divisions) as you define two types of services (ID vs.
-HIV) in the problem statement
-that clinicians in a single department can provide as $\mathcal{D}$.
+We denote the set of all services %or I think services is good (vs. services/divisions) as you define two types of services (ID vs. HIV) in the problem statement
+that clinicians in a single department can provide as $\mathcal{S}$.
 The following formulation assumes that all clinicians are able to provide all services.
@@
 The sets of blocks and weekends that clinicians will be assigned to are denoted as $\mathcal{B}$ and $\mathcal{W}$ respectively.
-The size of a block is not constrained and can be adapted %how?
-to the needs of the given department.
+The block size used in our experiments is 2 weeks, but the following LP formulation does not specify a particular size for blocks, and so it can be adapted to the needs of the department in question.
+%The size of a block is not constrained and can be adapted %how?
+%to the needs of the given department. 
 A subset of weekends are denoted $\mathcal{L}$, corresponding to established long/holiday weekends such as the Canadian Civic Day weekend in August.
@@
 		\textbf{Set}                         & \textbf{Index} & \textbf{Description}              \\ \hline
-		$\mathcal{D} = \{1, \ldots, D \}$    & $d$            & services/divisions                \\
+		$\mathcal{S} = \{1, \ldots, S \}$    & $s$            & services                          \\
 		$\mathcal{C} = \{1, \ldots, C \}$    & $c$            & clinicians                        \\
@@
 \subsection{Variables} \label{sec:meth-variables}
-Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as $X_{c, b, d}$.
-A value of 1 indicates that the given clinician $c$ is assigned to cover division $d$ during block $b$.
+Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as $X_{c, b, s}$.
+A value of 1 indicates that the given clinician $c$ is assigned to provide service $s$ during block $b$.
 Similarly, we define weekend assignments as $Y_{c, w}$, although these are not indexed by service, as clinicians are expected to provide all services during the weekends.
-To optimize the soft constraint Block-Weekend Adjacency, we maximize the product $X_{c, b, d} \cdot Y_{c, w}$ for adjacent blocks and weekends.
-To formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, d}$, with additional constraints on its range.
+To optimize the soft constraint Block-Weekend Adjacency, we maximize the product $X_{c, b, s} \cdot Y_{c, w}$ for adjacent blocks and weekends.
+To formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, s}$, with additional constraints on its range.
 Further details regarding this variable are described in Section \ref{sec:meth-objectives}.
-We then introduce a set of constants $m_{c, d}$ and $M_{c, d}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year.
+We then introduce a set of constants $m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year.
 Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.
@@
 		\textbf{Name}              & \textbf{Description}                                               \\ \hline
-		$X_{c, b, d} \in \{0, 1\}$ & assignment of clinician $c$ for service/division $d$ on block $b$    \\
+		$X_{c, b, s} \in \{0, 1\}$ & assignment of clinician $c$ to service $s$ on block $b$            \\
 		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                         \\
-		$Z_{c, b, d} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend adjacency               \\
-		$m_{c, d}$                 & minimum number of blocks clinician $c$ should cover for division $d$ \\
-		$M_{c, d}$                 & maximum number of blocks clinician $c$ should cover for division $d$
+		$Z_{c, b, s} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend adjacency             \\
+		$m_{c, s}$                 & minimum number of blocks clinician $c$ should cover on service $s$ \\
+		$M_{c, s}$                 & maximum number of blocks clinician $c$ should cover on service $s$
 	\end{tabular}
@@
 \begin{align}
-	&\sum_{c=1}^{C} X_{c, b, d} = 1 &&\forall b\in \mathcal{B}, d \in \mathcal{D} \tag{Block Coverage} \label{eqn:constr-block-cov} \\
+	&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S} \tag{Block Coverage} \label{eqn:constr-block-cov} \\
 	&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{Weekend Coverage} \label{eqn:constr-weekend-cov} \\
-	&m_{c, d} \leq \sum_{b=1}^{B} X_{c, b, d} \leq M_{c, d} &&\forall\ c\in\mathcal{C}, d\in\mathcal{D} \tag{Min/Max} \label{eqn:constr-min-max} \\
-	&X_{c, b, d} + X_{c, b + 1, d} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, d\in\mathcal{D} \tag{No Consecutive Blocks} \label{eqn:constr-no-consec-blocks} \\
+	&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\ c\in\mathcal{C}, s\in\mathcal{S} \tag{Min/Max} \label{eqn:constr-min-max} \\
+	&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, s\in\mathcal{S} \tag{No Consecutive Blocks} \label{eqn:constr-no-consec-blocks} \\
 	&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{No Consecutive Weekends} \label{eqn:constr-no-consec-weekends} \\
@@
 \begin{align}
-	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, d} \tag{Block Requests} \label{eqn:obj-block-requests}\\
+	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{Block Requests} \label{eqn:obj-block-requests}\\
 	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{Weekend Requests} \label{eqn:obj-weekend-requests}
@@
 In the above two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments.
-The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, d}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$.
+The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, s}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$.
 For instance, clinicians might want to be assigned during a weekend that falls within an assigned block.
 So if clinician $c$ is assigned to work during block 3, corresponding to weeks 5 and 6 assuming 2-week blocks, they might also want to be assigned to work during weekend 5.
-In that case, we would like $X_{c, b=3, d} \cdot Y_{c, w=5}$ to be 1, since that indicates both variables are assigned.
+In that case, we would like $X_{c, b=3, s} \cdot Y_{c, w=5}$ to be 1, since that indicates both variables are assigned.
 If at least one of the two variables is not assigned, the product will be 0.
@@
 \begin{align}
-	&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} X_{c, b, d}\cdot Y_{c, w=\varphi(b)} \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
+	&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} X_{c, b, s}\cdot Y_{c, w=\varphi(b)} \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
 \end{align}
@@
 An approach used to convert such functions into linear objectives involves introducing a helper variable and additional constraints [ref \ref{???}].
-In our case, introducing a variable $Z_{c, b, d}$ for every product $X_{c, b, d} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that
+In our case, introducing a variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that
 \begin{align}
-	&Z_{c, b, d} \leq X_{c, b, d} \\
-	&Z_{c, b, d} \leq Y_{c, w=\varphi(b)} &&\forall d\in\mathcal{D}
+	&Z_{c, b, s} \leq X_{c, b, s} \\
+	&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S}
 \end{align}
@@
 \begin{align}
-	&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{d=1}^{D} Z_{c, b, d}
+	&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} Z_{c, b, s}
 \end{align}
-Indeed, whenever $X_{c, b, d} \cdot Y_{c, w} = 1$, $Z_{c, b, d}$ can attain a maximum value of 1, and whenever $X_{c, b, d} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, d}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, d}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective.
+Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a maximum value of 1, and whenever $X_{c, b, s} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, s}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, s}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective.
 \\
@@
 This is the approach we decided to use in our problem, to ensure all three objectives are considered when finding a solution.
-Under the assumption that each clinician in $\mathcal{C}$ provides all types of services in $\mathcal{D}$, the normalized objectives can be written as follows,
+Under the assumption that each clinician in $\mathcal{C}$ provides all types of services in $\mathcal{S}$, the normalized objectives can be written as follows,
 \begin{align}
-	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot D} \tag{Block Requests} \label{eqn:norm-obj-block-requests}\\
+	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot S} \tag{Block Requests} \label{eqn:norm-obj-block-requests}\\
 	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests} \label{eqn:norm-obj-weekend-requests} \\
-	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot D} \tag{Block-Weekend Adjacency} \label{eqn:norm-obj-block-weekend-adj}
+	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot S} \tag{Block-Weekend Adjacency} \label{eqn:norm-obj-block-weekend-adj}
 \end{align}

### 2019-08-22 paper/paper.results.tex
Minor fixes to wording in simulations

@@
 
-%We begin by evaluating the runtime in a single-division department, with an increasing number of clinicians, over a 1-year time-span.
-Figure \ref{fig:avg-runtime-clinicians} shows the average runtime in seconds of 10 runs, for different numbers of clinicians, $C = \{5, 10, 15, 25, 50, 100\}$.
-From the graph, we can see that the LP formulation performs extremely well on a typical hospital department with no more than 30 clinicians [ref \ref{??}], and can even handle busier departments with upwards of 100 clinicians.
+The effect of an increasing number of requests per clinician on the runtime of the LP solver is shown in figure \ref{fig:avg-runtime-requests}.
+We executed the program on a department with 10 clinicians offering two services, similar to the department at St.
+Michael's Hospital.
+Each clinician had $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{???\}$ non-overlapping requests.
+The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can accommodate a lot of flexibility in clinician requests.
 \\
@@
 %	\centering
-%	\includegraphics[scale=.7]{fig/avg_runtime_clinicians}
-%	\caption{Plot of average runtime for LP with an increasing number of clinicians}
-%	\label{fig:avg-runtime-clinicians}
+%	\includegraphics[scale=.7]{fig/avg_runtime_requests}
+%	\caption{Average runtime for LP with an increasing number of requests per clinician}
+%	\label{fig:avg-runtime-requests}
 %\end{figure}
 
-%Next we evaluate the effect of increasing the number of departments.
-Figure \ref{fig:avg-runtime-divisions} shows the average runtime for a department with 10 clinicians, and a range of divisions, $D = \{1, 2, 3\}$.
-We can see that increasing the number of divisions even by 2 greatly affects the runtime.
-The LP solver had significant trouble when we increased the number of divisions further, making it impractical to quickly generate schedules for departments with many divisions.
-\\
-
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[scale=.7]{fig/avg_runtime_divisions}
-%	\caption{Plot of average runtime for LP with an increasing number of divisions}
-%	\label{fig:avg-runtime-divisions}
-%\end{figure}
-
-The effect of an increasing number of requests per clinician on the average runtime of the LP solver is shown in figure \ref{fig:avg-runtime-requests}.
-We executed the program on a single-division department with 10 clinicians, and $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{1, 2, 3, 5, 10\}$ non-overlapping requests per clinician.
-The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can accommodate a lot of flexibility in clinician requests.
-\\
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=.7]{fig/avg_runtime_requests}
-	\caption{Average runtime for LP with an increasing number of requests per clinician}
-	\label{fig:avg-runtime-requests}
-\end{figure}
-
-Figure \ref{fig:avg-runtime-blocks} presents the average runtime over 10 runs for an increasing number of 2-week blocks $B = \{26, 52, 78, 104\}$ to simulate a department that generates schedules for a few years in advance.
+Figure \ref{fig:avg-runtime-blocks} presents the change in runtime when increasing the number of 2-week blocks $B = \{26, 52, 78, 104\}$ in a department with 10 clinicians offering two services.
 Although here was an increasing trend in the runtime, the solver was able to find an optimal schedule for all time horizons in under XX amount of time.
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=.7]{fig/avg_runtime_blocks}
-	\caption{Average runtime for LP with an increasing number of 2-week blocks}
-	\label{fig:avg-runtime-blocks}
-\end{figure}
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[scale=.7]{fig/avg_runtime_blocks}
+%	\caption{Average runtime for LP with an increasing number of 2-week blocks}
+%	\label{fig:avg-runtime-blocks}
+%\end{figure}
 

### 2019-08-23 paper/paper.results.tex
Update plots and descriptions in results

@@
 
-The effect of an increasing number of requests per clinician on the runtime of the LP solver is shown in figure \ref{fig:avg-runtime-requests}.
+The effect of an increasing number of requests per clinician on the runtime of the LP solver is shown in figure \ref{fig:runtime-requests}.
 We executed the program on a department with 10 clinicians offering two services, similar to the department at St.
 Michael's Hospital.
-Each clinician had $R = \abs{\mathcal{WR}} + \abs{\mathcal{BR}} = \{???\}$ non-overlapping requests.
+For simplicity, each clinician was configured with $\abs{\mathcal{BR}} = \{0, 5, 10, 15\}$ non-overlapping block requests.
 The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can accommodate a lot of flexibility in clinician requests.
+Moreover, we see that all four runs were completed in under 1 second.
 \\
 
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[scale=.7]{fig/avg_runtime_requests}
-%	\caption{Average runtime for LP with an increasing number of requests per clinician}
-%	\label{fig:avg-runtime-requests}
-%\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=.5]{fig/runtime_requests}
+	\caption{Runtime of LP solver with an increasing number of requests per clinician}
+	\label{fig:runtime-requests}
+\end{figure}
 
-Figure \ref{fig:avg-runtime-blocks} presents the change in runtime when increasing the number of 2-week blocks $B = \{26, 52, 78, 104\}$ in a department with 10 clinicians offering two services.
-Although here was an increasing trend in the runtime, the solver was able to find an optimal schedule for all time horizons in under XX amount of time.
+Figure \ref{fig:runtime-blocks} presents the change in runtime when increasing the number of 2-week blocks $B = \{26, 52, 78, 104\}$ in a department with 10 clinicians offering two services.
+There does not seem to be a strong effect on runtime when we lengthen the time-horizon of the schedule, and in fact the solver was able to find an optimal schedule for all time horizons within 3 seconds.
 \\ % 'very reasonable amount of time" is vague.
@@
 
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[scale=.7]{fig/avg_runtime_blocks}
-%	\caption{Average runtime for LP with an increasing number of 2-week blocks}
-%	\label{fig:avg-runtime-blocks}
-%\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=.5]{fig/runtime_blocks}
+	\caption{Runtime of LP solver with an increasing number of 2-week blocks}
+	\label{fig:runtime-blocks}
+\end{figure}
 
@@
 We executed the algorithm for $S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in total across all services.
-For 2 or more services, a roster of more than XX clinicians becomes impractical to schedule, as searching requires more than 24 hours.
-However, when relaxing the `No Consecutive Blocks' constraint, we see a great improvement in runtime for 2 and 3 services, and we are able to generate a schedule with upwards of 50 clinicians in under XX amount of time.
-%what does division mean?
+For 2 concurrent services, a roster of 30 or more clinicians becomes impractical to schedule, as searching for a solution requires over 24 hours.
+We see similar issues for a roster of 20 or more clinicians assigned to a division with 3 concurrent services.
+However, when relaxing the `No Consecutive Blocks' constraint, we see a great improvement in runtime for divisions with 2 and 3 services, and we are able to generate a schedule with upwards of 50 clinicians in under 1.5 seconds.
+\\  %what does division mean?
 do you mean services?

### 2019-08-23 paper/tbl/2018_schedule_comparison.tex
Remove 2017 table + use adjustbox to scale table
Update plots and descriptions in results

@@
-% Table generated by Excel2LaTeX from sheet '2018'
-\begin{table}[h]
-	\tiny
+% Table generated by Excel2LaTeX from sheet '2018_oldLP'
+\begin{table}[htbp]
   \centering
+	\begin{adjustbox}{scale=0.8}
     \begin{tabular}{c||ccc||ccc}
     \multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} & \multicolumn{3}{c||}{LP Solution} & \multicolumn{3}{c}{Historical Data} \\
-    	                                                  &                  HIV                  &                                 ID                                 &                              Weekend                               &                  HIV                  &                                 ID                                 &                              Weekend                               \\ \midrule\midrule
+          & HIV   & ID    & Weekend & HIV   & ID    & Weekend \\
+    \midrule
+    \midrule
     1     & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .459,  .443,  .443}H \\
-    	                        2                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
+    2     & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A \\
     3     & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .518,  .592,  .69}G \\
@@
     6     & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ 1,  .851,  .4}C \\
-    	                        7                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .608,  .761,  .902}B                \\
-    	                        8                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
-    	                        9                         & \cellcolor[rgb]{ .788,  .788,  .788}D &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .608,  .761,  .902}B &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
-    	                       10                         & \cellcolor[rgb]{ .788,  .788,  .788}D &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
-    	                       11                         & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
-    	                       12                         & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
-    	                       13                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
-    	                       14                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
-    	                       15                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
-    	                       16                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
-    	                       17                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
-    	                       18                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
-    	                       19                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .663,  .816,  .557}A &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
-    	                       20                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
-    	                       21                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
-    	                       22                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
-    	                       23                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
-    	                       24                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
+    7     & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .608,  .761,  .902}B \\
+    8     & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .518,  .592,  .69}G \\
+    9     & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D \\
+    10    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .459,  .443,  .443}H \\
+    11    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F \\
+    12    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A \\
+    13    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .459,  .443,  .443}H \\
+    14    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
+    15    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ 1,  .851,  .4}C \\
+    16    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .957,  .69,  .518}E \\
+    17    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .788,  .788,  .788}D \\
+    18    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .957,  .69,  .518}E \\
+    19    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F \\
+    20    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C \\
+    21    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .663,  .816,  .557}A \\
+    22    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ 1,  .851,  .4}C \\
+    23    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .788,  .788,  .788}D \\
+    24    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ 1,  .851,  .4}C \\
     25    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .518,  .592,  .69}G \\
     26    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D \\
-    	                       27                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
-    	                       28                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
-    	                       29                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
-    	                       30                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
-    	                       31                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
-    	                       32                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
+    27    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E \\
+    28    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
+    29    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .788,  .788,  .788}D \\
+    30    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .663,  .816,  .557}A \\
+    31    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .957,  .69,  .518}E \\
+    32    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F \\
     33    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
-    	                       34                         & \cellcolor[rgb]{ .608,  .761,  .902}B &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
-    	                       35                         & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
+    34    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ 1,  .851,  .4}C \\
+    35    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G \\
     36    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
-    	                       37                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
-    	                       38                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .788,  .788,  .788}D &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
-    	                       39                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
-    	                       40                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
-    	                       41                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
-    	                       42                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .788,  .788,  .788}D &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
-    	                       43                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
-    	                       44                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
-    	                       45                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .788,  .788,  .788}D &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
-    	                       46                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
-    	                       47                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
-    	                       48                         &   \cellcolor[rgb]{ 1,  .851,  .4}C    &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
-    	                       49                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
-    	                       50                         & \cellcolor[rgb]{ .663,  .816,  .557}A &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
-    	                       51                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E
+    37    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .663,  .816,  .557}A \\
+    38    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E \\
+    39    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D \\
+    40    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
+    41    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .518,  .592,  .69}G \\
+    42    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F \\
+    43    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ 1,  .851,  .4}C \\
+    44    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
+    45    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .957,  .69,  .518}E \\
+    46    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A \\
+    47    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D \\
+    48    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .518,  .592,  .69}G \\
+    49    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .557,  .663,  .859}F \\
+    50    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G \\
+    51    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .957,  .69,  .518}E \\
     \end{tabular}%
+	\end{adjustbox}
 	\caption{Comparison of schedules for 2018}

### 2019-08-23 paper/paper.abstract.tex
Add abstract

Scheduling of personnel in a hospital environment is vital to improving the service provided to patients and balancing the workload assigned to clinicians. Many approaches have been tried and successfully applied to generate efficient schedules in such settings. However, due to the computational complexity of the scheduling problem in general, most approaches resort to heuristics to find a non-optimal solution in a reasonable amount of time. Our paper describes an integer linear programming formulation to finding an optimal schedule in a clinical division of a hospital. It mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt the formulation to a variety of clinical divisions. \\

A case study for our approach is presented, using data from the Infectious Diseases division at St. Michael's Hospital in Toronto, Canada. We analyze and compare the results of our approach to manually-created schedules at the hospital, and note much improved adherence to departmental constraints and clinician preferences. We also perform runtime analysis of our linear program for various parameters and observe reassuring results, signifying the practicality of our approach in different real-world scenarios.

### 2019-08-24 paper/paper.abstract.tex
SM edits abstract
Well done!

@@
 However, due to the computational complexity of the scheduling problem in general, most approaches resort to heuristics to find a non-optimal solution in a reasonable amount of time.
-Our paper describes an integer linear programming formulation to finding an optimal schedule in a clinical division of a hospital.
+We designed an integer linear programming formulation to find an optimal schedule in a clinical division of a hospital.
 It mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt the formulation to a variety of clinical divisions.
@@
 
-A case study for our approach is presented, using data from the Infectious Diseases division at St.
+We then conducted a case study for our approach using data from the Infectious Diseases division at St.
 Michael's Hospital in Toronto, Canada.
-We analyze and compare the results of our approach to manually-created schedules at the hospital, and note much improved adherence to departmental constraints and clinician preferences.
-We also perform runtime analysis of our linear program for various parameters and observe reassuring results, signifying the practicality of our approach in different real-world scenarios.
+We analyzed and compared the results of our approach to manually-created schedules at the hospital, and found improved adherence to departmental constraints and clinician preferences.
+We then used simulated data to examine the the sensitivity of the runtime of our linear program for various parameters and observed reassuring results, signifying the practicality and generazability of our approach in different real-world scenarios.

### 2019-08-26 paper/paper.discussion.tex
Work on intro and discussion sections
- Rewrite 3 goals in intro
- Add outline of paper to intro
- Address goals in separate paragraphs in discussion

@@
-In this paper our goal was to develop a rather simple, yet flexible, integer linear programming formulation to generate schedules for clinical consultation (?) departments at hospitals.
-The difficulty in applying a linear programming approach to this task lies in the fact that ILP is an NP-hard problem, and as such the time to find an optimal solution grows exponentially as we increase the size of the problem.
+In this paper our goal was to develop a rather simple, yet flexible, integer linear programming formulation to generate schedules for clinical departments at hospitals.
+The difficulty in applying such an approach to this task lies in the fact that ILP is an NP-hard problem, and as such, the time to find an optimal solution grows exponentially as we increase the size of the problem.
 As a result, the vast majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time.
@@
 Our formulation includes hard constraints to ensure the schedule satisfies hospital and logistics requirements.
-It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function.
-We are able to use the tool to generate schedules for a wide variety of simulated departments, supporting upwards of 50 total clinicians.
-Moreover, our formulation is able to accommodate the requests of clinicians without negatively affecting the runtime.
-The scenarios when the ILP formulation starts to have trouble are ones with multiple services offered within a single clinical department.
-Once there are more than 10 clinicians in total, providing 2 or more services, the search space becomes very complex, making it difficult for the tool to find an optimal solution in a reasonable amount of time.
-For these cases, we recommend allowing consecutive blocks, to simplify the problem and find a solution much faster.
+It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function [...] \\
+
+We developed a tool that implements the ILP formulation and generates an optimal schedule based on provided input data.
+It provides a simple and easy-to-use user interface that can be used by any hospital staff to generate schedules well in advance.
+We compared the optimal schedule generated with the help of the tool to the manually-created schedules at St.
+Michael's Hospital.
+We found that the ILP formulation was always able to find an optimal schedule satisfying all required hard constraints, unlike the manual schedule, that often did not take all constraints into account.
+Moreover, due to the multi-goal objective function in the ILP, the tool was able to maximize and fulfill the majority of clinician preferences and requests, more so than the manually-created schedule.
+These observations reinforce the idea that it is vital to employ automated tools when generating schedules in hospital departments, to reduce human error, balance the work-load of clinicians and improve the service provided to patients.
 \\
 
-Our tool is also employed at the Infectious Diseases Department at St.
-Michael's Hospital.
-It provides a simple and easy-to-use user interface that can be used by any hospital staff to generate schedules well in advance.
-We compare the manually created schedule at St.
-Michael's Hospital to the optimal schedule generated with the help of our tool.
-We see adherence to all required hard constraints in the generated schedule, while some of them were overlooked in the manual schedule.
-Moreover, we see a much better fulfillment of clinician requests than in the manual schedule.
-This reinforces the idea that it is vital to employ automated tools when generating schedules in hospital departments, to reduce human error, balance the work-load of clinicians and improve the service provided to patients.
+We analyzed the performance of our formulation on simulated data meant to resemble real-life clinician departments at different hospitals.
+Our results show that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences.
+Further, we saw that the algorithm can accommodate increasing the time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance.
+The most notable impact on the runtime was noticed when trying to increase the number of services offered in a single department.
+While such cases are unlikely to be encountered in the real-world, as most departments tend to focus on a single service [ref \ref{???}], the runtime issues can be mitigated by relaxing certain constraints.
 \\

### 2019-08-26 paper/paper.intro.tex
Work on intro and discussion sections
- Rewrite 3 goals in intro
- Add outline of paper to intro
- Address goals in separate paragraphs in discussion

@@
 
-%When generating these schedules, there are many variables and constraints that need to be taken into account, such as preventing multiple consecutive assignments for a given clinician or ensuring that assignments are spread out evenly throughout a certain period of time.
-It is very common for departments to opt for a manual method of generating these schedules [??].
-Unfortunately, such methods are unreliable and can lead to unfavourable assignments, since a manual method may miss certain constraints or forget to account for a certain variable during the process.
-\\
-
 Automated methods to optimize schedules have been studied and applied in many industries, including transportation [??], manufacturing [??], [...].
@@
 A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??].
+\\
 
@@
 
-[...] Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation.
+Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation.
 While these department-specific approaches allow end-users to find precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians in that department, they are difficult to readily adapt to other departments in the same hospital or other hospitals.
@@
 Michael's Hospital in Toronto, Canada.
-Our goal is to (1) present a simple formulation for the problem developed and tested at the hospital after switching from a manual approach to scheduling; and (2) analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach; and (3) describe the adaptability of the formulation as a basic framework for solving similar problems in other departments.
+Our goal is to (1) present a simple integer linear programming formulations for the scheduling problem as developed for the hospital, and describe the adaptability of the formulation to solving similar problems in other departments; (2) compare the performance of the ILP scheduler to the results of the manual approach; and (3) analyze the robustness of the algorithm in difficult instances of the problem.
+\\
+% present a simple formulation for the problem developed and tested at the hospital after switching from a manual approach to scheduling; and (2) analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach; and (3) describe the adaptability of the formulation as a basic framework for solving similar problems in other departments.
 \\ %make #3 an objective
-We begin by describing the problem, then...% list the main contents/elements of the paper.
 
-[...]
-
-%Automated methods for scheduling in hospital settings have been studied extensively, most notably in the context of the Nurse Scheduling Problem (NSP) in Operations Research.
-In an NSP, the goal is to assign nurses to shifts, while satisfying various constraints pertaining to both hospital regulations and nurse preferences.
-The constraints can be classified as soft or hard constraints, based on the importance of satisfying them.
-Typically, it is necessary for a solution to satisfy all of the hard constraints, while the soft constraints are considered optional.
-The NSP, in its general case, is known to be NP-complete, meaning that although it is possible to efficiently check whether a given schedule satisfies all the constraints using an algorithm, it is not yet known whether an efficient algorithm for finding a satisfying schedule exists, and it is in fact equivalent to the hardest computational problems in Computer Science, Operations Research, and \textit{etc...} [??].
-In order to tackle NP-complete problems such as NSP in real-life, algorithms either try to use various heuristics to approximate a potential solution rather than finding an exact solution or they focus on relatively small problem sizes, ensuring that the runtime of the algorithm is still practical.
-\textit{Some examples of heuristic approaches...} .
-\textit{Some examples of exact approaches (LP, etc.)} .
-A comprehensive review of various approaches is presented by Burke et al.
-[??].
-\\
-
-%In this paper we present the application of two methods, Network Flows and Integer Linear Programming (ILP), to a variation of the Nurse Scheduling Problem in order to generate a yearly schedule for clinicians working simultaneously at two different divisions of St.
-Michael's Hospital in Toronto, ON.
-The paper is organized as follows.
-Section [??] contains a detailed description of the scenario and a mathematical formalization of the problem.
-Section [??] describes first the Network Flow approach followed by the ILP approach to tackle the problem.
-Section [??] presents the results of both approaches when applied to the clinician data from St.
-Michael's Hospital.
-Section [??] presents results from simulations with a variety of problem sizes.
-We conclude the paper with a discussion of the real-world and simulation results in Section [??].
+We begin by describing the problem in Section \ref{sec:problem}, and presenting our ILP formulation in Section \ref{sec:methods}.
+Next, we compare the results of our formulation to manually-created schedules, and evaluate the performance of the algorithm on simulated data in Section \ref{sec:results}.
+Finally, we discuss and interpret the results in Section \ref{sec:discussion}.
+% list the main contents/elements of the paper.

### 2019-08-27 paper/paper.intro.tex
Work on intro + references
Add majority of references for automated methods/scheduling papers in intro

@@
 For example, it is important that a hospital department allocates its resources, such as the availability of a finite number of clinicians, optimally, to ensure the best possible service for its patients.
-Carefully allocated on-call schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes [ref].
+Carefully allocated schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes [ref].
 It is common practice for on-call schedules to be created manually.
@@
 First, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method may not provide an optimal schedule.
-Second, such methods are likely to overlook certain constraints that must be maintained to have an operational department, such as XXXX.
+Second, such methods are likely to overlook certain constraints that must be maintained to have an operational department, such as preventing many consecutive work blocks from being assigned or ensuring clinicians are allocated a specific amount of work blocks throughout the year.
 Third, manual scheduling is often time-consuming for the person developing the schedule.
@@
 
-Automated methods to optimize schedules have been studied and applied in many industries, including transportation [??], manufacturing [??], [...].
+Automated methods to optimize schedules have been studied and applied in many industries, including transportation \cite{aickelin_improved_2006, goel_truck_2012, gunther_combined_2010}, manufacturing \cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008, alfares_simulation_2007}, retail \cite{chapados_retail_2011, nissen_automatic_2010} and military \cite{horn_scheduling_2007, laguna_modeling_2005}.
 Of special interest to a clinician on-call scheduling problem are the approaches to schedule nurses, who often work in shifts.
 In the nurse scheduling problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences.
-A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??].
+A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming \cite{azaiez_0-1_2005, trilling_nurse_2006, widyastiti_nurses_2016}, network flows \cite{el_adoly_new_2018}, genetic algorithms \cite{aickelin_exploiting_2000, jan_evolutionary_2000, kawanaka_genetic_2001}, simulated annealing \cite{jaszkiewicz_metaheuristic_1997}, and artificial intelligence \cite{abdennadher_nurse_nodate, li_hybrid_2003}.
+A comprehensive literature review of these and other methods applied to nurse rostering is presented in \cite{burke_state_2004}.
 \\
@@
 what makes their generalizablility/adaptability limited?
-Moreover, the large number of variables and constraints also leads to computational complexity issues [ref], especially when using exact methods for finding the solution.
+Moreover, the large number of variables and constraints also leads to computational complexity issues \cite{goos_complexity_1996}, especially when using exact methods for finding the solution.
 In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease (ID) consults; and HIV consults service) at St.

### 2019-08-27 paper/paper.discussion.tex
Work on discussion

@@
 Our formulation includes hard constraints to ensure the schedule satisfies hospital and logistics requirements.
-It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function [...] \\
+It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function.
+While one of our goals was to keep a minimal set of constraints to ensure that the ILP formulation is usable in practice, it leaves room to adapt it to various clinical departments at different hospitals.
+The flexibility of our ILP allows changing the number of services provided in a division, the length of a work block, clinicians' preference for work block to weekend adjacency as well as clinicians' requests for time off.
+\\
 

### 2019-08-29 paper/paper.tex
Rewrite portions of introduction
Add bits of problem description
Add summary of constraints + more details
Move mathematical formulation to methods section, for now
Move \usepackage lines to separate file
Add all hard constraints to methods
Add very rough idea of results section w/table
Add section labels
Add abstract
Work on intro + references
Add majority of references for automated methods/scheduling papers in intro
Add preliminary title + author/affil list

@@
 \documentclass[]{article}
+\input{paper.includes.tex}
+\input{paper.definitions.tex}
+
+\title{A flexible integer linear programming formulation for scheduling clinician on-call service in hospitals}
+
+\author[1, 5]{David Landsman}
+\author[5]{Huiting Ma}
+\author[5]{Jesse Knight}
+\author[6]{Kevin Gough}
+\author[2, 3, 4, 5]{Sharmistha Mishra}
+
+\renewcommand\Affilfont{\itshape\small}
+\affil[1]{Department of Computer Science, University of Toronto}
+\affil[2]{Department of Medicine, Division of Infectious Disease, University of Toronto}
+\affil[3]{Institute of Health Policy, Management and Evaluation, Dalla Lana School of Public Health, University of Toronto}
+\affil[4]{Institute of Medical Sciences, University of Toronto}
+\affil[5]{MAP Centre for Urban Health Solutions, Unity Health Toronto}
+\affil[6]{Department of Medicine, St. Michael's Hospital}
 
 \begin{document}
+	\maketitle
+
 	\begin{abstract}
@@
 
-	\section{Introduction} 
+	\section{Introduction} \label{sec:introduction}
 		\input{paper.intro.tex}
-	\section{Problem}
+	\section{Problem} \label{sec:problem}
 		\input{paper.problem.tex}
-	\section{Methods}
+	\section{Methods} \label{sec:methods}
 		\input{paper.methods.tex}
-	\section{Results}
+	\section{Results} \label{sec:results}
 		\input{paper.results.tex}
-	\section{Simulations}
-		\input{paper.simulations.tex}
-	\section{Discussion}
+	\section{Discussion} \label{sec:discussion}
 		\input{paper.discussion.tex}
+	\bibliographystyle{unsrt}
+	\bibliography{references}
 \end{document}

### 2019-08-29 paper/paper.abstract.tex
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

@@
 We designed an integer linear programming formulation to find an optimal schedule in a clinical division of a hospital.
-It mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt the formulation to a variety of clinical divisions.
+Our formulation mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt it to a variety of clinical divisions.
 \\
@@
 We analyzed and compared the results of our approach to manually-created schedules at the hospital, and found improved adherence to departmental constraints and clinician preferences.
-We then used simulated data to examine the the sensitivity of the runtime of our linear program for various parameters and observed reassuring results, signifying the practicality and generazability of our approach in different real-world scenarios.
+We used simulated data to examine the sensitivity of the runtime of our linear program for various parameters and observed reassuring results, signifying the practicality and generazability of our approach in different real-world scenarios.

### 2019-08-29 paper/tbl/runtime_services_clinicians.tex
Rename divisions -> services throughout
Also additional changes to results section following comments
Update runtimes in services table
Minor fixes to wording in simulations
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

--- 2019-08-19 paper/tbl/runtime_divisions.tex
@@
-\begin{table}
+\begin{table}[htbp]
 	\centering
-	\begin{tabular}{|c|c|c|c|c|}
+	\begin{tabular}{|c|c||c|c||c|c|}
 		\hline
-		&       \multicolumn{4}{c|}{Number of Divisions}        \\ \hline
-		Number of Clinicians &  1   &   2    & 2 (Consecutive Blocks Allowed) &  3   \\ \hline
-		10          & 0.22 &  0.51  &              0.26              & 1.49 \\ \hline
-		20          & 0.42 & 139.42 &              0.50              &     \\ \hline
-		30          & 0.63 &   -    &              0.90              &     \\ \hline
-		50          & 1.06 &   -    &              1.30              &     \\ \hline
+		                                      &  \multicolumn{5}{c|}{Number of Services}  \\ \hline
+		\makecell[l]{Number of \\ Clinicians} &  1   &    2    & 2 (without NCB) &  3   & 3 (without NCB) \\ \hline
+		                 10                   & 0.16 &  0.74   &  0.16   & 1.40 &  0.23   \\ \hline
+		                 20                   & 0.25 & 7468.86 &  0.32   &  -   &  0.43   \\ \hline
+		                 30                   & 0.42 &    -    &  0.49   &  -   &  0.66   \\ \hline
+		                 50                   & 0.62 &    -    &  0.82   &  -   &  1.14   \\ \hline
 	\end{tabular}
-	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians. ``-'' indicates that a solution was not found after 24 hours.}
-	\label{tbl:runtime-divisions-comparison}%
+	\caption{Comparison of program runtime (in seconds) for different number of services and total clinicians in the division.
+`-' indicates that a solution was not found within 24 hours.
+`without NCB' indicates that the No Consecutive Blocks (NCB) constraint was relaxed.}
+	\label{tbl:runtime-services-clinicians-comparison}%
 \end{table}

### 2019-08-29 paper/paper.intro.tex
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

@@
-On-call schedules for a fixed number of health-care providers are central to the efficient running of hospitals.
 Hospital departments provide services where patient needs, and thus the system's demands, often exceed the available supply.
-For example, it is important that a hospital department allocates its resources, such as the availability of a finite number of clinicians, optimally, to ensure the best possible service for its patients.
+In particular, on-call schedules for a fixed number of health-care providers are central to the efficient running of hospitals.
 Carefully allocated schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes [ref].
-It is common practice for on-call schedules to be created manually.
-Yet manually-created schedules are prone to errors and potential for biases [ref].
+It is common practice for on-call schedules in hospitals to be created manually, yet manually-created schedules are prone to errors and potential for various biases [ref].
 First, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method may not provide an optimal schedule.
@@
 \\
+%For example, it is important that a hospital department allocates its resources, such as the availability of a finite number of clinicians, optimally, to ensure the best possible service for its patients.
 
@@
 
-Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation.
+Many of the approaches to nurse scheduling were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation.
 While these department-specific approaches allow end-users to find precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians in that department, they are difficult to readily adapt to other departments in the same hospital or other hospitals.

### 2019-08-29 paper/paper.methods.tex
Work on intro + references
Add majority of references for automated methods/scheduling papers in intro
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

@@
-In this section, we present an application of 0-1 Linear Programming to solve the clinician scheduling problem presented in Section \ref{sec:problem}.
+In this section, we present an application of integer linear programming to solve the clinician scheduling problem presented in Section \ref{sec:problem}.
 First, we describe the sets, indices and variables present in the formulation of the program.
@@
 \subsection{Variables} \label{sec:meth-variables}
-Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as $X_{c, b, s}$.
+Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as a binary variable $X_{c, b, s}$.
 A value of 1 indicates that the given clinician $c$ is assigned to provide service $s$ during block $b$.
-Similarly, we define weekend assignments as $Y_{c, w}$, although these are not indexed by service, as clinicians are expected to provide all services during the weekends.
-To optimize the soft constraint Block-Weekend Adjacency, we maximize the product $X_{c, b, s} \cdot Y_{c, w}$ for adjacent blocks and weekends.
+Weekend assignments are similarly defined using a binary variable $Y_{c, w}$, but without a service index, as clinicians are expected to provide all services during the weekends.
+We then introduce a set of constants $m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year.
+Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.
+%To optimize the soft constraint Block-Weekend Adjacency, we maximize the product $X_{c, b, s} \cdot Y_{c, w}$ for adjacent blocks and weekends.
 To formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, s}$, with additional constraints on its range.
 Further details regarding this variable are described in Section \ref{sec:meth-objectives}.
-We then introduce a set of constants $m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year.
-Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.
 
@@
 		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                         \\
-		$Z_{c, b, s} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend adjacency             \\
+%		$Z_{c, b, s} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend adjacency             \\
 		$m_{c, s}$                 & minimum number of blocks clinician $c$ should cover on service $s$ \\
@@
 \subsection{Constraints} \label{sec:meth-constraints}
-We now present the hard constraints in Table \ref{tbl:constraint-summary} using the variables defined above.
+We now formalize the hard constraints in Table \ref{tbl:constraint-summary} using the variables defined above.
 
 \begin{align}
-	&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S} \tag{Block Coverage} \label{eqn:constr-block-cov} \\
-	&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{Weekend Coverage} \label{eqn:constr-weekend-cov} \\
-	&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\ c\in\mathcal{C}, s\in\mathcal{S} \tag{Min/Max} \label{eqn:constr-min-max} \\
-	&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, s\in\mathcal{S} \tag{No Consecutive Blocks} \label{eqn:constr-no-consec-blocks} \\
-	&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{No Consecutive Weekends} \label{eqn:constr-no-consec-weekends} \\
-	&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}} &&\forall c\in\mathcal{C} \tag{Equal Weekends} \label{eqn:constr-equal-weekends} \\
-	&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq \ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C} \tag{Equal Holidays} \label{eqn:constr-equal-holidays}
+	&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S} \tag{BC} \label{eqn:constr-block-cov} \\
+	&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{WC} \label{eqn:constr-weekend-cov} \\
+	&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\ c\in\mathcal{C}, s\in\mathcal{S} \tag{MM} \label{eqn:constr-min-max} \\
+	&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, s\in\mathcal{S} \tag{NCB} \label{eqn:constr-no-consec-blocks} \\
+	&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{NCW} \label{eqn:constr-no-consec-weekends} \\
+	&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}} &&\forall c\in\mathcal{C} \tag{EW} \label{eqn:constr-equal-weekends} \\
+	&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq \ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C} \tag{EH} \label{eqn:constr-equal-holidays}
 \end{align}
@@
 \subsection{Objectives} \label{sec:meth-objectives}
-As described in Section \ref{sec:problem}, the soft constraints of the clinician scheduling problem include: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks.
+As described in Section \ref{sec:problem}, the soft constraints of the clinician scheduling problem include: satisfying clinician block off requests (BR), satisfying clinician weekend off requests (WR), and assigning weekends closer to blocks (BWA).
 We convert these soft constraints into linear objective functions of the binary variables defined in Section \ref{sec:meth-variables}.
-Objectives (1) and (2) can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to.
+Objectives (\ref{eqn:obj-block-requests}) and (\ref{eqn:obj-weekend-requests}) can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to.
 \begin{align}
-	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{Block Requests} \label{eqn:obj-block-requests}\\
-	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{Weekend Requests} \label{eqn:obj-weekend-requests}
+	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{BR} \label{eqn:obj-block-requests}\\
+	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{WR} \label{eqn:obj-weekend-requests}
 \end{align}
-where $\ind(\cdot)$ is the indicator function that has value 1 when the condition is met and 0 otherwise.
+where $\ind(P)$ is the indicator function that has value 1 when predicate $P$ holds and 0 otherwise.
 In the above two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments.
+\\
+
 The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, s}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$.
+This leads to the maximization objective %avoid editorial adjectives
+\begin{align}
+&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} X_{c, b, s}\cdot Y_{c, w=\varphi(b)} \tag{BWA} \label{eqn:obj-block-weekend-adj}
+\end{align}
+where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by some appropriate definition of adjacency.
 For instance, clinicians might want to be assigned during a weekend that falls within an assigned block.
-So if clinician $c$ is assigned to work during block 3, corresponding to weeks 5 and 6 assuming 2-week blocks, they might also want to be assigned to work during weekend 5.
+In this case, we will have $\varphi(b) = 2b-1$.
+\\
+
+%So if clinician $c$ is assigned to work during block 3, corresponding to weeks 5 and 6 assuming 2-week blocks, they might also want to be assigned to work during weekend 5.
 In that case, we would like $X_{c, b=3, s} \cdot Y_{c, w=5}$ to be 1, since that indicates both variables are assigned.
 If at least one of the two variables is not assigned, the product will be 0.
-This leads to the maximization objective  %avoid editorial adjectives
-\begin{align}
-	&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} X_{c, b, s}\cdot Y_{c, w=\varphi(b)} \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
-\end{align}
-where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by some appropriate definition of adjacency. For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\
 % I really liked the example explanation!
+%For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\
 
 However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework.
-An approach used to convert such functions into linear objectives involves introducing a helper variable and additional constraints [ref \ref{???}].
+An approach used to convert such functions into linear objectives involves introducing a helper variable and additional constraints \cite{hammer_boolean_1968}.
 In our case, introducing a variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that
 \begin{align}
-	&Z_{c, b, s} \leq X_{c, b, s} \\
-	&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S}
+	&Z_{c, b, s} \leq X_{c, b, s} \label{eqn:helper-x-constraint}\\
+	&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S} \label{eqn:helper-y-constraint}
 \end{align}
@@
 \end{align}
-Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a maximum value of 1, and whenever $X_{c, b, s} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, s}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, s}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective.
+Indeed from Eqns.
+(\ref{eqn:helper-x-constraint}) and (\ref{eqn:helper-y-constraint}),  whenever $X_{c, b, s} \cdot Y_{c, w} = 1$ (respectively, 0), $Z_{c, b, s}$ can attain a maximum value of 1 (respectively, 0), giving us the correct adjacency maximization objective.
+\\
+%Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a maximum value of 1, and whenever $X_{c, b, s} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, s}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, s}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective.
 \\
 
-Thus, our clinician scheduling problem is a multiple objective optimization problem.
+In order to optimize all objectives simultaneously, we optimize a weighted sum of the normalized objective functions,
+\begin{equation}
+\max_{X, Y, Z} \alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
+\end{equation}
+where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha, \beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal \cite{stanimirovic_linear_2011}. \\
+
+%Thus, our clinician scheduling problem is a multiple objective optimization problem.
 The most common approach to solving multiple objective optimization problems is by optimizing a weighted sum of the normalized objective functions, as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}].
@@
 Under the assumption that each clinician in $\mathcal{C}$ provides all types of services in $\mathcal{S}$, the normalized objectives can be written as follows,
-\begin{align}
-	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot S} \tag{Block Requests} \label{eqn:norm-obj-block-requests}\\
-	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests} \label{eqn:norm-obj-weekend-requests} \\
-	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot S} \tag{Block-Weekend Adjacency} \label{eqn:norm-obj-block-weekend-adj}
-\end{align}
-where we divide each of the original objective functions by the sum of the absolute values of its coefficients [ref \ref{???}]. The final weighted objective is given by
-\begin{equation}
-	\alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
-\end{equation}
-with $0 \leq \alpha, \beta \leq 1$. \\
-
-[...]
+%\begin{align}
+%	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot S} \tag{Block Requests} \label{eqn:norm-obj-block-requests}\\
+%	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests} \label{eqn:norm-obj-weekend-requests} \\
+%	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot S} \tag{Block-Weekend Adjacency} \label{eqn:norm-obj-block-weekend-adj}
+%\end{align}
+%where we divide each of the original objective functions by the sum of the absolute values of its coefficients [ref \ref{???}]. The final weighted objective is given by
+%\begin{equation}
+%	\alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
+%\end{equation}
+%with $0 \leq \alpha, \beta \leq 1$. \\

### 2019-08-29 paper/paper.problem.tex
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

@@
 These limits may change from year to year as the number of clinicians in the department changes.
-Moreover, the schedule should not assign a clinician to work for two blocks or two weekends in a row.
+Moreover, the schedule should not assign a clinician to work for two consecutive blocks or two consecutive weekends in a row.
 The schedule should also distribute both regular and holiday weekends equally among all clinicians.
@@
 		\textbf{Constraint Name} & \textbf{Description}                                                                                                           & \textbf{Type} \\ \hline
-		Block Coverage           & \makecell[l]{each service needs to have exactly \\ one clinician that covers any given block}                                 & Hard          \\ \hline
-		Weekend Coverage         & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              & Hard          \\ \hline
-		Min/Max                  & \makecell[l]{for a given service, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} & Hard          \\ \hline
-		No Consecutive Blocks    & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all services}                                    & Hard          \\ \hline
-		No Consecutive Weekends  & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        & Hard          \\ \hline
-		Equal Weekends           & \makecell[l]{weekends should be equally distributed \\ between clinicians}                                                     & Hard          \\ \hline
-		Equal Holidays           & \makecell[l]{long weekends should be equally distributed \\ between clinicians}                                                & Hard          \\ \hline
-		Block Requests           & \makecell[l]{each clinician can request to be off service \\ during certain blocks throughout the year}                        & Soft          \\ \hline
-		Weekend Requests         & \makecell[l]{each clinician can request to be off service \\ during certain weekends throughout the year}                      & Soft          \\ \hline
-		Block-Weekend Adjacency  & \makecell[l]{the block and weekend assignments of a given \\ clinician should be adjacent}                                     & Soft
+		Block Coverage (BC)           & \makecell[l]{each service needs to have exactly \\ one clinician that covers any given block}                                 & Hard          \\ \hline
+		Weekend Coverage (WC)        & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              & Hard          \\ \hline
+		Min/Max (MM)                  & \makecell[l]{for a given service, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} & Hard          \\ \hline
+		No Consecutive Blocks (NCB)   & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all services}                                    & Hard          \\ \hline
+		No Consecutive Weekends (NCW) & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        & Hard          \\ \hline
+		Equal Weekends (EW)          & \makecell[l]{weekends should be equally distributed \\ between clinicians}                                                     & Hard          \\ \hline
+		Equal Holidays (EH)          & \makecell[l]{long weekends should be equally distributed \\ between clinicians}                                                & Hard          \\ \hline
+		Block Requests (BR)         & \makecell[l]{each clinician can request to be off service \\ during certain blocks throughout the year}                        & Soft          \\ \hline
+		Weekend Requests (WR)        & \makecell[l]{each clinician can request to be off service \\ during certain weekends throughout the year}                      & Soft          \\ \hline
+		Block-Weekend Adjacency (BWA)  & \makecell[l]{the block and weekend assignments of a given \\ clinician should be adjacent}                                     & Soft
 	\end{tabular}
@@
 In most scheduling problems, the constraints can be divided into hard and soft constraints.
-Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions [ref].
-Typically, soft constraints are encoded as objective functions rather than actual constraints.
+Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions.
+Typically, soft constraints are encoded as objective functions rather than actual constraints [ref].
 These objective functions are maximized or minimized when finding a solution.
@@
 Though it is important to take clinician requests into account when constructing the schedule, it is crucial that the work-load of the schedule is balanced among all clinicians, and the needs of the patients are fulfilled.
-\\
-
 It is important to note that - irrespective of how the schedule is generated - clinicians may exchange certain weeks or days throughout the year after the schedule is implemented.
-The approach to solving the clinician scheduling problem does not account for these future exchanges, and only focuses on the full year time horizon.
+The approach to solving the clinician scheduling problem does not account for these future exchanges, and only focuses on the full time horizon.

### 2019-08-29 paper/paper.discussion.tex
Work on discussion

@@
 While one of our goals was to keep a minimal set of constraints to ensure that the ILP formulation is usable in practice, it leaves room to adapt it to various clinical departments at different hospitals.
-The flexibility of our ILP allows changing the number of services provided in a division, the length of a work block, clinicians' preference for work block to weekend adjacency as well as clinicians' requests for time off.
+The flexibility of our ILP allows changing the number of services provided in a division, the length of a work block, clinicians' preference for block to weekend adjacency as well as clinicians' requests for time off.
 \\
@@
 The most notable impact on the runtime was noticed when trying to increase the number of services offered in a single department.
-While such cases are unlikely to be encountered in the real-world, as most departments tend to focus on a single service [ref \ref{???}], the runtime issues can be mitigated by relaxing certain constraints.
+While such cases are unlikely to be encountered in the real-world, as most departments tend to focus on a single service [ref \ref{???}], the runtime issues can be mitigated by relaxing certain constraints, which can then be readjusted manually from the generated schedule.
+Overall, it is very encouraging to observe that an ILP formulation can be applied to schedule clinicians in practice, by providing optimal schedules rather than having to utilize heuristics.
 \\
@@
 On top of that, we can introduce new objectives that incorporate additional clinician preferences and allow for better distribution of work-load.
-However, since our ultimate goal is to maintain a ILP formulation, care must be taken not to overly complicate the model and risk running into additional time complexity issues.
+However, since our ultimate goal is to maintain a ILP formulation, care must be taken not to overly complicate the model and avoid running into additional time complexity issues.

### 2019-08-29 paper/paper.discussion.tex
SM edits/comments discussion

@@
-In this paper our goal was to develop a rather simple, yet flexible, integer linear programming formulation to generate schedules for clinical departments at hospitals.
-The difficulty in applying such an approach to this task lies in the fact that ILP is an NP-hard problem, and as such, the time to find an optimal solution grows exponentially as we increase the size of the problem.
+In this paper, we present a simple, yet flexible, integer linear programming formulation to generate schedules for clinical departments at hospitals. 
+The challenge in applying such an approach to this task lies in the fact that ILP is an NP-hard problem.
+%specify what 'such an approach' and 'this task' are...
+the first part of the sentence was a bit hard to follow.
+Define what you mean by NP-hard problem and cite --> i.e.
+NP-hard = the following sentence about how time to optimal solution grows exponentially?
+if yes, then instead of 'As such' - say 'That is,...or "An NP-hard problem means that..."
+As such, the time to find an optimal solution grows exponentially as we increase the size of the problem.
 As a result, the vast majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time.
-\\
+\\  %cite/reference statement (i.e.
+the 'vast majority'...) I can't remember if the introduction details the heuristics --> but if yes, then great.
+if no, then include that review of the prior literature here in discussion.
+I think as you said - the journals we are submitting to seem to have more of those details in the introduction.
+
+%double-check all abbreviations have been defined. I would prefer limiting abbreviations to no more than 3 at most if possible. We are ok for word count in this paper :).
 
@@
 It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function.
-While one of our goals was to keep a minimal set of constraints to ensure that the ILP formulation is usable in practice, it leaves room to adapt it to various clinical departments at different hospitals.
+Although we restricted our use of the formulation to a minimal set of constraints for the particular needs of the case study (St.
+Michael's Hospital Division of Infectious Diseases) - our formulation can be adapted to various clinical departments at different hospitals.
 The flexibility of our ILP allows changing the number of services provided in a division, the length of a work block, clinicians' preference for block to weekend adjacency as well as clinicians' requests for time off.
@@
 
-We developed a tool that implements the ILP formulation and generates an optimal schedule based on provided input data.
-It provides a simple and easy-to-use user interface that can be used by any hospital staff to generate schedules well in advance.
-We compared the optimal schedule generated with the help of the tool to the manually-created schedules at St.
-Michael's Hospital.
-We found that the ILP formulation was always able to find an optimal schedule satisfying all required hard constraints, unlike the manual schedule, that often did not take all constraints into account.
+When comparing the optimal schedule generated by tool to the manually-created schedules at St.
+Michael's Hospital, we found that the ILP formulation was always able to find an optimal schedule satisfying all required hard constraints, unlike the manual schedule, that often did not take all constraints into account.
 Moreover, due to the multi-goal objective function in the ILP, the tool was able to maximize and fulfill the majority of clinician preferences and requests, more so than the manually-created schedule.
-These observations reinforce the idea that it is vital to employ automated tools when generating schedules in hospital departments, to reduce human error, balance the work-load of clinicians and improve the service provided to patients.
-\\
+These observations reinforce the benefits of automated tools when generating schedules in hospital departments to balance the work-load of clinicians and improve the service provided to patients.
+\\  % how the finding compares to the wider literature when automated compared to manual?
+What does it mean re: "so what" - human error in generating tools, etc.
+or challenges in heuristics used manually by people?
+include citations this last sentence in particular - what dos it reinforce exactly - what are the benefits?
 
-We analyzed the performance of our formulation on simulated data meant to resemble real-life clinician departments at different hospitals.
-Our results show that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences.
+We also found that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences.
 Further, we saw that the algorithm can accommodate increasing the time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance.
-The most notable impact on the runtime was noticed when trying to increase the number of services offered in a single department.
-While such cases are unlikely to be encountered in the real-world, as most departments tend to focus on a single service [ref \ref{???}], the runtime issues can be mitigated by relaxing certain constraints, which can then be readjusted manually from the generated schedule.
-Overall, it is very encouraging to observe that an ILP formulation can be applied to schedule clinicians in practice, by providing optimal schedules rather than having to utilize heuristics.
-\\
-
-In future work, we can look first and foremost into improving the runtime of the tool by modifying constraints and objectives, so that it can be used in larger departments offering multiple services for their patients.
-On top of that, we can introduce new objectives that incorporate additional clinician preferences and allow for better distribution of work-load.
-However, since our ultimate goal is to maintain a ILP formulation, care must be taken not to overly complicate the model and avoid running into additional time complexity issues.
+One key limitation we identified was the sensitivity of the runtime to larger numbers of services offered in a single department.
+Such cases are unlikely to be encountered in the real-world because most clinical departments tend to provide a single service or at most two services by the same roster of clinicians [ref \ref{???}].
+One solution to mitigate the runtime issues generated by a larger number of services would be relaxing certain constraints, followed by manual readjustment from the generated schedule.
+%what do we mean by certain constraints?
+specify the other constraints?
+Thus, next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services i parallel for patients and other clients.
+Overall, our sensitivity analyses using simulated data provide reassurance that the ILP formulation can be applied to schedule clinicians across real-world variability between clinical departments.
+\\ %SM - will look for reference and we can ask Kevin Gough too.

### 2019-08-30 paper/paper.intro.tex
Add acks + funding sections
- Fix affiliation order, minor intro corrections
- TODO: add CPU specs to results section

@@
 In particular, on-call schedules for a fixed number of health-care providers are central to the efficient running of hospitals.
-Carefully allocated schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes [ref].
-It is common practice for on-call schedules in hospitals to be created manually, yet manually-created schedules are prone to errors and potential for various biases [ref].
-First, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method may not provide an optimal schedule.
+Carefully allocated schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes.
+It is common practice for on-call schedules in hospitals to be created manually, yet manually-created schedules are prone to errors and potential for various biases.
+First, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method may not provide a schedule that satisfies all clinicians.
 Second, such methods are likely to overlook certain constraints that must be maintained to have an operational department, such as preventing many consecutive work blocks from being assigned or ensuring clinicians are allocated a specific amount of work blocks throughout the year.

### 2019-08-30 paper/paper.tex
SM edits affiliations
Add acks + funding sections
- Fix affiliation order, minor intro corrections
- TODO: add CPU specs to results section

@@
 
-\author[1, 5]{David Landsman}
-\author[5]{Huiting Ma}
-\author[5]{Jesse Knight}
-\author[6]{Kevin Gough}
-\author[2, 3, 4, 5]{Sharmistha Mishra}
+\author[a, b]{David Landsman}
+\author[a]{Huiting Ma}
+\author[a]{Jesse Knight}
+\author[c]{Kevin Gough}
+\author[a, c, d, e]{Sharmistha Mishra}
 
 \renewcommand\Affilfont{\itshape\small}
-\affil[1]{Department of Computer Science, University of Toronto}
-\affil[2]{Department of Medicine, Division of Infectious Disease, University of Toronto}
-\affil[3]{Institute of Health Policy, Management and Evaluation, Dalla Lana School of Public Health, University of Toronto}
-\affil[4]{Institute of Medical Sciences, University of Toronto}
-\affil[5]{MAP Centre for Urban Health Solutions, Unity Health Toronto}
-\affil[6]{Department of Medicine, St. Michael's Hospital}
+\affil[a]{MAP Centre for Urban Health Solutions, Unity Health Toronto, Toronto, ON, Canada}
+\affil[b]{Department of Computer Science, University of Toronto, Toronto, ON, Canada}
+\affil[c]{Department of Medicine, Division of Infectious Disease, St. Michael's Hospital, University of Toronto, Toronto, ON, Canada}
+\affil[d]{Institute of Health Policy, Management and Evaluation, Dalla Lana School of Public Health, University of Toronto, Toronto, ON, Canada}
+\affil[e]{Institute of Medical Sciences, University of Toronto, Toronto, ON, Canada}
 
@@
 		\input{paper.discussion.tex}
+	\section{Acknowledgements} \label{sec:acknowledgements}
+		\input{paper.acknowledgements.tex}
+	\section{Funding} \label{sec:funding}
+		\input{paper.funding.tex}
+
 	\bibliographystyle{unsrt}

### 2019-08-30 paper/paper.intro.tex
Work on fixing discussion
- Remove/simplify discussion of NP-hardness
- Add more detail to impact of formulation on clinicians, hospital dept. w/citation
- Clean up ILP abbreviation throughout paper

@@
 Michael's Hospital in Toronto, Canada.
-Our goal is to (1) present a simple integer linear programming formulations for the scheduling problem as developed for the hospital, and describe the adaptability of the formulation to solving similar problems in other departments; (2) compare the performance of the ILP scheduler to the results of the manual approach; and (3) analyze the robustness of the algorithm in difficult instances of the problem.
+Our goal is to (1) present a simple integer linear programming (ILP) formulation for the scheduling problem as developed for the hospital, and describe the adaptability of the formulation to solving similar problems in other departments; (2) compare the performance of the ILP scheduler to the results of the manual approach; and (3) analyze the robustness of the algorithm in difficult instances of the problem.
 \\

### 2019-08-30 paper/paper.results.tex
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references
Add acks + funding sections
- Fix affiliation order, minor intro corrections
- TODO: add CPU specs to results section
Work on fixing discussion
- Remove/simplify discussion of NP-hardness
- Add more detail to impact of formulation on clinicians, hospital dept. w/citation
- Clean up ILP abbreviation throughout paper

@@
 \subsection{Software}
-We developed a software package with a user interface that implements the above LP problem and allows configuration of clinicians at [ref \ref{???}], to be used by the ID division at St.
+We developed a software package with a user interface that implements the above linear program and allows configuration of clinicians at [ref \ref{???}], to be used by the ID division at St.
 Michael's Hospital.
 The software was used to generate the results in the following sections, using real data as well as simulated data as input.
+All the following experiments were conducted on a [X-Core CPU] computer @ [Y GHz] and [Z GB] of RAM running 64bit Windows 10.
+The software was implemented using the COIN-OR Branch-and-Cut open source solver version 2.9.9 \cite{johnjforrest_coin-or/cbc:_2019}.
 
 \subsection{Infectious Diseases Division}  %main comment  = I think use past tense in writing
-We used clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the LP problem.
+We used clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the ILP problem.
 Table \ref{tbl:2018-schedule-comparison} presents the optimal schedule generated using the software as well as the manually-created schedule for data from 2018.
@@
 Evaluating the objectives, we see that LP outperforms the manually created schedule in all four years, by accommodating almost all time-off requests and ensuring that weekends are always assigned close to blocks.
+\\
 
@@
 \subsection{Simulations}
-We then examined the properties of the LP problem using simulated data.
+We then examined the properties of the ILP problem using simulated data.
 Specifically, we examined the influence of expanding the following parameters on the runtime of the algorithm: number of clinicians; number of services offered in the division; amount of time-off requests per clinician throughout the year.
@@
 
-The effect of an increasing number of requests per clinician on the runtime of the LP solver is shown in figure \ref{fig:runtime-requests}.
+The effect of an increasing number of requests per clinician on the runtime of the ILP solver is shown in figure \ref{fig:runtime-requests}.
 We executed the program on a department with 10 clinicians offering two services, similar to the department at St.
@@
 For simplicity, each clinician was configured with $\abs{\mathcal{BR}} = \{0, 5, 10, 15\}$ non-overlapping block requests.
-The increase in the number of requests did not affect the runtime of the LP solver, indicating that it can accommodate a lot of flexibility in clinician requests.
+The increase in the number of requests did not affect the runtime of the ILP solver, indicating that it can accommodate a lot of flexibility in clinician requests.
 Moreover, we see that all four runs were completed in under 1 second.
@@
 	\includegraphics[scale=.5]{fig/runtime_requests}
-	\caption{Runtime of LP solver with an increasing number of requests per clinician}
+	\caption{Runtime of ILP solver with an increasing number of requests per clinician}
 	\label{fig:runtime-requests}
@@
 	\includegraphics[scale=.5]{fig/runtime_blocks}
-	\caption{Runtime of LP solver with an increasing number of 2-week blocks}
+	\caption{Runtime of ILP solver with an increasing number of 2-week blocks}
 	\label{fig:runtime-blocks}

### 2019-08-30 paper/paper.discussion.tex
Work on fixing discussion
- Remove/simplify discussion of NP-hardness
- Add more detail to impact of formulation on clinicians, hospital dept. w/citation
- Clean up ILP abbreviation throughout paper

@@
 In this paper, we present a simple, yet flexible, integer linear programming formulation to generate schedules for clinical departments at hospitals.
-The challenge in applying such an approach to this task lies in the fact that ILP is an NP-hard problem.
+The challenge in applying ILP to the task of scheduling clinicians lies in the computational complexity of finding an optimal solution.
+As the size of our problem grows, due to a larger roster of clinicians or more complicated constraints, the time it takes to generate an optimal schedule may grow exponentially.
+%The challenge in applying such an approach to this task lies in the fact that ILP is an NP-hard problem.
 %specify what 'such an approach' and 'this task' are...
@@
 if yes, then instead of 'As such' - say 'That is,...or "An NP-hard problem means that..."
-As such, the time to find an optimal solution grows exponentially as we increase the size of the problem.
 As a result, the vast majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time.
@@
 
-Our formulation includes hard constraints to ensure the schedule satisfies hospital and logistics requirements.
+We presented a formulation that includes hard constraints to ensure the schedule satisfies hospital and logistics requirements, as well as a multi-goal objective function to satisfy the work preferences of clinicians.
+%Our formulation includes hard constraints to ensure the schedule satisfies hospital and logistics requirements.
 It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function.
@@
 When comparing the optimal schedule generated by tool to the manually-created schedules at St.
-Michael's Hospital, we found that the ILP formulation was always able to find an optimal schedule satisfying all required hard constraints, unlike the manual schedule, that often did not take all constraints into account.
+Michael's Hospital, we found that the ILP formulation was always able to find an optimal schedule satisfying all required hard constraints, unlike the manual schedule, which often did not take all constraints into account.
 Moreover, due to the multi-goal objective function in the ILP, the tool was able to maximize and fulfill the majority of clinician preferences and requests, more so than the manually-created schedule.
 These observations reinforce the benefits of automated tools when generating schedules in hospital departments to balance the work-load of clinicians and improve the service provided to patients.
+The use of automated tools alleviates the time spent on designing the schedule by hand, and provides clinical departments with a more fair distribution of work that helps improve the overall satisfaction of both employees and patients \cite{silvestro_evaluation_2000}.
 \\  % how the finding compares to the wider literature when automated compared to manual?
@@
 
-We also found that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences.
+In our simulations, we also found that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences.
 Further, we saw that the algorithm can accommodate increasing the time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance.
@@
 specify the other constraints?
-Thus, next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services i parallel for patients and other clients.
+Thus, next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services in parallel for patients and other clients.
 Overall, our sensitivity analyses using simulated data provide reassurance that the ILP formulation can be applied to schedule clinicians across real-world variability between clinical departments.
-\\ %SM - will look for reference and we can ask Kevin Gough too.
+%SM - will look for reference and we can ask Kevin Gough too.

### 2019-09-04 paper/paper.discussion.tex
Add CPU/RAM specs + fix up discussion

@@
 if yes, then instead of 'As such' - say 'That is,...or "An NP-hard problem means that..."
-As a result, the vast majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time.
+As a result, the majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time \cite{burke_state_2004}.
 \\  %cite/reference statement (i.e.
@@
 Such cases are unlikely to be encountered in the real-world because most clinical departments tend to provide a single service or at most two services by the same roster of clinicians [ref \ref{???}].
-One solution to mitigate the runtime issues generated by a larger number of services would be relaxing certain constraints, followed by manual readjustment from the generated schedule.
+One solution to mitigate the runtime issues generated by a larger number of services would be relaxing the constraints preventing assignment of consecutive blocks, followed by manual readjustment from the generated schedule.
 %what do we mean by certain constraints?

### 2019-09-04 paper/paper.discussion.tex
Minor typo and other fixes

@@
 The challenge in applying ILP to the task of scheduling clinicians lies in the computational complexity of finding an optimal solution.
-As the size of our problem grows, due to a larger roster of clinicians or more complicated constraints, the time it takes to generate an optimal schedule may grow exponentially.
+As the size of the scheduling problem grows, due to a larger roster of clinicians or more complicated constraints, the time it takes to generate an optimal schedule may grow exponentially.
 %The challenge in applying such an approach to this task lies in the fact that ILP is an NP-hard problem.
@@
 
-We presented a formulation that includes hard constraints to ensure the schedule satisfies hospital and logistics requirements, as well as a multi-goal objective function to satisfy the work preferences of clinicians.
+We presented a formulation that includes both hard constraints to ensure the schedule satisfies hospital and logistics requirements, and a multi-goal objective function to satisfy the work preferences of clinicians.
 %Our formulation includes hard constraints to ensure the schedule satisfies hospital and logistics requirements.
@@
 
-When comparing the optimal schedule generated by tool to the manually-created schedules at St.
+When comparing the optimal schedule generated by our tool to the manually-created schedules at St.
 Michael's Hospital, we found that the ILP formulation was always able to find an optimal schedule satisfying all required hard constraints, unlike the manual schedule, which often did not take all constraints into account.
@@
 In our simulations, we also found that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences.
-Further, we saw that the algorithm can accommodate increasing the time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance.
+Further, we saw that the algorithm can accommodate an increase in time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance.
 One key limitation we identified was the sensitivity of the runtime to larger numbers of services offered in a single department.
-Such cases are unlikely to be encountered in the real-world because most clinical departments tend to provide a single service or at most two services by the same roster of clinicians [ref \ref{???}].
-One solution to mitigate the runtime issues generated by a larger number of services would be relaxing the constraints preventing assignment of consecutive blocks, followed by manual readjustment from the generated schedule.
+Such cases are unlikely to be encountered in the real-world because most clinical departments tend to provide a single service or at most two services by the same roster of clinicians [{\color{red}SM/Kevin, could you help with reference here?}].
+One solution to mitigate the runtime issues created by a larger number of services would be relaxing the constraints that prevent assignment of consecutive blocks, followed by manual readjustment from the generated schedule.
 %what do we mean by certain constraints?
 specify the other constraints?
-Thus, next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services in parallel for patients and other clients.
 Overall, our sensitivity analyses using simulated data provide reassurance that the ILP formulation can be applied to schedule clinicians across real-world variability between clinical departments.
 %SM - will look for reference and we can ask Kevin Gough too.
+Next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services in parallel for patients and other clients.

### 2019-09-04 paper/paper.results.tex
Add CPU/RAM specs + fix up discussion
Minor typo and other fixes

@@
 The software was used to generate the results in the following sections, using real data as well as simulated data as input.
-All the following experiments were conducted on a [X-Core CPU] computer @ [Y GHz] and [Z GB] of RAM running 64bit Windows 10.
+All the following experiments were conducted on an Intel Core i7-4770k CPU @ 3.50 GHz and 16 GB of RAM running 64-bit Windows 10.
 The software was implemented using the COIN-OR Branch-and-Cut open source solver version 2.9.9 \cite{johnjforrest_coin-or/cbc:_2019}.
@@
 First, we evaluated the software-generated schedule by comparing it with the manually-created schedule.
-Specifically, we examined the adherence of each schedule to the constraints presented in table \ref{tbl:constraint-summary}.
-As shown in table \ref{tbl:constraints-comparison}, the optimal schedule generated using the software satisfied all mandatory constraints.
+Specifically, we examined the adherence of each schedule to the constraints presented in Table \ref{tbl:constraint-summary}.
+As shown in Table \ref{tbl:constraints-comparison}, the optimal schedule generated using the software satisfied all mandatory constraints.
 In contrast, the manually-created schedule was not able to satisfy all mandatory constraints.
@@
 
-The effect of an increasing number of requests per clinician on the runtime of the ILP solver is shown in figure \ref{fig:runtime-requests}.
+The effect of an increasing number of requests per clinician on the runtime of the ILP solver is shown in Figure \ref{fig:runtime-requests}.
 We executed the program on a department with 10 clinicians offering two services, similar to the department at St.
@@
 We executed the algorithm for $S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in total across all services.
-For 2 concurrent services, a roster of 30 or more clinicians becomes impractical to schedule, as searching for a solution requires over 24 hours.
-We see similar issues for a roster of 20 or more clinicians assigned to a division with 3 concurrent services.
-However, when relaxing the `No Consecutive Blocks' constraint, we see a great improvement in runtime for divisions with 2 and 3 services, and we are able to generate a schedule with upwards of 50 clinicians in under 1.5 seconds.
+For 2 concurrent services, a roster of 30 or more clinicians becomes impractical to schedule, as searching for a solution required over 24 hours.
+We saw similar issues for a roster of 20 or more clinicians assigned to a division with 3 concurrent services.
+However, when relaxing the NCB constraint, we saw a great improvement in runtime for divisions with 2 and 3 services, and we were able to generate a schedule with upwards of 50 clinicians in under 1.5 seconds.
 \\  %what does division mean?

### 2019-09-04 paper/paper.methods.tex
Work on fixing discussion
- Remove/simplify discussion of NP-hardness
- Add more detail to impact of formulation on clinicians, hospital dept. w/citation
- Clean up ILP abbreviation throughout paper
Minor typo and other fixes

@@
-In this section, we present an application of integer linear programming to solve the clinician scheduling problem presented in Section \ref{sec:problem}.
+In this section, we present an application of ILP to solve the clinician scheduling problem presented in Section \ref{sec:problem}.
 First, we describe the sets, indices and variables present in the formulation of the program.
@@
 \end{align}
-where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by some appropriate definition of adjacency.
+where $\varphi(b)$ is a one-to-one mapping of a block to an adjacent weekend, by some appropriate definition of adjacency.
 For instance, clinicians might want to be assigned during a weekend that falls within an assigned block.

### 2019-09-07 paper/paper.intro.tex
Add hard line breaks for easier editing

@@
-Hospital departments provide services where patient needs, and thus the system's demands, often exceed the available supply.
-In particular, on-call schedules for a fixed number of health-care providers are central to the efficient running of hospitals.
-Carefully allocated schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes.
-It is common practice for on-call schedules in hospitals to be created manually, yet manually-created schedules are prone to errors and potential for various biases.
-First, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method may not provide a schedule that satisfies all clinicians.
-Second, such methods are likely to overlook certain constraints that must be maintained to have an operational department, such as preventing many consecutive work blocks from being assigned or ensuring clinicians are allocated a specific amount of work blocks throughout the year.
-Third, manual scheduling is often time-consuming for the person developing the schedule.
-For these reasons, it is important to develop automated methods that can generate optimal schedules that satisfy the given constraints of the hospital department.
-\\
-%For example, it is important that a hospital department allocates its resources, such as the availability of a finite number of clinicians, optimally, to ensure the best possible service for its patients.
+Hospital departments provide services where patient needs, and thus the system's
+demands, often exceed the available supply. In particular, on-call schedules for
+a fixed number of health-care providers are central to the efficient running of
+hospitals. Carefully allocated schedules are meant to simultaneously ensure
+sufficient resources are provided to patients while not overworking clinicians
+to prevent costly mistakes. It is common practice for on-call schedules in
+hospitals to be created manually, yet manually-created schedules are prone to
+errors and potential for various biases. First, when there is a large number of
+clinicians in a single department, or the constraints that need to be satisfied
+by the department are very complex, a manual method may not provide a schedule
+that satisfies all clinicians. Second, such methods are likely to overlook
+certain constraints that must be maintained to have an operational department,
+such as preventing many consecutive work blocks from being assigned or ensuring
+clinicians are allocated a specific amount of work blocks throughout the year.
+Third, manual scheduling is often time-consuming for the person developing the
+schedule. For these reasons, it is important to develop automated methods that
+can generate optimal schedules that satisfy the given constraints of the
+hospital department. \\
+%For example, it is important that a hospital department allocates its
+%resources, such as the availability of a finite number of clinicians,
+%optimally, to ensure the best possible service for its patients.
 
-Automated methods to optimize schedules have been studied and applied in many industries, including transportation \cite{aickelin_improved_2006, goel_truck_2012, gunther_combined_2010}, manufacturing \cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008, alfares_simulation_2007}, retail \cite{chapados_retail_2011, nissen_automatic_2010} and military \cite{horn_scheduling_2007, laguna_modeling_2005}.
-Of special interest to a clinician on-call scheduling problem are the approaches to schedule nurses, who often work in shifts.
-In the nurse scheduling problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences.
-A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming \cite{azaiez_0-1_2005, trilling_nurse_2006, widyastiti_nurses_2016}, network flows \cite{el_adoly_new_2018}, genetic algorithms \cite{aickelin_exploiting_2000, jan_evolutionary_2000, kawanaka_genetic_2001}, simulated annealing \cite{jaszkiewicz_metaheuristic_1997}, and artificial intelligence \cite{abdennadher_nurse_nodate, li_hybrid_2003}.
-A comprehensive literature review of these and other methods applied to nurse rostering is presented in \cite{burke_state_2004}.
-\\
+Automated methods to optimize schedules have been studied and applied in many
+industries, including transportation \cite{aickelin_improved_2006,
+	goel_truck_2012, gunther_combined_2010}, manufacturing
+\cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008,
+	alfares_simulation_2007}, retail \cite{chapados_retail_2011,
+	nissen_automatic_2010} and military \cite{horn_scheduling_2007,
+	laguna_modeling_2005}. Of special interest to a clinician on-call scheduling
+problem are the approaches to schedule nurses, who often work in shifts. In the
+nurse scheduling problem, the goal is to find an optimal assignment of nurses to
+shifts that satisfies all of the hard constraints, such as hospital regulations,
+and as many soft constraints as possible, which may include nurse preferences. A
+wide variety of approaches, including exact and heuristic approaches, have been
+used to solve the nurse scheduling problem: integer linear programming
+\cite{azaiez_0-1_2005, trilling_nurse_2006, widyastiti_nurses_2016}, network
+flows \cite{el_adoly_new_2018}, genetic algorithms
+\cite{aickelin_exploiting_2000, jan_evolutionary_2000, kawanaka_genetic_2001},
+simulated annealing \cite{jaszkiewicz_metaheuristic_1997}, and artificial
+intelligence \cite{abdennadher_nurse_nodate, li_hybrid_2003}. A comprehensive
+literature review of these and other methods applied to nurse rostering is
+presented in \cite{burke_state_2004}. \\
 
 %An extensive literature review of these and other methods is presented by [??].
-We will briefly summarize the main ideas of some of these approaches.
-\\ SM - don't need to introduce that you will do this for this type of paper I think - would do in a thesis chapter though.
+%We will briefly summarize the main ideas of some of these approaches. \\ SM -
+%don't need to introduce that you will do this for this type of paper I think -
+%would do in a thesis chapter though.
 
-Many of the approaches to nurse scheduling were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation.
-While these department-specific approaches allow end-users to find precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians in that department, they are difficult to readily adapt to other departments in the same hospital or other hospitals.
-% explain why hard to adapt?...
-what makes their generalizablility/adaptability limited?
-Moreover, the large number of variables and constraints also leads to computational complexity issues \cite{goos_complexity_1996}, especially when using exact methods for finding the solution.
-In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease (ID) consults; and HIV consults service) at St.
-Michael's Hospital in Toronto, Canada.
-Our goal is to (1) present a simple integer linear programming (ILP) formulation for the scheduling problem as developed for the hospital, and describe the adaptability of the formulation to solving similar problems in other departments; (2) compare the performance of the ILP scheduler to the results of the manual approach; and (3) analyze the robustness of the algorithm in difficult instances of the problem.
-\\
-% present a simple formulation for the problem developed and tested at the hospital after switching from a manual approach to scheduling; and (2) analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach; and (3) describe the adaptability of the formulation as a basic framework for solving similar problems in other departments.
-\\ %make #3 an objective
+Many of the approaches to nurse scheduling were designed to satisfy the
+requirements of a specific hospital department which causes a large number of
+variables and constraints to be incorporated into the problem formulation. While
+these department-specific approaches allow end-users to find precise schedules
+that satisfy the needs of the department and the preferences of the nurses and
+clinicians in that department, they are difficult to readily adapt to other
+departments in the same hospital or other hospitals. % explain why hard to
+%adapt?... what makes their generalizablility/adaptability limited?
+Moreover, the large number of variables and constraints also leads to
+computational complexity issues \cite{goos_complexity_1996}, especially when
+using exact methods for finding the solution. In this paper, we tackle a version
+of the nurse scheduling problem arising from a case study of one clinical
+division, providing two different services simultaneously (general infectious
+disease (ID) consults; and HIV consults service) at St. Michael's Hospital in
+Toronto, Canada. Our goal is to (1) present a simple integer linear programming
+(ILP) formulation for the scheduling problem as developed for the hospital, and
+describe the adaptability of the formulation to solving similar problems in
+other departments; (2) compare the performance of the ILP scheduler to the
+results of the manual approach; and (3) analyze the robustness of the algorithm
+in difficult instances of the problem. \\
+% present a simple formulation for the problem developed and tested at the
+%hospital after switching from a manual approach to scheduling; and (2) analyze
+%the performance of integer linear programming in solving difficult instances of
+%the problem and compare the results with those of the manual approach; and (3)
+%describe the adaptability of the formulation as a basic framework for solving
+%similar problems in other departments. \\ %make #3 an objective
 
-We begin by describing the problem in Section \ref{sec:problem}, and presenting our ILP formulation in Section \ref{sec:methods}.
-Next, we compare the results of our formulation to manually-created schedules, and evaluate the performance of the algorithm on simulated data in Section \ref{sec:results}.
-Finally, we discuss and interpret the results in Section \ref{sec:discussion}.
-% list the main contents/elements of the paper.
+We begin by describing the problem in Section \ref{sec:problem}, and presenting
+our ILP formulation in Section \ref{sec:methods}. Next, we compare the results
+of our formulation to manually-created schedules, and evaluate the performance
+of the algorithm on simulated data in Section \ref{sec:results}. Finally, we
+discuss and interpret the results in Section \ref{sec:discussion}. % list the
+%main contents/elements of the paper.

### 2019-09-07 paper/paper.methods.tex
Add hard line breaks for easier editing

@@
-In this section, we present an application of ILP to solve the clinician scheduling problem presented in Section \ref{sec:problem}.
-First, we describe the sets, indices and variables present in the formulation of the program.
-We then convert the constraints given in Table \ref{tbl:constraint-summary} into mathematical terms, and outline the objective function of the linear program.
+In this section, we present an application of ILP to solve the clinician
+scheduling problem presented in Section \ref{sec:problem}. First, we describe
+the sets, indices and variables present in the formulation of the program. We
+then convert the constraints given in Table \ref{tbl:constraint-summary} into
+mathematical terms, and outline the objective function of the linear program.
 
 \subsection{Sets and Indices} \label{sec:meth-sets-indices}
-We denote the set of all services %or I think services is good (vs. services/divisions) as you define two types of services (ID vs. HIV) in the problem statement
-that clinicians in a single department can provide as $\mathcal{S}$.
-The following formulation assumes that all clinicians are able to provide all services.
-The set of all clinicians in the department is denoted as $\mathcal{C}$.
-The sets of blocks and weekends that clinicians will be assigned to are denoted as $\mathcal{B}$ and $\mathcal{W}$ respectively.
-The block size used in our experiments is 2 weeks, but the following LP formulation does not specify a particular size for blocks, and so it can be adapted to the needs of the department in question.
+We denote the set of all services %or I think services is good (vs.
+%services/divisions) as you define two types of services (ID vs. HIV) in the
+%problem statement
+that clinicians in a single department can provide as $\mathcal{S}$. The
+following formulation assumes that all clinicians are able to provide all
+services. The set of all clinicians in the department is denoted as
+$\mathcal{C}$. The sets of blocks and weekends that clinicians will be assigned
+to are denoted as $\mathcal{B}$ and $\mathcal{W}$ respectively. The block size
+used in our experiments is 2 weeks, but the following LP formulation does not
+specify a particular size for blocks, and so it can be adapted to the needs of
+the department in question.
 %The size of a block is not constrained and can be adapted %how?
 %to the needs of the given department. 
-A subset of weekends are denoted $\mathcal{L}$, corresponding to established long/holiday weekends such as the Canadian Civic Day weekend in August.
-Lastly, we denote with $\mathcal{BR}_c$ and $\mathcal{WR}_c$ the block and weekend requests of clinicians as subsets of all blocks and weekends, respectively.
-For instance, if clinician $c$'s requests intersect with blocks 1 and 2, and weekend 1, then $\mathcal{BR}_c = \{1, 2\}$ and $\mathcal{WR}_c = \{1\}$.
-Table \ref{tbl:sets-indices} presents a summary of the sets and indices described.
+A subset of weekends are denoted $\mathcal{L}$, corresponding to established
+long/holiday weekends such as the Canadian Civic Day weekend in August. Lastly,
+we denote with $\mathcal{BR}_c$ and $\mathcal{WR}_c$ the block and weekend
+requests of clinicians as subsets of all blocks and weekends, respectively. For
+instance, if clinician $c$'s requests intersect with blocks 1 and 2, and weekend
+1, then $\mathcal{BR}_c = \{1, 2\}$ and $\mathcal{WR}_c = \{1\}$. Table
+\ref{tbl:sets-indices} presents a summary of the sets and indices described.
 
@@
 		\hline
-		\textbf{Set}                         & \textbf{Index} & \textbf{Description}              \\ \hline
-		$\mathcal{S} = \{1, \ldots, S \}$    & $s$            & services                          \\
-		$\mathcal{C} = \{1, \ldots, C \}$    & $c$            & clinicians                        \\
-		$\mathcal{B} = \{1, \ldots, B \}$    & $b$            & blocks                            \\
-		$\mathcal{W} = \{1, \ldots, W \}$    & $w$            & weekends                          \\
-		$\mathcal{L} \subset \mathcal{W}$    &                & long weekends                     \\
-		$\mathcal{BR}_c \subset \mathcal{B}$ &                & block requests of clinician $c$   \\
-		$\mathcal{WR}_c \subset \mathcal{W}$ &                & weekend requests of clinician $c$
+		\textbf{Set}                         & \textbf{Index} & \textbf{Description}  
+		\\ \hline
+		$\mathcal{S} = \{1, \ldots, S \}$    & $s$            & services              
+		\\
+		$\mathcal{C} = \{1, \ldots, C \}$    & $c$            & clinicians            
+		\\
+		$\mathcal{B} = \{1, \ldots, B \}$    & $b$            & blocks                
+		\\
+		$\mathcal{W} = \{1, \ldots, W \}$    & $w$            & weekends              
+		\\
+		$\mathcal{L} \subset \mathcal{W}$    &                & long weekends         
+		\\
+		$\mathcal{BR}_c \subset \mathcal{B}$ &                & block requests of
+		clinician $c$   \\
+		$\mathcal{WR}_c \subset \mathcal{W}$ &                & weekend requests of
+		clinician $c$
 	\end{tabular}
@@
 \subsection{Variables} \label{sec:meth-variables}
-Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as a binary variable $X_{c, b, s}$.
-A value of 1 indicates that the given clinician $c$ is assigned to provide service $s$ during block $b$.
-Weekend assignments are similarly defined using a binary variable $Y_{c, w}$, but without a service index, as clinicians are expected to provide all services during the weekends.
-We then introduce a set of constants $m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year.
-Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.
-%To optimize the soft constraint Block-Weekend Adjacency, we maximize the product $X_{c, b, s} \cdot Y_{c, w}$ for adjacent blocks and weekends.
-To formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, s}$, with additional constraints on its range.
-Further details regarding this variable are described in Section \ref{sec:meth-objectives}.
+Since each clinician may be assigned to work on any service, during any block of
+the year, we denote such an assignment as a binary variable $X_{c, b, s}$. A
+value of 1 indicates that the given clinician $c$ is assigned to provide service
+$s$ during block $b$. Weekend assignments are similarly defined using a binary
+variable $Y_{c, w}$, but without a service index, as clinicians are expected to
+provide all services during the weekends. We then introduce a set of constants
+$m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks
+each clinician is allowed to work during the year. Table
+\ref{tbl:variables-constants} presents a summary of the constants and variables
+in the problem.
+%To optimize the soft constraint Block-Weekend Adjacency, we maximize the
+%product $X_{c, b, s} \cdot Y_{c, w}$ for adjacent blocks and weekends. To
+%formulate such an objective as a linear function of variables, we introduce
+%another set of variables, denoted by $Z_{c, b, s}$, with additional constraints
+%on its range. Further details regarding this variable are described in Section
+%\ref{sec:meth-objectives}.
 
@@
 		\hline
-		\textbf{Name}              & \textbf{Description}                                               \\ \hline
-		$X_{c, b, s} \in \{0, 1\}$ & assignment of clinician $c$ to service $s$ on block $b$            \\
-		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                         \\
-%		$Z_{c, b, s} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend adjacency             \\
-		$m_{c, s}$                 & minimum number of blocks clinician $c$ should cover on service $s$ \\
-		$M_{c, s}$                 & maximum number of blocks clinician $c$ should cover on service $s$
+		\textbf{Name}              & \textbf{Description}                             
+		\\ \hline
+		$X_{c, b, s} \in \{0, 1\}$ & assignment of clinician $c$ to service $s$ on
+		block $b$            \\
+		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$       
+		\\
+		%		$Z_{c, b, s} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend
+		%adjacency             \\
+		$m_{c, s}$                 & minimum number of blocks clinician $c$ should
+		cover on service $s$ \\
+		$M_{c, s}$                 & maximum number of blocks clinician $c$ should
+		cover on service $s$
 	\end{tabular}
@@
 \subsection{Constraints} \label{sec:meth-constraints}
-We now formalize the hard constraints in Table \ref{tbl:constraint-summary} using the variables defined above.
+We now formalize the hard constraints in Table \ref{tbl:constraint-summary}
+using the variables defined above.
 
 \begin{align}
-	&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S} \tag{BC} \label{eqn:constr-block-cov} \\
-	&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{WC} \label{eqn:constr-weekend-cov} \\
-	&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\ c\in\mathcal{C}, s\in\mathcal{S} \tag{MM} \label{eqn:constr-min-max} \\
-	&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, s\in\mathcal{S} \tag{NCB} \label{eqn:constr-no-consec-blocks} \\
-	&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{NCW} \label{eqn:constr-no-consec-weekends} \\
-	&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}} &&\forall c\in\mathcal{C} \tag{EW} \label{eqn:constr-equal-weekends} \\
-	&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq \ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C} \tag{EH} \label{eqn:constr-equal-holidays}
+&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S}
+\tag{BC} \label{eqn:constr-block-cov} \\
+&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{WC}
+\label{eqn:constr-weekend-cov} \\
+&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\
+c\in\mathcal{C}, s\in\mathcal{S} \tag{MM} \label{eqn:constr-min-max} \\
+&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1,
+s\in\mathcal{S} \tag{NCB} \label{eqn:constr-no-consec-blocks} \\
+&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1
+\tag{NCW} \label{eqn:constr-no-consec-weekends} \\
+&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}}
+&&\forall c\in\mathcal{C} \tag{EW} \label{eqn:constr-equal-weekends} \\
+&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq
+\ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C} \tag{EH}
+\label{eqn:constr-equal-holidays}
 \end{align}
@@
 \subsection{Objectives} \label{sec:meth-objectives}
-As described in Section \ref{sec:problem}, the soft constraints of the clinician scheduling problem include: satisfying clinician block off requests (BR), satisfying clinician weekend off requests (WR), and assigning weekends closer to blocks (BWA).
-We convert these soft constraints into linear objective functions of the binary variables defined in Section \ref{sec:meth-variables}.
-Objectives (\ref{eqn:obj-block-requests}) and (\ref{eqn:obj-weekend-requests}) can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to.
+As described in Section \ref{sec:problem}, the soft constraints of the clinician
+scheduling problem include: satisfying clinician block off requests (BR),
+satisfying clinician weekend off requests (WR), and assigning weekends closer to
+blocks (BWA). We convert these soft constraints into linear objective functions
+of the binary variables defined in Section \ref{sec:meth-variables}. Objectives
+(\ref{eqn:obj-block-requests}) and (\ref{eqn:obj-weekend-requests}) can be
+written as the following linear functions:  %minor suggestion - consider
+%numbering the objectives for ease later when saying things like "these two
+%objectives"...or writing the full name of the objectives instead of 'this' and
+%'these' as sometimes can get confusing what the 'this' and the 'these' refer
+%to. 
 \begin{align}
-	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{BR} \label{eqn:obj-block-requests}\\
-	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{WR} \label{eqn:obj-weekend-requests}
+&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S}
+(-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{BR}
+\label{eqn:obj-block-requests}\\
+&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W}
+(-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{WR}
+\label{eqn:obj-weekend-requests}
 \end{align}
-where $\ind(P)$ is the indicator function that has value 1 when predicate $P$ holds and 0 otherwise.
-In the above two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments.
-\\
+where $\ind(P)$ is the indicator function that has value 1 when predicate $P$
+holds and 0 otherwise. In the above two objectives, we penalize any assignments
+that conflict with a block or weekend request, and aim to maximize the
+non-conflicting assignments. \\
 
-The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, s}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$.
-This leads to the maximization objective %avoid editorial adjectives
+The Block-Weekend Adjacency is optimized by considering the product $X_{c, b,
+	s}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$. This
+leads
+to the maximization objective %avoid editorial adjectives
 \begin{align}
-&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} X_{c, b, s}\cdot Y_{c, w=\varphi(b)} \tag{BWA} \label{eqn:obj-block-weekend-adj}
+&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} X_{c, b, s}\cdot Y_{c,
+	w=\varphi(b)} \tag{BWA} \label{eqn:obj-block-weekend-adj}
 \end{align}
-where $\varphi(b)$ is a one-to-one mapping of a block to an adjacent weekend, by some appropriate definition of adjacency.
-For instance, clinicians might want to be assigned during a weekend that falls within an assigned block.
-In this case, we will have $\varphi(b) = 2b-1$.
-\\
+where $\varphi(b)$ is a one-to-one mapping of a block to an adjacent weekend, by
+some appropriate definition of adjacency. For instance, clinicians might want to
+be assigned during a weekend that falls within an assigned block. In this case,
+we will have $\varphi(b) = 2b-1$. \\
 
-%So if clinician $c$ is assigned to work during block 3, corresponding to weeks 5 and 6 assuming 2-week blocks, they might also want to be assigned to work during weekend 5.
-In that case, we would like $X_{c, b=3, s} \cdot Y_{c, w=5}$ to be 1, since that indicates both variables are assigned.
-If at least one of the two variables is not assigned, the product will be 0.
+%So if clinician $c$ is assigned to work during block 3, corresponding to weeks
+%5 and 6 assuming 2-week blocks, they might also want to be assigned to work
+%during weekend 5. In that case, we would like $X_{c, b=3, s} \cdot Y_{c, w=5}$
+%to be 1, since that indicates both variables are assigned. If at least one of
+%the two variables is not assigned, the product will be 0. 
 % I really liked the example explanation!
@@
 
-However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework.
-An approach used to convert such functions into linear objectives involves introducing a helper variable and additional constraints \cite{hammer_boolean_1968}.
-In our case, introducing a variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that
+However, as it is, $Q_3$ is not a linear function of its variables and cannot be
+optimized in a linear programming framework. An approach used to convert such
+functions into linear objectives involves introducing a helper variable and
+additional constraints \cite{hammer_boolean_1968}. In our case, introducing a
+variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w =
+\varphi(b)$, and constraining $Z$ such that 
 \begin{align}
 &Z_{c, b, s} \leq X_{c, b, s} \label{eqn:helper-x-constraint}\\
-	&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S} \label{eqn:helper-y-constraint}
+&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S}
+\label{eqn:helper-y-constraint}
 \end{align}
@@
 \end{align}
-Indeed from Eqns.
-(\ref{eqn:helper-x-constraint}) and (\ref{eqn:helper-y-constraint}),  whenever $X_{c, b, s} \cdot Y_{c, w} = 1$ (respectively, 0), $Z_{c, b, s}$ can attain a maximum value of 1 (respectively, 0), giving us the correct adjacency maximization objective.
-\\
-%Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a maximum value of 1, and whenever $X_{c, b, s} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, s}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, s}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective.
-\\
+Indeed from Eqns. (\ref{eqn:helper-x-constraint}) and
+(\ref{eqn:helper-y-constraint}),  whenever $X_{c, b, s} \cdot Y_{c, w} = 1$
+(respectively, 0), $Z_{c, b, s}$ can attain a maximum value of 1 (respectively,
+0), giving us the correct adjacency maximization objective. \\
+%Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a
+%maximum value of 1, and whenever $X_{c, b, s} \cdot Y_{c, w} = 0$, at least one
+%of $X_{c, b, s}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, s}$ will be constrained
+%to attain a maximum value of 0, giving us the correct adjacency maximization
+%objective. \\
 
-In order to optimize all objectives simultaneously, we optimize a weighted sum of the normalized objective functions,
+In order to optimize all objectives simultaneously, we optimize a weighted sum
+of the normalized objective functions,
 \begin{equation}
-\max_{X, Y, Z} \alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
+\max_{X, Y, Z} \alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta)
+\bar{Q}_3(Z)
 \end{equation}
-where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha, \beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal \cite{stanimirovic_linear_2011}. \\
+where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha,
+\beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal
+\cite{stanimirovic_linear_2011}. \\
 
-%Thus, our clinician scheduling problem is a multiple objective optimization problem.
-The most common approach to solving multiple objective optimization problems is by optimizing a weighted sum of the normalized objective functions, as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}].
-This is the approach we decided to use in our problem, to ensure all three objectives are considered when finding a solution.
-Under the assumption that each clinician in $\mathcal{C}$ provides all types of services in $\mathcal{S}$, the normalized objectives can be written as follows,
+%Thus, our clinician scheduling problem is a multiple objective optimization
+%problem. The most common approach to solving multiple objective optimization
+%problems is by optimizing a weighted sum of the normalized objective functions,
+%as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}].
+%This is the approach we decided to use in our problem, to ensure all three
+%objectives are considered when finding a solution. Under the assumption that
+%each clinician in $\mathcal{C}$ provides all types of services in
+%$\mathcal{S}$, the normalized objectives can be written as follows,
 %\begin{align}
-%	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot S} \tag{Block Requests} \label{eqn:norm-obj-block-requests}\\
-%	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests} \label{eqn:norm-obj-weekend-requests} \\
-%	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot S} \tag{Block-Weekend Adjacency} \label{eqn:norm-obj-block-weekend-adj}
+%	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot S} \tag{Block Requests}
+%\label{eqn:norm-obj-block-requests}\\
+%	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests}
+%\label{eqn:norm-obj-weekend-requests} \\
+%	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot S} \tag{Block-Weekend Adjacency}
+%\label{eqn:norm-obj-block-weekend-adj}
 %\end{align}
-%where we divide each of the original objective functions by the sum of the absolute values of its coefficients [ref \ref{???}]. The final weighted objective is given by
+%where we divide each of the original objective functions by the sum of the
+%absolute values of its coefficients [ref \ref{???}]. The final weighted
+%objective is given by
 %\begin{equation}

### 2019-09-07 paper/paper.results.tex
Add hard line breaks for easier editing

@@
 \subsection{Software}
-We developed a software package with a user interface that implements the above linear program and allows configuration of clinicians at [ref \ref{???}], to be used by the ID division at St.
-Michael's Hospital.
-The software was used to generate the results in the following sections, using real data as well as simulated data as input.
-All the following experiments were conducted on an Intel Core i7-4770k CPU @ 3.50 GHz and 16 GB of RAM running 64-bit Windows 10.
-The software was implemented using the COIN-OR Branch-and-Cut open source solver version 2.9.9 \cite{johnjforrest_coin-or/cbc:_2019}.
+We developed a software package with a user interface that implements the above
+linear program and allows configuration of clinicians at [ref \ref{???}], to be
+used by the ID division at St. Michael's Hospital. The software was used to
+generate the results in the following sections, using real data as well as
+simulated data as input. All the following experiments were conducted on an
+Intel Core i7-4770k CPU @ 3.50 GHz and 16 GB of RAM running 64-bit Windows 10.
+The software was implemented using the COIN-OR Branch-and-Cut open source solver
+version 2.9.9 \cite{johnjforrest_coin-or/cbc:_2019}.
 
-\subsection{Infectious Diseases Division}  %main comment  = I think use past tense in writing
-We used clinician time-off requests and minimum/maximum requirements from 2015-2018 as input data for the ILP problem.
-Table \ref{tbl:2018-schedule-comparison} presents the optimal schedule generated using the software as well as the manually-created schedule for data from 2018.
-The schedule is color-coded to distinguish between the different clinicians assigned.
-\\
+\subsection{Infectious Diseases Division}  %main comment  = I think use past
+%tense in writing
+We used clinician time-off requests and minimum/maximum requirements from
+2015-2018 as input data for the ILP problem. Table
+\ref{tbl:2018-schedule-comparison} presents the optimal schedule generated using
+the software as well as the manually-created schedule for data from 2018. The
+schedule is color-coded to distinguish between the different clinicians
+assigned. \\
 
@@
 
-First, we evaluated the software-generated schedule by comparing it with the manually-created schedule.
-Specifically, we examined the adherence of each schedule to the constraints presented in Table \ref{tbl:constraint-summary}.
-As shown in Table \ref{tbl:constraints-comparison}, the optimal schedule generated using the software satisfied all mandatory constraints.
-In contrast, the manually-created schedule was not able to satisfy all mandatory constraints.
-In particular, we see that the manual schedule assigned clinicians to multiple consecutive blocks in all four years.
-Moreover, the manual schedule did not have an equal distribution of weekends and holidays for all four years of data.
-Evaluating the objectives, we see that LP outperforms the manually created schedule in all four years, by accommodating almost all time-off requests and ensuring that weekends are always assigned close to blocks.
-\\
+First, we evaluated the software-generated schedule by comparing it with the
+manually-created schedule. Specifically, we examined the adherence of each
+schedule to the constraints presented in Table \ref{tbl:constraint-summary}. As
+shown in Table \ref{tbl:constraints-comparison}, the optimal schedule generated
+using the software satisfied all mandatory constraints. In contrast, the
+manually-created schedule was not able to satisfy all mandatory constraints. In
+particular, we see that the manual schedule assigned clinicians to multiple
+consecutive blocks in all four years. Moreover, the manual schedule did not have
+an equal distribution of weekends and holidays for all four years of data.
+Evaluating the objectives, we see that LP outperforms the manually created
+schedule in all four years, by accommodating almost all time-off requests and
+ensuring that weekends are always assigned close to blocks. \\
 
@@
 We then examined the properties of the ILP problem using simulated data.
-Specifically, we examined the influence of expanding the following parameters on the runtime of the algorithm: number of clinicians; number of services offered in the division; amount of time-off requests per clinician throughout the year.
-We also examined the effect of a longer time-horizon on the runtime.
-\\
+Specifically, we examined the influence of expanding the following parameters on
+the runtime of the algorithm: number of clinicians; number of services offered
+in the division; amount of time-off requests per clinician throughout the year.
+We also examined the effect of a longer time-horizon on the runtime. \\
 
-The effect of an increasing number of requests per clinician on the runtime of the ILP solver is shown in Figure \ref{fig:runtime-requests}.
-We executed the program on a department with 10 clinicians offering two services, similar to the department at St.
-Michael's Hospital.
-For simplicity, each clinician was configured with $\abs{\mathcal{BR}} = \{0, 5, 10, 15\}$ non-overlapping block requests.
-The increase in the number of requests did not affect the runtime of the ILP solver, indicating that it can accommodate a lot of flexibility in clinician requests.
-Moreover, we see that all four runs were completed in under 1 second.
-\\
+The effect of an increasing number of requests per clinician on the runtime of
+the ILP solver is shown in Figure \ref{fig:runtime-requests}. We executed the
+program on a department with 10 clinicians offering two services, similar to the
+department at St. Michael's Hospital. For simplicity, each clinician was
+configured with $\abs{\mathcal{BR}} = \{0, 5, 10, 15\}$ non-overlapping block
+requests. The increase in the number of requests did not affect the runtime of
+the ILP solver, indicating that it can accommodate a lot of flexibility in
+clinician requests. Moreover, we see that all four runs were completed in under
+1 second. \\
 
@@
 	\includegraphics[scale=.5]{fig/runtime_requests}
-	\caption{Runtime of ILP solver with an increasing number of requests per clinician}
+	\caption{Runtime of ILP solver with an increasing number of requests per
+		clinician}
 	\label{fig:runtime-requests}
@@
 
-Figure \ref{fig:runtime-blocks} presents the change in runtime when increasing the number of 2-week blocks $B = \{26, 52, 78, 104\}$ in a department with 10 clinicians offering two services.
-There does not seem to be a strong effect on runtime when we lengthen the time-horizon of the schedule, and in fact the solver was able to find an optimal schedule for all time horizons within 3 seconds.
-\\ % 'very reasonable amount of time" is vague.
-what is reasonable?
-better to give a range in value.
+Figure \ref{fig:runtime-blocks} presents the change in runtime when increasing
+the number of 2-week blocks $B = \{26, 52, 78, 104\}$ in a department with 10
+clinicians offering two services. There does not seem to be a strong effect on
+runtime when we lengthen the time-horizon of the schedule, and in fact the
+solver was able to find an optimal schedule for all time horizons within 3
+seconds. \\ % 'very reasonable amount of time" is vague. what is reasonable?
+%better to give a range in value.
 
@@
 
-The effect of increasing the number of services %divisions %do you mean services?
-on the runtime of the program is shown in Table \ref{tbl:runtime-services-clinicians-comparison}.
-We executed the algorithm for $S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in total across all services.
-For 2 concurrent services, a roster of 30 or more clinicians becomes impractical to schedule, as searching for a solution required over 24 hours.
-We saw similar issues for a roster of 20 or more clinicians assigned to a division with 3 concurrent services.
-However, when relaxing the NCB constraint, we saw a great improvement in runtime for divisions with 2 and 3 services, and we were able to generate a schedule with upwards of 50 clinicians in under 1.5 seconds.
-\\  %what does division mean?
-do you mean services?
+The effect of increasing the number of services %divisions %do you mean
+%services?
+on the runtime of the program is shown in Table
+\ref{tbl:runtime-services-clinicians-comparison}. We executed the algorithm for
+$S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in
+total across all services. For 2 concurrent services, a roster of 30 or more
+clinicians becomes impractical to schedule, as searching for a solution required
+over 24 hours. We saw similar issues for a roster of 20 or more clinicians
+assigned to a division with 3 concurrent services. However, when relaxing the
+NCB constraint, we saw a great improvement in runtime for divisions with 2 and 3
+services, and we were able to generate a schedule with upwards of 50 clinicians
+in under 1.5 seconds. \\  %what does division mean? do you mean services?
 

### 2019-09-07 paper/paper.discussion.tex
Add hard line breaks for easier editing

@@
-In this paper, we present a simple, yet flexible, integer linear programming formulation to generate schedules for clinical departments at hospitals.
-The challenge in applying ILP to the task of scheduling clinicians lies in the computational complexity of finding an optimal solution.
-As the size of the scheduling problem grows, due to a larger roster of clinicians or more complicated constraints, the time it takes to generate an optimal schedule may grow exponentially.
-%The challenge in applying such an approach to this task lies in the fact that ILP is an NP-hard problem.
-%specify what 'such an approach' and 'this task' are...
-the first part of the sentence was a bit hard to follow.
-Define what you mean by NP-hard problem and cite --> i.e.
-NP-hard = the following sentence about how time to optimal solution grows exponentially?
-if yes, then instead of 'As such' - say 'That is,...or "An NP-hard problem means that..."
-As a result, the majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time \cite{burke_state_2004}.
-\\  %cite/reference statement (i.e.
-the 'vast majority'...) I can't remember if the introduction details the heuristics --> but if yes, then great.
-if no, then include that review of the prior literature here in discussion.
-I think as you said - the journals we are submitting to seem to have more of those details in the introduction.
+In this paper, we present a simple, yet flexible, integer linear programming
+formulation to generate schedules for clinical departments at hospitals.
+The challenge in applying ILP to the task of scheduling clinicians lies in the
+computational complexity of finding an optimal solution. As the size of the
+scheduling problem grows, due to a larger roster of clinicians or more
+complicated constraints, the time it takes to generate an optimal schedule may
+grow exponentially.
+%The challenge in applying such an approach to this task lies in the fact that
+%ILP is an NP-hard problem. %specify what 'such an approach' and 'this task'
+%are... the first part of the sentence was a bit hard to follow. Define what you
+%mean by NP-hard problem and cite --> i.e. NP-hard = the following sentence
+%about how time to optimal solution grows exponentially? if yes, then instead of
+%'As such' - say 'That is,...or "An NP-hard problem means that..."
+As a result, the majority of approaches taken to create schedules for similar
+scenarios tend to use heuristics in order to find an approximately optimal
+solution in a shorter time \cite{burke_state_2004}. \\  %cite/reference
+%statement (i.e. the 'vast majority'...) I can't remember if the introduction
+%details the heuristics --> but if yes, then great. if no, then include that
+%review of the prior literature here in discussion. I think as you said - the
+%journals we are submitting to seem to have more of those details in the
+%introduction.
 
-%double-check all abbreviations have been defined. I would prefer limiting abbreviations to no more than 3 at most if possible. We are ok for word count in this paper :).
+%double-check all abbreviations have been defined. I would prefer limiting
+%abbreviations to no more than 3 at most if possible. We are ok for word count
+%in this paper :).
 
-We presented a formulation that includes both hard constraints to ensure the schedule satisfies hospital and logistics requirements, and a multi-goal objective function to satisfy the work preferences of clinicians.
-%Our formulation includes hard constraints to ensure the schedule satisfies hospital and logistics requirements.
-It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function.
-Although we restricted our use of the formulation to a minimal set of constraints for the particular needs of the case study (St.
-Michael's Hospital Division of Infectious Diseases) - our formulation can be adapted to various clinical departments at different hospitals.
-The flexibility of our ILP allows changing the number of services provided in a division, the length of a work block, clinicians' preference for block to weekend adjacency as well as clinicians' requests for time off.
-\\
+We presented a formulation that includes both hard constraints to ensure the
+schedule satisfies hospital and logistics requirements, and a multi-goal
+objective function to satisfy the work preferences of clinicians.
+%Our formulation includes hard constraints to ensure the schedule satisfies
+%hospital and logistics requirements. It also aims to satisfy the work
+%preferences of clinicians in a clinical department by optimizing a multi-goal
+%objective function. 
+Although we restricted our use of the formulation to a minimal set of
+constraints for the particular needs of the case study (St. Michael's Hospital
+Division of Infectious Diseases) - our formulation can be adapted to various
+clinical departments at different hospitals. The flexibility of our ILP allows
+changing the number of services provided in a division, the length of a work
+block, clinicians' preference for block to weekend adjacency as well as
+clinicians' requests for time off. \\
 
-When comparing the optimal schedule generated by our tool to the manually-created schedules at St.
-Michael's Hospital, we found that the ILP formulation was always able to find an optimal schedule satisfying all required hard constraints, unlike the manual schedule, which often did not take all constraints into account.
-Moreover, due to the multi-goal objective function in the ILP, the tool was able to maximize and fulfill the majority of clinician preferences and requests, more so than the manually-created schedule.
-These observations reinforce the benefits of automated tools when generating schedules in hospital departments to balance the work-load of clinicians and improve the service provided to patients.
-The use of automated tools alleviates the time spent on designing the schedule by hand, and provides clinical departments with a more fair distribution of work that helps improve the overall satisfaction of both employees and patients \cite{silvestro_evaluation_2000}.
-\\  % how the finding compares to the wider literature when automated compared to manual?
-What does it mean re: "so what" - human error in generating tools, etc.
-or challenges in heuristics used manually by people?
-include citations this last sentence in particular - what dos it reinforce exactly - what are the benefits?
+When comparing the optimal schedule generated by our tool to the
+manually-created schedules at St. Michael's Hospital, we found that the ILP
+formulation was always able to find an optimal schedule satisfying all required
+hard constraints, unlike the manual schedule, which often did not take all
+constraints into account. Moreover, due to the multi-goal objective function in
+the ILP, the tool was able to maximize and fulfill the majority of clinician
+preferences and requests, more so than the manually-created schedule. These
+observations reinforce the benefits of automated tools when generating schedules
+in hospital departments to balance the work-load of clinicians and improve the
+service provided to patients. The use of automated tools alleviates the time
+spent on designing the schedule by hand, and provides clinical departments with
+a more fair distribution of work that helps improve the overall satisfaction of
+both employees and patients \cite{silvestro_evaluation_2000}. \\  % how the
+%finding compares to the wider literature when automated compared to manual?
+%What does it mean re: "so what" - human error in generating tools, etc. or
+%challenges in heuristics used manually by people? include citations this last
+%sentence in particular - what dos it reinforce exactly - what are the benefits?
 
-In our simulations, we also found that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences.
-Further, we saw that the algorithm can accommodate an increase in time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance.
-One key limitation we identified was the sensitivity of the runtime to larger numbers of services offered in a single department.
-Such cases are unlikely to be encountered in the real-world because most clinical departments tend to provide a single service or at most two services by the same roster of clinicians [{\color{red}SM/Kevin, could you help with reference here?}].
-One solution to mitigate the runtime issues created by a larger number of services would be relaxing the constraints that prevent assignment of consecutive blocks, followed by manual readjustment from the generated schedule.
-%what do we mean by certain constraints?
-specify the other constraints?
-Overall, our sensitivity analyses using simulated data provide reassurance that the ILP formulation can be applied to schedule clinicians across real-world variability between clinical departments.
-%SM - will look for reference and we can ask Kevin Gough too.
-Next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services in parallel for patients and other clients.
+In our simulations, we also found that increasing the number of requests per
+clinician does not affect the runtime of the algorithm, highlighting the
+flexibility of the tool in incorporating clinician preferences. Further, we saw
+that the algorithm can accommodate an increase in time-horizon up to four years
+with little impact on runtime, which can be applied to departments that generate
+schedules far in advance. One key limitation we identified was the sensitivity
+of the runtime to larger numbers of services offered in a single department.
+Such cases are unlikely to be encountered in the real-world because most
+clinical departments tend to provide a single service or at most two services by
+the same roster of clinicians [{\color{red}SM/Kevin, could you help with
+	reference here?}]. One solution to mitigate the runtime issues created by a
+larger number of services would be relaxing the constraints that prevent
+assignment of consecutive blocks, followed by manual readjustment from the
+generated schedule. %what do we mean by certain constraints? specify the other
+%constraints?
+Overall, our sensitivity analyses using simulated data provide reassurance that
+the ILP formulation can be applied to schedule clinicians across real-world
+variability between clinical departments. %SM - will look for reference and we
+%can ask Kevin Gough too.
+Next steps include expanding the generalizability of the tool beyond smaller
+clinical departments to larger departments within and outside of health-care -
+especially those that provide multiple services in parallel for patients and
+other clients.

### 2019-09-07 paper/paper.problem.tex
Work on fixing discussion
- Remove/simplify discussion of NP-hardness
- Add more detail to impact of formulation on clinicians, hospital dept. w/citation
- Clean up ILP abbreviation throughout paper
Add hard line breaks for easier editing

@@
-At St.
-Michael's Hospital, the division of infectious diseases (ID) offers general ID and HIV consultation on inpatients as two parallel services.
-Each service provides clinical care throughout the year, during regular work weeks as well as weekends and holidays.
-The clinicians in the division typically receive a schedule in advance, outlining their on-call service dates for the full year.
-In the yearly schedule each clinician is assigned to blocks of regular work weeks and weekends.
-Each block corresponds to two consecutive work weeks.
-Apart from long (holiday) weekends, a work week starts on Monday at 8 A.M.
-and ends on Friday at 5 P.M.
-Accordingly, weekend service starts on Friday at 5 P.M., and ends on Monday at 8 A.M.
-During the weekend, ID and HIV consultation services are combined and provided by one clinician.
-During the regular work week the ID and HIV services are led by one clinician each.
-\\
+At St. Michael's Hospital, the division of infectious diseases (ID) offers
+general ID and HIV consultation on inpatients as two parallel services. Each
+service provides clinical care throughout the year, during regular work weeks as
+well as weekends and holidays. The clinicians in the division typically receive
+a schedule in advance, outlining their on-call service dates for the full year.
+In the yearly schedule each clinician is assigned to blocks of regular work
+weeks and weekends. Each block corresponds to two consecutive work weeks. Apart
+from long (holiday) weekends, a work week starts on Monday at 8 A.M. and ends on
+Friday at 5 P.M. Accordingly, weekend service starts on Friday at 5 P.M., and
+ends on Monday at 8 A.M. During the weekend, ID and HIV consultation services
+are combined and provided by one clinician. During the regular work week the ID
+and HIV services are led by one clinician each. \\
 
-Several constraints are placed on the on-call assignments.
-First, each clinician has limits on the number of blocks they can and must work during the year, depending on the type of consultation.
-For instance, a clinician might have to provide 3-5 blocks of general ID consultation as well as 2-3 blocks of HIV consultation throughout the year.
-These limits may change from year to year as the number of clinicians in the department changes.
-Moreover, the schedule should not assign a clinician to work for two consecutive blocks or two consecutive weekends in a row.
-The schedule should also distribute both regular and holiday weekends equally among all clinicians.
-\\
+Several constraints are placed on the on-call assignments. First, each clinician
+has limits on the number of blocks they can and must work during the year,
+depending on the type of consultation. For instance, a clinician might have to
+provide 3-5 blocks of general ID consultation as well as 2-3 blocks of HIV
+consultation throughout the year. These limits may change from year to year as
+the number of clinicians in the department changes. Moreover, the schedule
+should not assign a clinician to work for two consecutive blocks or two
+consecutive weekends in a row. The schedule should also distribute both regular
+and holiday weekends equally among all clinicians. \\
 
-In addition to maintaining a balanced work load among clinicians, the schedule should also accommodate their preferences.
-Clinicians provide their requests for time off ahead of schedule generation so that the requests may be integrated into the schedule.
-Individuals may specify days, weeks or weekends off, with the understanding that any blocks overlapping with their request will be assigned to a different clinician where possible.
-For example, if a clinician only requests a given Monday and Tuesday off, the schedule will generally avoid assigning the entire block to that clinician.
-Clinicians also prefer to have their weekend and block assignments close together, so the schedule should account for this when distributing assignments.
-A summary of the outlined constraints is given in Table \ref{tbl:constraint-summary}.
+In addition to maintaining a balanced work load among clinicians, the schedule
+should also accommodate their preferences. Clinicians provide their requests for
+time off ahead of schedule generation so that the requests may be integrated
+into the schedule. Individuals may specify days, weeks or weekends off, with the
+understanding that any blocks overlapping with their request will be assigned to
+a different clinician where possible. For example, if a clinician only requests
+a given Monday and Tuesday off, the schedule will generally avoid assigning the
+entire block to that clinician. Clinicians also prefer to have their weekend and
+block assignments close together, so the schedule should account for this when
+distributing assignments. A summary of the outlined constraints is given in
+Table \ref{tbl:constraint-summary}.
 
@@
 		\hline
-		\textbf{Constraint Name} & \textbf{Description}                                                                                                           & \textbf{Type} \\ \hline
-		Block Coverage (BC)           & \makecell[l]{each service needs to have exactly \\ one clinician that covers any given block}                                 & Hard          \\ \hline
-		Weekend Coverage (WC)        & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              & Hard          \\ \hline
-		Min/Max (MM)                  & \makecell[l]{for a given service, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} & Hard          \\ \hline
-		No Consecutive Blocks (NCB)   & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all services}                                    & Hard          \\ \hline
-		No Consecutive Weekends (NCW) & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        & Hard          \\ \hline
-		Equal Weekends (EW)          & \makecell[l]{weekends should be equally distributed \\ between clinicians}                                                     & Hard          \\ \hline
-		Equal Holidays (EH)          & \makecell[l]{long weekends should be equally distributed \\ between clinicians}                                                & Hard          \\ \hline
-		Block Requests (BR)         & \makecell[l]{each clinician can request to be off service \\ during certain blocks throughout the year}                        & Soft          \\ \hline
-		Weekend Requests (WR)        & \makecell[l]{each clinician can request to be off service \\ during certain weekends throughout the year}                      & Soft          \\ \hline
-		Block-Weekend Adjacency (BWA)  & \makecell[l]{the block and weekend assignments of a given \\ clinician should be adjacent}                                     & Soft
+		\textbf{Constraint Name} & \textbf{Description}                               
+		&
+		\textbf{Type} \\ \hline
+		Block Coverage (BC)           & \makecell[l]{each service needs to have
+			exactly \\ one clinician that covers any given block}                           
+		& Hard          \\ \hline
+		Weekend Coverage (WC)        & \makecell[l]{every weekend needs to have
+			exactly \\ one clinician that covers it}                                        
+		& Hard          \\ \hline
+		Min/Max (MM)                  & \makecell[l]{for a given service, each
+			clinician can only \\ work between the minimum and maximum \\ number of allowed
+			blocks} & Hard          \\ \hline
+		No Consecutive Blocks (NCB)   & \makecell[l]{any clinician should not work \\
+			two consecutive blocks, across all services}                                   
+		& Hard          \\ \hline
+		No Consecutive Weekends (NCW) & \makecell[l]{any clinician should not work two
+			consecutive \\ weekends}                                                       
+		& Hard          \\ \hline
+		Equal Weekends (EW)          & \makecell[l]{weekends should be equally
+			distributed \\ between clinicians}                                              
+		& Hard          \\ \hline
+		Equal Holidays (EH)          & \makecell[l]{long weekends should be equally
+			distributed \\ between clinicians}                                              
+		& Hard          \\ \hline
+		Block Requests (BR)         & \makecell[l]{each clinician can request to be
+			off service \\ during certain blocks throughout the year}                       
+		& Soft          \\ \hline
+		Weekend Requests (WR)        & \makecell[l]{each clinician can request to be
+			off service \\ during certain weekends throughout the year}                     
+		& Soft          \\ \hline
+		Block-Weekend Adjacency (BWA)  & \makecell[l]{the block and weekend
+			assignments of a given \\ clinician should be adjacent}                         
+		& Soft
 	\end{tabular}
@@
 
-In most scheduling problems, the constraints can be divided into hard and soft constraints.
-Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions.
-Typically, soft constraints are encoded as objective functions rather than actual constraints [ref].
-These objective functions are maximized or minimized when finding a solution.
-In the case of our clinician scheduling problem, we chose Block Requests, Weekend Requests and Block-Weekend Adjacency as our soft constraints, while the rest of the constraints are hard.
-Though it is important to take clinician requests into account when constructing the schedule, it is crucial that the work-load of the schedule is balanced among all clinicians, and the needs of the patients are fulfilled.
-It is important to note that - irrespective of how the schedule is generated - clinicians may exchange certain weeks or days throughout the year after the schedule is implemented.
-The approach to solving the clinician scheduling problem does not account for these future exchanges, and only focuses on the full time horizon.
+In most scheduling problems, the constraints can be divided into hard and soft
+constraints. Hard constraints must be satisfied by any candidate solution, while
+soft constraints can be used to select a more favourable solution from all
+candidate solutions. Typically, soft constraints are encoded as objective
+functions rather than actual constraints. These objective functions are
+maximized or minimized when finding a solution. In the case of our clinician
+scheduling problem, we chose Block Requests, Weekend Requests and Block-Weekend
+Adjacency as our soft constraints, while the rest of the constraints are hard.
+Though it is important to take clinician requests into account when constructing
+the schedule, it is crucial that the work-load of the schedule is balanced among
+all clinicians, and the needs of the patients are fulfilled. It is important to
+note that - irrespective of how the schedule is generated - clinicians may
+exchange certain weeks or days throughout the year after the schedule is
+implemented. The approach to solving the clinician scheduling problem does not
+account for these future exchanges, and only focuses on the full time horizon.

### 2019-09-07 paper/paper.discussion.tex
Fix most chktex warnings

@@
 scenarios tend to use heuristics in order to find an approximately optimal
-solution in a shorter time \cite{burke_state_2004}. \\  %cite/reference
+solution in a shorter time~\cite{burke_state_2004}. \\  %cite/reference
 %statement (i.e. the 'vast majority'...) I can't remember if the introduction
@@
 Although we restricted our use of the formulation to a minimal set of
-constraints for the particular needs of the case study (St. Michael's Hospital
+constraints for the particular needs of the case study (St.\ Michael's Hospital
 Division of Infectious Diseases) - our formulation can be adapted to various
@@
 When comparing the optimal schedule generated by our tool to the
-manually-created schedules at St. Michael's Hospital, we found that the ILP
+manually-created schedules at St.\ Michael's Hospital, we found that the ILP
 formulation was always able to find an optimal schedule satisfying all required
@@
 a more fair distribution of work that helps improve the overall satisfaction of
-both employees and patients \cite{silvestro_evaluation_2000}. \\  % how the
+both employees and patients~\cite{silvestro_evaluation_2000}. \\  % how the
 %finding compares to the wider literature when automated compared to manual?

### 2019-09-07 paper/paper.intro.tex
Fix most chktex warnings

@@
 Automated methods to optimize schedules have been studied and applied in many
-industries, including transportation \cite{aickelin_improved_2006,
-	goel_truck_2012, gunther_combined_2010}, manufacturing
-\cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008,
-	alfares_simulation_2007}, retail \cite{chapados_retail_2011,
-	nissen_automatic_2010} and military \cite{horn_scheduling_2007,
+industries, including transportation~\cite{aickelin_improved_2006,
+	goel_truck_2012, gunther_combined_2010}, manufacturing~\cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008,
+	alfares_simulation_2007}, retail~\cite{chapados_retail_2011,
+	nissen_automatic_2010} and military~\cite{horn_scheduling_2007,
 	laguna_modeling_2005}. Of special interest to a clinician on-call scheduling
@@
 wide variety of approaches, including exact and heuristic approaches, have been
-used to solve the nurse scheduling problem: integer linear programming
-\cite{azaiez_0-1_2005, trilling_nurse_2006, widyastiti_nurses_2016}, network
-flows \cite{el_adoly_new_2018}, genetic algorithms
-\cite{aickelin_exploiting_2000, jan_evolutionary_2000, kawanaka_genetic_2001},
-simulated annealing \cite{jaszkiewicz_metaheuristic_1997}, and artificial
-intelligence \cite{abdennadher_nurse_nodate, li_hybrid_2003}. A comprehensive
+used to solve the nurse scheduling problem: integer linear programming~\cite{azaiez_0-1_2005, trilling_nurse_2006, widyastiti_nurses_2016}, network
+flows~\cite{el_adoly_new_2018}, genetic algorithms~\cite{aickelin_exploiting_2000, jan_evolutionary_2000, kawanaka_genetic_2001},
+simulated annealing~\cite{jaszkiewicz_metaheuristic_1997}, and artificial
+intelligence~\cite{abdennadher_nurse_nodate, li_hybrid_2003}. A comprehensive
 literature review of these and other methods applied to nurse rostering is
-presented in \cite{burke_state_2004}. \\
+presented in~\cite{burke_state_2004}. \\
 
@@
 Moreover, the large number of variables and constraints also leads to
-computational complexity issues \cite{goos_complexity_1996}, especially when
+computational complexity issues~\cite{goos_complexity_1996}, especially when
 using exact methods for finding the solution. In this paper, we tackle a version
@@
 division, providing two different services simultaneously (general infectious
-disease (ID) consults; and HIV consults service) at St. Michael's Hospital in
+disease (ID) consults; and HIV consults service) at St.\ Michael's Hospital in
 Toronto, Canada. Our goal is to (1) present a simple integer linear programming
@@
 
-We begin by describing the problem in Section \ref{sec:problem}, and presenting
-our ILP formulation in Section \ref{sec:methods}. Next, we compare the results
+We begin by describing the problem in Section~\ref{sec:problem}, and presenting
+our ILP formulation in Section~\ref{sec:methods}. Next, we compare the results
 of our formulation to manually-created schedules, and evaluate the performance
-of the algorithm on simulated data in Section \ref{sec:results}. Finally, we
-discuss and interpret the results in Section \ref{sec:discussion}. % list the
+of the algorithm on simulated data in Section~\ref{sec:results}. Finally, we
+discuss and interpret the results in Section~\ref{sec:discussion}. % list the
 %main contents/elements of the paper.

### 2019-09-07 paper/paper.methods.tex
Fix most chktex warnings

@@
 In this section, we present an application of ILP to solve the clinician
-scheduling problem presented in Section \ref{sec:problem}. First, we describe
+scheduling problem presented in Section~\ref{sec:problem}. First, we describe
 the sets, indices and variables present in the formulation of the program. We
-then convert the constraints given in Table \ref{tbl:constraint-summary} into
+then convert the constraints given in Table~\ref{tbl:constraint-summary} into
 mathematical terms, and outline the objective function of the linear program.
 
-\subsection{Sets and Indices} \label{sec:meth-sets-indices}
+\subsection{Sets and Indices}\label{sec:meth-sets-indices}
 We denote the set of all services %or I think services is good (vs.
@@
 instance, if clinician $c$'s requests intersect with blocks 1 and 2, and weekend
-1, then $\mathcal{BR}_c = \{1, 2\}$ and $\mathcal{WR}_c = \{1\}$. Table
-\ref{tbl:sets-indices} presents a summary of the sets and indices described.
+1, then $\mathcal{BR}_c = \{1, 2\}$ and $\mathcal{WR}_c = \{1\}$. Table~\ref{tbl:sets-indices} presents a summary of the sets and indices described.
 
@@
 	\end{tabular}
-	\caption{Description of sets and indices in the problem}
-	\label{tbl:sets-indices}
+	\caption{Description of sets and indices in the problem}\label{tbl:sets-indices}
 \end{table}
 
-\subsection{Variables} \label{sec:meth-variables}
+\subsection{Variables}\label{sec:meth-variables}
 Since each clinician may be assigned to work on any service, during any block of
@@
 $m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks
-each clinician is allowed to work during the year. Table
-\ref{tbl:variables-constants} presents a summary of the constants and variables
+each clinician is allowed to work during the year. Table~\ref{tbl:variables-constants} presents a summary of the constants and variables
 in the problem.
@@
 	\end{tabular}
-	\caption{Description of variables and constants in the problem}
-	\label{tbl:variables-constants}
+	\caption{Description of variables and constants in the problem}\label{tbl:variables-constants}
 \end{table}
 
-\subsection{Constraints} \label{sec:meth-constraints}
-We now formalize the hard constraints in Table \ref{tbl:constraint-summary}
+\subsection{Constraints}\label{sec:meth-constraints}
+We now formalize the hard constraints in Table~\ref{tbl:constraint-summary}
 using the variables defined above.
@@
 
-\subsection{Objectives} \label{sec:meth-objectives}
-As described in Section \ref{sec:problem}, the soft constraints of the clinician
+\subsection{Objectives}\label{sec:meth-objectives}
+As described in Section~\ref{sec:problem}, the soft constraints of the clinician
 scheduling problem include: satisfying clinician block off requests (BR),
@@
 blocks (BWA). We convert these soft constraints into linear objective functions
-of the binary variables defined in Section \ref{sec:meth-variables}. Objectives
+of the binary variables defined in Section~\ref{sec:meth-variables}. Objectives
 (\ref{eqn:obj-block-requests}) and (\ref{eqn:obj-weekend-requests}) can be
@@
 &Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S}
-(-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{BR}
+{(-1)}^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{BR}
 \label{eqn:obj-block-requests}\\
 &Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W}
-(-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{WR}
+{(-1)}^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{WR}
 \label{eqn:obj-weekend-requests}
@@
 functions into linear objectives involves introducing a helper variable and
-additional constraints \cite{hammer_boolean_1968}. In our case, introducing a
+additional constraints~\cite{hammer_boolean_1968}. In our case, introducing a
 variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w =
@@
 where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha,
-\beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal
-\cite{stanimirovic_linear_2011}. \\
+\beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal~\cite{stanimirovic_linear_2011}. \\
 

### 2019-09-07 paper/paper.problem.tex
Fix most chktex warnings

@@
-At St. Michael's Hospital, the division of infectious diseases (ID) offers
+At St.\ Michael's Hospital, the division of infectious diseases (ID) offers
 general ID and HIV consultation on inpatients as two parallel services. Each
@@
 distributing assignments. A summary of the outlined constraints is given in
-Table \ref{tbl:constraint-summary}.
+Table~\ref{tbl:constraint-summary}.
 
@@
 	\end{tabular}
-	\caption{Summary of the constraints for the clinician scheduling problem}
-	\label{tbl:constraint-summary}
+	\caption{Summary of the constraints for the clinician scheduling problem}\label{tbl:constraint-summary}
 \end{table}

### 2019-09-07 paper/paper.results.tex
Fix most chktex warnings

@@
 linear program and allows configuration of clinicians at [ref \ref{???}], to be
-used by the ID division at St. Michael's Hospital. The software was used to
+used by the ID division at St.\ Michael's Hospital. The software was used to
 generate the results in the following sections, using real data as well as
@@
 The software was implemented using the COIN-OR Branch-and-Cut open source solver
-version 2.9.9 \cite{johnjforrest_coin-or/cbc:_2019}.
+version 2.9.9~\cite{johnjforrest_coin-or/cbc:_2019}.
 
@@
 We used clinician time-off requests and minimum/maximum requirements from
-2015-2018 as input data for the ILP problem. Table
-\ref{tbl:2018-schedule-comparison} presents the optimal schedule generated using
+2015-2018 as input data for the ILP problem. Table~\ref{tbl:2018-schedule-comparison} presents the optimal schedule generated using
 the software as well as the manually-created schedule for data from 2018. The
@@
 manually-created schedule. Specifically, we examined the adherence of each
-schedule to the constraints presented in Table \ref{tbl:constraint-summary}. As
-shown in Table \ref{tbl:constraints-comparison}, the optimal schedule generated
+schedule to the constraints presented in Table~\ref{tbl:constraint-summary}. As
+shown in Table~\ref{tbl:constraints-comparison}, the optimal schedule generated
 using the software satisfied all mandatory constraints. In contrast, the
@@
 The effect of an increasing number of requests per clinician on the runtime of
-the ILP solver is shown in Figure \ref{fig:runtime-requests}. We executed the
+the ILP solver is shown in Figure~\ref{fig:runtime-requests}. We executed the
 program on a department with 10 clinicians offering two services, similar to the
-department at St. Michael's Hospital. For simplicity, each clinician was
+department at St.\ Michael's Hospital. For simplicity, each clinician was
 configured with $\abs{\mathcal{BR}} = \{0, 5, 10, 15\}$ non-overlapping block
@@
 	\caption{Runtime of ILP solver with an increasing number of requests per
-		clinician}
-	\label{fig:runtime-requests}
+		clinician}\label{fig:runtime-requests}
 \end{figure}
 
-Figure \ref{fig:runtime-blocks} presents the change in runtime when increasing
+Figure~\ref{fig:runtime-blocks} presents the change in runtime when increasing
 the number of 2-week blocks $B = \{26, 52, 78, 104\}$ in a department with 10
@@
 	\includegraphics[scale=.5]{fig/runtime_blocks}
-	\caption{Runtime of ILP solver with an increasing number of 2-week blocks}
-	\label{fig:runtime-blocks}
+	\caption{Runtime of ILP solver with an increasing number of 2-week blocks}\label{fig:runtime-blocks}
 \end{figure}
@@
 %services?
-on the runtime of the program is shown in Table
-\ref{tbl:runtime-services-clinicians-comparison}. We executed the algorithm for
+on the runtime of the program is shown in Table~\ref{tbl:runtime-services-clinicians-comparison}. We executed the algorithm for
 $S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in

### 2019-09-13 paper/paper.problem.tex
paragraph spacing
tables & figures cleanup

@@
 are combined and provided by one clinician. During the regular work week the ID
-and HIV services are led by one clinician each. \\
+and HIV services are led by one clinician each.
 
@@
 consecutive weekends in a row. The schedule should also distribute both regular
-and holiday weekends equally among all clinicians. \\
+and holiday weekends equally among all clinicians.
 
@@
 	\centering
+	\caption{Summary of the constraints for the clinician scheduling problem}%
+  \label{tbl:constraint-summary}
+  % JK: table caption should go at the top.
+  %     The spacing can be funny so I use:
+  %     \usepackage{caption}\captionsetup[table]{skip=1em}
+  %     also labels should be on their own line.
+  % JK: below: I would consider making the 2-3 letter abbreviations its own column
 	\begin{tabular}{ l l l }
-		\hline
+		\toprule % JK: different line weights support table structure
 		\textbf{Constraint Name} & \textbf{Description}                               
 		&
-		\textbf{Type} \\ \hline
+		\textbf{Type} \\ \midrule
 		Block Coverage (BC)           & \makecell[l]{each service needs to have
@@
 		& Soft
+    \\\bottomrule % JK: I think usually table bottom also has a rule
 	\end{tabular}
-	\caption{Summary of the constraints for the clinician scheduling problem}\label{tbl:constraint-summary}
+  % JK: please define Hard & Soft constraints in a footnote
 \end{table}

### 2019-09-13 paper/paper.results.tex
paragraph spacing
tables & figures cleanup

@@
 schedule is color-coded to distinguish between the different clinicians
-assigned. \\
+assigned.
 
-\include{tbl/2018_schedule_comparison}
+\input{tbl/2018_schedule_comparison}
+% JK: I think use input not include here to avoid unnecessary page breaks
+%     https://tex.stackexchange.com/questions/246
 
@@
 schedule in all four years, by accommodating almost all time-off requests and
-ensuring that weekends are always assigned close to blocks. \\
+ensuring that weekends are always assigned close to blocks.
 
-\include{tbl/constraints_comparison}
+\input{tbl/constraints_comparison}
 
@@
 in the division; amount of time-off requests per clinician throughout the year.
-We also examined the effect of a longer time-horizon on the runtime. \\
+We also examined the effect of a longer time-horizon on the runtime.
 
@@
 clinician requests. Moreover, we see that all four runs were completed in under
-1 second. \\
+1 second.
 
@@
 	\centering
-	\includegraphics[scale=.5]{fig/runtime_requests}
-	\caption{Runtime of ILP solver with an increasing number of requests per
-		clinician}\label{fig:runtime-requests}
+	\includegraphics[scale=.5]{fig/runtime_requests} % JK: can you export this to PDF or SVG to avoid pixelation?
+	\caption{Runtime of ILP solver with an increasing number of requests per clinician}
+  \label{fig:runtime-requests}
+  % JK: The trend here is a bit confusing, since I think you're trying to say
+  %     that the runtime is not affected by the number of constraints,
+  %     but there is noise in these results.
+  %     I would re-run this 10-100 times and average the results so that (I expect)
+  %     the mean trend is actually flat. You could (should) then also present the runtimes
+  %     as box-plots to show that the trend is less than the variability.
 \end{figure}
@@
 solver was able to find an optimal schedule for all time horizons within 3
-seconds. \\ % 'very reasonable amount of time" is vague. what is reasonable?
+seconds. % 'very reasonable amount of time" is vague. what is reasonable?
 %better to give a range in value.
@@
 	\includegraphics[scale=.5]{fig/runtime_blocks}
-	\caption{Runtime of ILP solver with an increasing number of 2-week blocks}\label{fig:runtime-blocks}
+	\caption{Runtime of ILP solver with an increasing number of 2-week blocks}%
+  \label{fig:runtime-blocks}
+  % JK: same comment as above re. trends -- are you trying to show one?
+  % JK: Thess figures could also possible be Figure 1. a) and b) 
 \end{figure}
@@
 services, and we were able to generate a schedule with upwards of 50 clinicians
-in under 1.5 seconds. \\  %what does division mean? do you mean services?
+in under 1.5 seconds.  %what does division mean? do you mean services?
 
-\include{tbl/runtime_services_clinicians}
+\input{tbl/runtime_services_clinicians}

### 2019-09-13 paper/paper.methods.tex
paragraph spacing
tables & figures cleanup

@@
 	\centering
+  \caption{Description of sets and indices in the problem}%
+  \label{tbl:sets-indices}
 	\begin{tabular}{ c c l }
-		\hline
+		\toprule
 		\textbf{Set}                         & \textbf{Index} & \textbf{Description}  
-		\\ \hline
+		\\ \midrule
 		$\mathcal{S} = \{1, \ldots, S \}$    & $s$            & services              
@@
 		$\mathcal{WR}_c \subset \mathcal{W}$ &                & weekend requests of
-		clinician $c$
+		clinician $c$ \\
+    \bottomrule
 	\end{tabular}
-	\caption{Description of sets and indices in the problem}\label{tbl:sets-indices}
+
 \end{table}
@@
 	\centering
+  \caption{Description of variables and constants in the problem}%
+  \label{tbl:variables-constants}
 	\begin{tabular}{ c l }
-		\hline
+		\toprule
 		\textbf{Name}              & \textbf{Description}                             
-		\\ \hline
+		\\ \midrule
 		$X_{c, b, s} \in \{0, 1\}$ & assignment of clinician $c$ to service $s$ on
@@
 		$M_{c, s}$                 & maximum number of blocks clinician $c$ should
-		cover on service $s$
+		cover on service $s$ \\
+    \bottomrule
 	\end{tabular}
-	\caption{Description of variables and constants in the problem}\label{tbl:variables-constants}
 \end{table}
@@
 that conflict with a block or weekend request, and aim to maximize the
-non-conflicting assignments. \\
+non-conflicting assignments.
 
@@
 be assigned during a weekend that falls within an assigned block. In this case,
-we will have $\varphi(b) = 2b-1$. \\
+we will have $\varphi(b) = 2b-1$.
 
@@
 (respectively, 0), $Z_{c, b, s}$ can attain a maximum value of 1 (respectively,
-0), giving us the correct adjacency maximization objective. \\
+0), giving us the correct adjacency maximization objective.
 %Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a
@@
 where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha,
-\beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal~\cite{stanimirovic_linear_2011}. \\
+\beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal~\cite{stanimirovic_linear_2011}.
 

### 2019-09-13 paper/tbl/runtime_services_clinicians.tex
Fix most chktex warnings
tables & figures cleanup

@@
 	\centering
+ 	\caption{Comparison of program runtime (in seconds) for different numbers of services and total clinicians in the division.}%
+  \label{tbl:runtime-services-clinicians-comparison}%
 	\begin{tabular}{|c|c||c|c||c|c|}
-		\hline
-		                                      &  \multicolumn{5}{c|}{Number of Services}  \\ \hline
-		\makecell[l]{Number of \\ Clinicians} &  1   &    2    & 2 (without NCB) &  3   & 3 (without NCB) \\ \hline
+		\toprule
+		                                      &  \multicolumn{5}{c|}{Number of Services}  \\ \midrule
+		\makecell[l]{Number of \\ Clinicians} &  1   &    2    & 2 (without NCB) &  3   & 3 (without NCB) \\ \midrule
 		                 10                   & 0.16 &  0.74   &  0.16   & 1.40 &  0.23   \\ \hline
-		                 20                   & 0.25 & 7468.86 &  0.32   &  -   &  0.43   \\ \hline
-		                 30                   & 0.42 &    -    &  0.49   &  -   &  0.66   \\ \hline
-		                 50                   & 0.62 &    -    &  0.82   &  -   &  1.14   \\ \hline
-	\end{tabular}
-	\caption{Comparison of program runtime (in seconds) for different number of services and total clinicians in the division.
-`-' indicates that a solution was not found within 24 hours.
-`without NCB' indicates that the No Consecutive Blocks (NCB) constraint was relaxed.}
-	\label{tbl:runtime-services-clinicians-comparison}%
+		                 20                   & 0.25 & 7468.86 &  0.32   &  --  &  0.43   \\ \hline
+		                 30                   & 0.42 &    --   &  0.49   &  --  &  0.66   \\ \hline
+		                 50                   & 0.62 &    --   &  0.82   &  --  &  1.14   \\ \bottomrule
+	\end{tabular}\\[1em]
+  \footnotesize\raggedright
+  Notes:
+  ``--'' indicates that no solution was found within 24 hours;
+  ``without NCB'' indicates that the No Consecutive Blocks (NCB) constraint was relaxed.
 \end{table}

### 2019-09-13 paper/paper.intro.tex
paragraph spacing
intro edits

@@
 demands, often exceed the available supply. In particular, on-call schedules for
+% JK: Not sure I follow this sentence.
+%     I think we dont need to say demands > needs, can just say efficieny improves care.
 a fixed number of health-care providers are central to the efficient running of
-hospitals. Carefully allocated schedules are meant to simultaneously ensure
-sufficient resources are provided to patients while not overworking clinicians
-to prevent costly mistakes. It is common practice for on-call schedules in
-hospitals to be created manually, yet manually-created schedules are prone to
-errors and potential for various biases. First, when there is a large number of
-clinicians in a single department, or the constraints that need to be satisfied
-by the department are very complex, a manual method may not provide a schedule
-that satisfies all clinicians. Second, such methods are likely to overlook
+hospitals. Carefully allocated schedules should balance
+sufficient staff with clinician workload,
+to maximize quality of care. It is common for on-call schedules in
+hospitals to be created manually. However,
+manually-created schedules are subject to three problems.
+First, when there are a large number of clinicians,
+or the constraints that need to be satisfied
+by the schedule are complex, a manual method may not provide a schedule
+that satisfies all constraints. Second, manual methods may overlook
 certain constraints that must be maintained to have an operational department,
@@
 clinicians are allocated a specific amount of work blocks throughout the year.
-Third, manual scheduling is often time-consuming for the person developing the
-schedule. For these reasons, it is important to develop automated methods that
-can generate optimal schedules that satisfy the given constraints of the
-hospital department. \\
+% JK: Can you clarify why / how these may be overlooked?
+Third, manual scheduling is often time-consuming.
+For these reasons, it is important to develop automated methods that
+can efficiently generate schedules that satisfy the given constraints.
+% JK: haven't introduced what "optimal" means yet so I think maybe leave it out for now.
 %For example, it is important that a hospital department allocates its
@@
 %optimally, to ensure the best possible service for its patients.
+% JK: instead of using \\ everywhere for paragraph breaks, use:
+%     \setlength{\parskip}{1em}
+%     in the preamble
 
-Automated methods to optimize schedules have been studied and applied in many
-industries, including transportation~\cite{aickelin_improved_2006,
-	goel_truck_2012, gunther_combined_2010}, manufacturing~\cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008,
-	alfares_simulation_2007}, retail~\cite{chapados_retail_2011,
-	nissen_automatic_2010} and military~\cite{horn_scheduling_2007,
-	laguna_modeling_2005}. Of special interest to a clinician on-call scheduling
-problem are the approaches to schedule nurses, who often work in shifts. In the
+Automated methods to generate schedules have been studied and applied in many
+industries, including
+% JK: I find this is a nice way to organize these lists of topics & citations
+transportation~\cite{aickelin_improved_2006, goel_truck_2012, gunther_combined_2010},
+manufacturing~\cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008, alfares_simulation_2007},
+retail~\cite{chapados_retail_2011, nissen_automatic_2010}, and
+military~\cite{horn_scheduling_2007, laguna_modeling_2005}.
+Of special interest to a clinician on-call scheduling
+problem are the approaches to scheduling nurses, who often work in shifts. In the
 nurse scheduling problem, the goal is to find an optimal assignment of nurses to
@@
 and as many soft constraints as possible, which may include nurse preferences. A
+% JK: Is this specific to the nurse scheduling problem?
+%     This sounds like the more general definition of the scheduling problem.
+%     Can you define hard and soft constraints here and give
+%     one example of a hard constraint and one example of a soft constraint.
 wide variety of approaches, including exact and heuristic approaches, have been
-used to solve the nurse scheduling problem: integer linear programming~\cite{azaiez_0-1_2005, trilling_nurse_2006, widyastiti_nurses_2016}, network
-flows~\cite{el_adoly_new_2018}, genetic algorithms~\cite{aickelin_exploiting_2000, jan_evolutionary_2000, kawanaka_genetic_2001},
-simulated annealing~\cite{jaszkiewicz_metaheuristic_1997}, and artificial
-intelligence~\cite{abdennadher_nurse_nodate, li_hybrid_2003}. A comprehensive
-literature review of these and other methods applied to nurse rostering is
-presented in~\cite{burke_state_2004}. \\
+used to solve the nurse scheduling problem:
+integer linear programming~\cite{azaiez_0-1_2005, trilling_nurse_2006, widyastiti_nurses_2016},
+network flows~\cite{el_adoly_new_2018},
+genetic algorithms~\cite{aickelin_exploiting_2000, jan_evolutionary_2000, kawanaka_genetic_2001},
+simulated annealing~\cite{jaszkiewicz_metaheuristic_1997}, and
+artificial intelligence~\cite{abdennadher_nurse_nodate, li_hybrid_2003}.
+A comprehensive
+literature review of these and other methods applied to nurse scheduling is
+presented in~\cite{burke_state_2004}.
+% JK: rostering = scheduling? I think okay to re-use "scheduling" unless its getting extremely repetitive.
+%     but even then I think SM would agree clarity > flow/repetition.
 
@@
 Many of the approaches to nurse scheduling were designed to satisfy the
-requirements of a specific hospital department which causes a large number of
+requirements of a specific hospital department, which results in a large number of
 variables and constraints to be incorporated into the problem formulation. While
 these department-specific approaches allow end-users to find precise schedules
-that satisfy the needs of the department and the preferences of the nurses and
-clinicians in that department, they are difficult to readily adapt to other
-departments in the same hospital or other hospitals. % explain why hard to
+that satisfy the needs of that department and the preferences of nurses and
+clinicians in that department, they are difficult to adapt to other
+departments. % explain why hard to
 %adapt?... what makes their generalizablility/adaptability limited?
+% JK: I'm less bothered by this lack of explaination, since I know its
+%     hard to explain without getting into the nitty-gritty.
 Moreover, the large number of variables and constraints also leads to
 computational complexity issues~\cite{goos_complexity_1996}, especially when
-using exact methods for finding the solution. In this paper, we tackle a version
+% JK: I think good to clarify that these complextiy issues would imply that
+%     the problem is essentially impossible to solve in a limited amount of time.
+trying to find the most optimal solution. In this paper, we tackle a version
 of the nurse scheduling problem arising from a case study of one clinical
-division, providing two different services simultaneously (general infectious
-disease (ID) consults; and HIV consults service) at St.\ Michael's Hospital in
-Toronto, Canada. Our goal is to (1) present a simple integer linear programming
-(ILP) formulation for the scheduling problem as developed for the hospital, and
-describe the adaptability of the formulation to solving similar problems in
-other departments; (2) compare the performance of the ILP scheduler to the
-results of the manual approach; and (3) analyze the robustness of the algorithm
-in difficult instances of the problem. \\
+division, providing two different services (general infectious
+disease (ID) consults; and HIV consults) at St.\ Michael's Hospital in
+Toronto, Canada. Our goals are to (1) present a integer linear programming
+(ILP) formulation for the scheduling problem, and
+% JK: the nurse scheduling problem?
+%     I think we need to define the "scheduling problem"
+%     and how / if it differs from other scheduling problem.
+describe the flexibility of this formulation for solving similar problems;
+(2) compare the performance of an ILP-based scheduling tool to the
+results of a manual approach; and (3) analyze the robustness of this approach
+in difficult instances of the problem.
+% JK: what do you mean by different instances?  Can you be more specific?
+%     The results explore:
+%     (1) runtime vs number of requests
+%     (2) runtime vs number of blocks
+%     (3) satisfaction of hard and soft constraints.
 % present a simple formulation for the problem developed and tested at the
@@
 
-We begin by describing the problem in Section~\ref{sec:problem}, and presenting
+We begin by describing the details of the problem in Section~\ref{sec:problem}, and presenting
 our ILP formulation in Section~\ref{sec:methods}. Next, we compare the results

### 2019-09-14 paper/paper.problem.tex
editing problem

@@
 At St.\ Michael's Hospital, the division of infectious diseases (ID) offers
-general ID and HIV consultation on inpatients as two parallel services. Each
-service provides clinical care throughout the year, during regular work weeks as
-well as weekends and holidays. The clinicians in the division typically receive
-a schedule in advance, outlining their on-call service dates for the full year.
-In the yearly schedule each clinician is assigned to blocks of regular work
-weeks and weekends. Each block corresponds to two consecutive work weeks. Apart
+separate but concurrent services for general ID and HIV consultation. Each
+service provides clinical care throughout the year, during regular working hours
+and on weekends and holidays. The schedule is created in advance,
+outlining all work-week and weekend shifts for the full year.
+In the yearly schedule clinicians are assigned to ``blocks'' of
+two consecutive regular work weeks and individual weekends. Apart
+% JK: when I first read this I thought weekends were also in 2-week blocks
 from long (holiday) weekends, a work week starts on Monday at 8 A.M. and ends on
@@
 
-Several constraints are placed on the on-call assignments. First, each clinician
-has limits on the number of blocks they can and must work during the year,
-depending on the type of consultation. For instance, a clinician might have to
-provide 3-5 blocks of general ID consultation as well as 2-3 blocks of HIV
-consultation throughout the year. These limits may change from year to year as
-the number of clinicians in the department changes. Moreover, the schedule
+Two types of constraints are placed on the clinician assignments. 
+% JK: I think you need to introduce BC and WC constraints here,
+%     since they appear in Table 1 without text in the body.
+First, each clinician
+must work between a minimum and maximum number of blocks of each service during the year.
+For instance, one clinician might have to
+provide 3-5 blocks of general ID service and 2-3 blocks of HIV service
+% JK: trying to reduce synonyms "service" / "consultation"
+throughout the year. These limits may be different fo each clinician,
+and they may change from year to year as
+the number of clinicians in the department changes. Second, the schedule
 should not assign a clinician to work for two consecutive blocks or two
-consecutive weekends in a row. The schedule should also distribute both regular
-and holiday weekends equally among all clinicians.
+consecutive weekends. The schedule should also distribute regular
+weekends and holiday weekends each equally among all clinicians.
 
-In addition to maintaining a balanced work load among clinicians, the schedule
-should also accommodate their preferences. Clinicians provide their requests for
-time off ahead of schedule generation so that the requests may be integrated
-into the schedule. Individuals may specify days, weeks or weekends off, with the
+In addition to balancing the work load among clinicians, the schedule
+should accommodate their preferences. Clinicians provide their requests for
+time off before schedule generation so that the requests may be integrated
+into the schedule. Clinician may specify days, weeks or weekends off, with the
 understanding that any blocks overlapping with their request will be assigned to
 a different clinician where possible. For example, if a clinician only requests
-a given Monday and Tuesday off, the schedule will generally avoid assigning the
-entire block to that clinician. Clinicians also prefer to have their weekend and
-block assignments close together, so the schedule should account for this when
-distributing assignments. A summary of the outlined constraints is given in
+a given Monday and Tuesday off, the schedule should avoid assigning the
+entire block to that clinician. Clinicians also typically prefer to have their weekend and
+block assignments side by side, so the schedule should accommodate this where possible.
+A summary of the outlined constraints is given in
 Table~\ref{tbl:constraint-summary}.
@@
 constraints. Hard constraints must be satisfied by any candidate solution, while
-soft constraints can be used to select a more favourable solution from all
-candidate solutions. Typically, soft constraints are encoded as objective
-functions rather than actual constraints. These objective functions are
-maximized or minimized when finding a solution. In the case of our clinician
-scheduling problem, we chose Block Requests, Weekend Requests and Block-Weekend
+soft constraints can be used to select a more favourable solution from a set of
+candidate solutions. Typically, soft constraints are therefore encoded as objective
+functions which are
+maximized when finding a solution. In the case of the clinician
+% JK: the direction of max / minimization is arbitrary,
+%     I think sufficient to pick one here to avoid confusion.
+scheduling problem described above, we chose Block Requests, Weekend Requests and Block-Weekend
 Adjacency as our soft constraints, while the rest of the constraints are hard.
-Though it is important to take clinician requests into account when constructing
+While it is important to take clinician requests into account when constructing
 the schedule, it is crucial that the work-load of the schedule is balanced among
-all clinicians, and the needs of the patients are fulfilled. It is important to
-note that - irrespective of how the schedule is generated - clinicians may
-exchange certain weeks or days throughout the year after the schedule is
-implemented. The approach to solving the clinician scheduling problem does not
-account for these future exchanges, and only focuses on the full time horizon.
+all clinicians, and all services are assigned.
+After the schedule is generated, clinicians may
+exchange certain weeks or days throughout the year.
+Our approach to schedule generation does not
+consider such future exchanges, and instead generates on one full year at a time.
+% JK: I think this para (4) should go instead as the 2nd in this section.
+%     you could introduce the ideas of hard and soft constraints (this para) first,
+%     and then have the current paras (2) and (3) after this one, where the reader
+%     is already aware that (2) is hard constraints and (3) is soft.
+%     Oh, I see that they're not currently split up exactly, but I think it would be
+%     helpful to do so. Maybe double check with SM though.

### 2019-09-15 paper/paper.methods.tex
editing methods

@@
-In this section, we present an application of ILP to solve the clinician
+In this section, we present an application of integer linear programming to solve the clinician
+% JK: there are a few variations of
+%     "nurse scheduling problem", "scheduling problem" "clinician scheduling problem"
+%     floating around. Can you do a consistency check throughout?
 scheduling problem presented in Section~\ref{sec:problem}. First, we describe
-the sets, indices and variables present in the formulation of the program. We
-then convert the constraints given in Table~\ref{tbl:constraint-summary} into
-mathematical terms, and outline the objective function of the linear program.
+% JK: Also, while the intro gives a reference for ILP,
+%     I think you should define it here. e.g. It is not clear to the naive reader (me!)
+%     that ILP needs an objective function
+%     or (most importantly) how ILP works to satisfy multiple constraints.
+the sets, indices and variables necessary for the formulation of the problem. We
+then write the constraints given in Table~\ref{tbl:constraint-summary} as
+mathematical equations, and define the objective function of the linear program.
 
@@
 %problem statement
+% JK: same thing, consistency check with division / service
+%     I think maybe you already did this...
 that clinicians in a single department can provide as $\mathcal{S}$. The
 following formulation assumes that all clinicians are able to provide all
+% JK: but could you not also just set some clinician-service max-mins zero
+%     to deal with if some clinicians are not able to provide some services?
 services. The set of all clinicians in the department is denoted as
-$\mathcal{C}$. The sets of blocks and weekends that clinicians will be assigned
-to are denoted as $\mathcal{B}$ and $\mathcal{W}$ respectively. The block size
+$\mathcal{C}$. The sets of blocks and weekends
+% JK: removed "will be assigned to" since on first read I confused it with
+%     the binary assignment variable X below.
+are denoted as $\mathcal{B}$ and $\mathcal{W}$ respectively. The block size
 used in our experiments is 2 weeks, but the following LP formulation does not
-specify a particular size for blocks, and so it can be adapted to the needs of
-the department in question.
+require a particular size for blocks, and so it can be adapted for other cases.
 %The size of a block is not constrained and can be adapted %how?
@@
 long/holiday weekends such as the Canadian Civic Day weekend in August. Lastly,
-we denote with $\mathcal{BR}_c$ and $\mathcal{WR}_c$ the block and weekend
-requests of clinicians as subsets of all blocks and weekends, respectively. For
-instance, if clinician $c$'s requests intersect with blocks 1 and 2, and weekend
+% JK: I'm being picky, but maybe a non-Canadian holiday like Thanksgiving as the e.g.
+the block and weekend request-off dates of clinicians as are denoted as the following
+subsets of all blocks and weekends: $\mathcal{BR}_c$ and $\mathcal{WR}_c$, respectively.
+% JK: keeping "x is denoted as x" sentence structure
+% JK: Also, I'm generally not a fan of multiple-letter variables since they can be 
+%     easily confused for multiplication. Can you find another notation?
+For instance, if clinician $c$'s request-off dates intersect with blocks 1 and 2, and weekend
 1, then $\mathcal{BR}_c = \{1, 2\}$ and $\mathcal{WR}_c = \{1\}$. Table~\ref{tbl:sets-indices} presents a summary of the sets and indices described.
@@
 Since each clinician may be assigned to work on any service, during any block of
-the year, we denote such an assignment as a binary variable $X_{c, b, s}$. A
-value of 1 indicates that the given clinician $c$ is assigned to provide service
-$s$ during block $b$. Weekend assignments are similarly defined using a binary
-variable $Y_{c, w}$, but without a service index, as clinicians are expected to
-provide all services during the weekends. We then introduce a set of constants
-$m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks
-each clinician is allowed to work during the year. Table~\ref{tbl:variables-constants} presents a summary of the constants and variables
+the year, we denote such an assignment as a binary variable $X_{c, b, s} \in \{0,1\}$.
+% JK: I know its implied but if its easy to spell it out might as well.
+A value of 1 indicates that clinician $c$ is assigned to service $s$ during block $b$,
+while a value of 0 indicates they are not assigned.
+Weekend assignments are similarly defined using a binary
+variable $Y_{c, w} \in \{0,1\}$, but without a service index, as clinicians are expected to
+provide all services during the weekends. We then define
+$m_{c, s}$ and $M_{c, s}$ to represent the minimal and maximal number of blocks of service $s$
+that clinician $c$ is required to work during the year.
+Table~\ref{tbl:variables-constants} presents a summary of the constants and variables
 in the problem.
@@
 
+% JK: not sure how I feel about these letters as equation refs
 \begin{align}
@@
 0), giving us the correct adjacency maximization objective.
+% JK: not sure I follow these (respectively, 0).
 %Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a
@@
 of the normalized objective functions,
+% JK: This only considers the soft constraints.
+%     Should this maximization not also be subject to the Constraints in 3.3?
 \begin{equation}
@@
 \end{equation}
+% JK: I would use $\alpha_1, \alpha_2, \alpha_3$ here which could take any value,
+%     and then not necessary to have any bounds like \sum_i^3 \alpha_i = 1,
+%     though you could enforce it to compare across permutations of \alpha
 where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha,
 \beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal~\cite{stanimirovic_linear_2011}.
+% JK: How can ILP be used solve such problems? Can you briefly describe?
 

### 2019-09-15 paper/paper.results.tex
editing results

@@
-\subsection{Software}
+% JK: I think give an overview here of the experiments here
+%     and what you're going to show in this section (and why)
+\subsection{Implementation}
 We developed a software package with a user interface that implements the above
 linear program and allows configuration of clinicians at [ref \ref{???}], to be
+% JK: what is this ref? Do you want to mention it is in Python?
 used by the ID division at St.\ Michael's Hospital. The software was used to
@@
 The software was implemented using the COIN-OR Branch-and-Cut open source solver
+% JK: "software" here is not clear to me, is it the solver for the ILP formulation?
 version 2.9.9~\cite{johnjforrest_coin-or/cbc:_2019}.
 
-\subsection{Infectious Diseases Division}  %main comment  = I think use past
+\subsection{Comparison with Manually Generated Schedules}  %main comment  = I think use past
 %tense in writing
 We used clinician time-off requests and minimum/maximum requirements from
-2015-2018 as input data for the ILP problem. Table~\ref{tbl:2018-schedule-comparison} presents the optimal schedule generated using
-the software as well as the manually-created schedule for data from 2018. The
-schedule is color-coded to distinguish between the different clinicians
-assigned.
+2015-2018 as input data for the ILP problem.
+Table~\ref{tbl:2018-schedule-comparison} compares the optimal schedule generated using
+the software with the manually-created schedule for data from 2018. The
+schedule is color-coded to distinguish between the different clinicians.
 
@@
 
-First, we evaluated the software-generated schedule by comparing it with the
+First, we evaluated the automatically-generated schedule by comparing it with the
+% JK: I think it would be good to give a short name for manual and the proposed
+%     data -- e.g. LP Solution as in Table 4 and Manual Generation.
+%     Then check for consistency throughout.
+%     While lack of synonyms can sound repetitive, it can help avoid confusion.
 manually-created schedule. Specifically, we examined the adherence of each
@@
 using the software satisfied all mandatory constraints. In contrast, the
-manually-created schedule was not able to satisfy all mandatory constraints. In
+% JK: "mandatory" = "hard"? if possible just use one.
+manually-created schedule did not satisfy all mandatory constraints. In
 particular, we see that the manual schedule assigned clinicians to multiple
@@
 an equal distribution of weekends and holidays for all four years of data.
-Evaluating the objectives, we see that LP outperforms the manually created
+Considering all objectives, we see that LP outperforms the manually created
 schedule in all four years, by accommodating almost all time-off requests and
@@
 \input{tbl/constraints_comparison}
+% JK: If you want to be funny, you can estimate the approximate time to
+%     generate the schedule and include that too (16 hours vs 1 second)...
 
-\subsection{Simulations}
-We then examined the properties of the ILP problem using simulated data.
-Specifically, we examined the influence of expanding the following parameters on
-the runtime of the algorithm: number of clinicians; number of services offered
-in the division; amount of time-off requests per clinician throughout the year.
-We also examined the effect of a longer time-horizon on the runtime.
+\subsection{Influence of Problem Complexity of Runtime}
+Next, we examined the influence of the following four parameters on the
+runtime of the ILP solver using simulated data:
+number of clinicians;
+number of services offered;
+number of time-off requests per clinician per year;
+time-horizon of the schedule.
 
+% JK: Good to introduce the detault parameters first, then for each parameter explored,
+%     Describe what varied.
+We simulated a department with 10 clinicians offering two services, similar to the
+department at St.\ Michael's Hospital. Each clinician was
+configured with $[0, 5, 10, 15]$ non-overlapping block
+% JK: \abs{BR} does not clearly imply number of blocks (could be confused with week indices),
+%     so just leave out the LHS here I think.
+% JK: What do you mean by non-overlapping? And why only increments of 5?
+requests.
 The effect of an increasing number of requests per clinician on the runtime of
-the ILP solver is shown in Figure~\ref{fig:runtime-requests}. We executed the
-program on a department with 10 clinicians offering two services, similar to the
-department at St.\ Michael's Hospital. For simplicity, each clinician was
-configured with $\abs{\mathcal{BR}} = \{0, 5, 10, 15\}$ non-overlapping block
-requests. The increase in the number of requests did not affect the runtime of
+the ILP solver is shown in Figure~\ref{fig:runtime-requests}.
+The increase in the number of requests did not affect the runtime of
 the ILP solver, indicating that it can accommodate a lot of flexibility in
@@
 1 second.
+% JK: See somments below about trends and repeating simulations.
 
@@
 
+% JK: Introduce what you did first and then the figure showing results (see above)
+% JK: Try to introduce these results in the same order as they appear in the
+%     first paragraph of this section.
 Figure~\ref{fig:runtime-blocks} presents the change in runtime when increasing
@@
 %better to give a range in value.
+% JK: same comment as above re. trends
 
@@
   \label{fig:runtime-blocks}
-  % JK: same comment as above re. trends -- are you trying to show one?
+  % JK: same comment as above re. trends
   % JK: Thess figures could also possible be Figure 1. a) and b) 
@@
 
+% JK: Introduce what you did first and then the figure showing results (see above)
 The effect of increasing the number of services %divisions %do you mean
 %services?
-on the runtime of the program is shown in Table~\ref{tbl:runtime-services-clinicians-comparison}. We executed the algorithm for
+on the runtime of the program is shown in Table~\ref{tbl:runtime-services-clinicians-comparison}.
+We executed the algorithm for
 $S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in

### 2019-09-15 paper/paper.discussion.tex
paragraph spacing
editing discussion

@@
 computational complexity of finding an optimal solution. As the size of the
+% JK: Is this a challenge of applying ILP? Or of the problem in general?
+%     I got the impression that ILP was a nice way to overcome this problem.
 scheduling problem grows, due to a larger roster of clinicians or more
@@
 %'As such' - say 'That is,...or "An NP-hard problem means that..."
+% JK: How come only "may"? Under what conditions?
 As a result, the majority of approaches taken to create schedules for similar
 scenarios tend to use heuristics in order to find an approximately optimal
-solution in a shorter time~\cite{burke_state_2004}. \\  %cite/reference
+solution in a shorter time~\cite{burke_state_2004}. %cite/reference
 %statement (i.e. the 'vast majority'...) I can't remember if the introduction
@@
 schedule satisfies hospital and logistics requirements, and a multi-goal
-objective function to satisfy the work preferences of clinicians.
+objective function to satisfy soft constraints (work preferences of clinicians).
 %Our formulation includes hard constraints to ensure the schedule satisfies
@@
 %objective function. 
-Although we restricted our use of the formulation to a minimal set of
+Although we restricted our application of the formulation to a set of
 constraints for the particular needs of the case study (St.\ Michael's Hospital
-Division of Infectious Diseases) - our formulation can be adapted to various
+Division of Infectious Diseases), our formulation can be adapted to various
 clinical departments at different hospitals. The flexibility of our ILP allows
@@
 block, clinicians' preference for block to weekend adjacency as well as
-clinicians' requests for time off. \\
+clinicians' requests for time off.
+% JK: Nice!
 
@@
 formulation was always able to find an optimal schedule satisfying all required
-hard constraints, unlike the manual schedule, which often did not take all
-constraints into account. Moreover, due to the multi-goal objective function in
-the ILP, the tool was able to maximize and fulfill the majority of clinician
+hard constraints, unlike the manual schedule, which often did not satisfy all
+constraints. Moreover, due to the multi-goal objective function in
+the ILP, the algorithm was able to fulfill the majority of clinician
 preferences and requests, more so than the manually-created schedule. These
@@
 a more fair distribution of work that helps improve the overall satisfaction of
-both employees and patients~\cite{silvestro_evaluation_2000}. \\  % how the
+both employees and patients~\cite{silvestro_evaluation_2000}.  % how the
 %finding compares to the wider literature when automated compared to manual?
@@
 In our simulations, we also found that increasing the number of requests per
-clinician does not affect the runtime of the algorithm, highlighting the
-flexibility of the tool in incorporating clinician preferences. Further, we saw
+clinician did not affect the runtime of the algorithm, highlighting the
+flexibility of the tool to incorporate clinician preferences. Further, we saw
 that the algorithm can accommodate an increase in time-horizon up to four years
-with little impact on runtime, which can be applied to departments that generate
+with little impact on runtime, suggesting the algorithm can be used generate
 schedules far in advance. One key limitation we identified was the sensitivity
-of the runtime to larger numbers of services offered in a single department.
+of the runtime to larger numbers of services offered by a single department.
 Such cases are unlikely to be encountered in the real-world because most
@@
 %constraints?
-Overall, our sensitivity analyses using simulated data provide reassurance that
+Overall, our sensitivity analyses using simulated data provided reassurance that
 the ILP formulation can be applied to schedule clinicians across real-world
@@
 Next steps include expanding the generalizability of the tool beyond smaller
-clinical departments to larger departments within and outside of health-care -
+clinical departments to larger departments within and outside of health-care --
 especially those that provide multiple services in parallel for patients and

### 2019-09-16 paper/paper.intro.tex
Fix intro

@@
-Hospital departments provide services where patient needs, and thus the system's
-demands, often exceed the available supply. In particular, on-call schedules for
+Hospital departments must allocate and use their limited resources efficiently
+in order to provide a high quality of care for their patients.
+%Hospital departments provide services where patient needs, and thus the system's
+%demands, often exceed the available supply. 
 % JK: Not sure I follow this sentence.
 %     I think we dont need to say demands > needs, can just say efficieny improves care.
+In particular, on-call schedules for
 a fixed number of health-care providers are central to the efficient running of
@@
 or the constraints that need to be satisfied
-by the schedule are complex, a manual method may not provide a schedule
-that satisfies all constraints. Second, manual methods may overlook
-certain constraints that must be maintained to have an operational department,
-such as preventing many consecutive work blocks from being assigned or ensuring
-clinicians are allocated a specific amount of work blocks throughout the year.
+by the schedule are complex, it becomes infeasible
+to find a satisfying schedule by hand.
+%a manual method may not provide a schedule
+%that satisfies all constraints. 
+Second, when creating a schedule manually it is difficult to ensure
+that all constraints are met while also trying to satisfy all staff
+preferences.
+%manual methods may overlook
+%certain constraints that must be maintained to have an operational department,
+%such as preventing many consecutive work blocks from being assigned or ensuring
+%clinicians are allocated a specific amount of work blocks throughout the year,
+%especially if there are many complex constraints to keep track of.
 % JK: Can you clarify why / how these may be overlooked?
-Third, manual scheduling is often time-consuming.
+Third, manual scheduling is often time-consuming even for relatively small 
+departments, and can take up time and resources that are better used
+for improving patient care.
 For these reasons, it is important to develop automated methods that
@@
 military~\cite{horn_scheduling_2007, laguna_modeling_2005}.
-Of special interest to a clinician on-call scheduling
+Of special interest to a clinician scheduling
 problem are the approaches to scheduling nurses, who often work in shifts. In the
 nurse scheduling problem, the goal is to find an optimal assignment of nurses to
-shifts that satisfies all of the hard constraints, such as hospital regulations,
-and as many soft constraints as possible, which may include nurse preferences. A
+shifts that satisfies all hard constraints (such as hospital regulations),
+and as many soft constraints (such as nurse preferences) as possible.
 % JK: Is this specific to the nurse scheduling problem?
@@
 %     one example of a hard constraint and one example of a soft constraint.
-wide variety of approaches, including exact and heuristic approaches, have been
+Hard constraints must be satisfied by any candidate solution to the nurse scheduling problem,
+while soft constraints can be used to rank the candidate solutions. 
+For instance, a nurse scheduling problem may include a hard constraint to assign
+at most a single shift for each nurse per day. It can also incorporate 
+nurse preference for shift time (that is, day versus night shifts) as a soft constraint
+that is meant to optimize the schedule, but is not guaranteed to be fulfilled.
+A wide variety of approaches, including exact and heuristic approaches, have been
 used to solve the nurse scheduling problem:
@@
 %     the problem is essentially impossible to solve in a limited amount of time.
-trying to find the most optimal solution. In this paper, we tackle a version
-of the nurse scheduling problem arising from a case study of one clinical
+trying to find the most optimal solution. In particular, difficult instances of these formulations
+become impossible to solve in a reasonable amount of time. 
+In this paper, we tackle the clinician 
+scheduling problem arising from a case study of one clinical
 division, providing two different services (general infectious
 disease (ID) consults; and HIV consults) at St.\ Michael's Hospital in
-Toronto, Canada. Our goals are to (1) present a integer linear programming
-(ILP) formulation for the scheduling problem, and
+Toronto, Canada. The clinician scheduling problem involves creating
+a yearly schedule that assigns clinicians to on-call work on a weekly basis,
+while ensuring a fair and balanced work-load.
 % JK: the nurse scheduling problem?
@@
 %     and how / if it differs from other scheduling problem.
+Our goals in this paper are to (1) present an integer linear programming 
+(ILP) formulation for our problem, and
 describe the flexibility of this formulation for solving similar problems;
-(2) compare the performance of an ILP-based scheduling tool to the
-results of a manual approach; and (3) analyze the robustness of this approach
-in difficult instances of the problem.
+(2) compare the performance of our tool for solving the ILP formulation to the
+results of a manual approach; 
+and (3) analyze the robustness of this approach
+in difficult instances of the problem, by exploring the effect of increasing
+the number of clinicians, requests per clinician, and services provided by
+the department on the runtime of the tool.
 % JK: what do you mean by different instances?  Can you be more specific?

### 2019-09-16 paper/paper.discussion.tex
Fix problem

@@
 observations reinforce the benefits of automated tools when generating schedules
-in hospital departments to balance the work-load of clinicians and improve the
+in hospital departments to balance the workload of clinicians and improve the
 service provided to patients. The use of automated tools alleviates the time

### 2019-09-16 paper/paper.intro.tex
Fix problem

@@
 a yearly schedule that assigns clinicians to on-call work on a weekly basis,
-while ensuring a fair and balanced work-load.
+while ensuring a fair and balanced workload.
 % JK: the nurse scheduling problem?

### 2019-09-16 paper/paper.problem.tex
one last note on problem
Fix problem

@@
 and HIV services are led by one clinician each.
+% JK: I thik we need something like:
+%     Therefore, our objective is to: assign one clinician to ...
+%     while additionally considering the following constraints ...
+Therefore, our objective is to assign a single clinician to cover
+each service for each block and each weekend of the year, while 
+additionally ensuring a balanced workload.
 
-Two types of constraints are placed on the clinician assignments. 
+In most scheduling problems, the constraints can be divided into hard and soft
+constraints. Hard constraints must be satisfied by any candidate solution, while
+soft constraints can be used to select a more favourable solution from a set of
+candidate solutions. Typically, soft constraints are therefore encoded as objective
+functions which are
+maximized when finding a solution. In the case of the clinician
+% JK: the direction of max / minimization is arbitrary,
+%     I think sufficient to pick one here to avoid confusion.
+scheduling problem, we chose departmental regulations and workload balance as hard constraints,
+while clinician preference and requests serve as soft constraints.
+%we chose Block Requests, Weekend Requests and Block-Weekend
+%Adjacency as our soft constraints, while the rest of the constraints are hard.
+Although it is important to take clinician requests into account when constructing
+the schedule, it is crucial that the workload of the schedule is balanced among
+all clinicians, and all services are assigned.
+After the schedule is generated, clinicians may
+exchange certain weeks or days throughout the year, to fulfill any missed requests.
+%Our approach to schedule generation does not
+%consider such future exchanges, and instead generates on one full year at a time.
+% JK: I think this para (4) should go instead as the 2nd in this section.
+%     you could introduce the ideas of hard and soft constraints (this para) first,
+%     and then have the current paras (2) and (3) after this one, where the reader
+%     is already aware that (2) is hard constraints and (3) is soft.
+%     Oh, I see that they're not currently split up exactly, but I think it would be
+%     helpful to do so. Maybe double check with SM though.
+
+The following are the departmental and workload constraints placed on the clinician assignments. 
 % JK: I think you need to introduce BC and WC constraints here,
@@
 % JK: trying to reduce synonyms "service" / "consultation"
-throughout the year. These limits may be different fo each clinician,
+throughout the year. These limits may be different for each clinician,
 and they may change from year to year as
@@
 
-In addition to balancing the work load among clinicians, the schedule
+In addition to balancing the workload among clinicians, the schedule
 should accommodate their preferences. Clinicians provide their requests for
 time off before schedule generation so that the requests may be integrated
-into the schedule. Clinician may specify days, weeks or weekends off, with the
+into the schedule. Clinicians may specify days, weeks or weekends off, with the
 understanding that any blocks overlapping with their request will be assigned to
@@
   % JK: below: I would consider making the 2-3 letter abbreviations its own column
-	\begin{tabular}{ l l l }
-		\toprule % JK: different line weights support table structure
-		\textbf{Constraint Name} & \textbf{Description}                               
-		&
-		\textbf{Type} \\ \midrule
-		Block Coverage (BC)           & \makecell[l]{each service needs to have
-			exactly \\ one clinician that covers any given block}                           
-		& Hard          \\ \hline
-		Weekend Coverage (WC)        & \makecell[l]{every weekend needs to have
-			exactly \\ one clinician that covers it}                                        
-		& Hard          \\ \hline
-		Min/Max (MM)                  & \makecell[l]{for a given service, each
+	\begin{tabular}{ l c l l }
+		\toprule
+		% JK: different line weights support table structure
+		\textbf{Constraint Name} & \textbf{Abbreviation} & \textbf{Description}                                                                                                                 & \textbf{Type} \\ \midrule
+		Block Coverage                                                                  & BC                    & \makecell[l]{each service needs to have
+			exactly \\ one clinician that covers any given block}                                     & Hard          \\ \hline
+		Weekend Coverage                                                                & WC                    & \makecell[l]{every weekend needs to have
+			exactly \\ one clinician that covers it}                                                 & Hard          \\ \hline
+		Min/Max                                                                         & MM                    & \makecell[l]{for a given service, each
 			clinician can only \\ work between the minimum and maximum \\ number of allowed
 			blocks} & Hard          \\ \hline
-		No Consecutive Blocks (NCB)   & \makecell[l]{any clinician should not work \\
-			two consecutive blocks, across all services}                                   
-		& Hard          \\ \hline
-		No Consecutive Weekends (NCW) & \makecell[l]{any clinician should not work two
-			consecutive \\ weekends}                                                       
-		& Hard          \\ \hline
-		Equal Weekends (EW)          & \makecell[l]{weekends should be equally
-			distributed \\ between clinicians}                                              
-		& Hard          \\ \hline
-		Equal Holidays (EH)          & \makecell[l]{long weekends should be equally
-			distributed \\ between clinicians}                                              
-		& Hard          \\ \hline
-		Block Requests (BR)         & \makecell[l]{each clinician can request to be
-			off service \\ during certain blocks throughout the year}                       
-		& Soft          \\ \hline
-		Weekend Requests (WR)        & \makecell[l]{each clinician can request to be
-			off service \\ during certain weekends throughout the year}                     
-		& Soft          \\ \hline
-		Block-Weekend Adjacency (BWA)  & \makecell[l]{the block and weekend
-			assignments of a given \\ clinician should be adjacent}                         
-		& Soft
-    \\\bottomrule % JK: I think usually table bottom also has a rule
+		No Consecutive Blocks                                                           & NCB                   & \makecell[l]{any clinician should not work \\
+			two consecutive blocks, across all services}                                        & Hard          \\ \hline
+		No Consecutive Weekends                                                         & NCW                   & \makecell[l]{any clinician should not work two
+			consecutive \\ weekends}                                                           & Hard          \\ \hline
+		Equal Weekends                                                                  & EW                    & \makecell[l]{weekends should be equally
+			distributed \\ between clinicians}                                                        & Hard          \\ \hline
+		Equal Holidays                                                                  & EH                    & \makecell[l]{long weekends should be equally
+			distributed \\ between clinicians}                                                   & Hard          \\ \hline
+		Block Requests                                                                  & BR                    & \makecell[l]{each clinician can request to be
+			off service \\ during certain blocks throughout the year}                           & Soft          \\ \hline
+		Weekend Requests                                                                & WR                    & \makecell[l]{each clinician can request to be
+			off service \\ during certain weekends throughout the year}                         & Soft          \\ \hline
+		Block-Weekend Adjacency                                                         & BWA                   & \makecell[l]{the block and weekend
+			assignments of a given \\ clinician should be adjacent}                                        & Soft          \\ \bottomrule
+		% JK: I think usually table bottom also has a rule                              &                       &                                                                                                                                      &
 	\end{tabular}
   % JK: please define Hard & Soft constraints in a footnote
+\footnotesize\raggedright
+Hard constraints must be satisfied by any candidate schedule. 
+Soft constraints are optionally satisfied, and are used to rank the set of candidate solutions.
 \end{table}
-
-In most scheduling problems, the constraints can be divided into hard and soft
-constraints. Hard constraints must be satisfied by any candidate solution, while
-soft constraints can be used to select a more favourable solution from a set of
-candidate solutions. Typically, soft constraints are therefore encoded as objective
-functions which are
-maximized when finding a solution. In the case of the clinician
-% JK: the direction of max / minimization is arbitrary,
-%     I think sufficient to pick one here to avoid confusion.
-scheduling problem described above, we chose Block Requests, Weekend Requests and Block-Weekend
-Adjacency as our soft constraints, while the rest of the constraints are hard.
-While it is important to take clinician requests into account when constructing
-the schedule, it is crucial that the work-load of the schedule is balanced among
-all clinicians, and all services are assigned.
-After the schedule is generated, clinicians may
-exchange certain weeks or days throughout the year.
-Our approach to schedule generation does not
-consider such future exchanges, and instead generates on one full year at a time.
-% JK: I think this para (4) should go instead as the 2nd in this section.
-%     you could introduce the ideas of hard and soft constraints (this para) first,
-%     and then have the current paras (2) and (3) after this one, where the reader
-%     is already aware that (2) is hard constraints and (3) is soft.
-%     Oh, I see that they're not currently split up exactly, but I think it would be
-%     helpful to do so. Maybe double check with SM though.

### 2019-09-16 paper/paper.intro.tex
Clarify experiments in intro

@@
 and (3) analyze the robustness of this approach
-in difficult instances of the problem, by exploring the effect of increasing
-the number of clinicians, requests per clinician, and services provided by
-the department on the runtime of the tool.
+in difficult instances of the problem, by exploring the change in runtime with
+changes to:
+the number of clinicians,
+the number of services provided by the department,
+the number of requests per clinician, and
+the time-horizon of the schedule.
 % JK: what do you mean by different instances?  Can you be more specific?

### 2019-09-16 paper/paper.methods.tex
Fix methods section

@@
 %     floating around. Can you do a consistency check throughout?
-scheduling problem presented in Section~\ref{sec:problem}. First, we describe
+scheduling problem presented in Section~\ref{sec:problem}. 
+An integer linear program (ILP) consists of a linear objective function and linear constraints,
+with integer-valued variables. The optimal solution to the ILP must lie within the space
+defined by the constraints while also maximizing the objective value.
+First, we describe
 % JK: Also, while the intro gives a reference for ILP,
@@
 then write the constraints given in Table~\ref{tbl:constraint-summary} as
-mathematical equations, and define the objective function of the linear program.
+linear functions of the variables, and define the objective function of the ILP.
 
@@
 %     I think maybe you already did this...
-that clinicians in a single department can provide as $\mathcal{S}$. The
-following formulation assumes that all clinicians are able to provide all
+that clinicians in a single department can provide as $\mathcal{S}$. 
+%The following formulation assumes that all clinicians are able to provide all services. 
 % JK: but could you not also just set some clinician-service max-mins zero
 %     to deal with if some clinicians are not able to provide some services?
-services. The set of all clinicians in the department is denoted as
+The set of all clinicians in the department is denoted as
 $\mathcal{C}$. The sets of blocks and weekends
@@
 %to the needs of the given department. 
-A subset of weekends are denoted $\mathcal{L}$, corresponding to established
-long/holiday weekends such as the Canadian Civic Day weekend in August. Lastly,
+A subset of weekends are denoted $\mathcal{L} \subset \mathcal{W}$, corresponding to established
+long/holiday weekends such as the Thanksgiving weekend. Lastly,
 % JK: I'm being picky, but maybe a non-Canadian holiday like Thanksgiving as the e.g.
-the block and weekend request-off dates of clinicians as are denoted as the following
-subsets of all blocks and weekends: $\mathcal{BR}_c$ and $\mathcal{WR}_c$, respectively.
+the block and weekend time-off requests of clinicians are denoted as the following
+subsets of all blocks and weekends: $\mathcal{U}_c \subset \mathcal{B}$ and $\mathcal{V}_c \subset \mathcal{W}$, respectively.
 % JK: keeping "x is denoted as x" sentence structure
@@
 %     easily confused for multiplication. Can you find another notation?
-For instance, if clinician $c$'s request-off dates intersect with blocks 1 and 2, and weekend
-1, then $\mathcal{BR}_c = \{1, 2\}$ and $\mathcal{WR}_c = \{1\}$. Table~\ref{tbl:sets-indices} presents a summary of the sets and indices described.
+For instance, if clinician $c$'s time-off requests intersect with blocks 1 and 2, and weekend
+1, then $\mathcal{U}_c = \{1, 2\}$ and $\mathcal{V}_c = \{1\}$. Table~\ref{tbl:sets-indices} presents a summary of the sets and indices described.
 
@@
 		\\
-		$\mathcal{BR}_c \subset \mathcal{B}$ &                & block requests of
+		$\mathcal{U}_c \subset \mathcal{B}$ &                & block requests of
 		clinician $c$   \\
-		$\mathcal{WR}_c \subset \mathcal{W}$ &                & weekend requests of
+		$\mathcal{V}_c \subset \mathcal{W}$ &                & weekend requests of
 		clinician $c$ \\
@@
 &Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S}
-{(-1)}^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{BR}
+{(-1)}^{\ind(b\,\in\,\mathcal{U}_c)}\cdot X_{c, b, s} \tag{BR}
 \label{eqn:obj-block-requests}\\
 &Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W}
-{(-1)}^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{WR}
+{(-1)}^{\ind(w\,\in\,\mathcal{V}_c)}\cdot Y_{c, w} \tag{WR}
 \label{eqn:obj-weekend-requests}
@@
 \begin{equation}
-\max_{X, Y, Z} \alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta)
+\max_{X, Y, Z} \alpha_1 \bar{Q}_1(X) + \alpha_2 \bar{Q}_2(Y) + \alpha_3
 \bar{Q}_3(Z)
@@
 %     though you could enforce it to compare across permutations of \alpha
-where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha,
-\beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal~\cite{stanimirovic_linear_2011}.
+subject to the constraints defined in Section \ref{sec:meth-constraints},
+where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha_i \leq 1$. 
+This method guarantees an optimal solution to be Pareto optimal~\cite{stanimirovic_linear_2011}.
+
+Currently, the most efficient approach to finding an exact solution for 
+an ILP is called Branch-and-Cut. This method involves iteratively solving 
+relaxed versions of the ILP, then constraining the relaxed problems and
+considering various sub-problems until it finds integral solutions to the 
+original ILP~\cite{mitchell_branch-and-cut_2002}.
+The intermediate relaxations allow the variables to take on real values, 
+and can be solved efficiently using the Simplex method~\cite{shamir_efficiency_1987}. The 
+complexity of finding an optimal integral solution thus lies in the 
+branching search structure of the approach.
+
 % JK: How can ILP be used solve such problems? Can you briefly describe?

### 2019-09-17 paper/tbl/runtime_services_clinicians.tex
Start fixing results (experiments) section

@@
   ``--'' indicates that no solution was found within 24 hours;
-  ``without NCB'' indicates that the No Consecutive Blocks (NCB) constraint was relaxed.
+  ``without NCB'' indicates that the No Consecutive Blocks (NCB) constraint was removed.
 \end{table}

### 2019-09-17 paper/paper.experiments.tex
Start fixing results (experiments) section

--- 2019-09-15 paper/paper.results.tex
@@
 %     and what you're going to show in this section (and why)
+In this section we compare the schedules created by solving the ILP formulation
+given in Section~\ref{sec:methods} to schedules that were manually generated,
+in terms of adherence to the hard and soft constraints outlined in Section~\ref{sec:problem}.
+We also want to analyze the efficiency of the ILP approach in generating schedules
+by its runtime on a variety of instances that may be found in the real-world.
+Our goal is to show that the schedules provided by solving the ILP successfully
+enforce all hard constraints and improve satisfiability of soft constraints, 
+as well as to demonstrate that our ILP formulation can be used for a wide range
+of configurations.
+
 \subsection{Implementation}
-We developed a software package with a user interface that implements the above
-linear program and allows configuration of clinicians at [ref \ref{???}], to be
+We developed a Python software package with a user interface that implements the above
+linear program and allows configuration of clinicians at~\cite{...}, to be
 % JK: what is this ref? Do you want to mention it is in Python?
+% TODO: I will add a ref to our Github repo here, once I clean it up and open source it
 used by the ID division at St.\ Michael's Hospital. The software was used to
@@
 simulated data as input. All the following experiments were conducted on an
-Intel Core i7-4770k CPU @ 3.50 GHz and 16 GB of RAM running 64-bit Windows 10.
-The software was implemented using the COIN-OR Branch-and-Cut open source solver
+Intel Core i7-4770k CPU @ 3.50 GHz with 16 GB of RAM running 64-bit Windows 10.
+Our software package uses COIN-OR Branch-and-Cut open source solver
 % JK: "software" here is not clear to me, is it the solver for the ILP formulation?
@@
 
-First, we evaluated the automatically-generated schedule by comparing it with the
+First, we evaluated the ILP solution by comparing it with the
 % JK: I think it would be good to give a short name for manual and the proposed
@@
 %     While lack of synonyms can sound repetitive, it can help avoid confusion.
-manually-created schedule. Specifically, we examined the adherence of each
+manual generation, as in Table~\ref{tbl:2018-schedule-comparison}. Specifically, we examined the adherence of each
 schedule to the constraints presented in Table~\ref{tbl:constraint-summary}. As
-shown in Table~\ref{tbl:constraints-comparison}, the optimal schedule generated
-using the software satisfied all mandatory constraints. In contrast, the
+shown in Table~\ref{tbl:constraints-comparison}, the ILP solution satisfied all 
+hard constraints. In contrast,
 % JK: "mandatory" = "hard"? if possible just use one.
-manually-created schedule did not satisfy all mandatory constraints. In
-particular, we see that the manual schedule assigned clinicians to multiple
-consecutive blocks in all four years. Moreover, the manual schedule did not have
+manual generation did not satisfy all hard constraints. In
+particular, we see that the manual generation assigned clinicians to multiple
+consecutive blocks in all four years. Moreover, the manual generation did not have
 an equal distribution of weekends and holidays for all four years of data.
-Considering all objectives, we see that LP outperforms the manually created
-schedule in all four years, by accommodating almost all time-off requests and
+Considering all objectives, we see that the ILP solution outperforms manual generation
+in all four years, by accommodating almost all time-off requests and
 ensuring that weekends are always assigned close to blocks.
@@
 
-\subsection{Influence of Problem Complexity of Runtime}
+\subsection{Influence of Problem Complexity on Runtime}
 Next, we examined the influence of the following four parameters on the
@@
 
+% JK: Introduce what you did first and then the figure showing results (see above)
+The effect of increasing the number of clinicians and number of services %divisions %do you mean
+%services?
+on the runtime of the program is shown in Table~\ref{tbl:runtime-services-clinicians-comparison}.
+We executed the algorithm for
+$S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in
+total across all services. 
+In a department providing a single service, increasing the number of clinicians
+did not affect the runtime, and we were able to find an ILP solution in all four cases
+within 1 second.
+For 2 concurrent services, a roster of 30 or more
+clinicians becomes impractical to schedule, as searching for a solution required
+over 24 hours. We saw similar issues for a roster of 20 or more clinicians
+assigned to a division with 3 concurrent services. However, when removing the
+NCB constraint, we saw a great improvement in runtime for divisions with 2 and 3
+services, and we were able to generate a schedule with upwards of 50 clinicians
+in under 1.5 seconds.  %what does division mean? do you mean services?
+
+\input{tbl/runtime_services_clinicians}
+
 % JK: Good to introduce the detault parameters first, then for each parameter explored,
 %     Describe what varied.
-We simulated a department with 10 clinicians offering two services, similar to the
-department at St.\ Michael's Hospital. Each clinician was
-configured with $[0, 5, 10, 15]$ non-overlapping block
+For the remaining experiments, we simulated a department with 10 clinicians offering 
+two services, similar to the department at St.\ Michael's Hospital. 
+Each clinician was
+configured with $\{0, 5, 10, 15\}$ non-overlapping block
 % JK: \abs{BR} does not clearly imply number of blocks (could be confused with week indices),
@@
 \end{figure}
-
-% JK: Introduce what you did first and then the figure showing results (see above)
-The effect of increasing the number of services %divisions %do you mean
-%services?
-on the runtime of the program is shown in Table~\ref{tbl:runtime-services-clinicians-comparison}.
-We executed the algorithm for
-$S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in
-total across all services. For 2 concurrent services, a roster of 30 or more
-clinicians becomes impractical to schedule, as searching for a solution required
-over 24 hours. We saw similar issues for a roster of 20 or more clinicians
-assigned to a division with 3 concurrent services. However, when relaxing the
-NCB constraint, we saw a great improvement in runtime for divisions with 2 and 3
-services, and we were able to generate a schedule with upwards of 50 clinicians
-in under 1.5 seconds.  %what does division mean? do you mean services?
-
-\input{tbl/runtime_services_clinicians}

### 2019-09-17 paper/tbl/2018_schedule_comparison.tex
Fix most chktex warnings
tables & figures cleanup
Start fixing results (experiments) section

@@
   \centering
+  \caption{Comparison of automatically generated (ILP solution) and manually generated schedules for 2018.
+In the ILP solution there are rigid 2-week block assignments, unlike manual generation that often assigns 3-4 weeks in a row to a single clinician.
+Moreover, in the ILP solution we see for each block either the HIV or ID clinician was assigned weekend coverage, indicating improved Block-Weekend Adjacency.}%
+  \label{tbl:2018-schedule-comparison}%
+  % JK: could you clarify in the caption what we should take away from this table?
 	\begin{adjustbox}{scale=0.8}
     \begin{tabular}{c||ccc||ccc}
-    \multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} & \multicolumn{3}{c||}{LP Solution} & \multicolumn{3}{c}{Historical Data} \\
+    \multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} & \multicolumn{3}{c||}{ILP Solution} & \multicolumn{3}{c}{Manual Generation} \\
           & HIV   & ID    & Weekend & HIV   & ID    & Weekend \\
@@
     \end{tabular}%
-	\end{adjustbox}
-	\caption{Comparison of schedules for 2018}
-	\label{tbl:2018-schedule-comparison}%
+	\end{adjustbox}\\[1em]
+	\footnotesize\raggedright
+	The ``HIV/ID'' columns represent the assignments for the two concurrent services offered at the department.
+The ``Weekend'' column represents the assignments for weekend coverage in both services.
+Different colours and letters are used to distinguish different clinicians.
+  % JK: should use a footnote to define the meaning of
+  %     - HIV / ID
+  %     - Weekend
+  %     - Colors & Letters
+  %     so that the reader can understand the table basically without reading the paper.
 \end{table}%

### 2019-09-17 paper/paper.discussion.tex
Minor fix to discussion

@@
 complicated constraints, the time it takes to generate an optimal schedule may
-grow exponentially.
+grow exponentially in the worst case.
 %The challenge in applying such an approach to this task lies in the fact that

### 2019-09-19 paper/paper.methods.tex
Work on methods fixes

@@
 (\ref{eqn:obj-block-requests}) and (\ref{eqn:obj-weekend-requests}) can be
-written as the following linear functions:  %minor suggestion - consider
+written as the following linear functions of $X$ and $Y$:  %minor suggestion - consider
 %numbering the objectives for ease later when saying things like "these two
@@
 
-However, as it is, $Q_3$ is not a linear function of its variables and cannot be
-optimized in a linear programming framework. An approach used to convert such
+However, as it is, $Q_3$ is not a linear function of the assignment variables
+$X$ and $Y$, and cannot be optimized in a linear programming framework. 
+An approach used to convert such
 functions into linear objectives involves introducing a helper variable and
-additional constraints~\cite{hammer_boolean_1968}. In our case, introducing a
-variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w =
-\varphi(b)$, and constraining $Z$ such that 
+additional constraints~\cite{hammer_boolean_1968}. 
+In our case, we introduce a variable $Z_{c, b, s}$ 
+for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w = \varphi(b)$, and
+constraining $Z$ such that 
 \begin{align}
@@
 \end{align}
-allows us to rewrite $Q_3$ as a linear function of $Z$,
+Since $X_{c, b, s}$ and $Y_{c, w}$ are binary variables, $Z_{c, b, s}$ will be constrained
+to 0, unless both $X_{c, b, s}$ and $Y_{c, w}$ are 1. Therefore, it suffices to maximize
+the following linear function of $Z$,
 \begin{align}
@@
 \end{align}
-Indeed from Eqns. (\ref{eqn:helper-x-constraint}) and
-(\ref{eqn:helper-y-constraint}),  whenever $X_{c, b, s} \cdot Y_{c, w} = 1$
-(respectively, 0), $Z_{c, b, s}$ can attain a maximum value of 1 (respectively,
-0), giving us the correct adjacency maximization objective.
+to get the correct adjacency maximization objective.
+%In our case, introducing a
+%variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w =
+%\varphi(b)$, and constraining $Z$ such that 
+%\begin{align}
+%&Z_{c, b, s} \leq X_{c, b, s} \label{eqn:helper-x-constraint}\\
+%&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S}
+%\label{eqn:helper-y-constraint}
+%\end{align}
+%allows us to rewrite $Q_3$ as a linear function of $Z$,
+%\begin{align}
+%&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} Z_{c, b, s}
+%\end{align}
+%Indeed from Eqns. (\ref{eqn:helper-x-constraint}) and
+%(\ref{eqn:helper-y-constraint}),  whenever $X_{c, b, s} \cdot Y_{c, w} = 1$
+%(respectively, 0), $Z_{c, b, s}$ can attain a maximum value of 1 (respectively,
+%0), giving us the correct adjacency maximization objective.
 % JK: not sure I follow these (respectively, 0).
@@
 Currently, the most efficient approach to finding an exact solution for 
-an ILP is called Branch-and-Cut. This method involves iteratively solving 
-relaxed versions of the ILP, then constraining the relaxed problems and
+an ILP is called Branch-and-Cut~\cite{mitchell_branch-and-cut_2002}. 
+This method involves iteratively solving 
+LP relaxations versions of the ILP, then constraining the relaxed problems and
 considering various sub-problems until it finds integral solutions to the 
-original ILP~\cite{mitchell_branch-and-cut_2002}.
+original ILP.
 The intermediate relaxations allow the variables to take on real values, 

### 2019-09-19 paper/paper.discussion.tex
Work on methods fixes

@@
 	reference here?}]. One solution to mitigate the runtime issues created by a
-larger number of services would be relaxing the constraints that prevent
+larger number of services would be removing the constraints that prevent
 assignment of consecutive blocks, followed by manual readjustment from the

### 2019-09-19 paper/paper.discussion.tex
Weaken statement about other approaches

@@
 % JK: How come only "may"? Under what conditions?
-As a result, the majority of approaches taken to create schedules for similar
-scenarios tend to use heuristics in order to find an approximately optimal
-solution in a shorter time~\cite{burke_state_2004}. %cite/reference
+Many previous approaches to creating schedules in similar scenarios have avoided
+this problem by using heuristics to find an approximately optimal solution 
+in a shorter amount of time~\cite{burke_state_2004}
+%As a result, the majority of approaches taken to create schedules for similar
+%scenarios tend to use heuristics in order to find an approximately optimal
+%solution in a shorter time~\cite{burke_state_2004}. %cite/reference
 %statement (i.e. the 'vast majority'...) I can't remember if the introduction

### 2019-09-24 paper/paper.problem.tex
Remove unnecessary detail in problem
Add a few other corrections

@@
-At St.\ Michael's Hospital, the division of infectious diseases (ID) offers
+At St.\ Michael's Hospital, the division of infectious diseases offers
 separate but concurrent services for general ID and HIV consultation. Each
@@
 %Adjacency as our soft constraints, while the rest of the constraints are hard.
-Although it is important to take clinician requests into account when constructing
-the schedule, it is crucial that the workload of the schedule is balanced among
-all clinicians, and all services are assigned.
 After the schedule is generated, clinicians may

### 2019-10-03 paper/paper.methods.tex
Clarify Branch-and-Cut relaxations

@@
 This method involves iteratively solving 
-LP relaxations versions of the ILP, then constraining the relaxed problems and
+LP relaxations of the ILP, then constraining the relaxed problems and
 considering various sub-problems until it finds integral solutions to the 
 original ILP.
-The intermediate relaxations allow the variables to take on real values, 
-and can be solved efficiently using the Simplex method~\cite{shamir_efficiency_1987}. The 
+In the intermediate relaxations, the integer assignment variables can
+take on real values, allowing the problem to be solved efficiently using 
+the Simplex method~\cite{shamir_efficiency_1987}. The 
 complexity of finding an optimal integral solution thus lies in the 
-branching search structure of the approach.
+branching search structure of Branch-and-Cut.
 

### 2019-10-03 paper/paper.experiments.tex
Replace figures in experiments section

@@
 two services, similar to the department at St.\ Michael's Hospital. 
-Each clinician was
-configured with $\{0, 5, 10, 15\}$ non-overlapping block
 % JK: \abs{BR} does not clearly imply number of blocks (could be confused with week indices),
@@
 % JK: What do you mean by non-overlapping? And why only increments of 5?
-requests.
 The effect of an increasing number of requests per clinician on the runtime of
 the ILP solver is shown in Figure~\ref{fig:runtime-requests}.
-The increase in the number of requests did not affect the runtime of
-the ILP solver, indicating that it can accommodate a lot of flexibility in
-clinician requests. Moreover, we see that all four runs were completed in under
-1 second.
+In this experiment, each clinician was configured with 1 to 15 total block requests.
+The runtime of the algorithm is constant with respect to the number of requests,
+indicating that it can accommodate a lot of flexibility in
+clinician requests. Moreover, we see that all runs were completed in under
+2 seconds.
 % JK: See somments below about trends and repeating simulations.
@@
 	\centering
-	\includegraphics[scale=.5]{fig/runtime_requests} % JK: can you export this to PDF or SVG to avoid pixelation?
+	\def\svgwidth{\columnwidth}
 	\caption{Runtime of ILP solver with an increasing number of requests per clinician}
+	\input{fig/runtime-requests.pdf_tex} % JK: can you export this to PDF or SVG to avoid pixelation?
   \label{fig:runtime-requests}
@@
 Figure~\ref{fig:runtime-blocks} presents the change in runtime when increasing
-the number of 2-week blocks $B = \{26, 52, 78, 104\}$ in a department with 10
-clinicians offering two services. There does not seem to be a strong effect on
-runtime when we lengthen the time-horizon of the schedule, and in fact the
-solver was able to find an optimal schedule for all time horizons within 3
-seconds. % 'very reasonable amount of time" is vague. what is reasonable?
+the number of 2-week blocks in a department with 10
+clinicians offering two services. In this experiment, we investigated time horizons
+from 5 to 110 blocks. This is equivalent to generating a schedule for up to 4 years ahead.
+The trend in the graph indicates a linear growth in runtime with respect to the time-horizon.
+Notably, the ILP solver was able to find all solutions in under 6 seconds, indicating very 
+good performance for long-term scheduling. 
+% 'very reasonable amount of time" is vague. what is reasonable?
 %better to give a range in value.
@@
 	\centering
-	\includegraphics[scale=.5]{fig/runtime_blocks}
+	\def\svgwidth{\textwidth}
 	\caption{Runtime of ILP solver with an increasing number of 2-week blocks}%
+	\input{fig/runtime-blocks.pdf_tex}
   \label{fig:runtime-blocks}

### 2019-10-04 paper/data/generate_configs.py
Add configs + requests to replicate paper results

@@ -1,0 +1,28 @@
+import json
+
+def main(num_clin, num_services):
+    divisions = {}
+    for i in range(1, num_services + 1):
+        divisions[str(i)] = {
+            "min": 0, "max": 1000
+        }
+
+    data = {}
+    for i in range(1, num_clin + 1):
+        data[str(i)] = {
+            "name": str(i),
+            "email": "",
+            "divisions": divisions
+        }
+
+    with open('./configs/sim_{}srv_{}clin.json'.format(num_services, num_clin), 'w') as f:
+        json.dump(data, f)
+
+if __name__ == "__main__":
+    import argparse
+    parser = argparse.ArgumentParser()
+    parser.add_argument('num_clinicians', type=int)
+    parser.add_argument('num_services', type=int)
+
+    args = parser.parse_args()
+    main(args.num_clinicians, args.num_services)

### 2019-10-04 paper/data/generate_requests.py

@@ -1,0 +1,37 @@
+from openpyxl import Workbook
+from datetime import datetime, timedelta
+import itertools
+
+num_clin = 10
+req_per_clin = 10
+
+def main(num_clinicians, req_per_clin):
+    start = datetime(2019, 1, 1)
+    next_date = datetime(2019, 1, 1)
+    blocks = []
+    clin_requests = {i: list() for i in range(1, num_clin + 1)}
+
+    while (next_date - start).days < 360:
+        blocks.append(next_date)
+        next_date = next_date + timedelta(weeks=2)
+
+    wb = Workbook()
+    block_iterator = itertools.cycle(blocks)
+    for clin in range(num_clinicians):
+        ws = wb.create_sheet(str(clin + 1))
+        for row in range(1, req_per_clin + 1):
+            val = next(block_iterator).strftime('%d-%b-%Y')
+            ws.cell(column=1, row=row, value=val)
+            ws.cell(column=2, row=row, value=val)
+
+    wb.remove(wb.active)
+    wb.save('{}clin_{}requests.xlsx'.format(num_clinicians, req_per_clin))
+
+if __name__ == '__main__':
+    import argparse
+    parser = argparse.ArgumentParser()
+    parser.add_argument('num_clinicians', type=int)
+    parser.add_argument('req_per_clinician', type=int)
+
+    args = parser.parse_args()
+    main(args.num_clinicians, args.req_per_clinician)

### 2019-10-04 paper/paper.discussion.tex
Add a few other corrections
Remove sentence about services in hospital depts

@@
 of the runtime to larger numbers of services offered by a single department. 
-Such cases are unlikely to be encountered in the real-world because most
-clinical departments tend to provide a single service or at most two services by
-the same roster of clinicians [{\color{red}SM/Kevin, could you help with
-	reference here?}]. One solution to mitigate the runtime issues created by a
-larger number of services would be removing the constraints that prevent
+One solution to mitigate the runtime issues created by a
+larger number of services would be removing the constraint that prevents
 assignment of consecutive blocks, followed by manual readjustment from the
@@
 especially those that provide multiple services in parallel for patients and
-other clients.
+other clients. As well, additional work can be done to incorporate other
+clinician preferences, such as the ability to request time-on slots or preference
+for time of year.

### 2019-10-04 paper/paper.intro.tex
Remove sentence about services in hospital depts

@@
 simulated annealing~\cite{jaszkiewicz_metaheuristic_1997}, and
-artificial intelligence~\cite{abdennadher_nurse_nodate, li_hybrid_2003}.
+artificial intelligence~\cite{abdennadher_nurse_1999, li_hybrid_2003}.
 A comprehensive

### 2019-10-08 paper/paper.discussion.tex
Minor fixes

@@
 this problem by using heuristics to find an approximately optimal solution 
-in a shorter amount of time~\cite{burke_state_2004}
+in a shorter amount of time~\cite{burke_state_2004}.
 %As a result, the majority of approaches taken to create schedules for similar

### 2019-10-08 paper/paper.intro.tex
Minor fixes

@@
 of our formulation to manually-created schedules, and evaluate the performance
-of the algorithm on simulated data in Section~\ref{sec:results}. Finally, we
+of the algorithm on simulated data in Section~\ref{sec:experiments}. Finally, we
 discuss and interpret the results in Section~\ref{sec:discussion}. % list the

### 2019-10-09 paper/paper.abstract.tex
Update paper.abstract.tex
abstract edits

@@
 We designed an integer linear programming formulation to find an optimal schedule in a clinical division of a hospital.
-Our formulation mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt it to a variety of clinical divisions.
+Our formulation mitigates issues related to computational complexity via a minimum the set of constraints, yet retains sufficient flexibility so that it can be adapted to a variety of clinical divisions.
 \\

### 2019-10-09 paper/paper.intro.tex
Update paper.intro.tex
Intro reviewed - terrific suggestions by JK!!

@@
 % JK: Not sure I follow this sentence.
-%     I think we dont need to say demands > needs, can just say efficieny improves care.
+%     I think we dont need to say demands > needs, can just say efficieny improves care. SM: i really like JK suggestions and edits!
 In particular, on-call schedules for
@@
 hospitals. Carefully allocated schedules should balance
-sufficient staff with clinician workload,
+sufficient staff with clinician workload  %SM: are staff and clinician the same thing? or are we balancing the two. clarify this sentence - what is balanced with what?
 to maximize quality of care. It is common for on-call schedules in
@@
 industries, including
-% JK: I find this is a nice way to organize these lists of topics & citations
+% JK: I find this is a nice way to organize these lists of topics & citations. SM: I like this 
 transportation~\cite{aickelin_improved_2006, goel_truck_2012, gunther_combined_2010},
@@
 % JK: rostering = scheduling? I think okay to re-use "scheduling" unless its getting extremely repetitive.
-%     but even then I think SM would agree clarity > flow/repetition.
+%     but even then I think SM would agree clarity > flow/repetition.			SM: agree!! :) use same terminology throughout
 
@@
 the time-horizon of the schedule.
-% JK: what do you mean by different instances?  Can you be more specific?
+% JK: what do you mean by different instances?  Can you be more specific? SM: nice, i like the clarity of what is being explored up front like this
 %     The results explore:

### 2019-10-09 paper/paper.experiments.tex
Minor fixes in abstract/acks/experiments
Add citation for scheduling repo

@@
 We developed a Python software package with a user interface that implements the above
-linear program and allows configuration of clinicians at~\cite{...}, to be
+linear program and allows configuration of clinicians, to be
 % JK: what is this ref? Do you want to mention it is in Python?
-% TODO: I will add a ref to our Github repo here, once I clean it up and open source it
-used by the ID division at St.\ Michael's Hospital. The software was used to
+used by the ID division at St.\ Michael's Hospital~\cite{landsman_scheduling}. 
+The software was used to
 generate the results in the following sections, using real data as well as

### 2019-10-09 paper/paper.methods.tex
Update paper.methods.tex
SM reviewed. Main residual comment --> change from abbreviations (e.g. WC) to numbers for equations.

@@
 %     "nurse scheduling problem", "scheduling problem" "clinician scheduling problem"
-%     floating around. Can you do a consistency check throughout?
+%     floating around. Can you do a consistency check throughout?  SM: agree
 scheduling problem presented in Section~\ref{sec:problem}. 
@@
 %     that ILP needs an objective function
-%     or (most importantly) how ILP works to satisfy multiple constraints.
+%     or (most importantly) how ILP works to satisfy multiple constraints.  %SM: great JK comments + how you addrssed them DL
 the sets, indices and variables necessary for the formulation of the problem. We
@@
 %to the needs of the given department. 
-A subset of weekends are denoted $\mathcal{L} \subset \mathcal{W}$, corresponding to established
-long/holiday weekends such as the Thanksgiving weekend. Lastly,
-% JK: I'm being picky, but maybe a non-Canadian holiday like Thanksgiving as the e.g.
+A subset of weekends are denoted $\mathcal{L} \subset \mathcal{W}$, corresponding to statutory
+long/holiday weekends such as the Thanksgiving weekend in the United States and Canada. Lastly,
+% JK: I'm being picky, but maybe a non-Canadian holiday like Thanksgiving as the e.g.  %SM: but not in Europe yak!
 the block and weekend time-off requests of clinicians are denoted as the following
@@
 % JK: not sure how I feel about these letters as equation refs
+% SM: agree. on reading in full, it felt odd enough for me to suggest using number for equations...
+% e.g. The WC (weekend coverage) constraint is given by .... [equation]
 \begin{align}
@@
 %objectives"...or writing the full name of the objectives instead of 'this' and
-%'these' as sometimes can get confusing what the 'this' and the 'these' refer
+%'these' as sometimes can get confusing what the 'this' and the 'these' refer  %SM: agree!
 %to. 
@@
 Currently, the most efficient approach to finding an exact solution for 
-an ILP is called Branch-and-Cut~\cite{mitchell_branch-and-cut_2002}. 
+an ILP is called Branch-and-Cut~\cite{mitchell_branch-and-cut_2002}.   %SM: good
 This method involves iteratively solving 

### 2019-10-09 paper/paper.abstract.tex

--- 2019-08-29 paper/paper.abstract.tex
@@
-Scheduling of personnel in a hospital environment is vital to improving the service provided to patients and balancing the workload assigned to clinicians.
-Many approaches have been tried and successfully applied to generate efficient schedules in such settings.
-However, due to the computational complexity of the scheduling problem in general, most approaches resort to heuristics to find a non-optimal solution in a reasonable amount of time.
-We designed an integer linear programming formulation to find an optimal schedule in a clinical division of a hospital.
-Our formulation mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt it to a variety of clinical divisions.
-\\
+Scheduling of personnel in a hospital environment is vital to improving the
+service provided to patients and balancing the workload assigned to clinicians.
+Many approaches have been tried and successfully applied to generate efficient
+schedules in such settings. However, due to the computational complexity of the
+scheduling problem in general, most approaches resort to heuristics to find a
+non-optimal solution in a reasonable amount of time. We designed an integer
+linear programming formulation to find an optimal schedule in a clinical
+division of a hospital. Our formulation mitigates issues related to
+computational complexity by minimizing the set of constraints, yet retains
+sufficient flexibility so that it can be adapted to a variety of clinical
+divisions. \\
 
-We then conducted a case study for our approach using data from the Infectious Diseases division at St.
-Michael's Hospital in Toronto, Canada.
-We analyzed and compared the results of our approach to manually-created schedules at the hospital, and found improved adherence to departmental constraints and clinician preferences.
-We used simulated data to examine the sensitivity of the runtime of our linear program for various parameters and observed reassuring results, signifying the practicality and generazability of our approach in different real-world scenarios.
+We then conducted a case study for our approach using data from the Infectious
+Diseases division at St. Michael's Hospital in Toronto, Canada. We analyzed and
+compared the results of our approach to manually-created schedules at the
+hospital, and found improved adherence to departmental constraints and clinician
+preferences. We used simulated data to examine the sensitivity of the runtime of
+our linear program for various parameters and observed reassuring results,
+signifying the practicality and generalizability of our approach in different
+real-world scenarios.

### 2019-10-09 paper/paper.experiments.tex
Update paper.experiments.tex
SM reviewed experiments - not finished yet

@@
 %     and what you're going to show in this section (and why)
-In this section we compare the schedules created by solving the ILP formulation
+% check verb tense - use past tense I think given flow
+Our goal was to determine if the schedules provided by solving the ILP 
+could successfully
+enforce all hard constraints; improve satisfiability of soft constraints compared with ??; %SM: found "improve satisfiability" a bit hard to follow. revise this phrase for clarity
+and assess if our ILP formulation can be used for a wide range
+of configurations.  %list objectives with numbers (i, ii, iii) or separated by semi-colons - helps the reader re: flow
+First, we compared the schedules created by solving the ILP formulation
 given in Section~\ref{sec:methods} to schedules that were manually generated,
-in terms of adherence to the hard and soft constraints outlined in Section~\ref{sec:problem}.
-We also want to analyze the efficiency of the ILP approach in generating schedules
+with respect to adherence to the hard and soft constraints outlined in Section~\ref{sec:problem}.
+We then examined the efficiency of the ILP approach in generating schedules
 by its runtime on a variety of instances that may be found in the real-world.
-Our goal is to show that the schedules provided by solving the ILP successfully
-enforce all hard constraints and improve satisfiability of soft constraints, 
-as well as to demonstrate that our ILP formulation can be used for a wide range
-of configurations.
 
@@
 % JK: I think it would be good to give a short name for manual and the proposed
-%     data -- e.g. LP Solution as in Table 4 and Manual Generation.
+%     data -- e.g. LP Solution as in Table 4 and Manual Generation.		%SM: great suggestion!
 %     Then check for consistency throughout.
-%     While lack of synonyms can sound repetitive, it can help avoid confusion.
+%     While lack of synonyms can sound repetitive, it can help avoid confusion.  %SM: agree
 manual generation, as in Table~\ref{tbl:2018-schedule-comparison}. Specifically, we examined the adherence of each
@@
 services, and we were able to generate a schedule with upwards of 50 clinicians
-in under 1.5 seconds.  %what does division mean? do you mean services?
+in under 1.5 seconds.
+%what does division mean?
+do you mean services?
+%SM: make sure division vs.
+services defined in introduction, but may need to repeat again in problem statement to help the reader
 

### 2019-10-09 paper/paper.experiments.tex
Update paper.experiments.tex
SM reviewed experiments section

@@
 could successfully
-enforce all hard constraints; improve satisfiability of soft constraints compared with ??; %SM: found "improve satisfiability" a bit hard to follow. revise this phrase for clarity
-and assess if our ILP formulation can be used for a wide range
+(i) enforce all hard constraints; 
+(ii) improve fulfillment of soft constraints compared to the manual approach; %SM: found "improve satisfiability" a bit hard to follow. revise this phrase for clarity
+and (iii) assess whether our ILP formulation can be used for a wide range
 of configurations.  %list objectives with numbers (i, ii, iii) or separated by semi-colons - helps the reader re: flow
@@
   %     the mean trend is actually flat. You could (should) then also present the runtimes
-  %     as box-plots to show that the trend is less than the variability.
+  %     as box-plots to show that the trend is less than the variability.  %SM: great suggestion! Did you do this (re-run 10-100 times?) - was not clear from description of experiments above
 \end{figure}
@@
 \end{figure}
+
+%SM: Table 6. change column heading "number of services" to "number of services within 1 division"
+%SM: Table 6. without NCB is a double-negative (without no consecutive blocks). Can this be revised?

### 2019-10-09 paper/paper.methods.tex

@@
 using the variables defined above. 
-
+The BC (block coverage) and WC (weekend coverage) constraints, 
+are given by Eqns. (\ref{eqn:constr-block-cov}) and (\ref{eqn:constr-weekend-cov}), respectively. 
+\begin{align}
+&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S} \label{eqn:constr-block-cov} \\
+&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \label{eqn:constr-weekend-cov}
+\end{align}
+The MM (min/max) constraint is given by:
+\begin{align}
+&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\
+c\in\mathcal{C}, s\in\mathcal{S} \label{eqn:constr-min-max}
+\end{align}
+The NCB (no consecutive blocks) and NCW (no consecutive weekends) constraints are 
+given by Eqns. (\ref{eqn:constr-no-consec-blocks}) and (\ref{eqn:constr-no-consec-weekends}), respectively.
+\begin{align}
+&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1,
+s\in\mathcal{S} \label{eqn:constr-no-consec-blocks} \\
+&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \label{eqn:constr-no-consec-weekends}
+\end{align}
+The EW (equal weekend) and EH (equal holidays) constraints are given by Eqns. (\ref{eqn:constr-equal-weekends})
+and (\ref{eqn:constr-equal-holidays}), respectively.
+\begin{align}
+&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}}
+&&\forall c\in\mathcal{C} \label{eqn:constr-equal-weekends} \\
+&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq
+\ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C}
+\label{eqn:constr-equal-holidays}
+\end{align}
 % JK: not sure how I feel about these letters as equation refs
@@
 % e.g. The WC (weekend coverage) constraint is given by .... [equation]
-\begin{align}
-&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S}
-\tag{BC} \label{eqn:constr-block-cov} \\
-&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{WC}
-\label{eqn:constr-weekend-cov} \\
-&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\
-c\in\mathcal{C}, s\in\mathcal{S} \tag{MM} \label{eqn:constr-min-max} \\
-&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1,
-s\in\mathcal{S} \tag{NCB} \label{eqn:constr-no-consec-blocks} \\
-&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1
-\tag{NCW} \label{eqn:constr-no-consec-weekends} \\
-&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}}
-&&\forall c\in\mathcal{C} \tag{EW} \label{eqn:constr-equal-weekends} \\
-&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq
-\ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C} \tag{EH}
-\label{eqn:constr-equal-holidays}
-\end{align}
 
@@
 blocks (BWA). We convert these soft constraints into linear objective functions
-of the binary variables defined in Section~\ref{sec:meth-variables}. Objectives
-(\ref{eqn:obj-block-requests}) and (\ref{eqn:obj-weekend-requests}) can be
-written as the following linear functions of $X$ and $Y$:  %minor suggestion - consider
+of the binary variables defined in Section~\ref{sec:meth-variables}. 
+Objectives BR and WR are given in Eqns. (\ref{eqn:obj-block-requests}) and (\ref{eqn:obj-weekend-requests})
+as linear functions of $X$ and $Y$:
+%minor suggestion - consider
 %numbering the objectives for ease later when saying things like "these two
@@
 &Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S}
-{(-1)}^{\ind(b\,\in\,\mathcal{U}_c)}\cdot X_{c, b, s} \tag{BR}
+{(-1)}^{\ind(b\,\in\,\mathcal{U}_c)}\cdot X_{c, b, s}
 \label{eqn:obj-block-requests}\\
 &Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W}
-{(-1)}^{\ind(w\,\in\,\mathcal{V}_c)}\cdot Y_{c, w} \tag{WR}
+{(-1)}^{\ind(w\,\in\,\mathcal{V}_c)}\cdot Y_{c, w}
 \label{eqn:obj-weekend-requests}
@@
 
-The Block-Weekend Adjacency is optimized by considering the product $X_{c, b,
-	s}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$. This
-leads
-to the maximization objective %avoid editorial adjectives
+The BWA objective is optimized by considering the product $X_{c, b, s}\cdot Y_{c, w}$ 
+for values of $w$ ``adjacent'' to the value of $b$. This
+leads to the maximization objective:%avoid editorial adjectives
 \begin{align}
 &Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} X_{c, b, s}\cdot Y_{c,
-	w=\varphi(b)} \tag{BWA} \label{eqn:obj-block-weekend-adj}
+	w=\varphi(b)} \label{eqn:obj-block-weekend-adj}
 \end{align}

### 2019-10-09 paper/tbl/runtime_services_clinicians.tex

@@
 		                                      &  \multicolumn{5}{c|}{Number of Services}  \\ \midrule
-		\makecell[l]{Number of \\ Clinicians} &  1   &    2    & 2 (without NCB) &  3   & 3 (without NCB) \\ \midrule
+		\makecell[l]{Number of \\ Clinicians} &  1   &    2    & 2 (*) &  3   & 3 (*) \\ \midrule
 		                 10                   & 0.16 &  0.74   &  0.16   & 1.40 &  0.23   \\ \hline
@@
   ``--'' indicates that no solution was found within 24 hours;
-  ``without NCB'' indicates that the No Consecutive Blocks (NCB) constraint was removed.
+  ``(*)'' indicates that the No Consecutive Blocks (NCB) constraint was removed.
 \end{table}

### 2019-10-09 paper/paper.tex
Add hard line breaks for easier editing
Fix most chktex warnings
editing results
Merge includes + definitions
Update front page

--- 2019-06-27 paper/paper.tex
@@
 \documentclass[]{article}
+\input{paper.includes.tex}
+%\input{paper.definitions.tex}
+
+\title{A flexible integer linear programming formulation for scheduling
+	clinician on-call service in hospitals}
+
+\makeatletter
+\newcommand{\printfnsymbol}[1]{%
+	\textsuperscript{\@fnsymbol{#1}}%
+}
+\makeatother
+
+\author[a, b]{David Landsman}
+\author[a]{Huiting Ma\thanks{Contributed equally to this work.}}
+\author[a, c]{Jesse Knight\printfnsymbol{1}}
+\author[d]{Kevin Gough}
+\author[a, c, d, e]{Sharmistha Mishra}
+
+\renewcommand\Affilfont{\itshape\small}
+\affil[a]{MAP Centre for Urban Health Solutions, St.\ Michael's Hospital, 
+	Unity Health Toronto, Toronto, ON, Canada}
+\affil[b]{Department of Computer Science, University of Toronto, Toronto, ON,
+	Canada}
+\affil[c]{Institute of Medical Sciences, University of Toronto, Toronto, ON,
+	Canada}
+\affil[d]{Department of Medicine, Division of Infectious Disease, St.\ Michael's
+	Hospital, Unity Health Toronto, Toronto, ON, Canada}
+\affil[e]{Institute of Health Policy, Management and Evaluation, Dalla Lana
+	School of Public Health, University of Toronto, Toronto, ON, Canada}
 
 \begin{document}
+	\maketitle
+
 	\begin{abstract}
@@
 
-	\section{Introduction} 
+	\section{Introduction}\label{sec:introduction}
 	\input{paper.intro.tex}
-	\section{Problem}
+	\section{Problem}\label{sec:problem}
 	\input{paper.problem.tex}
-	\section{Methods}
+	\section{Methods}\label{sec:methods}
 	\input{paper.methods.tex}
-	\section{Results}
-		\input{paper.results.tex}
-	\section{Simulations}
-		\input{paper.simulations.tex}
-	\section{Discussion}
+	\section{Experiments}\label{sec:experiments}
+  % JK: not sure the journal requirements, but I would title this section Experiments
+  %     since the methods do not introduce any experiments and then its not clear what
+  %     results this are refering to.
+	\input{paper.experiments.tex}
+	\section{Discussion}\label{sec:discussion}
 	\input{paper.discussion.tex}
+	\section{Acknowledgements}\label{sec:acknowledgements}
+	\input{paper.acknowledgements.tex}
+	\section{Funding}\label{sec:funding}
+	\input{paper.funding.tex}
+
+	\bibliographystyle{unsrt}
+	\bibliography{references}
 \end{document}

### 2019-10-11 paper/paper.discussion.tex
Update paper.discussion.tex
SM reviewed discussion
Cleanup latex comments

@@
 computational complexity of finding an optimal solution. As the size of the
-% JK: Is this a challenge of applying ILP? Or of the problem in general?
-%     I got the impression that ILP was a nice way to overcome this problem.
 scheduling problem grows, due to a larger roster of clinicians or more
@@
 grow exponentially in the worst case.
-%The challenge in applying such an approach to this task lies in the fact that
-%ILP is an NP-hard problem. %specify what 'such an approach' and 'this task'
-%are... the first part of the sentence was a bit hard to follow. Define what you
-%mean by NP-hard problem and cite --> i.e. NP-hard = the following sentence
-%about how time to optimal solution grows exponentially? if yes, then instead of
-%'As such' - say 'That is,...or "An NP-hard problem means that..."
-% JK: How come only "may"? Under what conditions?
 Many previous approaches to creating schedules in similar scenarios have avoided
@@
 in a shorter amount of time~\cite{burke_state_2004}.
-%As a result, the majority of approaches taken to create schedules for similar
-%scenarios tend to use heuristics in order to find an approximately optimal
-%solution in a shorter time~\cite{burke_state_2004}. %cite/reference
-%statement (i.e. the 'vast majority'...) I can't remember if the introduction
-%details the heuristics --> but if yes, then great. if no, then include that
-%review of the prior literature here in discussion. I think as you said - the
-%journals we are submitting to seem to have more of those details in the
-%introduction.
-
-%double-check all abbreviations have been defined. I would prefer limiting
-%abbreviations to no more than 3 at most if possible. We are ok for word count
-%in this paper :).
 
@@
 objective function to satisfy soft constraints (work preferences of clinicians).
-%Our formulation includes hard constraints to ensure the schedule satisfies
-%hospital and logistics requirements. It also aims to satisfy the work
-%preferences of clinicians in a clinical department by optimizing a multi-goal
-%objective function. 
 Although we restricted our application of the formulation to a set of
@@
 clinicians' requests for time off.
-% JK: Nice!
 
@@
 a more fair distribution of work that helps improve the overall satisfaction of
-both employees and patients~\cite{silvestro_evaluation_2000}.  % how the
-%finding compares to the wider literature when automated compared to manual?
-%What does it mean re: "so what" - human error in generating tools, etc. or
-%challenges in heuristics used manually by people? include citations this last
-%sentence in particular - what dos it reinforce exactly - what are the benefits?
+both employees and patients~\cite{silvestro_evaluation_2000}.
 
@@
 with little impact on runtime, suggesting the algorithm can be used generate
-schedules far in advance. One key limitation we identified was the sensitivity
+schedules far in advance. A key limitation we identified was the sensitivity
 of the runtime to larger numbers of services offered by a single department. 
@@
 assignment of consecutive blocks, followed by manual readjustment from the
-generated schedule. %what do we mean by certain constraints? specify the other
-%constraints?
+generated schedule.
 Overall, our sensitivity analyses using simulated data provided reassurance that
 the ILP formulation can be applied to schedule clinicians across real-world
-variability between clinical departments. %SM - will look for reference and we
-%can ask Kevin Gough too.
+variability between clinical departments.
 Next steps include expanding the generalizability of the tool beyond smaller

### 2019-10-11 paper/paper.experiments.tex
Cleanup latex comments

--- 2019-10-09 paper/paper.experiments.tex
@@
-% JK: I think give an overview here of the experiments here
-%     and what you're going to show in this section (and why)
-In this section we compare the schedules created by solving the ILP formulation
+Our goal was to determine if the schedules provided by solving the ILP 
+could successfully
+(i) enforce all hard constraints; 
+(ii) improve fulfillment of soft constraints compared to the manual approach; 
+and (iii) assess whether our ILP formulation can be used for a wide range
+of configurations.
+First, we compared the schedules created by solving the ILP formulation
 given in Section~\ref{sec:methods} to schedules that were manually generated,
-in terms of adherence to the hard and soft constraints outlined in Section~\ref{sec:problem}.
-We also want to analyze the efficiency of the ILP approach in generating schedules
+with respect to adherence to the hard and soft constraints outlined in Section~\ref{sec:problem}.
+We then examined the efficiency of the ILP approach in generating schedules
 by its runtime on a variety of instances that may be found in the real-world.
-Our goal is to show that the schedules provided by solving the ILP successfully
-enforce all hard constraints and improve satisfiability of soft constraints, 
-as well as to demonstrate that our ILP formulation can be used for a wide range
-of configurations.
 
@@
 linear program and allows configuration of clinicians, to be
-% JK: what is this ref? Do you want to mention it is in Python?
 used by the ID division at St.\ Michael's Hospital~\cite{landsman_scheduling}. 
@@
 Our software package uses COIN-OR Branch-and-Cut open source solver
-% JK: "software" here is not clear to me, is it the solver for the ILP formulation?
 version 2.9.9~\cite{johnjforrest_coin-or/cbc:_2019}.
 
-\subsection{Comparison with Manually Generated Schedules}  %main comment  = I think use past
-%tense in writing
+\subsection{Comparison with Manually Generated Schedules}
 We used clinician time-off requests and minimum/maximum requirements from
@@
 \input{tbl/2018_schedule_comparison}
-% JK: I think use input not include here to avoid unnecessary page breaks
-%     https://tex.stackexchange.com/questions/246
 
 First, we evaluated the ILP solution by comparing it with the
-% JK: I think it would be good to give a short name for manual and the proposed
-%     data -- e.g. LP Solution as in Table 4 and Manual Generation.
-%     Then check for consistency throughout.
-%     While lack of synonyms can sound repetitive, it can help avoid confusion.
 manual generation, as in Table~\ref{tbl:2018-schedule-comparison}. Specifically, we examined the adherence of each
@@
 hard constraints. In contrast,
-% JK: "mandatory" = "hard"? if possible just use one.
 manual generation did not satisfy all hard constraints. In
@@
 \input{tbl/constraints_comparison}
-% JK: If you want to be funny, you can estimate the approximate time to
-%     generate the schedule and include that too (16 hours vs 1 second)...
 
@@
 
-% JK: Introduce what you did first and then the figure showing results (see above)
-The effect of increasing the number of clinicians and number of services %divisions %do you mean
-%services?
+The effect of increasing the number of clinicians and number of services 
 on the runtime of the program is shown in Table~\ref{tbl:runtime-services-clinicians-comparison}.
@@
 services, and we were able to generate a schedule with upwards of 50 clinicians
-in under 1.5 seconds.  %what does division mean? do you mean services?
+in under 1.5 seconds.
 
@@
 
-% JK: Good to introduce the detault parameters first, then for each parameter explored,
-%     Describe what varied.
 For the remaining experiments, we simulated a department with 10 clinicians offering 
 two services, similar to the department at St.\ Michael's Hospital. 
-% JK: \abs{BR} does not clearly imply number of blocks (could be confused with week indices),
-%     so just leave out the LHS here I think.
-% JK: What do you mean by non-overlapping? And why only increments of 5?
 The effect of an increasing number of requests per clinician on the runtime of
@@
 2 seconds.
-% JK: See somments below about trends and repeating simulations.
 
@@
 	\caption{Runtime of ILP solver with an increasing number of requests per clinician}
-	\input{fig/runtime-requests.pdf_tex} % JK: can you export this to PDF or SVG to avoid pixelation?
+	\input{fig/runtime-requests.pdf_tex}
   \label{fig:runtime-requests}
-  % JK: The trend here is a bit confusing, since I think you're trying to say
-  %     that the runtime is not affected by the number of constraints,
-  %     but there is noise in these results.
-  %     I would re-run this 10-100 times and average the results so that (I expect)
-  %     the mean trend is actually flat. You could (should) then also present the runtimes
-  %     as box-plots to show that the trend is less than the variability.
 \end{figure}
 
-% JK: Introduce what you did first and then the figure showing results (see above)
-% JK: Try to introduce these results in the same order as they appear in the
-%     first paragraph of this section.
 Figure~\ref{fig:runtime-blocks} presents the change in runtime when increasing
@@
 good performance for long-term scheduling.
-% 'very reasonable amount of time" is vague. what is reasonable?
-%better to give a range in value.
-% JK: same comment as above re. trends
 
@@
   \label{fig:runtime-blocks}
-  % JK: same comment as above re. trends
-  % JK: Thess figures could also possible be Figure 1. a) and b) 
 \end{figure}

### 2019-10-11 paper/paper.intro.tex
Cleanup latex comments

--- 2019-10-08 paper/paper.intro.tex
@@
 in order to provide a high quality of care for their patients.
-%Hospital departments provide services where patient needs, and thus the system's
-%demands, often exceed the available supply. 
-% JK: Not sure I follow this sentence.
-%     I think we dont need to say demands > needs, can just say efficieny improves care.
 In particular, on-call schedules for
@@
 hospitals. Carefully allocated schedules should balance
-sufficient staff with clinician workload,
+sufficient staff with workload
 to maximize quality of care. It is common for on-call schedules in
@@
 to find a satisfying schedule by hand. 
-%a manual method may not provide a schedule
-%that satisfies all constraints. 
 Second, when creating a schedule manually it is difficult to ensure
@@
 preferences.
-%manual methods may overlook
-%certain constraints that must be maintained to have an operational department,
-%such as preventing many consecutive work blocks from being assigned or ensuring
-%clinicians are allocated a specific amount of work blocks throughout the year,
-%especially if there are many complex constraints to keep track of.
-% JK: Can you clarify why / how these may be overlooked?
 Third, manual scheduling is often time-consuming even for relatively small 
@@
 can efficiently generate schedules that satisfy the given constraints.
-% JK: haven't introduced what "optimal" means yet so I think maybe leave it out for now.
-%For example, it is important that a hospital department allocates its
-%resources, such as the availability of a finite number of clinicians,
-%optimally, to ensure the best possible service for its patients.
-% JK: instead of using \\ everywhere for paragraph breaks, use:
-%     \setlength{\parskip}{1em}
-%     in the preamble
 
@@
 industries, including
-% JK: I find this is a nice way to organize these lists of topics & citations
 transportation~\cite{aickelin_improved_2006, goel_truck_2012, gunther_combined_2010},
@@
 and as many soft constraints (such as nurse preferences) as possible.
-% JK: Is this specific to the nurse scheduling problem?
-%     This sounds like the more general definition of the scheduling problem.
-%     Can you define hard and soft constraints here and give
-%     one example of a hard constraint and one example of a soft constraint.
 Hard constraints must be satisfied by any candidate solution to the nurse scheduling problem,
@@
 presented in~\cite{burke_state_2004}.
-% JK: rostering = scheduling? I think okay to re-use "scheduling" unless its getting extremely repetitive.
-%     but even then I think SM would agree clarity > flow/repetition.
-
-%An extensive literature review of these and other methods is presented by [??].
-%We will briefly summarize the main ideas of some of these approaches. \\ SM -
-%don't need to introduce that you will do this for this type of paper I think -
-%would do in a thesis chapter though.
 
@@
 clinicians in that department, they are difficult to adapt to other
-departments. % explain why hard to
-%adapt?... what makes their generalizablility/adaptability limited?
-% JK: I'm less bothered by this lack of explaination, since I know its
-%     hard to explain without getting into the nitty-gritty.
+departments.
 Moreover, the large number of variables and constraints also leads to
 computational complexity issues~\cite{goos_complexity_1996}, especially when
-% JK: I think good to clarify that these complextiy issues would imply that
-%     the problem is essentially impossible to solve in a limited amount of time.
 trying to find the most optimal solution. In particular, difficult instances of these formulations
@@
 while ensuring a fair and balanced workload.
-% JK: the nurse scheduling problem?
-%     I think we need to define the "scheduling problem"
-%     and how / if it differs from other scheduling problem.
 Our goals in this paper are to (1) present an integer linear programming 
@@
 the time-horizon of the schedule.
-% JK: what do you mean by different instances?  Can you be more specific?
-%     The results explore:
-%     (1) runtime vs number of requests
-%     (2) runtime vs number of blocks
-%     (3) satisfaction of hard and soft constraints.
-% present a simple formulation for the problem developed and tested at the
-%hospital after switching from a manual approach to scheduling; and (2) analyze
-%the performance of integer linear programming in solving difficult instances of
-%the problem and compare the results with those of the manual approach; and (3)
-%describe the adaptability of the formulation as a basic framework for solving
-%similar problems in other departments. \\ %make #3 an objective
 
@@
 of the algorithm on simulated data in Section~\ref{sec:experiments}. Finally, we
-discuss and interpret the results in Section~\ref{sec:discussion}. % list the
-%main contents/elements of the paper.
+discuss and interpret the results in Section~\ref{sec:discussion}.

### 2019-10-11 paper/paper.problem.tex
Cleanup latex comments

@@
 At St.\ Michael's Hospital, the division of infectious diseases offers
-separate but concurrent services for general ID and HIV consultation. Each
+separate but concurrent services for general ID consultations and for HIV consultations. Each
 service provides clinical care throughout the year, during regular working hours
@@
 two consecutive regular work weeks and individual weekends. Apart
-% JK: when I first read this I thought weekends were also in 2-week blocks
 from long (holiday) weekends, a work week starts on Monday at 8 A.M. and ends on
@@
 and HIV services are led by one clinician each.
-% JK: I thik we need something like:
-%     Therefore, our objective is to: assign one clinician to ...
-%     while additionally considering the following constraints ...
 Therefore, our objective is to assign a single clinician to cover
@@
 maximized when finding a solution. In the case of the clinician
-% JK: the direction of max / minimization is arbitrary,
-%     I think sufficient to pick one here to avoid confusion.
 scheduling problem, we chose departmental regulations and workload balance as hard constraints,
 while clinician preference and requests serve as soft constraints.
-%we chose Block Requests, Weekend Requests and Block-Weekend
-%Adjacency as our soft constraints, while the rest of the constraints are hard.
 After the schedule is generated, clinicians may
 exchange certain weeks or days throughout the year, to fulfill any missed requests.
-%Our approach to schedule generation does not
-%consider such future exchanges, and instead generates on one full year at a time.
-% JK: I think this para (4) should go instead as the 2nd in this section.
-%     you could introduce the ideas of hard and soft constraints (this para) first,
-%     and then have the current paras (2) and (3) after this one, where the reader
-%     is already aware that (2) is hard constraints and (3) is soft.
-%     Oh, I see that they're not currently split up exactly, but I think it would be
-%     helpful to do so. Maybe double check with SM though.
 
 The following are the departmental and workload constraints placed on the clinician assignments. 
-% JK: I think you need to introduce BC and WC constraints here,
-%     since they appear in Table 1 without text in the body.
 First, each clinician
@@
 provide 3-5 blocks of general ID service and 2-3 blocks of HIV service
-% JK: trying to reduce synonyms "service" / "consultation"
 throughout the year. These limits may be different for each clinician,
@@
   \label{tbl:constraint-summary}
-  % JK: table caption should go at the top.
-  %     The spacing can be funny so I use:
-  %     \usepackage{caption}\captionsetup[table]{skip=1em}
-  %     also labels should be on their own line.
-  % JK: below: I would consider making the 2-3 letter abbreviations its own column
 	\begin{tabular}{ l c l l }
 		\toprule
-		% JK: different line weights support table structure
 		\textbf{Constraint Name} & \textbf{Abbreviation} & \textbf{Description}                                                                                                                 & \textbf{Type} \\ \midrule
@@
 			assignments of a given \\ clinician should be adjacent}                                        & Soft          \\ \bottomrule
-		% JK: I think usually table bottom also has a rule                              &                       &                                                                                                                                      &
 	\end{tabular}
-  % JK: please define Hard & Soft constraints in a footnote
 \footnotesize\raggedright

### 2019-10-17 paper/blinded/paper.blinded.tex
Add blinded version of paper

--- 2019-08-29 paper/paper.tex
@@
 \documentclass[]{article}
-\input{paper.includes.tex}
-\input{paper.definitions.tex}
+\input{paper.includes.blinded.tex}
 
-\title{A flexible integer linear programming formulation for scheduling clinician on-call service in hospitals}
-
-\author[1, 5]{David Landsman}
-\author[5]{Huiting Ma}
-\author[5]{Jesse Knight}
-\author[6]{Kevin Gough}
-\author[2, 3, 4, 5]{Sharmistha Mishra}
-
-\renewcommand\Affilfont{\itshape\small}
-\affil[1]{Department of Computer Science, University of Toronto}
-\affil[2]{Department of Medicine, Division of Infectious Disease, University of Toronto}
-\affil[3]{Institute of Health Policy, Management and Evaluation, Dalla Lana School of Public Health, University of Toronto}
-\affil[4]{Institute of Medical Sciences, University of Toronto}
-\affil[5]{MAP Centre for Urban Health Solutions, Unity Health Toronto}
-\affil[6]{Department of Medicine, St. Michael's Hospital}
+\title{A flexible integer linear programming formulation for scheduling
+	clinician on-call service in hospitals}
 
@@
 	\maketitle
+	\begin{abstract}
+		\input{paper.abstract.blinded.tex}
+	\end{abstract}
+	\keywords{Clinician scheduling, Integer linear programming, Exact method, Optimization, Clinician requests}
 
-	\begin{abstract}
-		\input{paper.abstract.tex}
-	\end{abstract}
+	\section{Introduction}\label{sec:introduction}
+	\input{paper.intro.blinded.tex}
+	\section{Problem}\label{sec:problem}
+	\input{paper.problem.blinded.tex}
+	\section{Methods}\label{sec:methods}
+	\input{paper.methods.blinded.tex}
+	\section{Experiments}\label{sec:experiments}
+	\input{paper.experiments.blinded.tex}
+	\section{Discussion}\label{sec:discussion}
+	\input{paper.discussion.blinded.tex}
 
-	\section{Introduction} \label{sec:introduction}
-		\input{paper.intro.tex}
-	\section{Problem} \label{sec:problem}
-		\input{paper.problem.tex}
-	\section{Methods} \label{sec:methods}
-		\input{paper.methods.tex}
-	\section{Results} \label{sec:results}
-		\input{paper.results.tex}
-	\section{Discussion} \label{sec:discussion}
-		\input{paper.discussion.tex}
 	\bibliographystyle{unsrt}
-	\bibliography{references}
+	\bibliography{references.blinded}
 \end{document}

### 2019-10-17 paper/blinded/paper.discussion.blinded.tex
Add blinded version of paper

--- 2019-10-11 paper/paper.discussion.tex
@@
 Although we restricted our application of the formulation to a set of
-constraints for the particular needs of the case study (St.\ Michael's Hospital
-Division of Infectious Diseases), our formulation can be adapted to various
+constraints for the particular needs of the case study 
+(the Division of Infectious Diseases at our candidate hospital),
+our formulation can be adapted to various
 clinical departments at different hospitals. The flexibility of our ILP allows
@@
 When comparing the optimal schedule generated by our tool to the
-manually-created schedules at St.\ Michael's Hospital, we found that the ILP
+manually-created schedules the hospital, we found that the ILP
 formulation was always able to find an optimal schedule satisfying all required

### 2019-10-17 paper/blinded/paper.experiments.blinded.tex
Add blinded version of paper

--- 2019-10-11 paper/paper.experiments.tex
@@
 linear program and allows configuration of clinicians, to be
-used by the ID division at St.\ Michael's Hospital~\cite{landsman_scheduling}. 
+used by the ID division at our candidate hospital~\cite{landsman_scheduling}. 
 The software was used to
@@
 For the remaining experiments, we simulated a department with 10 clinicians offering 
-two services, similar to the department at St.\ Michael's Hospital. 
+two services, similar to the department at the hospital. 
 The effect of an increasing number of requests per clinician on the runtime of

### 2019-10-17 paper/blinded/paper.intro.blinded.tex
Add blinded version of paper

--- 2019-10-11 paper/paper.intro.tex
@@
 division, providing two different services (general infectious
-disease (ID) consults; and HIV consults) at St.\ Michael's Hospital in
-Toronto, Canada. The clinician scheduling problem involves creating
+disease (ID) consults; and HIV consults) at our candidate hospital. 
+The clinician scheduling problem involves creating
 a yearly schedule that assigns clinicians to on-call work on a weekly basis,

### 2019-10-17 paper/blinded/paper.methods.blinded.tex
Add blinded version of paper

--- 2019-10-09 paper/paper.methods.tex
@@
 In this section, we present an application of integer linear programming to solve the clinician
-% JK: there are a few variations of
-%     "nurse scheduling problem", "scheduling problem" "clinician scheduling problem"
-%     floating around. Can you do a consistency check throughout?  SM: agree
 scheduling problem presented in Section~\ref{sec:problem}. 
@@
 First, we describe
-% JK: Also, while the intro gives a reference for ILP,
-%     I think you should define it here. e.g. It is not clear to the naive reader (me!)
-%     that ILP needs an objective function
-%     or (most importantly) how ILP works to satisfy multiple constraints.  %SM: great JK comments + how you addrssed them DL
 the sets, indices and variables necessary for the formulation of the problem. We
@@
 \subsection{Sets and Indices}\label{sec:meth-sets-indices}
-We denote the set of all services %or I think services is good (vs.
-%services/divisions) as you define two types of services (ID vs. HIV) in the
-%problem statement
-% JK: same thing, consistency check with division / service
-%     I think maybe you already did this...
+We denote the set of all services
 that clinicians in a single department can provide as $\mathcal{S}$. 
-%The following formulation assumes that all clinicians are able to provide all services. 
-% JK: but could you not also just set some clinician-service max-mins zero
-%     to deal with if some clinicians are not able to provide some services?
 The set of all clinicians in the department is denoted as
 $\mathcal{C}$. The sets of blocks and weekends
-% JK: removed "will be assigned to" since on first read I confused it with
-%     the binary assignment variable X below.
 are denoted as $\mathcal{B}$ and $\mathcal{W}$ respectively. The block size
@@
 require a particular size for blocks, and so it can be adapted for other cases.
-%The size of a block is not constrained and can be adapted %how?
-%to the needs of the given department. 
 A subset of weekends are denoted $\mathcal{L} \subset \mathcal{W}$, corresponding to statutory
 long/holiday weekends such as the Thanksgiving weekend in the United States and Canada. Lastly,
-% JK: I'm being picky, but maybe a non-Canadian holiday like Thanksgiving as the e.g.  %SM: but not in Europe yak!
 the block and weekend time-off requests of clinicians are denoted as the following
 subsets of all blocks and weekends: $\mathcal{U}_c \subset \mathcal{B}$ and $\mathcal{V}_c \subset \mathcal{W}$, respectively.
-% JK: keeping "x is denoted as x" sentence structure
-% JK: Also, I'm generally not a fan of multiple-letter variables since they can be 
-%     easily confused for multiplication. Can you find another notation?
 For instance, if clinician $c$'s time-off requests intersect with blocks 1 and 2, and weekend
@@
 in the problem.
-%To optimize the soft constraint Block-Weekend Adjacency, we maximize the
-%product $X_{c, b, s} \cdot Y_{c, w}$ for adjacent blocks and weekends. To
-%formulate such an objective as a linear function of variables, we introduce
-%another set of variables, denoted by $Z_{c, b, s}$, with additional constraints
-%on its range. Further details regarding this variable are described in Section
-%\ref{sec:meth-objectives}.
 
@@
 		\\
-		%		$Z_{c, b, s} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend
-		%adjacency             \\
 		$m_{c, s}$                 & minimum number of blocks clinician $c$ should
@@
 \end{align}
-% JK: not sure how I feel about these letters as equation refs
-% SM: agree. on reading in full, it felt odd enough for me to suggest using number for equations...
-% e.g. The WC (weekend coverage) constraint is given by .... [equation]
 
@@
 as linear functions of $X$ and $Y$:
-%minor suggestion - consider
-%numbering the objectives for ease later when saying things like "these two
-%objectives"...or writing the full name of the objectives instead of 'this' and
-%'these' as sometimes can get confusing what the 'this' and the 'these' refer  %SM: agree!
-%to. 
 \begin{align}
@@
 
-%So if clinician $c$ is assigned to work during block 3, corresponding to weeks
-%5 and 6 assuming 2-week blocks, they might also want to be assigned to work
-%during weekend 5. In that case, we would like $X_{c, b=3, s} \cdot Y_{c, w=5}$
-%to be 1, since that indicates both variables are assigned. If at least one of
-%the two variables is not assigned, the product will be 0. 
-% I really liked the example explanation!
-%For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\
-
 However, as it is, $Q_3$ is not a linear function of the assignment variables
@@
 to get the correct adjacency maximization objective.
-%In our case, introducing a
-%variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w =
-%\varphi(b)$, and constraining $Z$ such that 
-%\begin{align}
-%&Z_{c, b, s} \leq X_{c, b, s} \label{eqn:helper-x-constraint}\\
-%&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S}
-%\label{eqn:helper-y-constraint}
-%\end{align}
-%allows us to rewrite $Q_3$ as a linear function of $Z$,
-%\begin{align}
-%&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} Z_{c, b, s}
-%\end{align}
-%Indeed from Eqns. (\ref{eqn:helper-x-constraint}) and
-%(\ref{eqn:helper-y-constraint}),  whenever $X_{c, b, s} \cdot Y_{c, w} = 1$
-%(respectively, 0), $Z_{c, b, s}$ can attain a maximum value of 1 (respectively,
-%0), giving us the correct adjacency maximization objective.
-% JK: not sure I follow these (respectively, 0).
-%Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a
-%maximum value of 1, and whenever $X_{c, b, s} \cdot Y_{c, w} = 0$, at least one
-%of $X_{c, b, s}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, s}$ will be constrained
-%to attain a maximum value of 0, giving us the correct adjacency maximization
-%objective. \\
 
@@
 of the normalized objective functions,
-% JK: This only considers the soft constraints.
-%     Should this maximization not also be subject to the Constraints in 3.3?
 \begin{equation}
@@
 \end{equation}
-% JK: I would use $\alpha_1, \alpha_2, \alpha_3$ here which could take any value,
-%     and then not necessary to have any bounds like \sum_i^3 \alpha_i = 1,
-%     though you could enforce it to compare across permutations of \alpha
 subject to the constraints defined in Section \ref{sec:meth-constraints},
@@
 Currently, the most efficient approach to finding an exact solution for 
-an ILP is called Branch-and-Cut~\cite{mitchell_branch-and-cut_2002}.   %SM: good
+an ILP is called Branch-and-Cut~\cite{mitchell_branch-and-cut_2002}.
 This method involves iteratively solving 
@@
 branching search structure of Branch-and-Cut.
-
-% JK: How can ILP be used solve such problems? Can you briefly describe?
-
-%Thus, our clinician scheduling problem is a multiple objective optimization
-%problem. The most common approach to solving multiple objective optimization
-%problems is by optimizing a weighted sum of the normalized objective functions,
-%as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}].
-%This is the approach we decided to use in our problem, to ensure all three
-%objectives are considered when finding a solution. Under the assumption that
-%each clinician in $\mathcal{C}$ provides all types of services in
-%$\mathcal{S}$, the normalized objectives can be written as follows,
-%\begin{align}
-%	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot S} \tag{Block Requests}
-%\label{eqn:norm-obj-block-requests}\\
-%	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests}
-%\label{eqn:norm-obj-weekend-requests} \\
-%	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot S} \tag{Block-Weekend Adjacency}
-%\label{eqn:norm-obj-block-weekend-adj}
-%\end{align}
-%where we divide each of the original objective functions by the sum of the
-%absolute values of its coefficients [ref \ref{???}]. The final weighted
-%objective is given by
-%\begin{equation}
-%	\alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
-%\end{equation}
-%with $0 \leq \alpha, \beta \leq 1$. \\

### 2019-10-17 paper/blinded/paper.problem.blinded.tex
Add blinded version of paper

--- 2019-10-11 paper/paper.problem.tex
@@
-At St.\ Michael's Hospital, the division of infectious diseases offers
+At the hospital, the division of infectious diseases offers
 separate but concurrent services for general ID consultations and for HIV consultations. Each

### 2019-11-11 paper/paper.tex
Cleanup latex comments
Add corresponding author + keywords
Add note about preprint

@@
 \input{paper.includes.tex}
-%\input{paper.definitions.tex}
 
@@
 	clinician on-call service in hospitals}
-
-\makeatletter
-\newcommand{\printfnsymbol}[1]{%
-	\textsuperscript{\@fnsymbol{#1}}%
-}
-\makeatother
+\date{}
 
@@
 \author[d]{Kevin Gough}
-\author[a, c, d, e]{Sharmistha Mishra}
+\author[a, c, d, e]{Sharmistha Mishra\thanks{Corresponding author (\texttt{sharmistha.mishra@utoronto.ca})}}
 
@@
 	\end{abstract}
+	\keywords{Clinician scheduling, Integer linear programming, Exact method, Optimization, Clinician requests}
+	\blfootnote{Preprint submitted to arXiv: \href{https://arxiv.org/abs/1910.08526}{https://arxiv.org/abs/1910.08526}}
 
@@
 	\section{Experiments}\label{sec:experiments}
-  % JK: not sure the journal requirements, but I would title this section Experiments
-  %     since the methods do not introduce any experiments and then its not clear what
-  %     results this are refering to.
 	\input{paper.experiments.tex}

### 2020-01-23 docs/source/conf.py
Update HTML docs
* New theme: RTD
* Add description of Settings tab
* Add page about installation

@@ -39,7 +39,8 @@ release = '2.0'
 # extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
 # ones.
 extensions = [
-    'sphinx.ext.githubpages'
+    'sphinx.ext.githubpages',
+    'sphinx_rtd_theme'
 ]
 
 # Add any paths that contain templates here, relative to this directory.
@@ -75,7 +76,7 @@ pygments_style = 'sphinx'
 # The theme to use for HTML and HTML Help pages.  See the documentation for
 # a list of builtin themes.
 #
-html_theme = 'alabaster'
+html_theme = 'sphinx_rtd_theme'
 
 # Theme options are theme-specific and customize the look and feel of a theme
 # further.  For a list of options available for each theme, see the

### 2020-01-23 docs/build/latex/ClinicianScheduler.tex
Update holidays & requests images in manual
Update PDF user manual
Fix some issues with latex generation

* Change to \usepackage[noconfigs]{babel} to solve package download issue
    ref: https://tex.stackexchange.com/a/313559
* Add preamble to prevent handling of quoted links
    This was causing problems with links that did not have quotes
    ref: https://tex.stackexchange.com/a/512877

@@
 \ifdefined\DeclareUnicodeCharacter
+% support both utf8 and utf8x syntaxes
   \ifdefined\DeclareUnicodeCharacterAsOptional
-  \DeclareUnicodeCharacter{"00A0}{\nobreakspace}
-  \DeclareUnicodeCharacter{"2500}{\sphinxunichar{2500}}
-  \DeclareUnicodeCharacter{"2502}{\sphinxunichar{2502}}
-  \DeclareUnicodeCharacter{"2514}{\sphinxunichar{2514}}
-  \DeclareUnicodeCharacter{"251C}{\sphinxunichar{251C}}
-  \DeclareUnicodeCharacter{"2572}{\textbackslash}
+    \def\sphinxDUC#1{\DeclareUnicodeCharacter{"#1}}
   \else
-  \DeclareUnicodeCharacter{00A0}{\nobreakspace}
-  \DeclareUnicodeCharacter{2500}{\sphinxunichar{2500}}
-  \DeclareUnicodeCharacter{2502}{\sphinxunichar{2502}}
-  \DeclareUnicodeCharacter{2514}{\sphinxunichar{2514}}
-  \DeclareUnicodeCharacter{251C}{\sphinxunichar{251C}}
-  \DeclareUnicodeCharacter{2572}{\textbackslash}
+    \let\sphinxDUC\DeclareUnicodeCharacter
   \fi
+  \sphinxDUC{00A0}{\nobreakspace}
+  \sphinxDUC{2500}{\sphinxunichar{2500}}
+  \sphinxDUC{2502}{\sphinxunichar{2502}}
+  \sphinxDUC{2514}{\sphinxunichar{2514}}
+  \sphinxDUC{251C}{\sphinxunichar{251C}}
+  \sphinxDUC{2572}{\textbackslash}
 \fi
@@
 \usepackage{amsmath,amssymb,amstext}
-\usepackage{babel}
+\usepackage[noconfigs]{babel}
+
 \usepackage{times}
+\expandafter\ifx\csname T@LGR\endcsname\relax
+\else
+% LGR was declared as font encoding
+  \substitutefont{LGR}{\rmdefault}{cmr}
+  \substitutefont{LGR}{\sfdefault}{cmss}
+  \substitutefont{LGR}{\ttdefault}{cmtt}
+\fi
+\expandafter\ifx\csname T@X2\endcsname\relax
+  \expandafter\ifx\csname T@T2A\endcsname\relax
+  \else
+  % T2A was declared as font encoding
+    \substitutefont{T2A}{\rmdefault}{cmr}
+    \substitutefont{T2A}{\sfdefault}{cmss}
+    \substitutefont{T2A}{\ttdefault}{cmtt}
+  \fi
+\else
+% X2 was declared as font encoding
+  \substitutefont{X2}{\rmdefault}{cmr}
+  \substitutefont{X2}{\sfdefault}{cmss}
+  \substitutefont{X2}{\ttdefault}{cmtt}
+\fi
+
 \usepackage[Bjarne]{fncychap}
@@
 
+\fvset{fontsize=\small}
 \usepackage{geometry}
@@
 
-\addto\captionsenglish{\renewcommand{\figurename}{Fig.}}
-\addto\captionsenglish{\renewcommand{\tablename}{Table}}
-\addto\captionsenglish{\renewcommand{\literalblockname}{Listing}}
+\usepackage{sphinxmessages}
 
-\addto\captionsenglish{\renewcommand{\literalblockcontinuedname}{continued from previous page}}
-\addto\captionsenglish{\renewcommand{\literalblockcontinuesname}{continues on next page}}
-
-\addto\extrasenglish{\def\pageautorefname{page}}
+\makeatletter
+\def\set@curr@file#1{\def\@curr@file{#1}}
+\makeatother
 
 \title{Clinician Scheduler Documentation}
-\date{Jul 09, 2019}
-\release{2.0}
+\date{Jan 23, 2020}
+\release{2.3}
 \author{David Landsman}
@@
 \makeindex
-
 \begin{document}
 
-\maketitle
+\pagestyle{empty}
+\sphinxmaketitle
+\pagestyle{plain}
 \sphinxtableofcontents
+\pagestyle{normal}
 \phantomsection\label{\detokenize{manual::doc}}
-
-\begin{sphinxShadowBox}
-\sphinxstyletopictitle{Table of Contents}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id4}}{\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id5}}{\hyperref[\detokenize{manual:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id6}}{\hyperref[\detokenize{manual:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id7}}{\hyperref[\detokenize{manual:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id8}}{\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id9}}{\hyperref[\detokenize{manual:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id10}}{\hyperref[\detokenize{manual:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
-
-\end{itemize}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id11}}{\hyperref[\detokenize{manual:requests-holiday-weekends}]{\sphinxcrossref{Requests \& Holiday Weekends}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id12}}{\hyperref[\detokenize{manual:clinician-requests-excel-file}]{\sphinxcrossref{Clinician Requests Excel File}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id13}}{\hyperref[\detokenize{manual:holiday-weekends-excel-file}]{\sphinxcrossref{Holiday Weekends Excel File}}}
-
-\end{itemize}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id14}}{\hyperref[\detokenize{manual:scheduling}]{\sphinxcrossref{Scheduling}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id15}}{\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id16}}{\hyperref[\detokenize{manual:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}
-
-\end{itemize}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id17}}{\hyperref[\detokenize{manual:sample-output}]{\sphinxcrossref{Sample Output}}}
-\begin{itemize}
-\item {} 
-\phantomsection\label{\detokenize{manual:id18}}{\hyperref[\detokenize{manual:yearly-excel-format}]{\sphinxcrossref{Yearly Excel Format}}}
-
-\item {} 
-\phantomsection\label{\detokenize{manual:id19}}{\hyperref[\detokenize{manual:monthly-excel-format}]{\sphinxcrossref{Monthly Excel Format}}}
-
-\end{itemize}
-
-\end{itemize}
-\end{sphinxShadowBox}
 
@@
 \centering
-\sphinxhref{\_static/images/configuration/new\_config.png}{\sphinxincludegraphics{{new_config}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/new\_config.png}{\sphinxincludegraphics{{new_config}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/configuration/save\_config.png}{\sphinxincludegraphics{{save_config}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/save\_config.png}{\sphinxincludegraphics{{save_config}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/configuration/load\_config.png}{\sphinxincludegraphics{{load_config}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/load\_config.png}{\sphinxincludegraphics{{load_config}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/configuration/add\_clinician/step1\_new\_clinician.png}{\sphinxincludegraphics{{step1_new_clinician}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/add\_clinician/step1\_new\_clinician.png}{\sphinxincludegraphics{{step1_new_clinician}.png}}
+\end{figure}
 \begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
 \setcounter{enumi}{1}
@@
 \centering
-\sphinxhref{\_static/images/configuration/add\_clinician/step2\_add\_division.png}{\sphinxincludegraphics{{step2_add_division}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/add\_clinician/step2\_add\_division.png}{\sphinxincludegraphics{{step2_add_division}.png}}
+\end{figure}
 \begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
 \setcounter{enumi}{2}
@@
 \centering
-\sphinxhref{\_static/images/configuration/add\_clinician/step3\_remove\_division.png}{\sphinxincludegraphics{{step3_remove_division}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/add\_clinician/step3\_remove\_division.png}{\sphinxincludegraphics{{step3_remove_division}.png}}
+\end{figure}
 \begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
 \setcounter{enumi}{3}
@@
 \centering
-\sphinxhref{\_static/images/configuration/add\_clinician/step4\_add\_clinician.png}{\sphinxincludegraphics{{step4_add_clinician}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/add\_clinician/step4\_add\_clinician.png}{\sphinxincludegraphics{{step4_add_clinician}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/configuration/delete\_clinician.png}{\sphinxincludegraphics{{delete_clinician}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/delete\_clinician.png}{\sphinxincludegraphics{{delete_clinician}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/configuration/edit\_clinician.png}{\sphinxincludegraphics{{edit_clinician}.png}}\end{figure}
+
+\sphinxhref{\_static/images/configuration/edit\_clinician.png}{\sphinxincludegraphics{{edit_clinician}.png}}
+\end{figure}
 
@@
 \label{\detokenize{manual:requests-holiday-weekends}}\label{\detokenize{manual:requests-and-holidays}}
-Clinician requests for time-off and the dates of holiday weekends are
+Clinician requests for time\sphinxhyphen{}off and the dates of holiday weekends are
 supplied using a pair of excel files.
@@
 \centering
-\sphinxhref{\_static/images/requests\_holidays/excel\_requests.png}{\sphinxincludegraphics{{excel_requests}.png}}\end{figure}
+
+\sphinxhref{\_static/images/requests\_holidays/excel\_requests.png}{\sphinxincludegraphics{{excel_requests}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/requests\_holidays/excel\_holidays.png}{\sphinxincludegraphics{{excel_holidays}.png}}\end{figure}
+
+\sphinxhref{\_static/images/requests\_holidays/excel\_holidays.png}{\sphinxincludegraphics{{excel_holidays}.png}}
+\end{figure}
 
@@
 \begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
 \item {} 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step1\_load.png}{\sphinxincludegraphics{{step1_load}.png}}\end{figure}
+
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step1\_load.png}{\sphinxincludegraphics{{step1_load}.png}}
+\end{figure}
 \begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
 \setcounter{enumi}{1}
@@
 Set the starting calendar year for the schedule, as well as the number
-of 2-week blocks that you want the schedule to cover. By default, the calendar
+of 2\sphinxhyphen{}week blocks that you want the schedule to cover. By default, the calendar
 year is the upcoming year, and the number of blocks is 26, to cover a full
@@
 \centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step2\_year\_blocks.png}{\sphinxincludegraphics{{step2_year_blocks}.png}}\end{figure}
+
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step2\_year\_blocks.png}{\sphinxincludegraphics{{step2_year_blocks}.png}}
+\end{figure}
 \begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
 \setcounter{enumi}{2}
@@
 \centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step3\_optional\_shuffle.png}{\sphinxincludegraphics{{step3_optional_shuffle}.png}}\end{figure}
+
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step3\_optional\_shuffle.png}{\sphinxincludegraphics{{step3_optional_shuffle}.png}}
+\end{figure}
 \begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
 \setcounter{enumi}{3}
@@
 \centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step4\_optional\_verbose.png}{\sphinxincludegraphics{{step4_optional_verbose}.png}}\end{figure}
+
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step4\_optional\_verbose.png}{\sphinxincludegraphics{{step4_optional_verbose}.png}}
+\end{figure}
 \begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
 \setcounter{enumi}{4}
@@
 
+\begin{figure}[H]
+\centering
+
+\sphinxhref{\_static/images/scheduling/generate\_schedule/step5\_generate.png}{\sphinxincludegraphics{{step5_generate}.png}}
+\end{figure}
+
 \begin{sphinxadmonition}{warning}{Warning:}
@@
 
-\begin{figure}[H]
-\centering
-\sphinxhref{\_static/images/scheduling/generate\_schedule/step5\_generate.png}{\sphinxincludegraphics{{step5_generate}.png}}\end{figure}
-
 \begin{sphinxadmonition}{warning}{Warning:}
 It is possible that the scheduler will not be able to come up with a
-schedule that satisfies your constraints. This can be a result of either
-of the following reasons:
+schedule that satisfies your constraints. This can be a result of one
+or more of the following reasons:
 \begin{itemize}
@@
 There are not enough clinicians to distribute evenly throughout the year
+
+\item {} 
+The specified constraints are too restrictive
 
@@
 
-Try changing the configuration file by adding new clinicians, or changing
-the min/max blocks of different clinicians to allow for more flexibility.
+In these situations try the following:
+\begin{enumerate}
+\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
+\item {} 
+Adjust the constraints in the \sphinxtitleref{Settings} tab, starting from the
+most restrictive constraints. See {\hyperref[\detokenize{manual:changing-settings}]{\sphinxcrossref{\DUrole{std,std-ref}{Changing settings}}}}
 
-See {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}} for more information on modifying the configuration file.
+\item {} 
+Try changing the configuration file by adding new clinicians, or
+changing the min/max blocks of existing clinicians to allow for
+more flexibility. See {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
+
+\end{enumerate}
 \end{sphinxadmonition}
+
+\section{Changing settings}
+\label{\detokenize{manual:changing-settings}}\label{\detokenize{manual:id4}}
+The \sphinxtitleref{Settings} tab allows you to modify the constraints applied when
+generating a potential schedule.
+
+In order to disable/enable a constraint, uncheck/check the
+corresponding checkbox and re\sphinxhyphen{}generate the schedule in the \sphinxtitleref{Scheduler}
+tab (see {\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{\DUrole{std,std-ref}{Generating a schedule}}}}).
+
+\begin{figure}[H]
+\centering
+
+\sphinxhref{\_static/images/scheduling/settings\_tab.png}{\sphinxincludegraphics{{settings_tab}.png}}
+\end{figure}
+
+\subsection{St. Michael’s Hospital (SMH) Constraints}
+\label{\detokenize{manual:st-michael-s-hospital-smh-constraints}}
+The following descriptions are specific to the SMH version of the
+scheduler (that is, the \sphinxcode{\sphinxupquote{scheduler\sphinxhyphen{}smh.exe}} executable file)
+
+\begin{sphinxadmonition}{note}{Note:}
+The following constraints are presented in order of restrictiveness,
+from least restrictive to most restrictive.
+
+If you are struggling to generate a schedule, try disabling them,
+starting from the latter constraints.
+\end{sphinxadmonition}
+\begin{itemize}
+\item {} \begin{description}
+\item[{\sphinxstylestrong{Cover All Block \& Weekends}}] \leavevmode
+This constraint enforces that a clinician is assigned to every block
+and weekend, in every department.
+
+\end{description}
+
+\item {} \begin{description}
+\item[{\sphinxstylestrong{Restrict Min/Max Blocks}}] \leavevmode
+This constraint enforces that a clinician is assigned to a number
+of blocks no less than their specified minimum and no more than
+their specified maximum.
+
+\end{description}
+
+\item {} \begin{description}
+\item[{\sphinxstylestrong{Balance Weekends}}] \leavevmode
+This constraint enforces that all clinicians are assigned to an
+approximately equal number of weekends
+
+\end{description}
+
+\item {} \begin{description}
+\item[{\sphinxstylestrong{Balance Long Weekends}}] \leavevmode
+This constraint enforces that all clinicians are assigned to an
+approximately equal number of long (holiday) weekends
+
+\end{description}
+
+\item {} \begin{description}
+\item[{\sphinxstylestrong{Prevent Consecutive Blocks}}] \leavevmode
+This constraint enforces that a clinician is not assigned to two
+consecutive blocks in a row
+
+\end{description}
+
+\item {} \begin{description}
+\item[{\sphinxstylestrong{Spread Out Blocks}}] \leavevmode
+This constraint enforces that a clinician is not assigned to more than
+2 blocks in every 5 consecutive blocks.
+
+\end{description}
+
+\item {} \begin{description}
+\item[{\sphinxstylestrong{Spread Out Weekends}}] \leavevmode
+This constraints enforces that a clinician is not assigned to more than
+1 block in every 4 consecutive blocks.
+
+\end{description}
+
+\end{itemize}
 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/export\_schedule/export\_yearly.png}{\sphinxincludegraphics{{export_yearly}.png}}\end{figure}
+
+\sphinxhref{\_static/images/scheduling/export\_schedule/export\_yearly.png}{\sphinxincludegraphics{{export_yearly}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/export\_schedule/export\_monthly.png}{\sphinxincludegraphics{{export_monthly}.png}}\end{figure}
+
+\sphinxhref{\_static/images/scheduling/export\_schedule/export\_monthly.png}{\sphinxincludegraphics{{export_monthly}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/sample\_yearly.png}{\sphinxincludegraphics{{sample_yearly}.png}}\end{figure}
+
+\sphinxhref{\_static/images/scheduling/sample\_yearly.png}{\sphinxincludegraphics{{sample_yearly}.png}}
+\end{figure}
 
@@
 \centering
-\sphinxhref{\_static/images/scheduling/sample\_monthly.png}{\sphinxincludegraphics{{sample_monthly}.png}}\end{figure}
+
+\sphinxhref{\_static/images/scheduling/sample\_monthly.png}{\sphinxincludegraphics{{sample_monthly}.png}}
+\end{figure}
 

### 2020-01-23 docs/source/conf.py

@@ -24,9 +24,9 @@ copyright = '2019, David Landsman'
 author = 'David Landsman'
 
 # The short X.Y version
-version = ''
+version = '2.3'
 # The full version, including alpha/beta/rc tags
-release = '2.0'
+release = '2.3'
 
 
 # -- General configuration ---------------------------------------------------
@@ -119,11 +119,14 @@ latex_elements = {
 
     # Additional stuff for the LaTeX preamble.
     #
-    # 'preamble': '',
+    'preamble': r'''\makeatletter
+\def\set@curr@file#1{\def\@curr@file{#1}}
+\makeatother''',
 
     # Latex figure (float) alignment
     #
     # 'figure_align': 'htbp',
+    'babel': '\\usepackage[noconfigs]{babel}',
     'figure_align': 'H'
 }
 
@@ -132,7 +135,7 @@ latex_elements = {
 #  author, documentclass [howto, manual, or own class]).
 latex_documents = [
     ('manual', 'ClinicianScheduler.tex', 'Clinician Scheduler Documentation',
-     'David Landsman', 'manual'),
+     author, 'manual'),
 ]
 
 

### 2020-05-25 src/services/scheduler.py
Add button to export LpProblem as file

@@ -273,6 +273,9 @@ class Scheduler:
 
             return (divAssignments, weekendAssignments, self.holiday_map)
 
+    def get_problem(self):
+        return self.problem
+
     def setup_solver(self):
         if getattr(sys, 'frozen', False):
             # running in a bundle

### 2020-05-25 src/ui/mainwindow.py

@@ -85,6 +85,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         self.generateScheduleButton.clicked.connect(self.generateSchedule)
         self.exportScheduleButton.clicked.connect(self.exportSchedule)
         self.exportMonthlyButton.clicked.connect(self.exportMonthlySchedule)
+        self.exportLpButton.clicked.connect(self.exportLpProblem)
 
         self.calendarYearSpinBox.setValue(datetime.now().year + 1)
 
@@ -219,9 +220,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
             del self.configuration[clinName]
             UiHelper.syncTreeView(self.treeView, self.model, self.configuration)
 
-    def generateSchedule(self):
-        self.clearScheduleTable()
-
+    def setupScheduler(self):
         numClinicians = len(self.configuration.keys())
         if numClinicians <= 0:
             self._logger.write_line('No clinicians configured! Please load a configuration file with at least one clinician!', level='ERROR')
@@ -229,8 +228,6 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
         numBlocks = self.numberOfBlocksSpinBox.value()
         calendarYear = self.calendarYearSpinBox.value()
-        shuffle = self.shuffleCheckBox.isChecked()
-        verbose = self.verboseCheckBox.isChecked()
         constraints = []
 
         # pull enabled constraints from UI checkboxes
@@ -239,12 +236,51 @@ class MainWindow(QMainWindow, Ui_MainWindow):
             if widget.isChecked():
                 constraints.append(widget.objectName().replace('CheckBox', ''))
 
-        # init scheduler with all the given data
-        schedule = scheduler \
-            .Scheduler(logger=self._logger, num_blocks=numBlocks, clin_data=self.configuration, \
-                 request_dict=self._requests, holidays=self._holidays,
-                 constraints=constraints) \
-            .generate(verbose=verbose, shuffle=shuffle)
+        return scheduler.Scheduler(
+            logger=self._logger, num_blocks=numBlocks, clin_data=self.configuration,
+            request_dict=self._requests, holidays=self._holidays,
+            constraints=constraints
+        )
+        
+    def exportLpProblem(self):
+        shuffle = self.shuffleCheckBox.isChecked()
+
+        scheduler = self.setupScheduler()
+        if scheduler is None:
+            self._logger.write_line('Could not setup scheduler!', level='ERROR')
+            return
+
[…]

### 2020-05-25 src/ui/ui_mainwindow.py

@@ -1,6 +1,6 @@
 # -*- coding: utf-8 -*-
 
-# Form implementation generated from reading ui file 'src\designer\mainwindow.ui'
+# Form implementation generated from reading ui file 'designer\mainwindow.ui'
 #
 # Created by: PyQt5 UI code generator 5.11.3
 #
@@ -92,6 +92,12 @@ class Ui_MainWindow(object):
         self.loadRequestsButton = QtWidgets.QPushButton(self.setupGroupBox)
         self.loadRequestsButton.setObjectName("loadRequestsButton")
         self.setupForm.setWidget(1, QtWidgets.QFormLayout.FieldRole, self.loadRequestsButton)
+        self.holidaysLabel = QtWidgets.QLabel(self.setupGroupBox)
+        self.holidaysLabel.setObjectName("holidaysLabel")
+        self.setupForm.setWidget(2, QtWidgets.QFormLayout.LabelRole, self.holidaysLabel)
+        self.loadHolidaysButton = QtWidgets.QPushButton(self.setupGroupBox)
+        self.loadHolidaysButton.setObjectName("loadHolidaysButton")
+        self.setupForm.setWidget(2, QtWidgets.QFormLayout.FieldRole, self.loadHolidaysButton)
         self.calendarYearLabel = QtWidgets.QLabel(self.setupGroupBox)
         self.calendarYearLabel.setObjectName("calendarYearLabel")
         self.setupForm.setWidget(3, QtWidgets.QFormLayout.LabelRole, self.calendarYearLabel)
@@ -109,12 +115,9 @@ class Ui_MainWindow(object):
         self.numberOfBlocksSpinBox.setProperty("value", 26)
         self.numberOfBlocksSpinBox.setObjectName("numberOfBlocksSpinBox")
         self.setupForm.setWidget(4, QtWidgets.QFormLayout.FieldRole, self.numberOfBlocksSpinBox)
-        self.holidaysLabel = QtWidgets.QLabel(self.setupGroupBox)
-        self.holidaysLabel.setObjectName("holidaysLabel")
-        self.setupForm.setWidget(2, QtWidgets.QFormLayout.LabelRole, self.holidaysLabel)
-        self.loadHolidaysButton = QtWidgets.QPushButton(self.setupGroupBox)
-        self.loadHolidaysButton.setObjectName("loadHolidaysButton")
-        self.setupForm.setWidget(2, QtWidgets.QFormLayout.FieldRole, self.loadHolidaysButton)
[…]

### 2020-05-25 src/ui/mainwindow.py
Add MPS export option

@@ -3,6 +3,7 @@ import ast
 import json
 import re
 from datetime import datetime, timedelta
+from functools import partial
 
 from PyQt5.QtCore import *
 from PyQt5.QtGui import *
@@ -10,8 +11,8 @@ from PyQt5.QtWidgets import *
 
 from constants import WEEK_HOURS, WEEKEND_HOURS
 from helpers.excelhelper import ExcelHelper
-from helpers.uihelper import UiHelper
 from helpers.logger import Logger
+from helpers.uihelper import UiHelper
 from services import scheduler
 
 from .dialog import DialogWindow
@@ -85,7 +86,8 @@ class MainWindow(QMainWindow, Ui_MainWindow):
         self.generateScheduleButton.clicked.connect(self.generateSchedule)
         self.exportScheduleButton.clicked.connect(self.exportSchedule)
         self.exportMonthlyButton.clicked.connect(self.exportMonthlySchedule)
-        self.exportLpButton.clicked.connect(self.exportLpProblem)
+        self.exportLpButton.clicked.connect(partial(self.exportLpProblem, mps=False))
+        self.exportMpsButton.clicked.connect(partial(self.exportLpProblem, mps=True))
 
         self.calendarYearSpinBox.setValue(datetime.now().year + 1)
 
@@ -242,7 +244,7 @@ class MainWindow(QMainWindow, Ui_MainWindow):
             constraints=constraints
         )
         
-    def exportLpProblem(self):
+    def exportLpProblem(self, mps=False):
         shuffle = self.shuffleCheckBox.isChecked()
 
         scheduler = self.setupScheduler()
@@ -255,19 +257,28 @@ class MainWindow(QMainWindow, Ui_MainWindow):
 
         problem = scheduler.get_problem()
 
+        if mps:
+            dialogName = "Save MPS file"
+            dialogExt = "MPS File (*.mps)"
+            export = problem.writeMPS
+        else:
+            dialogName = "Save LP file"
+            dialogExt = "LP File (*.lp)"
+            export = problem.writeLP
+
         # choose save location
         path, _ = QFileDialog.getSaveFileName(
-            self, "Save LP file", "", "LP File (*.lp)", "LP File (*.lp)"
+            self, dialogName, "", dialogExt, dialogExt
         )
 
         if path != '':
             try:
-                problem.writeLP(path)
-                self._logger.write_line('Saved LP file: {}'.format(path))
+                export(path)
[…]

### 2020-05-25 src/ui/ui_mainwindow.py

@@ -11,7 +11,7 @@ from PyQt5 import QtCore, QtGui, QtWidgets
 class Ui_MainWindow(object):
     def setupUi(self, MainWindow):
         MainWindow.setObjectName("MainWindow")
-        MainWindow.resize(800, 600)
+        MainWindow.resize(800, 602)
         font = QtGui.QFont()
         font.setPointSize(10)
         MainWindow.setFont(font)
@@ -118,6 +118,9 @@ class Ui_MainWindow(object):
         self.exportLpButton = QtWidgets.QPushButton(self.setupGroupBox)
         self.exportLpButton.setObjectName("exportLpButton")
         self.setupForm.setWidget(5, QtWidgets.QFormLayout.LabelRole, self.exportLpButton)
+        self.exportMpsButton = QtWidgets.QPushButton(self.setupGroupBox)
+        self.exportMpsButton.setObjectName("exportMpsButton")
+        self.setupForm.setWidget(5, QtWidgets.QFormLayout.FieldRole, self.exportMpsButton)
         self.horizontalLayout_6.addLayout(self.setupForm)
         self.setupScheduleGroupBox.addWidget(self.setupGroupBox)
         self.scheduleActions = QtWidgets.QGroupBox(self.schedulerTab)
@@ -270,6 +273,7 @@ class Ui_MainWindow(object):
         self.calendarYearSpinBox.setToolTip(_translate("MainWindow", "<html><head/><body><p>Set the calendar year to be used by the scheduler</p></body></html>"))
         self.numberOfBlocksLabel.setText(_translate("MainWindow", "Number of Blocks"))
         self.exportLpButton.setText(_translate("MainWindow", "Export as LP"))
+        self.exportMpsButton.setText(_translate("MainWindow", "Export as MPS"))
         self.scheduleActions.setTitle(_translate("MainWindow", "Schedule"))
         self.shuffleCheckBox.setText(_translate("MainWindow", "Shuffle?"))
         self.generateScheduleButton.setToolTip(_translate("MainWindow", "<html><head/><body><p>Generate a new schedule using the given configuration file and calendar information</p></body></html>"))

### 2020-05-27 paper/paper.comparison.tex
Work on comparison/lit review section

In recent years, there has been a substantial amount of literature published in
the nurse rostering area aiming to organize the types of problems and approaches
commonly
addressed in the real-world. This research serves to highlight the similarities
and differences
among computationally easy and hard instances of nurse rostering, evaluating
which algorithms
are best geared towards which types of problems, and [...]

De Causmaecker and Berghe [ref] outlined a categorization
of nurse rostering problems using a similar approach used in personnel
scheduling,
based around the personnel environment, work characteristics and optimization
objective. [...]

[a few more papers on categorization, brief paragraph about each...]

We follow the work in [ref-De Causmaecker and Berghe] to place our clinician
scheduling problem
in the realm of work for nurse rostering. According to their specification, our
problem includes 
personnel environment constraints targeting the availability of clinicians
([list which constraints...]),
sequences of assignments ([...]), and balance of rosters ([...]) 
which fall under the categories $\alpha : A$, $\alpha : S$ and $\alpha : B$,
respectively. 
For work characteristics, the relevant category for our problem are the range
constraints, $\alpha : R$, 
as we restrict the workload of each clinician between a minimum and maximum
allowed blocks ([constraint...]).
Lastly, for the optimization objective our problem corresponds to the category
$\gamma : P$, as our optimization objective
aim to adhere to clinician preferences. Therefore the appropriate categorization
for our problem is $ASB/R/P$.

[outline and compare to papers with a similar categorization...]

### 2020-05-29 src/services/scheduler.py
Fix invalid LP variable names

@@ -62,7 +62,7 @@ class WeekendVariable(Variable):
         self.clinician = clinician
         self.week_num = week_num
 
-        Variable.__init__(self, '{},weekend:{}'.format(
+        Variable.__init__(self, '{},weekend_{}'.format(
             self.clinician.name, self.week_num
         ), 0, 1)
 
@@ -84,7 +84,7 @@ class BlockVariable(Variable):
         self.block_num = block_num
         self.division = division
 
-        Variable.__init__(self, '{},div:{},block:{}'.format(
+        Variable.__init__(self, '{},div_{},block_{}'.format(
             self.clinician.name, self.division, self.block_num
         ), 0, 1)
 
@@ -575,7 +575,7 @@ class Scheduler:
                         lambda x, w=week_num: x.week_num == w)[0].get_var()
 
                     var_ = Variable(
-                        '{}:adjacency,div:{},block:{}*weekend:{}'.format(
+                        '{}_adjacency,div_{},block_{}__weekend_{}'.format(
                             clinician.name,
                             div.name,
                             block_num,

### 2020-06-02 paper/paper.comparison.tex
Work on comparison section

@@
 
-De Causmaecker and Berghe [ref] outlined a categorization
-of nurse rostering problems using a similar approach used in personnel
-scheduling,
+De Causmaecker and Berghe [ref] outline a categorization
+of nurse rostering problems using a similar approach used in scheduling,
 based around the personnel environment, work characteristics and optimization
-objective. [...]
+objective. Previously, there was a lack of a central categorization for
+nurse rostering problems found in the literature. Their work attempts
+to address this, by creating a classification that takes into account
+hard and soft constraints related to nurse preferences, consecutive 
+assignments, coverage and various flexibilities in terms of shift and
+skill types. They go on to describe where recent examples of the nurse rostering
+literature can fit in their classification.
 
-[a few more papers on categorization, brief paragraph about each...]
+% DL - maybe this should be in complexity section?
+Vanhoucke and Maenhout [ref] analyze and compare nurse rostering instances
+based on complexity indicators arising from problem constraints. They consider
+problem size (in terms of number of nurses, shifts and days), preference distribution
+of the nurses, coverage distribution and time-related constraints (such as weekend
+and consecutive constraints) as the four indicators for problem complexity.
+Then, they generate sample problems corresponding to all four complexity indicators
+and measure the effect of increasing complexity on CPU time required for solving each
+problem.
+
+[maybe 1 additional paragraph on categorization?]
 
@@
 
-[outline and compare to papers with a similar categorization...]
+Using the classification in [ref-De Causmaecker and Berghe], we compare our clinician
+scheduling problem and solution approach with existing literature. Azaiez and Al Sharif [ref]
+describe a nurse rostering problem to schedule nurses on a per-day basis, for a total period
+of 28 days. In this problem, they are required to assign a minimum number of staff members
+per day, unlike in our clinician scheduling problem. Similar to our problem, they include
+hard constraints to prevent consecutive assignments, and enforce a minimum and maximum 
+number of shifts for each nurse. When formulating the optimization goal, Azaiez and Al Sharif
+focus on balancing the workload among nurses, preventing day-night consecutive shifts and avoiding
+off-on-off assignments. Their formulation does not incorporate nurse preferences
+for time off into the model.
+
+[...]

### 2020-06-03 paper/paper.comparison.tex
Outline Bilgin2008 paper in comparisons

@@
 describe a nurse rostering problem to schedule nurses on a per-day basis, for a total period
-of 28 days. In this problem, they are required to assign a minimum number of staff members
+of 28 days. De Causmaecker and Berghe classify this problem as $ASBN/V2/PL$.
+In this problem, they are required to assign a minimum number of staff members
 per day, unlike in our clinician scheduling problem. Similar to our problem, they include
@@
 
+In [ref-Bilgin 2008], Bilgin et al address a nurse rostering problem using local search. 
+The problem presented is classified as $ASBI/RVNO/PLR$.
+Some of the constraints they incorporate include coverage constraints, overlapping assignment
+constraints and horizontal constraints, which include constraints arising from employment contracts
+(such as satisfying minimum and maximum thresholds and limiting consecutive assignments). Unlike
+our problem, Bilgin et al consider coverage constraints to be soft constraints, in order to prevent
+over-constraining of the problem. They incorporate nurse requests into the model formulation,
+similar to our problem, but they allow requests for both time off and time on.
 [...]

### 2020-06-04 paper/paper.comparison.tex
Minor fixes to comparison section

@@
 and differences
-among computationally easy and hard instances of nurse rostering, evaluating
+among computationally easy and hard instances of nurse rostering and evaluating
 which algorithms
-are best geared towards which types of problems, and [...]
+are best geared towards which types of problems thus allowing practitioners
+to choose which approach will work best for their case.
 
 De Causmaecker and Berghe [ref] outline a categorization
-of nurse rostering problems using a similar approach used in scheduling,
-based around the personnel environment, work characteristics and optimization
-objective. Previously, there was a lack of a central categorization for
-nurse rostering problems found in the literature. Their work attempts
-to address this, by creating a classification that takes into account
+of nurse rostering problems using a similar approach used in theoretic scheduling problems,
+based around the machine environment, job characteristics and constraints,
+and optimization objective. Previously, there was a lack of a central categorization system
+for nurse rostering problems found in the literature. Their work attempts
+to address this by creating a classification that takes into account
 hard and soft constraints related to nurse preferences, consecutive 
@@
 
-[maybe 1 additional paragraph on categorization?]
+[maybe 1 additional paragraph on categorization systems?]
 
 We follow the work in [ref-De Causmaecker and Berghe] to place our clinician
-scheduling problem
-in the realm of work for nurse rostering. According to their specification, our
-problem includes 
+scheduling problem in the realm of work for nurse rostering. 
+According to their specification, our problem includes 
 personnel environment constraints targeting the availability of clinicians
@@
 For work characteristics, the relevant category for our problem are the range
-constraints, $\alpha : R$, 
+constraints, $\beta : R$, 
 as we restrict the workload of each clinician between a minimum and maximum
@@
 similar to our problem, but they allow requests for both time off and time on.
+
 [...]

### 2020-07-06 paper/paper.comparison.tex
SM edits to comparison section

@@
-In recent years, there has been a substantial amount of literature published in
-the nurse rostering area aiming to organize the types of problems and approaches
-commonly
-addressed in the real-world. This research serves to highlight the similarities
+Prior work by [authors] have   %avoid editorial type statements without citations or numbers, best to state as objective statements
+provided categories to organize the types of problems and approaches
+commonly used for real-world nurse rostering.
+These categories highlight the similarities  %SM what does "this" refer to? this paper or the prior literature? [always use a concrete noun after "this or these" to help the reader :)
 and differences
-among computationally easy and hard instances of nurse rostering and evaluating
+among computationally easy and hard instances of nurse rostering and 
+help end-users determine
 which algorithms
-are best geared towards which types of problems thus allowing practitioners
-to choose which approach will work best for their case.
+are best suited to each type of problem.
 
@@
 assignments, coverage and various flexibilities in terms of shift and
-skill types. They go on to describe where recent examples of the nurse rostering
-literature can fit in their classification.
+skill types. De Causmaecker and Berghe go on to describe where recent examples of the nurse rostering
+literature can fit in their classification. %SM: such as?
 

### 2020-07-06 paper/paper.comparison.tex
Update paper.comparison.tex

@@
 
-De Causmaecker and Berghe [ref] outline a categorization
-of nurse rostering problems using a similar approach used in theoretic scheduling problems,
+In XXXyear, De Causmaecker and Berghe [ref] addressed a gap in the literature 
+surrounding a central categorization system for the nurse rostering problems identified in the literature.
+The authors categorized
+nurse rostering problems using a similar approach used in theoretic scheduling problems,
 based around the machine environment, job characteristics and constraints,
-and optimization objective. Previously, there was a lack of a central categorization system
-for nurse rostering problems found in the literature. Their work attempts
-to address this by creating a classification that takes into account
+and optimization objective. They also created a classification that takes into account
 hard and soft constraints related to nurse preferences, consecutive 
 assignments, coverage and various flexibilities in terms of shift and
-skill types. De Causmaecker and Berghe go on to describe where recent examples of the nurse rostering
-literature can fit in their classification. %SM: such as?
+skill types. De Causmaecker and Berghe went on to describe where recent examples of the nurse rostering
+literature fit into their classification. %SM: give examples of their recent examples.
 
-% DL - maybe this should be in complexity section?
+% DL - maybe this should be in complexity section? %SM: agree, move to complexity section.
 Vanhoucke and Maenhout [ref] analyze and compare nurse rostering instances
@@
 
-[maybe 1 additional paragraph on categorization systems?]
+[maybe 1 additional paragraph on categorization systems?] %SM: yes, describe a bit more or a table/box as the list of categories.
+That way - flows into next paragraph on where to place the SMH solution.
+%SM i.e. describe/define each category
 
-We follow the work in [ref-De Causmaecker and Berghe] to place our clinician
-scheduling problem in the realm of work for nurse rostering. 
-According to their specification, our problem includes 
+We follow the categorization by [ref-De Causmaecker and Berghe] to situate our clinician
+scheduling problem. Our problem includes 
 personnel environment constraints targeting the availability of clinicians
@@
 sequences of assignments ([...]), and balance of rosters ([...]) 
-which fall under the categories $\alpha : A$, $\alpha : S$ and $\alpha : B$,
+which fall under the categories $\alpha : A$, $\alpha : S$ and $\alpha : B$,  %are these the categories? think other way around -- these fall into the named categories from De Causeameker, etc.?
 respectively. 
@@
 
-Using the classification in [ref-De Causmaecker and Berghe], we compare our clinician
+Using the classification in [ref-De Causmaecker and Berghe], we compare our clinician  %use the term classification or categorization, but not both
 scheduling problem and solution approach with existing literature. Azaiez and Al Sharif [ref]
@@
 In this problem, they are required to assign a minimum number of staff members
-per day, unlike in our clinician scheduling problem. Similar to our problem, they include
+per day, unlike in our clinician scheduling problem. Similar to our problem, they include  %who are 'they'? 
 hard constraints to prevent consecutive assignments, and enforce a minimum and maximum 
@@
 In [ref-Bilgin 2008], Bilgin et al address a nurse rostering problem using local search. 
-The problem presented is classified as $ASBI/RVNO/PLR$.
-Some of the constraints they incorporate include coverage constraints, overlapping assignment
+The problem presented is classified as $ASBI/RVNO/PLR$.  %problem presented by whom?
+Some of the constraints they incorporate include coverage constraints, overlapping assignment  %who are 'they'?
 constraints and horizontal constraints, which include constraints arising from employment contracts
@@
 our problem, Bilgin et al consider coverage constraints to be soft constraints, in order to prevent
-over-constraining of the problem. They incorporate nurse requests into the model formulation,
-similar to our problem, but they allow requests for both time off and time on.
+over-constraining of the problem. They incorporate nurse requests into the model formulation,  %SM: tell us who 'they' are :) 
+similar to our problem, but they allow requests for both time off and time on.   %rephrase for clarity - not sure I followed?
 

### 2020-07-19 paper/paper.comparison.tex
Clarify and cleanup comparison section

--- 2020-07-06 paper/paper.comparison.tex
@@
 commonly used for real-world nurse rostering.
-These categories highlight the similarities  %SM what does "this" refer to? this paper or the prior literature? [always use a concrete noun after "this or these" to help the reader :)
+These categories highlight the similarities
 and differences
@@
 
-De Causmaecker and Berghe [ref] outline a categorization
-of nurse rostering problems using a similar approach used in theoretic scheduling problems,
-based around the machine environment, job characteristics and constraints,
-and optimization objective. Previously, there was a lack of a central categorization system
-for nurse rostering problems found in the literature. Their work attempts
-to address this by creating a classification that takes into account
+In 2010, De Causmaecker and Berghe [ref] addressed a gap in the literature 
+surrounding a central categorization system for the nurse rostering problems identified in the literature.
+The authors created a categorization framework that takes into account
 hard and soft constraints related to nurse preferences, consecutive
 assignments, coverage and various flexibilities in terms of shift and 
-skill types. De Causmaecker and Berghe go on to describe where recent examples of the nurse rostering
-literature can fit in their classification. %SM: such as?
+skill types. The described framework is inspired by the categorization used
+in theoretic scheduling problems which focuses on three main aspects of scheduling:
+(1) the machine environment;
+(2) the job characteristics;
+and (3) the objective function.
+De Causmaecker and Berghe use these three aspects as a basis for their categorization,
+and adjust them to fit the realm of nurse rostering. A description of the categories defined
+by the authors is given in Table \ref{tbl:nurse-rostering-categorization}.
 
-% DL - maybe this should be in complexity section?
-Vanhoucke and Maenhout [ref] analyze and compare nurse rostering instances
-based on complexity indicators arising from problem constraints. They consider
-problem size (in terms of number of nurses, shifts and days), preference distribution
-of the nurses, coverage distribution and time-related constraints (such as weekend
-and consecutive constraints) as the four indicators for problem complexity.
-Then, they generate sample problems corresponding to all four complexity indicators
-and measure the effect of increasing complexity on CPU time required for solving each
-problem.
+\begin{table}[h]
+	\centering
+	\caption{Description of De Causmaecker and Berghe's categorization system for nurse rostering problems.}%
+	\label{tbl:nurse-rostering-categorization}
+	\begin{tabular}{l|l|l}
+		Category                                          & \multicolumn{2}{l}{Sub-categories}                                                                                                                                                         \\ \hline
+		\multirow{2}{*}{$\alpha$: Personnel environment}  & Personnel Constraints                                                                                      & Skill Interactions                                                            \\ \cline{2-3}
+		                                                  & \makecell[l]{(A) Availability \\ (S) Sequences \\ (B) Balance \\ (C) Chaperoning}                                          & \makecell[l]{Fixed number \\ Variable number \\ Individual skill definitions} \\ \hline
+		\multirow{2}{*}{$\beta$: Work characteristics}    & Coverage Constraints                                                                                       & Shift Type                                                                    \\ \cline{2-3}
+		                                                  & \makecell[l]{(R) Range \\ (T) Time intervals \\ (V) Fluctuating}                                                       & \makecell[l]{Fixed number \\ Variable number \\ Overlapping}                  \\ \hline
+		\multirow{2}{*}{$\gamma$: Optimization objective} & Objective                                                                                                  & Mode                                                                          \\ \cline{2-3}
+		                                                  & \makecell[l]{(P) Personnel constraints \\ (L) Coverage constraints\\ (X) Number of personnel \\ (R) Robustness \\ (G) General} & \makecell[l]{Multi-objective}                                                 \\ \hline
+	\end{tabular}
+\end{table}
 
-[maybe 1 additional paragraph on categorization systems?]
+%The authors categorized
+%nurse rostering problems using a similar approach used in theoretic scheduling problems,
+%based around the machine environment, job characteristics and constraints,
+%and optimization objective. They also created a classification that takes into account
+%hard and soft constraints related to nurse preferences, consecutive 
+%assignments, coverage and various flexibilities in terms of shift and
+%skill types.
 
-We follow the work in [ref-De Causmaecker and Berghe] to place our clinician
-scheduling problem in the realm of work for nurse rostering. 
-According to their specification, our problem includes 
-personnel environment constraints targeting the availability of clinicians
-([list which constraints...]),
-sequences of assignments ([...]), and balance of rosters ([...]) 
-which fall under the categories $\alpha : A$, $\alpha : S$ and $\alpha : B$,
+%De Causmaecker and Berghe went on to describe where recent examples of the nurse rostering
+%literature fit into their classification. %SM: give examples of their recent examples.
+
+% DL - maybe this should be in complexity section? %SM: agree, move to complexity section.
+%Vanhoucke and Maenhout [ref] analyze and compare nurse rostering instances
+%based on complexity indicators arising from problem constraints. They consider
+%problem size (in terms of number of nurses, shifts and days), preference distribution
+%of the nurses, coverage distribution and time-related constraints (such as weekend
+%and consecutive constraints) as the four indicators for problem complexity.
+%Then, they generate sample problems corresponding to all four complexity indicators
+%and measure the effect of increasing complexity on CPU time required for solving each
+%problem.
+
+%[maybe 1 additional paragraph on categorization systems?] %SM: yes, describe a bit more or a table/box as the list of categories.
+That way - flows into next paragraph on where to place the SMH solution.
+%SM i.e. describe/define each category
+
+We follow the categorization in Table \ref{tbl:nurse-rostering-categorization} to situate our clinician
+scheduling problem. Our problem includes 
+personnel environment constraints targeting the availability and preferences of clinicians
+(constraints BR and WR in Table \ref{tbl:constraint-summary}),
+sequences of assignments (constraints NCB and NCW), and balance of rosters (constraints EW and EH) 
+which fall under the categories $\alpha : A$, $\alpha : S$ and $\alpha : B$,  %are these the categories? think other way around -- these fall into the named categories from De Causeameker, etc.?
 respectively. 
@@
 as we restrict the workload of each clinician between a minimum and maximum
-allowed blocks ([constraint...]).
+allowed blocks (constraint MM).
 Lastly, for the optimization objective our problem corresponds to the category
 $\gamma : P$, as our optimization objective
-aim to adhere to clinician preferences. Therefore the appropriate categorization
+aims to adhere to clinician preferences. Therefore the appropriate categorization
 for our problem is $ASB/R/P$.
 
-Using the classification in [ref-De Causmaecker and Berghe], we compare our clinician
+De Causmaecker and Berghe also present examples from the nurse rostering literature
+and their appropriate categorization in their framework. We use some of these examples
+to compare our clinician
 scheduling problem and solution approach with existing literature. Azaiez and Al Sharif [ref]
 describe a nurse rostering problem to schedule nurses on a per-day basis, for a total period
-of 28 days. De Causmaecker and Berghe classify this problem as $ASBN/V2/PL$.
-In this problem, they are required to assign a minimum number of staff members
-per day, unlike in our clinician scheduling problem. Similar to our problem, they include
-hard constraints to prevent consecutive assignments, and enforce a minimum and maximum 
-number of shifts for each nurse. When formulating the optimization goal, Azaiez and Al Sharif
-focus on balancing the workload among nurses, preventing day-night consecutive shifts and avoiding
-off-on-off assignments. Their formulation does not incorporate nurse preferences
+of 28 days. Azaiez and Al Sharif's problem is categorized as $ASBN/V2/PL$.
+Similar to our clinician scheduling problem, the authors include hard constraints
+to prevent consecutive assignments, and enforce a minimum and maximum number of shifts for each nurse.
+Unlike our problem, Azaiez and Al Sharif choose an optimization goal that focuses on
+balancing the workload among nurses, preventing day-night consecutive shifts and 
+avoiding off-on-off assignments. The authors do not attempt to incorporate nurse preferences
 for time off into the model.
@@
 In [ref-Bilgin 2008], Bilgin et al address a nurse rostering problem using local search.
-The problem presented is classified as $ASBI/RVNO/PLR$.
-Some of the constraints they incorporate include coverage constraints, overlapping assignment
-constraints and horizontal constraints, which include constraints arising from employment contracts
-(such as satisfying minimum and maximum thresholds and limiting consecutive assignments). Unlike
-our problem, Bilgin et al consider coverage constraints to be soft constraints, in order to prevent
-over-constraining of the problem. They incorporate nurse requests into the model formulation,
-similar to our problem, but they allow requests for both time off and time on.
+De Causmaecker and Berghe categorize the problem in [ref-Bilgin] as $ASBI/RVNO/PLR$.
+Bilgin et al incorporate coverage constraints, overlapping assignment constraints
+and horizontal constraints, which include constraints arising from employment contracts
+(such as minimum and maximum thresholds).
+Unlike our approach for these constraints, the authors consider coverage constraints to be soft
+constraints, in order to prevent over-constraining of the problem.
+Bilgin et al also incorporate nurse requests into the model formulation while allowing
+nurses to request request days off and days on.
 

### 2020-07-21 paper/paper.comparison.tex
Add another paper for comparison + minor fixes

@@
-Prior work by [authors] have   %avoid editorial type statements without citations or numbers, best to state as objective statements
-provided categories to organize the types of problems and approaches
-commonly used for real-world nurse rostering.
-These categories highlight the similarities
-and differences
-among computationally easy and hard instances of nurse rostering and 
-help end-users determine
-which algorithms
-are best suited to each type of problem.
+%Prior work by [authors] have   %avoid editorial type statements without citations or numbers, best to state as objective statements
+%provided categories to organize the types of problems and approaches
+%commonly used for real-world nurse rostering.
+%These categories highlight the similarities
+%and differences
+%among computationally easy and hard instances of nurse rostering and 
+%help end-users determine
+%which algorithms
+%are best suited to each type of problem.
 
-In 2010, De Causmaecker and Berghe [ref] addressed a gap in the literature 
+In 2010, De Causmaecker and Berghe~\cite{de_causmaecker_categorisation_2011} addressed a gap in the literature 
 surrounding a central categorization system for the nurse rostering problems identified in the literature.
@@
 
-%The authors categorized
-%nurse rostering problems using a similar approach used in theoretic scheduling problems,
-%based around the machine environment, job characteristics and constraints,
-%and optimization objective. They also created a classification that takes into account
-%hard and soft constraints related to nurse preferences, consecutive 
-%assignments, coverage and various flexibilities in terms of shift and
-%skill types.
-
-%De Causmaecker and Berghe went on to describe where recent examples of the nurse rostering
-%literature fit into their classification. %SM: give examples of their recent examples.
-
 % DL - maybe this should be in complexity section? %SM: agree, move to complexity section.
@@
 %problem.
-
-%[maybe 1 additional paragraph on categorization systems?] %SM: yes, describe a bit more or a table/box as the list of categories.
-That way - flows into next paragraph on where to place the SMH solution.
-%SM i.e. describe/define each category
 
@@
 to compare our clinician
-scheduling problem and solution approach with existing literature. Azaiez and Al Sharif [ref]
+scheduling problem and solution approach with existing literature. Azaiez and Al Sharif~\cite{azaiez_0-1_2005}
 describe a nurse rostering problem to schedule nurses on a per-day basis, for a total period
@@
 
-In [ref-Bilgin 2008], Bilgin et al address a nurse rostering problem using local search.
-De Causmaecker and Berghe categorize the problem in [ref-Bilgin] as $ASBI/RVNO/PLR$.
+In~\cite{bilgin_local_2012}, Bilgin et al address a nurse rostering problem using local search.
+De Causmaecker and Berghe categorize the problem in~\cite{bilgin_local_2012} as $ASBI/RVNO/PLR$.
 Bilgin et al incorporate coverage constraints, overlapping assignment constraints
@@
 constraints, in order to prevent over-constraining of the problem.
-Bilgin et al also incorporate nurse requests into the model formulation while allowing
-nurses to request request days off and days on.
+Bilgin et al also incorporate nurse requests into the model formulation by allowing
+nurses to request days off and days on.
 
-[...]
+Burke et al~\cite{burke_metaheuristics_2006} present a flexible framework for nurse rostering based on various approaches
+observed from Belgin hospitals. 
+In~\cite{de_causmaecker_categorisation_2011}, Burke et al's work is categorized as $ASBCI/RVNTO/PR$.
+The authors use a two-step algorithm comprised of an initialization procedure
+which provides a solution satisfying only the hard constraints of their problem,
+and a metaheuristic procedure, a hybrid tabu search, which aim to satisfy as many soft constraints as possible.
+Unlike our clinician scheduling problem, 
+Burke et al's hard constraints address only coverage requirements in an effort to make the first
+step of their algorithm efficient.
+In their metaheurstic procedure, the authors optimize soft constraints including assignment of free weekends
+and minimum number of assignments per staff member.
+Notably, Burke et al's approach make use of ``time interval personnel requirements'', which describe
+the amount of required staff per skill category as a varying number throughout the work day.

### 2020-07-21 paper/paper.intro.tex
Add another paper for comparison + minor fixes

@@
 We begin by describing the details of the problem in Section~\ref{sec:problem}, and presenting
-our ILP formulation in Section~\ref{sec:methods}. Next, we compare the results
-of our formulation to manually-created schedules, and evaluate the performance
+our ILP formulation in Section~\ref{sec:methods}. We then categorize our nurse rostering
+problem and solution approach in the categorization scheme of~\cite{de_causmaecker_categorisation_2011}
+and compare our problem to recent studies in Section~\ref{sec:comparison}.
+Next, we analyze the effectiveness of our automated approach when compared
+to manually-created schedules, and evaluate the performance
 of the algorithm on simulated data in Section~\ref{sec:experiments}. Finally, we

### 2020-07-21 paper/paper.intro.tex
Address minor comments from reviewers

@@
 transportation~\cite{aickelin_improved_2006, goel_truck_2012, gunther_combined_2010},
-manufacturing~\cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008, alfares_simulation_2007},
+%manufacturing~\cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008, alfares_simulation_2007},
 retail~\cite{chapados_retail_2011, nissen_automatic_2010}, and

### 2020-07-21 paper/paper.methods.tex
Address minor comments from reviewers

@@
 \begin{align}
-&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1,
-s\in\mathcal{S} \label{eqn:constr-no-consec-blocks} \\
+&\sum_{s=1}^S \left(X_{c, b, s} + X_{c, b + 1, s}\right) \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1 \label{eqn:constr-no-consec-blocks} \\
 &Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \label{eqn:constr-no-consec-weekends}
@@
 
-Currently, the most efficient approach to finding an exact solution for 
-an ILP is called Branch-and-Cut~\cite{mitchell_branch-and-cut_2002}.   %SM: good
+One of the efficient approaches to finding an exact solution for an 
+ILP is Branch-and-Cut~\cite{mitchell_branch-and-cut_2002}.
 This method involves iteratively solving 

### 2021-01-27 paper/paper.discussion.tex
Work on new 'Practical Considerations' section

@@
 that the algorithm can accommodate an increase in time-horizon up to four years
-with little impact on runtime, suggesting the algorithm can be used generate
+with little impact on runtime, suggesting the algorithm can be used to generate
 schedules far in advance. A key limitation we identified was the sensitivity

### 2021-01-27 paper/paper.practical.tex
Work on new 'Practical Considerations' section

In this section we discuss the implications of our approach as deployed in the hospital,
as well as the lessons learned from developing the scheduling tool.
The tool is used by administrative staff in the division in order to generate a potential schedule
for the ID and HIV departments. 
The staff provides the following data as input to the scheduler:
(1) maximum and minimum required number of blocks for each clinician, for each division;
(2) a set of time-off requests for each clinician;
(3) a set of holiday dates for the current calendar year.
The supplied time-off requests in (2) are mapped to an appropriate block time-off request by the scheduler.
For example, if a clinician requests time-off during weeks 2 and 3 of the year, the scheduler
will interpret the request as time-off during block 1 (covering weeks 1-2) and 2 (covering weeks 3-4).
The scheduler allows the user to disable certain constraints when generating a schedule
in order to allow a broader range of solutions and prevent issues with infeasibility. 
It also allows the user to generate multiple schedules by randomly shuffling the input list of clinicians.
Once a candidate schedule has been created, the department head will manually adjust it based on
additional [qualitative-?] constraints to create a final schedule for the division.
Since the development of the tool, it has been used to generate schedules for the years 2019-21.

Figure XX presents a comparison of the generated schedule and the final schedule after manual adjustments
for the year 20XX. We can see...

[TODO-lessons learned]

### 2021-01-27 paper/paper.tex
Work on comparison/lit review section
Work on comparison section
Add conflict of interest statement
Work on new 'Practical Considerations' section

@@
 \input{paper.includes.tex}
+\addbibresource{references.bib}
 
-\title{A flexible integer linear programming formulation for scheduling
-	clinician on-call service in hospitals}
+%\title{A flexible integer linear programming formulation for scheduling
+%	clinician on-call service in hospitals}
+\title{
+	Application of integer linear programming to scheduling clinician
+	on-call service in Toronto hospital: case study
+}
 \date{}
@@
 	\input{paper.intro.tex}
-	\section{Problem}\label{sec:problem}
+	\section{Problem Description}\label{sec:problem}
 	\input{paper.problem.tex}
-	\section{Methods}\label{sec:methods}
+	\section{Mathematical Formulation}\label{sec:methods}
 	\input{paper.methods.tex}
-	\section{Experiments}\label{sec:experiments}
+	\section{Categorization and Comparison with Other Models}\label{sec:comparison}
+	\input{paper.comparison.tex}
+	\section{Computational Experiments}\label{sec:experiments}
 	\input{paper.experiments.tex}
+	\section{Practical Considerations}\label{sec:practical}
+	\input{paper.practical.tex}
 	\section{Discussion}\label{sec:discussion}
@@
 	\input{paper.funding.tex}
+	\section{Conflict of Interest}
+	On behalf of all authors, the corresponding author states that there is no conflict of interest.
 
-	\bibliographystyle{unsrt}
-	\bibliography{references}
+	\printbibliography
 \end{document}

### 2021-02-01 paper/paper.practical.tex
Work on lessons learned paragraph in sec6

@@
 Since the development of the tool, it has been used to generate schedules for the years 2019-21.
+Table XX presents a comparison of the generated schedule and the final schedule after manual adjustments
+for the year 2019. We see that the manual adjustments were likely as a result of additional clinicians
+and/or fellows being assigned to cover a few of the blocks, as well as certain blocks that were 
+split into individual weeks. Overall, the generated schedule provides a good global template,
+and local adjustments can be performed post-hoc.
 
-Figure XX presents a comparison of the generated schedule and the final schedule after manual adjustments
-for the year 20XX. We can see...
-
-[TODO-lessons learned]
+The development of the scheduling solution started from a relaxed version of the problem
+(without constraints on consecutive blocks/weekends).
+We first investigated a maximum flow formulation for solving the problem.
+In this formulation, the problem is represented as a graph $G=(V, E)$ where
+the vertices comprise clinicians and blocks/weekends, and the edges comprise
+potential assignments of clinician to a block/weekend.
+A similar approach using maximum flow has been used in [XX, XX, XX].
+The advantages of a maximum flow formulation are in the computational efficiency
+of the algorithm and the interpretability of the constraints and solution in terms of the graph [???].
+However, a max flow formulation makes it difficult to incorporate new constraints.
+For instance, in order to accommodate the NCB and NCW constraints,
+additional edges and nodes need to be added to the graph, potentially reducing 
+the computational efficiency and interpretability.
+As a result, we opted to reformulate the problem as an ILP, in which the additional constraints
+can be incorporated more easily.
+[...]

### 2021-02-02 paper/paper.practical.tex
Work on practical section

@@
 can be incorporated more easily.
-[...]
+
+The ILP formulation went through several stages of design and tweaking.
+For instance, we originally aimed to optimize only the BR and WR constraints.
+However, after receiving feedback on some of the test schedules, we learned that 
+clinicians would prefer to have some of their weekend assignments closer to the block assignments.
+Once we were satisfied with the schedules produced by the algorithm, we targeted the
+user interface. Our goal was to develop an easy-to-use interface for the administrative staff to use,
+without requiring knowledge of the algorithm. [DL - do we need to elaborate more on this?]

### 2021-02-05 paper/paper.practical.tex
Rearrange section 6 + add references

@@
 as well as the lessons learned from developing the scheduling tool.
-The tool is used by administrative staff in the division in order to generate a potential schedule
-for the ID and HIV departments. 
-The staff provides the following data as input to the scheduler:
-(1) maximum and minimum required number of blocks for each clinician, for each division;
-(2) a set of time-off requests for each clinician;
-(3) a set of holiday dates for the current calendar year.
-The supplied time-off requests in (2) are mapped to an appropriate block time-off request by the scheduler.
-For example, if a clinician requests time-off during weeks 2 and 3 of the year, the scheduler
-will interpret the request as time-off during block 1 (covering weeks 1-2) and 2 (covering weeks 3-4).
-The scheduler allows the user to disable certain constraints when generating a schedule
-in order to allow a broader range of solutions and prevent issues with infeasibility. 
-It also allows the user to generate multiple schedules by randomly shuffling the input list of clinicians.
-Once a candidate schedule has been created, the department head will manually adjust it based on
-additional [qualitative-?] constraints to create a final schedule for the division.
-Since the development of the tool, it has been used to generate schedules for the years 2019-21.
-Table XX presents a comparison of the generated schedule and the final schedule after manual adjustments
-for the year 2019. We see that the manual adjustments were likely as a result of additional clinicians
-and/or fellows being assigned to cover a few of the blocks, as well as certain blocks that were 
-split into individual weeks. Overall, the generated schedule provides a good global template,
-and local adjustments can be performed post-hoc.
 
@@
 potential assignments of clinician to a block/weekend.
-A similar approach using maximum flow has been used in [XX, XX, XX].
+A similar approach using maximum flow has been used in~\cite{brucker_personnel_2011, smet_polynomially_2016, el_adoly_new_2018, moz_integer_2003}.
 The advantages of a maximum flow formulation are in the computational efficiency
@@
 without requiring knowledge of the algorithm. [DL - do we need to elaborate more on this?]
+
+Since the development of the tool, it has been used to generate schedules for the years 2019-21.
+The tool is used by administrative staff in the division in order to generate a potential schedule
+for the ID and HIV departments. 
+The staff provides the following data as input to the scheduler:
+(1) maximum and minimum required number of blocks for each clinician, for each division;
+(2) a set of time-off requests for each clinician;
+(3) a set of holiday dates for the current calendar year.
+The supplied time-off requests in (2) are mapped to an appropriate block time-off request by the scheduler.
+For example, if a clinician requests time-off during weeks 2 and 3 of the year, the scheduler
+will interpret the request as time-off during block 1 (covering weeks 1-2) and 2 (covering weeks 3-4).
+The scheduler allows the user to disable certain constraints when generating a schedule
+in order to allow a broader range of solutions and prevent issues with infeasibility. 
+It also allows the user to generate multiple schedules by randomly shuffling the input list of clinicians.
+Once a candidate schedule has been created, the department head may manually
+adjust the schedule to better accommodate some of the time-off requests
+and any potential [extraneous/qualitative-?] [requirements-?]
+(although at the cost of breaking some of the hard constraints).
+Table XX presents a comparison of the generated schedule and the final schedule after manual adjustments
+for the year 2019. We see that the manual adjustments were likely as a result of additional clinicians
+and/or fellows being assigned to cover a few of the blocks, as well as certain blocks that were 
+split into individual weeks. Overall, the generated schedule provides a good global template,
+and local adjustments can be performed post-hoc.

### 2021-02-10 paper/paper.practical.tex
Add 2019 comparisons and reword parts of sec 6

@@
-In this section we discuss the implications of our approach as deployed in the hospital,
-as well as the lessons learned from developing the scheduling tool.
-
+In this section we discuss the lessons learned from developing the scheduling tool,
+as well as the implications of our approach as deployed in the hospital.
 The development of the scheduling solution started from a relaxed version of the problem
@@
 
-Since the development of the tool, it has been used to generate schedules for the years 2019-21.
+Since the development of the tool, it has been used at St. Michael's Hospital
+to generate schedules for the years 2019-21.
 The tool is used by administrative staff in the division in order to generate a potential schedule
@@
 adjust the schedule to better accommodate some of the time-off requests
-and any potential [extraneous/qualitative-?] [requirements-?]
+and any additional requirements
 (although at the cost of breaking some of the hard constraints).
-Table XX presents a comparison of the generated schedule and the final schedule after manual adjustments
-for the year 2019. We see that the manual adjustments were likely as a result of additional clinicians
-and/or fellows being assigned to cover a few of the blocks, as well as certain blocks that were 
-split into individual weeks. Overall, the generated schedule provides a good global template,
-and local adjustments can be performed post-hoc.
+Table \ref{tbl:2019-schedule-comparison} presents a comparison of the generated schedule and the final schedule after manual adjustments
+for the year 2019. We see that the manual adjustments included splitting or merging of certain blocks
+and local swaps between clinicians.
+Additionally, blocks and weekends may be adjusted manually due to newly hired clinicians or fellows being 
+assigned to cover some of the assignments.
+For the year 2020, the generated schedules provided a good starting point for the schedule.
+However, due to the COVID-19 pandemic and an increase in the demand and workload of ID clinicians in the department,
+the schedules had to be adjusted manually as new requirements emerged.
+This is something that could be improved in future iterations of the model
+-- specifically, the ability to (re-)generate shorter term schedules which are appropriately
+tuned based on a partial completed schedule provided as input.
+Overall, the generated schedule from the ILP model provide a good global template
+to which local adjustments can be easily performed post-hoc.
+
+\input{tbl/2019_schedule_comparison}

### 2021-02-10 paper/tbl/2018_schedule_comparison.tex
Cleanup latex comments
Add 2019 comparisons and reword parts of sec 6

@@
-% Table generated by Excel2LaTeX from sheet '2018_oldLP'
 \begin{table}[htbp]
@@
   \label{tbl:2018-schedule-comparison}%
-  % JK: could you clarify in the caption what we should take away from this table?
 	\begin{adjustbox}{scale=0.8}
@@
     	\multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} &                                                                       \multicolumn{3}{c||}{ILP Solution}                                                                        &                                                                      \multicolumn{3}{c}{Manual Generation}                                                                      \\
-          & HIV   & ID    & Weekend & HIV   & ID    & Weekend \\
-    \midrule
-    \midrule
+    	                                                  &                  HIV                  &                                 ID                                 &                              Weekend                               &                  HIV                  &                                 ID                                 &                              Weekend                               \\ \midrule\midrule
     	                        1                         & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .663,  .816,  .557}A &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
@@
     	                       50                         & \cellcolor[rgb]{ .788,  .788,  .788}D &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
-    51    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .957,  .69,  .518}E \\
+    	                       51                         & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .608,  .761,  .902}B &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E
     \end{tabular}%
@@
 Different colours and letters are used to distinguish different clinicians.
-  % JK: should use a footnote to define the meaning of
-  %     - HIV / ID
-  %     - Weekend
-  %     - Colors & Letters
-  %     so that the reader can understand the table basically without reading the paper.
 \end{table}%

### 2021-02-10 paper/tbl/2019_schedule_comparison.tex
Add 2019 comparisons and reword parts of sec 6

\begin{table}[htbp]
  \centering
  \caption{Comparison of the automatically generated (ILP solution) schedule and the final schedule after manual adjustments for 2019.}%
  \label{tbl:2019-schedule-comparison}%
	\begin{adjustbox}{scale=0.8}
	\begin{tabular}{c||ccc||ccc}
		\multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} &                                                                                      \multicolumn{3}{c||}{ILP Solution}                                                                                      &                                                                                    \multicolumn{3}{c}{Manual Adjustment}                                                                                     \\
		                                                  &                                HIV                                 &                                 ID                                 &                              Weekend                               &                                HIV                                 &                                 ID                                 &                              Weekend                               \\ \midrule\midrule
		                        1                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                        2                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                        3                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                        4                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                        5                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                        6                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                        7                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                        8                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                        9                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                       10                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       11                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       12                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       13                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       14                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       15                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       16                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ 1,  .6,  .8}J                   \\
		                       17                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                       18                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       19                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       20                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       21                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       22                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       23                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       24                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       25                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       26                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       27                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       28                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                       29                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       30                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ 1,  .6,  .8}J                   \\
		                       31                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       32                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       33                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       34                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       35                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       36                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                       37                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       38                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       39                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       40                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       41                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       42                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       43                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       44                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                       45                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       46                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                       47                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       48                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       49                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                       50                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       51                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .459,  .443,  .443}H
	\end{tabular}%
\end{adjustbox}\\[1em]
  \label{tab:addlabel}%
\end{table}%
